---
title: "Data Flow"
---

```{r}
#| include: false
library(tabviz)
```

Trace one row of a `data.frame` from R into the rendered Svelte widget. This page explains how column specs resolve, when formula expressions evaluate, and how the result becomes JSON for the frontend.

## The Pipeline

```{mermaid}
flowchart LR
    subgraph Input
        A["data.frame"]
    end

    subgraph Processing
        B["Column Resolution"]
        C["Style Evaluation"]
        D["JSON Serialization"]
    end

    subgraph Output
        E["Svelte Component"]
        F["Interactive Widget"]
    end

    A --> B --> C --> D --> E --> F

    style D fill:#4f46e5,color:#fff
    style E fill:#ff3e00,color:#fff
```

### Stage 1: Column Resolution

tabviz maps your column specifications to data columns:

```{r}
#| eval: false
data <- data.frame(
  study = c("Trial A", "Trial B"),
  hr = c(0.72, 0.85),
  lo = c(0.55, 0.70),
  hi = c(0.95, 1.03)
)

tabviz(data,
  label = "study",  # Maps to data$study
  columns = list(
    viz_forest(point = "hr", lower = "lo", upper = "hi")  # Maps to data$hr, data$lo, data$hi
  )
)
```

At this stage:
- String arguments become column lookups
- Missing columns trigger errors
- Data types are validated

### Stage 2: Style Evaluation

Formula expressions are evaluated against your data:

```{r}
#| eval: false
tabviz(data,
  ...,
  row_bold = ~ pval < 0.05,  # Evaluates for each row
  marker_color = ~ ifelse(upper < 1, "#16a34a", "#64748b")
)
```

The formula `~ pval < 0.05` becomes a vector of `TRUE`/`FALSE` values—one per row.

### Stage 3: JSON Serialization

The configured specification converts to JSON for the Svelte frontend. `serialize_spec()` (in `R/utils-serialize.R`) emits a top-level object with these keys:

```jsonc
{
  "data":          { "rows": [...], "groups": [...], "summaries": [...] },
  "columns":       [...],
  "extraColumns":  [...],
  "theme":         { ... },
  "interaction":   { ... },
  "labels":        { "title": ..., "subtitle": ..., ... },
  "paginate":      { "rows": 30, "pages": [...], "nPages": 3, ... },  // when set
  "layout":        { "plotWidth": "auto" }
}
```

Per-row styling lives **inside each row's object**, not in a parallel array — so a `row_bold` flag computed by `~ pval < 0.05` lands on `row.style.bold`:

```jsonc
// One element of data.rows
{
  "id":       "row_1",
  "label":    "Smith 2020",
  "groupId":  null,
  "metadata": { "hr": 0.72, "lower": 0.55, "upper": 0.95, "is_sig": true },
  "style":    { "bold": true }
}
```

When `paginate = paginate_spec(...)` is set, breakpoints are computed once R-side over the same flattened display rows the frontend renders, then shipped as 0-based row ranges so the viewer and the PDF export agree on page boundaries:

```jsonc
"paginate": {
  "rows":         30,
  "breakOn":      "split",
  "keepGroups":   true,
  "pageLabel":    "x_of_y",
  "pages":        [{ "startIdx": 0,  "endIdx": 29 },
                   { "startIdx": 30, "endIdx": 59 }],
  "nPages":       2
}
```

### Stage 4: Svelte Rendering

The Svelte component receives the JSON and renders the interactive table with:
- Sortable columns (default; opt out per column with `sortable = FALSE`)
- Collapsible groups
- Hover tooltips (opt-in via `web_interaction(tooltip_fields = ...)`)
- Page controls when `paginate` is set
- Static export to SVG / PNG / PDF

## Data Requirements

### Required Structure

Your data.frame needs:
- **One row per item** to display
- **Columns referenced** by your column specifications

```{r}
# Minimal forest plot data
meta_data <- data.frame(
  study = c("Smith 2020", "Jones 2021"),
  hr = c(0.72, 0.85),
  lower = c(0.55, 0.70),
  upper = c(0.95, 1.03)
)
```

#### Row kinds via `row_type`

By default every row is a data row with a marker. Pass `row_type = "<col>"` to `tabviz()` (with a column whose values are `"data"`, `"header"`, `"summary"`, or `"spacer"`) to mix in:

| `row_type` | Behavior | Marker |
|---|---|---|
| `"data"` | Standard row | Point with CI line |
| `"header"` | Section header (point/lower/upper should be `NA`) | None |
| `"summary"` | Pooled / subtotal row | Diamond |
| `"spacer"` | Empty separator (point/lower/upper should be `NA`) | None |

See [Row Styling](../guide/row-styling.qmd#row-types) for the full specification.

### Column Types

`col_*()` family covers cell content; `viz_*()` family adds shared-axis charts in their own column. tabviz infers a `field_category` from each data column (`numeric`, `integer`, `string`, `logical`, `date`, `array-numeric`, `other`) which the interactive column picker uses to filter compatible types.

| R Type | Cell columns (`col_*`) | Visualization columns (`viz_*`) |
|---|---|---|
| character | `col_text`, `col_label`, `col_badge`, `col_icon`, `col_reference` | — |
| numeric | `col_numeric`, `col_n`, `col_percent`, `col_currency`, `col_pvalue`, `col_progress`, `col_heatmap`, `col_stars`, `col_pictogram`, `col_ring`, `col_bar` | `viz_bar` (one value), `viz_forest` (point + CI), `viz_boxplot` / `viz_violin` (per-row arrays via list-columns or pre-computed stats) |
| integer | `col_n` | `viz_bar` |
| logical | Row / cell semantic flags | — (read indirectly via `row_*` / cell `style_*`) |
| date | `col_date` | — |
| list of numeric | `col_sparkline` | `viz_boxplot` (raw or quartiles), `viz_violin` (KDE) |
| paired numeric | `col_interval`, `col_events`, `col_range` | — |

### Missing Values

NA handling varies by column type:

| Context | Behavior |
|---------|----------|
| Forest points | Row displays as label-only (no marker) |
| Numeric columns | Displays as empty cell |
| Text columns | Displays as empty cell |
| Row styling | NA rows skip that style |

## Grouping Data

Grouping creates collapsible sections:

```{mermaid}
flowchart TD
    A["Flat data"] --> B["group = 'category'"]
    B --> C["Grouped display"]

    subgraph "Grouped display"
        D["Group A Header"]
        E["  Row 1"]
        F["  Row 2"]
        G["Group B Header"]
        H["  Row 3"]
    end
```

### Single-Level Grouping

```{r}
#| eval: false
data <- data.frame(
  study = c("Trial A", "Trial B", "Trial C", "Trial D"),
  category = c("Drug X", "Drug X", "Drug Y", "Drug Y"),
  hr = c(0.72, 0.85, 0.91, 0.78),
  lo = c(0.55, 0.70, 0.75, 0.60),
  hi = c(0.95, 1.03, 1.10, 1.00)
)

tabviz(data,
  label = "study",
  group = "category",  # Groups by Drug X, Drug Y
  columns = list(viz_forest(point = "hr", lower = "lo", upper = "hi"))
)
```

### Hierarchical Grouping

```{r}
#| eval: false
data$region <- c("North", "North", "South", "South")

tabviz(data,
  label = "study",
  group = c("region", "category"),  # Region > Category hierarchy
  columns = list(viz_forest(...))
)
```

## Styling Data Flow

Styles resolve in order of specificity:

```{mermaid}
flowchart TD
    A["Theme defaults"] --> B["Row styling"]
    B --> C["Cell styling"]
    C --> D["Marker styling"]
    D --> E["Final appearance"]

    style E fill:#059669,color:#fff
```

See [Styling Cascade](styling-cascade.qmd) for details.

## Common Patterns

### Pre-computing Style Columns

```{r}
library(dplyr)

styled_data <- meta_data |>
  mutate(
    is_significant = upper < 1,
    direction_color = case_when(
      upper < 1 ~ "#16a34a",  # Green: benefit
      lower > 1 ~ "#dc2626",  # Red: harm
      TRUE ~ "#64748b"        # Gray: NS
    )
  )

tabviz(styled_data,
  label = "study",
  columns = list(viz_forest(point = "hr", lower = "lower", upper = "upper",
                            scale = "log", null_value = 1)),
  row_bold = "is_significant",
  marker_color = "direction_color"
)
```

### Using Formula Expressions

Equivalent approach without pre-computing:

```{r}
tabviz(meta_data,
  label = "study",
  columns = list(viz_forest(point = "hr", lower = "lower", upper = "upper",
                            scale = "log", null_value = 1)),
  row_bold = ~ upper < 1,
  marker_color = ~ case_when(
    upper < 1 ~ "#16a34a",
    lower > 1 ~ "#dc2626",
    TRUE ~ "#64748b"
  )
)
```

Both approaches produce identical output. Use formulas for concise one-offs; use pre-computed columns when reusing the same logic.
