---
title: "Data Flow"
---

```{r}
#| include: false
library(tabviz)
```

Understanding how data flows through tabviz helps you structure your data correctly and debug issues.

## The Pipeline

```{mermaid}
flowchart LR
    subgraph Input
        A["data.frame"]
    end

    subgraph Processing
        B["Column Resolution"]
        C["Style Evaluation"]
        D["JSON Serialization"]
    end

    subgraph Output
        E["Svelte Component"]
        F["Interactive Widget"]
    end

    A --> B --> C --> D --> E --> F

    style D fill:#4f46e5,color:#fff
    style E fill:#ff3e00,color:#fff
```

### Stage 1: Column Resolution

tabviz maps your column specifications to data columns:

```{r}
#| eval: false
data <- data.frame(
  study = c("Trial A", "Trial B"),
  hr = c(0.72, 0.85),
  lo = c(0.55, 0.70),
  hi = c(0.95, 1.03)
)

tabviz(data,
  label = "study",  # Maps to data$study
  columns = list(
    col_forest(point = "hr", lower = "lo", upper = "hi")  # Maps to data$hr, data$lo, data$hi
  )
)
```

At this stage:
- String arguments become column lookups
- Missing columns trigger errors
- Data types are validated

### Stage 2: Style Evaluation

Formula expressions are evaluated against your data:

```{r}
#| eval: false
tabviz(data,
  ...,
  row_bold = ~ pval < 0.05,  # Evaluates for each row
  marker_color = ~ ifelse(upper < 1, "#16a34a", "#64748b")
)
```

The formula `~ pval < 0.05` becomes a vector of `TRUE`/`FALSE` values—one per row.

### Stage 3: JSON Serialization

The configured specification converts to JSON for the JavaScript frontend:

```json
{
  "data": [...],
  "columns": [...],
  "theme": {...},
  "styles": {
    "row_bold": [true, false, true, ...],
    "marker_color": ["#16a34a", "#64748b", ...]
  }
}
```

### Stage 4: Svelte Rendering

The Svelte component receives the JSON and renders the interactive table with:
- Sortable columns
- Collapsible groups
- Hover tooltips
- Export functionality

## Data Requirements

### Required Structure

Your data.frame needs:
- **One row per item** to display
- **Columns referenced** by your column specifications

```{r}
# Minimal forest plot data
meta_data <- data.frame(
  study = c("Smith 2020", "Jones 2021"),
  hr = c(0.72, 0.85),
  lower = c(0.55, 0.70),
  upper = c(0.95, 1.03)
)
```

### Column Types

tabviz handles standard R types:

| R Type | Column Functions | Notes |
|--------|-----------------|-------|
| character | `col_text()`, `col_badge()` | Text display |
| numeric | `col_numeric()`, `col_n()`, `col_percent()` | Formatted numbers |
| integer | `col_n()` | Sample sizes |
| logical | Row/marker styling | TRUE/FALSE conditions |
| list | `col_sparkline()` | Nested numeric vectors |

### Missing Values

NA handling varies by column type:

| Context | Behavior |
|---------|----------|
| Forest points | Row displays as label-only (no marker) |
| Numeric columns | Displays as empty cell |
| Text columns | Displays as empty cell |
| Row styling | NA rows skip that style |

## Grouping Data

Grouping creates collapsible sections:

```{mermaid}
flowchart TD
    A["Flat data"] --> B["group = 'category'"]
    B --> C["Grouped display"]

    subgraph "Grouped display"
        D["Group A Header"]
        E["  Row 1"]
        F["  Row 2"]
        G["Group B Header"]
        H["  Row 3"]
    end
```

### Single-Level Grouping

```{r}
#| eval: false
data <- data.frame(
  study = c("Trial A", "Trial B", "Trial C", "Trial D"),
  category = c("Drug X", "Drug X", "Drug Y", "Drug Y"),
  hr = c(0.72, 0.85, 0.91, 0.78),
  lo = c(0.55, 0.70, 0.75, 0.60),
  hi = c(0.95, 1.03, 1.10, 1.00)
)

tabviz(data,
  label = "study",
  group = "category",  # Groups by Drug X, Drug Y
  columns = list(col_forest(point = "hr", lower = "lo", upper = "hi"))
)
```

### Hierarchical Grouping

```{r}
#| eval: false
data$region <- c("North", "North", "South", "South")

tabviz(data,
  label = "study",
  group = c("region", "category"),  # Region > Category hierarchy
  columns = list(col_forest(...))
)
```

## Styling Data Flow

Styles resolve in order of specificity:

```{mermaid}
flowchart TD
    A["Theme defaults"] --> B["Row styling"]
    B --> C["Cell styling"]
    C --> D["Marker styling"]
    D --> E["Final appearance"]

    style E fill:#059669,color:#fff
```

See [Styling Cascade](styling-cascade.qmd) for details.

## Common Patterns

### Pre-computing Style Columns

```{r}
library(dplyr)

styled_data <- meta_data |>
  mutate(
    is_significant = upper < 1,
    direction_color = case_when(
      upper < 1 ~ "#16a34a",  # Green: benefit
      lower > 1 ~ "#dc2626",  # Red: harm
      TRUE ~ "#64748b"        # Gray: NS
    )
  )

tabviz(styled_data,
  label = "study",
  columns = list(col_forest(point = "hr", lower = "lower", upper = "upper")),
  row_bold = "is_significant",
  marker_color = "direction_color"
)
```

### Using Formula Expressions

Equivalent approach without pre-computing:

```{r}
tabviz(meta_data,
  label = "study",
  columns = list(col_forest(point = "hr", lower = "lower", upper = "upper")),
  row_bold = ~ upper < 1,
  marker_color = ~ case_when(
    upper < 1 ~ "#16a34a",
    lower > 1 ~ "#dc2626",
    TRUE ~ "#64748b"
  )
)
```

Both approaches produce identical output. Use formulas for concise one-offs; use pre-computed columns when reusing the same logic.
