---
title: "Styling Cascade"
---

```{r}
#| include: false
library(tabviz)
```

tabviz has **two cascades, working at different layers**.

* The **theme cascade** is internal to a `WebTheme`. It turns a small set of customer-facing inputs into a fully specified visual surface. Runs at theme construction; you don't normally see it.
* The **runtime cascade** is what happens at render time. The theme provides defaults, then row-level styles override them, then cell-level styles override those, then marker-level styles override those.

Knowing which cascade a property lives in tells you exactly where to set it — and which other layers it will override.

```{mermaid}
flowchart LR
    subgraph TC["Theme cascade (build time)"]
      T1[Inputs] --> T2[Semantic roles] --> T3[Component bindings]
    end

    subgraph RC["Runtime cascade (render time)"]
      D[Theme defaults] --> R[Row styling] --> C[Cell styling] --> M[Marker styling]
    end

    T3 --> D

    style T1 fill:#dbeafe
    style T2 fill:#bfdbfe
    style T3 fill:#93c5fd
    style D fill:#e5e7eb
    style R fill:#86efac
    style C fill:#6ee7b7
    style M fill:#fbbf24
```

The theme cascade hands a fully resolved visual specification to the runtime cascade. The runtime cascade then specializes per row / cell / marker.

---

## Cascade A — the theme

### What a theme is, ontologically

A theme is **a small set of customer choices** plus the rules for turning those choices into a fully specified visual surface. Think of it as having three concentric rings:

```
       ┌─ Tier 3: Component bindings (header, row, cell, plot, marks…) ─┐
       │   ┌─ Tier 2: Semantic roles (surface, content, divider…) ──┐   │
       │   │   ┌─ Tier 1: Inputs (12 customer choices) ──────────┐  │   │
       │   │   │   neutral, brand, accent, series, fonts, …      │  │   │
       │   │   └──────────────────────────────────────────────────┘  │   │
       │   └─────────────────────────────────────────────────────────┘   │
       └─────────────────────────────────────────────────────────────────┘
                                      ↓
                          resolve_theme(theme)
                                      ↓
                       fully populated WebTheme
```

Only Tier 1 is authored. Tier 2 and Tier 3 are *derived* by `resolve_theme()`. Customers (and presets) can also reach into Tier 2 / Tier 3 to pin individual fields when the derivation isn't quite right; resolution leaves those overrides alone.

### Two ontologies, sharing only Tier 1

The cascade splits into two independent vocabularies:

| Subsystem | Inputs | Tier 2 roles | What they describe |
|-----------|--------|--------------|--------------------|
| **Chrome** | `neutral`, `brand`, `brand_deep`, `accent`, `accent_deep` | `surface`, `content`, `divider`, `accent.*`, `status` | The table's scaffolding — the surface a row sits on, the tone of body text, the rule between cells |
| **Data** | `series_anchors`, `summary_anchor`, `status_*` | per-series `SlotBundle`, `summary` SlotBundle | The data marks — what a forest point looks like, how a row's "muted" state mutes its mark |

**They never cross.** Changing `brand` cannot affect a series fill. Changing `series_anchors[3]` cannot affect a row-group divider. The two cascades meet only at theme reassembly. This isolation is what makes themes *predictable*: you can swap a brand color and know the data marks will not silently shift hue, and vice versa.

### The vocabulary at Tier 2

Tier 2 names matter because they're how Tier 3 (and overrides) refer to the system. Chrome:

| Role | Use |
|------|-----|
| `surface.base` | Row background, default canvas |
| `surface.muted` | Banding partner, sunken regions |
| `surface.raised` | Cards, popovers |
| `content.primary` | Body text, cell values |
| `content.secondary` | Subtitles, captions |
| `content.muted` | Axis labels, footnotes |
| `content.inverse` | Text on dark fills (bold-mode header) |
| `divider.subtle` | Cell hairlines, gridlines |
| `divider.strong` | Header rule, group L1 underline |
| `accent.default` | Selection bar, hover focus |
| `accent.muted` | Hover row fill, selected row fill |
| `accent.tint.subtle` | L1 group bar (light header) |
| `accent.tint.medium` | Cascading bold-mode tints |

Data:

| Role | Use |
|------|-----|
| `series[i].fill` | Mark body for the i-th effect |
| `series[i].stroke` | Companion outline / CI line |
| `series[i].fill_muted` | The same effect on a muted row |
| `series[i].fill_emphasis` | The same effect on an emphasized row |
| `series[i].text_fg` | Data labels associated with the slot |
| `summary.{fill, stroke, ...}` | Pooled-effect diamond |

A `SlotBundle` carries every state for one slot. The renderer always reads the right field for the current row state — you don't author each state by hand.

### Variants pick, they don't synthesize

`header_style ∈ {light, bold}` and `first_column_style ∈ {default, bold}` are per-table toggles. Both variants are *always pre-resolved* on every theme: `theme$header$light` and `theme$header$bold` both carry full bg/fg/rule values. Flipping the variant at render time is a cheap dispatch — `theme$variants$header_style == "bold" ? theme$header$bold : theme$header$light`.

`density ∈ {compact, comfortable, spacious}` is similar but generates a full `SpacingTokens` table at resolve time. Per-token `set_spacing()` overrides land *on top* of the density preset.

### Resolution is a function

`resolve_theme(theme) -> WebTheme` is idempotent and pure. Run it twice, get the same answer. Override any field, re-run resolution, your override survives (resolution only fills NA-valued fields).

This means the cascade is debuggable. Inspect any resolved field to see exactly what it is:

```{r}
#| eval: false
t <- web_theme_lancet()
t@inputs@brand        # "#00407A"        (Tier 1)
t@accent@tint_subtle  # derived blend     (Tier 2)
t@row_group@L1@bg     # = accent.tint_subtle (Tier 3)
t@series[[1]]@fill    # series anchor 1
t@series[[1]]@stroke  # OKLCH-darkened companion
```

OKLCH is the underlying color algebra. All blends, lightenings, and darkenings happen in OKLCH (perceptually uniform), with a final clip to sRGB. Quantitative claim: a 12% blend toward `surface.base` looks 12% blended whether the starting hue is navy, gold, or magenta.

### What this is *not*

It's worth saying what the model deliberately does not do:

* **No global "muted" or "primary" color.** v1 had a single `colors$muted` that meant both "subdued text" and "deemphasized data." v2 splits these: `content.muted` for text, `slot.fill_muted` for data marks. They share Tier 1 (the brand/accent inputs) but never share a Tier 2 role.
* **No mode-pair within one theme.** A theme is single-mode (typically light). Dark themes are paired *presets* (`web_theme_dark()`), not a `theme.dark` slot.
* **No variant baked into the inputs.** Density, header style, first-column style are orthogonal axes. The same brand can ship as comfortable-light-default OR compact-bold-default OR any other combination.
* **No customer-tunable tints.** `accent.tint_subtle` is derived; it's not in the input list. If you need a specific tint, override the Tier 2 field directly with `set_theme_field()`. Don't add it as a Tier 1 knob.

---

## Cascade B — runtime style application

The theme provides defaults. The runtime cascade then specializes them per row, cell, and marker.

```{mermaid}
flowchart TB
    A["Theme defaults"] --> B["Row styling"]
    B --> C["Cell styling"]
    C --> D["Marker styling"]
    D --> E["Final appearance"]

    style A fill:#e5e7eb
    style B fill:#86efac
    style C fill:#6ee7b7
    style D fill:#fbbf24
    style E fill:#4f46e5,color:#fff
```

Later styles override earlier ones. A cell style beats a row style, which beats a theme default.

### Layer 1: Theme defaults

The theme establishes the baseline appearance for all elements:

```{r}
#| eval: false
tabviz(data,
  columns = list(viz_forest(point = "hr", lower = "lo", upper = "hi",
                           scale = "log", null_value = 1)),
  theme = web_theme_jama()
)
```

The theme defaults cover:

* **Chrome**: surfaces, content tones, dividers, accent
* **Data marks**: per-series slot bundles, summary slot
* **Typography**: 8 named text-role bundles
* **Spacing**: 12 density-derived tokens

Modify any of them with the v2 modifiers:

```{r}
#| eval: false
web_theme_jama() |>
  set_inputs(
    brand          = "#2563EB",
    series_anchors = c("#2563EB", "#16A34A", "#DC2626")
  ) |>
  set_variants(density = "comfortable") |>
  set_spacing(row_height = 28)
```

### Layer 2: Row Styling

Row styling applies to entire rows, overriding theme defaults:

```{r}
meta_data <- data.frame(
  study = c("Primary Endpoint", "  Subgroup A", "  Subgroup B", "Summary"),
  hr = c(NA, 0.72, 0.85, 0.78),
  lo = c(NA, 0.55, 0.70, 0.65),
  hi = c(NA, 0.95, 1.03, 0.93),
  rtype = c("header", "data", "data", "summary"),
  rbold = c(TRUE, FALSE, FALSE, TRUE)
)

tabviz(meta_data,
  label = "study",
  columns = list(viz_forest(point = "hr", lower = "lo", upper = "hi",
                           scale = "log", null_value = 1)),
  row_type = "rtype",
  row_bold = "rbold"
)
```

| Argument | Effect |
|----------|--------|
| `row_type` | Controls row behavior (header, data, summary, spacer) |
| `row_bold` | Bold text |
| `row_italic` | Italic text |
| `row_color` | Text color |
| `row_bg` | Background color |
| `row_indent` | Indentation level |
| `row_emphasis` | Theme-aware emphasis (reads `theme$row$emphasis`) |
| `row_muted` | Theme-aware de-emphasis (reads `theme$row$muted`) |
| `row_accent` | Theme-aware accent (reads `theme$row$accent`) |

The three semantic flags (`row_emphasis`, `row_muted`, `row_accent`) reach into the theme's `row.{emphasis,muted,accent}` cluster. That cluster is itself a Tier 3 binding derived from Tier 2 — so these flags pick up theme-appropriate colors automatically, regardless of the active preset.

### Layer 3: Cell Styling

Cell styling targets specific columns, overriding row styles:

```{r}
#| eval: false
tabviz(data,
  label = "study",
  columns = list(
    col_pvalue("pval",
      bold  = ~ .x < 0.05,
      color = ~ ifelse(.x < 0.05, "#16a34a", "#64748b")
    ),
    col_numeric("hr",
      color = ~ ifelse(.x < 1, "#16a34a", "#dc2626")
    ),
    viz_forest(point = "hr", lower = "lo", upper = "hi",
               scale = "log", null_value = 1)
  )
)
```

Cell styling uses `.x` to reference the column's own values:

```r
col_numeric("hr", color = ~ ifelse(.x < 1, "green", "red"))
```

### `set_column_style()`

Apply styling to any column after creation:

```{r}
#| eval: false
tabviz(data, label = "study", columns = list(...)) |>
  set_column_style("hr",
    bold = ~ .x < 1,
    color = ~ case_when(
      .x < 0.8 ~ "#16a34a",
      .x < 1.0 ~ "#86efac",
      TRUE     ~ "#64748b"
    )
  )
```

### Layer 4: Marker Styling

Marker styling controls forest plot points and lines, overriding all other layers:

```{r}
library(dplyr)

marker_data <- data.frame(
  study = c("Trial A", "Trial B", "Trial C", "Trial D"),
  hr = c(0.72, 0.95, 0.68, 1.15),
  lo = c(0.55, 0.78, 0.52, 0.90),
  hi = c(0.94, 1.15, 0.89, 1.47),
  design = c("RCT", "RCT", "Observational", "Observational")
) |>
  mutate(
    marker_color = case_when(
      hi < 1 ~ "#16a34a",
      lo > 1 ~ "#dc2626",
      TRUE   ~ "#64748b"
    ),
    marker_shape = ifelse(design == "RCT", "circle", "square")
  )

tabviz(marker_data,
  label = "study",
  columns = list(viz_forest(point = "hr", lower = "lo", upper = "hi",
                           scale = "log", null_value = 1)),
  marker_color = "marker_color",
  marker_shape = "marker_shape"
)
```

| Argument | Effect |
|----------|--------|
| `marker_color` | Fill color |
| `marker_shape` | Shape (circle, square, diamond, triangle) |
| `marker_opacity` | Transparency (0-1) |
| `marker_size` | Size multiplier |

## The viz-cell color cascade

Inside any viz column (`viz_forest`, `viz_bar`, `viz_boxplot`, `viz_violin`), the fill color of a single glyph is resolved by walking a four-layer precedence stack. **Most specific wins; NA passes through.**

| Layer | Source | Behavior |
|---|---|---|
| 4. Per-row literal | `tabviz(marker_color = "col" \| ~ formula)` | Wins when value is non-`NA`. `NA` falls through to Layer 3. |
| 3. Row semantic class | `row_accent` / `row_emphasis` / `row_muted = "col" \| ~ formula` | **Single-effect viz:** fill replaced with the theme's `row.{accent,emphasis,muted}.marker_fill`. **Multi-effect viz:** per-effect fill preserved, outline (stroke) added in the theme color. `FALSE`/absent falls through to Layer 2. |
| 2. Per-effect literal | `effect_forest(color = "#hex")` (and `effect_bar`/`effect_boxplot`/`effect_violin`) | Sets fill for that effect on all rows. |
| 1. Theme palette | `theme$series[[i]]@fill` | Invisible default; effect index cycles. |

Semantic class → theme field mapping (v2):

| Class | Theme path |
|---|---|
| `row_accent` | `theme@row@accent@marker_fill` |
| `row_emphasis` | `theme@row@emphasis@marker_fill` |
| `row_muted` | `theme@row@muted@marker_fill` |

The cascade applies uniformly to all four viz_* columns. The single-effect vs multi-effect distinction is automatic — based on `length(effects)`.

### Why the multi-effect rule?

In a single-effect viz, the marker color isn't carrying any per-effect information, so replacing it with the row's semantic color is unambiguous. In a multi-effect viz, each effect's color carries identity (e.g. blue = ITT, green = Per-Protocol, orange = Safety). Replacing the fill would erase that identity for accented rows. The "outline augment" rule keeps the per-effect color encoding intact while still flagging the row visually.

### Examples

The following examples render the cascade across each viz type:

| Example | Demonstrates |
|---|---|
| `gallery_17_cascade_single_forest` | Single-effect forest: row_accent / row_emphasis / row_muted replace marker fill |
| `gallery_18_cascade_multi_forest`  | Multi-effect forest: row_accent adds outline; per-effect fills preserved |

## Cascade Examples

These examples show how multiple cascade layers compose:

```{r}
#| eval: false
tabviz(meta_data,
  label = "study",
  columns = list(
    col_pvalue("pval",
      bold  = ~ .x < 0.05,                          # cell layer
      color = ~ ifelse(.x < 0.05, "#16a34a", NA)    # cell layer (NA inherits)
    ),
    viz_forest(point = "hr", lower = "lo", upper = "hi",
               scale = "log", null_value = 1)
  ),
  row_emphasis = ~ pval < 0.001,    # row layer
  theme = web_theme_jama()          # theme layer (chrome + data)
)
```

In this example:
1. `web_theme_jama()` sets all theme defaults (Tier 1 + Tier 2 + Tier 3 internally).
2. `row_emphasis = ~ pval < 0.001` reads `theme@row@emphasis` for matching rows.
3. `col_pvalue(..., bold = ~ .x < 0.05)` overrides text weight on specific cells.
4. `col_pvalue(..., color = ~ ...)` overrides cell color where the formula returns non-NA.

## Debugging Styles

If something doesn't look right, walk the cascade in reverse:

```{r}
#| eval: false
# Inspect the resolved theme
t <- web_theme_jama()
t@row@emphasis@marker_fill   # what semantic emphasis evaluates to
t@series[[1]]@fill           # default marker fill
t@cell@fg                    # default cell color

# Debug a formula expression
debug_color <- function(x) {
  res <- ifelse(x < 0.05, "#16a34a", "#64748b")
  cat("Input:", x, "\nOutput:", res, "\n")
  res
}

col_numeric("hr", color = ~ debug_color(.x))
```
