---
title: "Styling Cascade"
---

```{r}
#| include: false
library(tabviz)
```

tabviz applies styles in four layers, from general to specific: theme defaults, then row-level styles, then per-cell styles, then forest-marker styles. Knowing which layer a property lives in tells you exactly where to set it — and which other layers it will override.

## The Cascade

```{mermaid}
flowchart TB
    A["Theme defaults"] --> B["Row styling"]
    B --> C["Cell styling"]
    C --> D["Marker styling"]
    D --> E["Final appearance"]

    style A fill:#e5e7eb
    style B fill:#93c5fd
    style C fill:#6ee7b7
    style D fill:#fbbf24
    style E fill:#4f46e5,color:#fff
```

Later styles override earlier ones. A cell style beats a row style, which beats a theme default.

## Layer 1: Theme Defaults

Themes establish the baseline appearance for all elements:

```{r}
#| eval: false
# Theme sets default colors, fonts, spacing
tabviz(data,
  columns = list(viz_forest(point = "hr", lower = "lo", upper = "hi",
                           scale = "log", null_value = 1)),
  theme = web_theme_jama()  # Sets primary color, font, spacing, etc.
)
```

Theme defaults include:
- **Colors**: Primary, accent, text, background
- **Typography**: Font family, sizes, weights
- **Spacing**: Row height, padding, margins
- **Shapes**: Point size, line width

### Modifying Theme Defaults

```{r}
#| eval: false
web_theme_jama() |>
  set_colors(
    primary = "#2563eb",
    ci_marker_positive = "#16a34a",
    ci_marker_negative = "#dc2626",
    ci_line = "#475569"
  ) |>
  set_typography(font_size_base = "0.9rem") |>
  set_spacing(row_height = 28)
```

:::{.callout-note}
## Color argument names
The user-facing arguments of `set_colors()` are `ci_marker_fill`, `ci_marker_positive`, `ci_marker_negative`, `ci_marker_neutral`, `ci_line`. The corresponding `ColorPalette` S7 slots are still named `interval`, `interval_positive`, `interval_negative`, `interval_neutral`, `interval_line` — only the argument names changed.
:::

## Layer 2: Row Styling

Row styling applies to entire rows, overriding theme defaults:

```{r}
meta_data <- data.frame(
  study = c("Primary Endpoint", "  Subgroup A", "  Subgroup B", "Summary"),
  hr = c(NA, 0.72, 0.85, 0.78),
  lo = c(NA, 0.55, 0.70, 0.65),
  hi = c(NA, 0.95, 1.03, 0.93),
  rtype = c("header", "data", "data", "summary"),
  rbold = c(TRUE, FALSE, FALSE, TRUE)
)

tabviz(meta_data,
  label = "study",
  columns = list(viz_forest(point = "hr", lower = "lo", upper = "hi",
                           scale = "log", null_value = 1)),
  row_type = "rtype",
  row_bold = "rbold"
)
```

Row styling options:

| Argument | Effect |
|----------|--------|
| `row_type` | Controls row behavior (header, data, summary, spacer) |
| `row_bold` | Bold text |
| `row_italic` | Italic text |
| `row_color` | Text color |
| `row_bg` | Background color |
| `row_indent` | Indentation level |
| `row_emphasis` | Theme-aware emphasis |
| `row_muted` | Theme-aware de-emphasis |
| `row_accent` | Theme accent color |

## Layer 3: Cell Styling

Cell styling targets specific columns, overriding row styles:

```{r}
#| eval: false
tabviz(data,
  label = "study",
  columns = list(
    col_pvalue("pval",
      bold = ~ .x < 0.05,        # Bold if p < 0.05
      color = ~ ifelse(.x < 0.05, "#16a34a", "#64748b")
    ),
    col_numeric("hr",
      color = ~ ifelse(.x < 1, "#16a34a", "#dc2626")
    ),
    viz_forest(point = "hr", lower = "lo", upper = "hi",
               scale = "log", null_value = 1)
  )
)
```

Cell styling uses `.x` to reference the column's own values:

```r
# In cell context, .x = the column's value for that row
col_numeric("hr", color = ~ ifelse(.x < 1, "green", "red"))

# Equivalent to:
# for each row: if (row$hr < 1) "green" else "red"
```

### set_column_style()

Apply styling to any column after creation:

```{r}
#| eval: false
tabviz(data, label = "study", columns = list(...)) |>
  set_column_style("hr",
    bold = ~ .x < 1,
    color = ~ case_when(
      .x < 0.8 ~ "#16a34a",
      .x < 1.0 ~ "#86efac",
      TRUE ~ "#64748b"
    )
  )
```

## Layer 4: Marker Styling

Marker styling controls the forest plot points and lines, overriding all other layers:

```{r}
library(dplyr)

marker_data <- data.frame(
  study = c("Trial A", "Trial B", "Trial C", "Trial D"),
  hr = c(0.72, 0.95, 0.68, 1.15),
  lo = c(0.55, 0.78, 0.52, 0.90),
  hi = c(0.94, 1.15, 0.89, 1.47),
  design = c("RCT", "RCT", "Observational", "Observational")
) |>
  mutate(
    marker_color = case_when(
      hi < 1 ~ "#16a34a",
      lo > 1 ~ "#dc2626",
      TRUE ~ "#64748b"
    ),
    marker_shape = ifelse(design == "RCT", "circle", "square")
  )

tabviz(marker_data,
  label = "study",
  columns = list(viz_forest(point = "hr", lower = "lo", upper = "hi",
                           scale = "log", null_value = 1)),
  marker_color = "marker_color",
  marker_shape = "marker_shape"
)
```

Marker styling options:

| Argument | Effect |
|----------|--------|
| `marker_color` | Fill color |
| `marker_shape` | Shape (circle, square, diamond, triangle) |
| `marker_opacity` | Transparency (0-1) |
| `marker_size` | Size multiplier |

## The viz-cell color cascade

Inside any viz column (`viz_forest`, `viz_bar`, `viz_boxplot`, `viz_violin`), the fill color of a single glyph is resolved by walking a four-layer precedence stack. **Most specific wins; NA passes through.**

| Layer | Source | Behavior |
|---|---|---|
| 4. Per-row literal | `tabviz(marker_color = "col" \| ~ formula)` | Wins when value is non-`NA`. `NA` falls through to Layer 3. |
| 3. Row semantic class | `row_accent` / `row_emphasis` / `row_muted = "col" \| ~ formula` | **Single-effect viz:** fill replaced with theme color. **Multi-effect viz:** per-effect fill preserved, outline (stroke) added in theme color. `FALSE`/absent falls through to Layer 2. |
| 2. Per-effect literal | `effect_forest(color = "#hex")` (and `effect_bar`/`effect_boxplot`/`effect_violin`) | Sets fill for that effect on all rows. |
| 1. Theme palette | `theme.shapes.effect_colors[i]` | Invisible default; effect index cycles. |

Semantic class → theme color mapping:

| Class | Theme color |
|---|---|
| `row_accent` | `theme.colors.accent` |
| `row_emphasis` | `theme.colors.foreground` |
| `row_muted` | `theme.colors.muted` |

The cascade applies uniformly to all four viz_* columns. The single-effect vs multi-effect distinction is automatic — based on `length(effects)`.

### Why the multi-effect rule?

In a single-effect viz, the marker color isn't carrying any per-effect information, so replacing it with the row's semantic color is unambiguous. In a multi-effect viz, each effect's color carries identity (e.g. blue = ITT, green = Per-Protocol, orange = Safety). Replacing the fill would erase that identity for accented rows. The "outline augment" rule keeps the per-effect color encoding intact while still flagging the row visually.

### Examples

The following examples render the cascade across each viz type:

| Example | Demonstrates |
|---|---|
| `gallery_17_cascade_single_forest` | Single-effect forest: row_accent / row_emphasis / row_muted replace marker fill |
| `gallery_18_cascade_multi_forest`  | Multi-effect forest: row_accent adds outline; per-effect fills preserved |
| `gallery_19_cascade_bar`           | viz_bar single-effect (fill) vs multi-effect (outline) side-by-side |
| `gallery_20_cascade_dist`          | viz_boxplot + viz_violin: existing strokes pick up semantic color |
| `gallery_21_cascade_precedence`    | All four layers + NA passthrough on a single forest |

### NA passthrough

When a Layer-4 formula returns `NA` for a given row, that row "doesn't have an opinion" at Layer 4 and falls through to Layer 3:

```r
tabviz(data,
  marker_color = ~ ifelse(pval < 0.001, "darkred",      # very strong → exact color
                   ifelse(pval < 0.005, "indianred",    # strong       → exact color
                          NA_character_)),               # otherwise   → fall through
  row_accent   = ~ pval < 0.05,                          # rest of "significant" rows → accent
  ...
)
```

Rows with `pval < 0.005` get the exact red color (Layer 4 wins). Rows with `0.005 ≤ pval < 0.05` get the theme accent color (Layer 4 returned NA, Layer 3 takes over). Other rows get the palette default.

## Cascade Examples

### Example 1: Theme + Row Override

```{r}
#| eval: false
tabviz(data,
  label = "study",
  columns = list(viz_forest(point = "hr", lower = "lo", upper = "hi",
                           scale = "log", null_value = 1)),
  theme = web_theme_jama(),           # Sets default text color
  row_color = ~ ifelse(pval < 0.05, "#16a34a", NULL)  # Overrides for significant rows
)
```

- Theme sets default text color (black)
- Row styling overrides to green for significant rows
- Non-significant rows keep theme default

### Example 2: Row + Cell Override

```{r}
#| eval: false
tabviz(data,
  label = "study",
  columns = list(
    col_numeric("hr", color = ~ ifelse(.x < 1, "green", "red"))  # Cell override
  ),
  row_color = "#333333"  # Row default
)
```

- Row styling sets all text to dark gray
- Cell styling overrides HR column to green/red based on value
- Other columns keep the row style

### Example 3: Full Cascade

```{r}
#| eval: false
tabviz(data,
  label = "study",
  columns = list(
    col_pvalue("pval", bold = ~ .x < 0.05),
    viz_forest(point = "hr", lower = "lo", upper = "hi",
               scale = "log", null_value = 1)
  ),
  theme = web_theme_jama(),                    # Layer 1: Theme
  row_bold = "is_summary",                     # Layer 2: Row
  # col_pvalue bold is                         # Layer 3: Cell
  marker_color = ~ ifelse(sig, "green", "gray") # Layer 4: Marker
) |>
  set_colors(primary = "#2563eb")              # Modifies Layer 1
```

## Debugging Styles

When styles don't appear as expected:

1. **Check layer order**: Is a later layer overriding?
2. **Check NA values**: NA styling values fall through to the next layer
3. **Inspect formula results**: Print `data |> mutate(result = ~ your_formula)` to see evaluated values
4. **Simplify**: Remove styling layers one at a time to find conflicts

```{r}
#| eval: false
# Debug a formula expression
library(dplyr)
data |>
  mutate(
    debug_bold = pval < 0.05,  # See what row_bold = ~ pval < 0.05 produces
    debug_color = ifelse(upper < 1, "green", "gray")
  ) |>
  select(study, pval, upper, debug_bold, debug_color)
```
