---
title: "Theming Manifesto"
subtitle: "Why tabviz themes are shaped the way they are"
---

```{r}
#| include: false
library(tabviz)
```

::: {.callout-note}
**Living document — last updated 2026-04-29.**
The cascade reference in [Styling Cascade](styling-cascade.qmd) tells you *what*
the tokens are. This document tells you *why*. Read this before proposing a
fourth identity tier, a "trend" slot, a charm recipe, or otherwise expanding
the surface area.

Decisions get appended to the [Decision log](#decision-log) at the bottom —
never rewritten — so future maintainers can diagnose drift against
original intent.
:::

## Why this document exists

A cascade specification is a *what*. It enumerates inputs, derived tokens,
component bindings, and the formulas that connect them. It can't tell you
why the inputs are *those* inputs and not different ones, or why the cascade
ends at three tiers and not four, or why a class of feature requests has
been considered and rejected over and over again.

This document is the *why* companion. It's also a record: every substantive
shape decision is dated and reasoned through here, so when someone proposes
re-litigating one of them, the prior conversation is on the page.

Three audiences:

- **Future contributors** evaluating a feature request that touches the
  cascade. Read the [anti-patterns](#anti-patterns) section first.
- **Theme authors** building a new preset. The
  [theme authoring recipe](#authoring-a-new-theme) is the practical
  starting point.
- **Curious users** who want to know what they're using. The
  [two-knob principle](#the-two-knob-principle) and the
  [three-axis contract](#the-three-axis-contract) are the entry points.

If you want the formal contract, go to [Styling Cascade](styling-cascade.qmd).

---

## Publication-quality as a constraint

tabviz's product wedge is **publication-quality output**. Not "looks
professional," not "modern aesthetic" — *publication-quality*: tables that
could appear unmodified in a peer-reviewed journal article, a regulatory
submission, an annual report.

This is a constraint, not a label. Publication-quality is downstream of a
specific aesthetic discipline:

- **Restraint over expressivity.** A fourth color knob, an emoji recipe, a
  decorative glyph slot — each one is an opportunity for a user to make
  their table louder than it should be.
- **Fewer dimensions, used well.** Two coordinated colors that read as
  intentional beat five independent colors that read as crowded.
- **Chrome stays neutral.** Banding, hairlines, gridlines, axis text —
  these are scaffolding. Pull them toward the brand at low strength for
  cohesion; never at high strength for prominence.
- **Defaults that work everywhere.** A user who pins primary and writes
  one line of R should get a table that wouldn't look out of place in
  *The Lancet*. Anything that requires editorial intervention to look
  acceptable has failed.

Most of the choices documented below are downstream of this single frame.
If you find yourself defending an addition with "but the user could choose
to use it well," check that frame again. Publication-quality is what most
users get *by default*, not what the most disciplined users *can achieve
with effort*.

---

## The two-knob principle

The theme contract is two knobs: **identity** and **engagement**.

Identity is the brand axis — the tone the table opens with, the tone it
returns to under any structural emphasis. It mirrors down the chain:
secondary defaults to primary unless the author explicitly pins a second
color. In a mono theme (most journals), there is one identity color; in a
two-color editorial theme (LoTR-style), there are two.

Engagement is the **accent** — the color that fires when something extra is
happening. Hover. Selected. A semantic row callout. The status.info fallback.
The selection edge bar. Accent is orthogonal to identity; it never enters
the mirror chain. It's also reserved: it does *not* land on default content
(badges, glyph fills, ornament). Those belong on identity, where they
coordinate with the rest of the table.

Why two knobs and not more:

- A single identity knob makes mono themes trivial: pin primary, done.
- A second identity knob (secondary, mirroring primary) makes two-color
  editorial themes coherent without adding any complexity to mono themes.
- A separate engagement knob (accent) keeps the "look here" signal
  semantically distinct from the brand surface, so a hovered row can't
  read the same color as a regular badge two rows down.
- Adding any knob beyond these reopens the question of which knob a new
  feature should land on. The contract collapses.

Why accent is reserved for engagement:

- When accent doubled as both "the engagement color" and "the default fill
  for content-bearing glyph columns" (badges, pictograms, rings, stars), the
  same hue did two unrelated jobs and the emphasis signal lost its meaning.
  A hovered row read the same color as a regular badge. The 2026-04-28
  redesign fixed this: glyph col defaults moved to the secondary identity
  axis (which mirrors primary in mono themes — visually identical to before
  for journal users — and lights up secondary in two-color themes).
- The mental model "identity for what is, accent for what's happening"
  is the entire reason themes feel coherent.

If a new role doesn't cleanly map to one of these two axes, the answer is
almost always to refine the role's purpose, not to add a third axis.

---

## The three-axis contract

Identity is two tiers, but the actual *roles* are three. Each axis owns
exactly one job:

| Axis | Owns | Default behavior |
|---|---|---|
| **primary** (+ `primary_deep`) | **Prominence**: title text, bold leaf-header band, container border, series[0] anchor | Required input. `_deep` auto-darkens by 15% (OKLCH) if NA. |
| **secondary** (+ `secondary_deep`) | **Structure**: bold col-group band, row-group L1/L2/L3 bg, glyph col defaults (badge/pictogram/ring/stars), AND chrome texture (surface.muted, divider.subtle/strong, alt-row banding, gridline, axis/tick) | NA mirrors primary. `secondary_deep` mirrors `primary_deep` when secondary is NA; auto-darkens otherwise. |
| **accent** (+ `accent_deep`) | **Engagement**: hover, selected, row.accent.fg, accent.tint_subtle/medium, semantic.fill, selection edge color, status.info fallback | Required input. Orthogonal — never enters the identity mirror chain. |

The single rule for every theme: **pin what's distinct, mirror the rest.**
Mono themes pin primary only. Two-color editorial themes pin primary +
secondary. Accent is always pinned (engagement is never silent).

The structural axis is doing the most work. In mono themes, you don't
notice — secondary mirrors primary, so chrome and groupings read as
primary-tinted variants of the canvas. In two-color themes, secondary
lights up and you get coordinated movement across:

- The bold column-group band reads in secondary
- Row-group bars read in a 16% (or 24% under tint/bold header) mix of
  secondary into surface
- Cell hairlines and dividers carry a faint secondary tint
- Alt-row banding picks up the same secondary chrome
- Glyph cells (badges, rings, stars, pictograms) default to secondary

All from one pin. No per-theme overrides. That's the dividend the contract
pays.

---

## Why tertiary was added (2026-04-28) and removed (2026-04-29)

Honest record. Tertiary was a 24-hour decision and the postmortem teaches
the most important lesson in this document.

### The 2026-04-28 decision

The previous architecture was `brand` / `brand_deep` + `accent`. Editorial
themes (LoTR-style demos) wanted a second color for chrome texture —
specifically, the umber/moss/wheat tones that distinguish the three LoTR
themes from each other. Without a knob for it, theme authors were either
hardcoding hex values into every chrome token (high maintenance, drifts on
input change) or accepting a primary-tinted chrome that looked wrong
against the bold identity bands.

The proposal: rename `brand` → `primary`, add `secondary` for column-group
bg + row-group bg + glyph defaults, add `tertiary` for chrome texture
(surface.muted, dividers, banding, gridlines, axis/tick). Three identity
tiers, each mirroring up the chain. Accent stays orthogonal.

This shipped to main on 2026-04-28. Memory entry saved. Themes migrated.
Validation green.

### The 2026-04-29 reversal

Within 24 hours, two problems surfaced when stress-testing editorial
themes:

1. **Three colors usually communicate worse than two.** When designers
   could pin three identity hues, they did — even when the third hue's
   only job was "tint the chrome differently from the band." The result
   was tables with too much going on. Editorial restraint requires that
   the architecture *not invite* the third pin in the first place.

2. **Chrome wants to read structurally, not ornamentally.** Routing chrome
   through its own knob made it feel like a third identity color (because
   it was). Folding chrome onto the secondary axis — same axis that owns
   col-group bg, row-group bg, glyph defaults — collapsed it into "the
   structural color" instead of "the chrome color." The mental model
   simplified: secondary is *everything that supports the primary band*.

The reversal: `tertiary` and `tertiary_deep` removed entirely. Chrome
texture migrated onto `secondary_deep` at the same mix strengths. Mono
themes are visually unchanged (secondary mirrors primary). Two-color
themes get one structural axis instead of two.

The lesson:

- **Adding expressivity is reversible only if you add it carefully.**
  Tertiary was added with full validation, full migration, full memory
  trail — which made the reversal cheap. A messier addition would have
  cost weeks.
- **A 24-hour postmortem on a fresh decision is healthier than living
  with the wrong shape for a quarter.** The reversal entry is in this
  log not as embarrassment but as evidence that the loop is short.
- **The right number of identity tiers is two. Probably forever.**

The LoTR themes lost nothing essential. Dwarven kept bronze (primary)
+ gold (secondary, now tinting chrome too) + ember (accent). The third
umber pin folded into the gold cascade — and the chrome in Dwarven now
reads as warm-burnished-metallurgy, which is *more* on-theme than the
original cool umber against warm bronze ever was.

---

## Mirror chain semantics

The identity chain is a single rule:

> If `secondary` is NA at construction, it mirrors `primary` during
> resolution. If `secondary_deep` is NA, it mirrors `primary_deep` (when
> secondary itself is NA) or auto-darkens from secondary (when secondary
> is pinned).

A few things to know about how this behaves:

**Mirroring happens at resolve time, not at construction.**
`ThemeInputs(primary = "#0891b2")` stores `secondary = NA` literally; the
NA only resolves to `"#0891b2"` after `resolve_theme()` has run. This
means re-resolving an already-resolved theme is idempotent: the resolver
only writes into NA-valued fields, so a previously-mirrored secondary
stays at primary's value, and a pinned secondary stays at the user's pin.

**Pinning a tier breaks the mirror for that tier and downstream tiers.**
If you pin `secondary`, `secondary_deep` no longer follows `primary_deep`
— it auto-derives from your pinned secondary instead. This is the
intentional break: the moment you say "I want a second color," the deep
companion tracks the second color, not the first.

**Accent never mirrors.** It's a required input precisely because
engagement is never silent. Even in a mono theme, accent is a deliberate
choice (often a deliberate counterpoint — Cochrane pairs cool teal primary
with warm coral accent specifically because the contrast lights up the
engagement signal).

**The mirror is not just visual.** It propagates through every derived
token. In a mono theme, `divider.strong` reads `secondary_deep` which
mirrors `primary_deep` which auto-darkens primary — a single edit to
primary cascades through dividers, surface.muted, banding, gridlines, all
of it. This is the dividend that justifies the implicit-NA design over
having theme authors restate primary in every cascade slot.

---

## Cascade tiers

Three tiers, each with one job:

- **T1 — Inputs.** What the customer chooses. `primary`, `secondary`,
  `accent`, plus `_deep` companions, plus `slot_style` (enum),
  `header_style` (enum), neutral ramp, status colors, series anchors,
  fonts. ~15 fields. This is the only tier most users edit directly.

- **T2 — Semantic roles.** Derived tokens with named semantics:
  `surface.base/muted/raised`, `content.primary/secondary/muted/inverse`,
  `divider.subtle/strong`, `accent.default/muted/tint_subtle/tint_medium`,
  `semantic.fill`, `status.*`, plus per-slot `SlotBundle` for series
  marks. The resolver fills these from T1 via OKLCH formulas; users
  override individual fields to pin a tone.

- **T3 — Component bindings.** Per-cluster overrides: `header.{light,
  tint, bold}.{bg, fg, rule}`, `column_group.*`, `row_group.L1/L2/L3.bg`,
  `cell.border`, `plot.axis_line/tick_mark/gridline/reference`. These
  default to T2 values with cluster-local computations (e.g., bold rule
  is a 40% mix of inverse against the band's own bg). Users can pin
  individual leaves; the rest cascade from T2.

Three tiers of overrides means three chances for a user to break a
required relationship. Construction-time validation
(`validate_resolved_theme`) catches the obvious failures (bold-band
contrast, body-text contrast) at theme construction — *not* at render —
so the error message points at the source of the bad pin, not at a
downstream symptom. WCAG AA Large (3.0) for chrome bold text; AA Normal
(4.5) for body text. Tests can opt out with `.validate = FALSE`.

The cascade is fully introspectable from R:

```{r}
#| eval: false
inspect_resolved(web_theme_cochrane(), "header.bold.bg")
#> $path: "header.bold.bg"
#> $value: "#006B9C"
#> $cascade: c("inputs$primary_deep", "inputs$primary")
#> $derivation: "inputs$primary_deep (auto: oklch_darken(primary, 0.15))"
```

The same provenance feeds the widget's Theme tab tooltips — single
source of truth for "where did this hex come from?".

---

## Anti-patterns

These are not "things we haven't gotten around to." These are things tabviz
will *refuse to do*. If you encounter one in a feature request, push back.

### Don't add a fourth identity tier (or revive tertiary)

Two-knob is canonical. Tertiary was added on 2026-04-28 and removed on
2026-04-29 — see the [postmortem](#why-tertiary-was-added-2026-04-28-and-removed-2026-04-29).
Any future proposal to add a third identity color must either (a)
demonstrate a use case that the secondary mirror chain genuinely cannot
serve, or (b) explain why the prior reversal's reasoning no longer
applies. Aesthetic preference for three colors is not sufficient — the
reversal happened *because* designers reached for three when two would
have read better.

### Don't introduce slot-typed Tier 1 anchors

A previous architectural proposal added `slot.trend`, `slot.distribution`,
`slot.range`, etc. as Tier 1 anchors keyed to recipe types. This was
**considered and retracted**. The two-knob model handles every case the
slot-typed proposal addressed (via accent-tint and identity-deep
derivations). Reintroducing slot-typed anchors adds dimensionality that
the cascade has to thread through every component binding for
near-zero practical gain.

### Don't mix accent into default content

Accent owns engagement. Hover, selected, callouts, status.info, semantic
row treatments. It does *not* own:

- Default badge fills
- Glyph cell defaults (rings, stars, pictograms)
- Ornament that fires by default

Those belong on the secondary identity axis. The 2026-04-28 redesign
documented this explicitly: glyph col defaults moved from accent to
secondary precisely so the engagement signal stays meaningful. If you find
yourself wanting accent on a content surface that's always rendered, the
correct fix is almost certainly identity, not accent.

### Don't tint chrome toward accent

Chrome (banding, dividers, gridlines, axis text) is structural. It tints
toward `secondary_deep` — which in mono themes mirrors `primary_deep`, so
chrome stays cohesive with the brand. Tinting chrome toward accent makes
hovered/selected rows merge into the chrome band. This was wrong before
the cascade rework and would still be wrong after.

### Don't add charm recipes

Pictogram arrays, emoji marks, decorative glyph recipes, isotype-style
visualizations. These are charming in stress-test demos but distract from
publication-quality positioning. The recipe library is intentionally
small: it grows when a serious publication-quality use case demonstrably
isn't covered (e.g., gauge-against-target, dots-of-N, stacked-bar
proportion compositions are on the roadmap), not when a demo would be
fun.

The boundary check: would this recipe appear in *The Lancet*? If no, hold
the line.

### Don't bypass the cascade

Every now and then a feature request asks for a theme override that
"skips" the cascade — a one-off color that doesn't flow through any of
the derivations, doesn't participate in mirroring, can't be inspected.
This signal almost always means the cascade is missing a case. The right
response is to extend the cascade (add the slot to T2 or T3, document its
derivation, run validation), not to introduce a side door.

If genuinely no T1/T2/T3 location fits, that's a decision-quality moment.
Stop and re-examine the role's semantics before committing to a new slot.

---

## Authoring a new theme

The recipe is short. Most themes need only Tier 1 inputs.

**Mono theme** (the journal default — Cochrane, Lancet, JAMA fall here):

```{r}
#| eval: false
my_theme <- web_theme(
  name = "my_journal",
  inputs = list(
    primary = "#0F4C81",     # identity hero
    accent  = "#C8553D",     # warm engagement counterpoint
    series_anchors = c("#0F4C81", "#C8553D", "#5C8A3F", "#7E5A99", "#D49A3A"),
    font_body = "Georgia, 'Times New Roman', serif"
  )
)
```

That's it. Secondary mirrors primary; chrome reads as primary-tinted;
glyph defaults read primary; engagement (hover/selected/info) reads
accent. The contract gives you a coherent table with two pinned values.

**Two-color editorial theme** (LoTR-style):

```{r}
#| eval: false
my_editorial <- web_theme(
  name = "my_editorial",
  inputs = list(
    primary   = "#7A4E22",   # bronze
    secondary = "#D4A955",   # gold — drives col-group bands, row-group bars,
                             #        glyph defaults, AND chrome texture
    accent    = "#E0673A",   # ember — hover, selected, status.info
    series_anchors = c("#D4A955", "#A8A8A0", "#B86C42", "#7A4E22", "#5C7AA0"),
    font_body = "'EB Garamond', Georgia, serif"
  ),
  variants = list(header_style = "bold")
)
```

One additional pin (secondary) and one variant flip (header_style) — the
rest cascades.

**Verifying your theme:**

```{r}
#| eval: false
# Check what a specific leaf resolved to + why.
inspect_resolved(my_theme, "header.bold.bg")
inspect_resolved(my_theme, "divider.strong")
inspect_resolved(my_theme, "row_group.L1.bg")

# Construction-time validation runs automatically — broken contrast
# aborts at web_theme() with a structured error pointing at the leaf.
# To bypass for experimental / synthetic colors:
my_test <- web_theme(..., .validate = FALSE)
```

**What to override at T2/T3 (and when):**

- Override at T1 first. 90% of themes never need to touch T2 or T3.
- Override at T2 when a *family* of derived tokens needs adjustment —
  e.g., `divider.strong` to pin the rule color across header/group/axis.
- Override at T3 only for cluster-specific tweaks — e.g., a particular
  preset wants the bold-mode header band darker than its primary_deep
  would give.

The wider you cast the override, the higher up the cascade you should
pin it. T3 overrides are bandages; T1 overrides are intent.

---

## Open questions

The architecture is mature enough that big redesigns are unlikely.
Outstanding questions are tactical:

- **Recipe fallbacks at narrow widths.** Whether violins below a width
  threshold should automatically degrade to density strip / boxplot.
  Could be guidance or runtime fallback. No current opinion.
- **Slot.style additional variants.** Currently three values
  (`fill_with_darker_stroke`, `flat_fill`, `outlined`). A "stroke_only"
  variant has been suggested for very dense plots. Not yet implemented;
  awaiting a real-world need.
- **Header tint strength.** The tint variant is a 12% mix of primary_deep
  into surface.base. This was chosen by eye; whether 10% or 15% reads
  better is taste-dependent and may benefit from per-theme override.

When new questions arise, append to the [Decision log](#decision-log).
When questions get resolved, move them out of this section and into the
log entry that resolved them.

---

## Decision log

Append-only. Newest first. Each entry: date, decision, alternatives
considered, reasoning. Future maintainers should append, not rewrite.

### 2026-04-29 — Tertiary removed; chrome migrates to secondary

**Decision.** Drop `tertiary` and `tertiary_deep` from `ThemeInputs`.
Chrome texture (surface.muted, divider.subtle/strong, alt-row banding,
gridline, axis/tick) now reads from `secondary_deep` at the same mix
strengths. Identity is two tiers (primary + secondary mirror chain) plus
orthogonal accent.

**Alternatives considered.**
- Keep three tiers, lower the bar for what tertiary owns. Rejected:
  doesn't solve the "designers reach for three colors" problem.
- Add a separate `chrome_tint` knob (not in identity chain). Rejected:
  same expressivity creep with extra surface area.
- Keep three tiers but make tertiary "advanced-only" (hidden from default
  panels). Rejected: API surface stays bloated; documentation cost is
  high; the prompt to use three colors is structural, not UI.

**Reasoning.** See [the postmortem](#why-tertiary-was-added-2026-04-28-and-removed-2026-04-29).
Two-knob is canonical. Bundled with this rework: `slot_style` lifted to
T1; `header_style` expanded with `tint` variant; row-group L1 bg gains
header_style-driven subtle/strong strength; construction-time contrast
validation; `inspect_resolved()` introspection API; 7 presets migrated
(LoTR trio rebalanced as 2-color).

**Migration cost.** ~3 hours of focused refactoring. Caught a real bug
along the way: dark theme's auto-derived `primary_deep` (#5C85C8) had
2.57:1 contrast against its inverse text — pinned darker to #2E5290
(5.33:1).

### 2026-04-28 — Tertiary added (later reversed)

**Decision.** Add `tertiary` and `tertiary_deep` to `ThemeInputs`. Three
identity tiers in a mirror chain (primary → secondary → tertiary).
Tertiary owns chrome texture.

**Reasoning at the time.** Editorial themes wanted a third color for
chrome texture distinct from the structural secondary; pinning it kept
chrome cohesive without per-theme hardcoded hex values.

**Reversed within 24 hours.** See the 2026-04-29 entry above.

### 2026-04-28 — Accent reserved for layered emphasis

**Decision.** Glyph col defaults (badge, pictogram, ring, stars) move
from `accent` to `secondary` identity. Accent is reserved strictly for
**layered emphasis**: roles that fire when something extra is happening
(hover, selected, semantic row callouts, selection edge, status.info
fallback).

**Reasoning.** When accent doubled as both "the engagement color" and
"the default fill for content-bearing glyph columns," the same hue did
two unrelated jobs and the emphasis signal lost meaning — a hovered row
read the same color as a regular badge two rows down.

This decision survives the 2026-04-29 tertiary reversal unchanged. It's
the most important constraint on what accent does.

### Earlier — Slot-typed Tier 1 anchors retracted

**Decision.** Do not introduce `slot.trend`, `slot.distribution`,
`slot.range`, etc. as Tier 1 anchors keyed to recipe types.

**Reasoning.** The two-knob theme architecture (identity + accent with
cascade-pinning + element overrides) handles the cases the slot-typed
proposal addressed. Reintroducing slot-typed anchors adds dimensionality
that the cascade has to thread through every component binding for
near-zero practical gain.

This is documented in `engineering-priorities.md` §"Anti-priorities" as
a binding constraint on future architecture work.

---

## See also

- [Styling Cascade](styling-cascade.qmd) — the formal contract (what)
- `?inspect_resolved` — the introspection API used in examples above
- `R/utils-theme-validate.R` — construction-time validation invariants
- `R/utils-theme-resolve.R` — the resolver that materializes T2/T3
- `R/themes.R`, `R/themes-lotr.R` — preset themes as canonical examples
