---
title: "Columns"
---

```{r}
#| include: false
library(tabviz)
library(dplyr)
```

Columns are the unit of display in tabviz. You compose a table by passing a list of column specifications to `tabviz(columns = ...)`. Each column helper (`col_*()`, `viz_*()`) picks a display type and exposes formatting knobs; the underlying `web_col()` constructor is there if you need something unusual.

## Universal arguments

Every `col_*()` helper accepts the same arguments for layout and styling. These behave identically regardless of column type.

| Argument | Default | Purpose |
|---|---|---|
| *Primary data* (see per-helper table) | required | Which column(s) in `data` provide values |
| `header` | `NULL` (auto) | Column header text; defaults to the field name or a type-specific default |
| `width` | `NULL` (auto) | Column width in pixels |
| `align` | `"left"` (numeric types default to `"right"`) | Body cell alignment: `"left"`, `"center"`, `"right"` |
| `header_align` | follows `align` | Header alignment; override to decouple |
| `show_header` | auto | `TRUE`/`FALSE` to force; `NULL` shows when `header` is non-empty |
| `wrap` | `FALSE` | Multi-line cell text. `FALSE`/`0` = single line (ellipsis on overflow). `TRUE`/`1` = up to 2 lines (1 extra). Integer `n` = up to `n + 1` lines. Author `\n` is honoured first, then long segments word-wrap within the column width. |
| `sortable` | `TRUE` | Allow sort on header click |
| `na_text` | `NULL` | Text to display for NA/missing values (default is blank) |
| `tooltip` | `NULL` | Column whose values become per-cell hover text |
| `formatter` | `NULL` | R function `function(x) ...` applied to values; output rendered as text. See [Custom formatters](#custom-formatters) below. |
| `bold`, `italic`, `color`, `bg`, `badge`, `icon`, `emphasis`, `muted`, `accent` | `NULL` | Per-cell styling driven by another column — see [Cell Styling](cell-styling.qmd) |

:::{.callout-tip}
## Everything routes through `web_col()`
The `col_*()` helpers are thin, opinionated wrappers over [`web_col()`](../reference/columns.html#web_col). When a helper doesn't cover your case (custom type, non-standard option), call `web_col()` directly.
:::

## Catalog at a glance

| Helper | Use for | Type-specific arguments |
|---|---|---|
| `col_text(field)` | Plain text, categoricals | `max_chars` |
| `col_label(field)` | Row-identifier column | `max_chars` |
| `col_numeric(field)` | General numbers | `decimals`, `digits`, `thousands_sep`, `abbreviate` |
| `col_n(field)` | Sample sizes | `decimals`, `digits`, `thousands_sep`, `abbreviate` |
| `col_percent(field)` | Proportions / percentages | `decimals`, `digits`, `multiply`, `symbol` |
| `col_currency(field)` | Money | `symbol`, `position`, `decimals`, `digits`, `thousands_sep`, `abbreviate` |
| `col_pvalue(field)` | P-values | `stars`, `thresholds`, `format`, `digits`, `exp_threshold`, `abbrev_threshold` |
| `col_interval(point, lower, upper)` | Point estimate + CI | `decimals`, `separator`, `imprecise_threshold` |
| `col_events(events, n)` | Events / N ratio | `separator`, `show_pct`, `thousands_sep`, `abbreviate` |
| `col_range(low, high)` | Low – High range | `separator`, `decimals`, `show_bar` |
| `col_date(field)` | Dates | `format` |
| `col_bar(field)` | Inline magnitude bar (no axis) | `max_value`, `show_label`, `color`, `scale` |
| `col_sparkline(field)` | Trend charts | `type`, `height`, `color` |
| `col_heatmap(field)` | Color-scale values | `palette`, `min_value`, `max_value`, `decimals`, `show_value`, `scale` |
| `col_progress(field)` | Percentage progress bar | `min_value`, `max_value`, `show_label`, `height`, `color` |
| `col_icon(field)` | Icons / emoji | `mapping`, `size`, `color` |
| `col_badge(field)` | Status chips | `variants`, `colors`, `size` |
| `col_stars(field)` | Rating (1–N stars) | `max_stars`, `color`, `empty_color`, `half_stars`, `min_value`, `max_value`, `size` |
| `col_img(field)` | Inline images | `height`, `max_width`, `fallback`, `shape` |
| `col_reference(field)` | Truncated text + optional link | `href_field`, `max_chars`, `show_icon` |
| `col_group(header, ...)` | Group child columns under one header | — |

For the four axis-backed visualization columns (`viz_forest`, `viz_bar`, `viz_boxplot`, `viz_violin`), see [Visualizations](visualizations/index.qmd).

## Composing columns

A `tabviz()` call takes a list of column specs in display order. The special `label` argument produces an implicit leftmost identifier column; you can also declare it explicitly with `col_label()` if you want finer control.

```{r}
data(glp1_trials)

meta <- glp1_trials |>
  filter(row_type == "data", group == "Main Trials") |>
  head(4)

tabviz(meta,
  label = "study",
  columns = list(
    col_text("drug", "Drug"),
    col_n("n"),
    col_events("events", "n"),
    col_pvalue("pvalue", "P")
  )
)
```

## Headers and widths

Headers default to the field name if omitted; pass a second positional argument (or named `header`) to override, and an empty string or `show_header = FALSE` to hide.

Widths are auto-sized by default. Pass a pixel value to fix a column.

```{r}
#| eval: false
col_text("drug")                          # header: "drug", auto width
col_text("drug", "Drug")                  # header: "Drug"
col_text("drug", header = "", width = 80) # no header, 80px wide
col_text("drug", show_header = FALSE)     # header cell hidden
```

## Alignment

`col_*` helpers that render numeric content right-align by default (`col_numeric`, `col_n`, `col_percent`, `col_currency`, `col_pvalue`, `col_range`). Everything else left-aligns. Override with `align`; header alignment follows the body unless you set `header_align` explicitly.

```{r}
#| eval: false
col_numeric("value", align = "left")                      # force body left
col_numeric("value", header_align = "center")             # center header only
```

## Numeric formatting

The numeric formatting vocabulary is shared across `col_numeric`, `col_n`, `col_percent`, `col_currency`, `col_pvalue`, `col_heatmap`, `col_events`, and `col_interval`. Use whichever helper fits your data — you get the same formatting primitives.

| Argument | Meaning |
|---|---|
| `decimals` | Fixed number of decimal places. Mutually exclusive with `digits`. |
| `digits` | Significant figures. Mutually exclusive with `decimals`. |
| `thousands_sep` | Separator string (`","`, `" "`, `"."`, …) or `FALSE` to disable |
| `abbreviate` | `TRUE` to shorten large numbers: `1,234 → "1.2K"`, `2.5e6 → "2.5M"` |

```{r}
options_data <- data.frame(
  item  = c("Small", "Medium", "Large"),
  n     = c(125, 4520, 1250000),
  rate  = c(0.156, 0.892, 0.034),
  value = c(1.234, 12.34, 123.4)
)

tabviz(options_data, label = "item",
  columns = list(
    col_n("n",                              "Count"),
    col_n("n",        abbreviate = TRUE,    "Abbreviated", id = "n_abbr"),
    col_numeric("value", decimals = 2,      "2 decimals"),
    col_numeric("value", digits   = 3,      "3 sig figs",   id = "value_sigfig"),
    col_percent("rate",                     "Percent")
  )
)
```

:::{.callout-note}
## Decimals vs. digits
`decimals = 2` always prints 2 places (`0.10`, `3.14`, `125.00`). `digits = 3` prints 3 meaningful figures (`0.156`, `3.14`, `125`). Use `digits` when values span orders of magnitude.
:::

## Handling missing data

Set `na_text` on any column to replace `NA` cells with custom text.

```{r}
#| eval: false
col_numeric("response", na_text = "—")
col_pvalue("pvalue", na_text = "n/a")
col_events("events", "n", na_text = "—")
```

The default is a blank cell.

## Per-cell tooltips

Pass `tooltip = "<col>"` to any column helper to drive the cell's hover-text from another column in the data. This overrides the default behavior of using the cell's own value as the tooltip.

```{r}
#| eval: false
tabviz(data, label = "study",
  columns = list(
    col_text("status", tooltip = "status_explanation"),  # custom hover text
    col_n("n", tooltip = "n_source")                      # e.g. "Per-protocol N"
  )
)
```

`tooltip` accepts a column name; the value at that cell becomes the rendered `title` attribute.

## Custom formatters

When the built-in formatting vocabulary doesn't fit (`decimals`/`digits`/`abbreviate`/`thousands_sep`), pass a `formatter` function. The function is applied to the column's values at render time and the output is treated as text — the column is rendered verbatim.

```{r}
#| eval: false
data <- data.frame(
  study = c("A", "B", "C"),
  estimate = c(1.234, 0.85, 2.4),
  se       = c(0.12, 0.08, 0.31)
)

tabviz(data, label = "study",
  columns = list(
    web_col("estimate",
      formatter = function(x) sprintf("%.2f ± %.2f", x, data$se))
  )
)
```

The formatter receives the raw column vector; return a character vector of the same length. Use this for unit suffixes, conditional formatting, or composite values that don't fit any single `col_*()` helper.

:::{.callout-note}
## When to reach for `formatter`
For routine formatting (decimals, sig figs, percentages, currencies, abbreviations), prefer the built-in arguments — they're discoverable, declarative, and fast. Reach for `formatter` only when the desired output can't be expressed via the built-in vocabulary.
:::

## Cell-level styling

All the cell-styling arguments (`bold`, `italic`, `color`, `bg`, `badge`, `icon`, `emphasis`, `muted`, `accent`) accept either a column name or a formula. See [Cell Styling](cell-styling.qmd) for the full formula-pronoun rules and recipes.

```{r}
glp1_trials |>
  filter(row_type == "data", group == "Main Trials") |>
  head(4) |>
  mutate(
    is_sig   = upper < 1,
    hr_color = ifelse(upper < 1, "#166534", "#64748b")
  ) |>
  tabviz(
    label = "study",
    columns = list(
      col_interval("hr", "lower", "upper", header = "HR (95% CI)",
                   bold = "is_sig", color = "hr_color"),
      col_pvalue("pvalue", "P", bold = "is_sig")
    )
  )
```

## Column groups

`col_group(header, ...)` wraps a set of columns under a shared header row — useful for multiple outcomes, time points, or analyses.

```{r}
outcomes <- data.frame(
  study          = c("PARADIGM-HF", "DAPA-HF", "EMPEROR-Reduced"),
  primary_hr     = c(0.80, 0.74, 0.75),
  primary_lower  = c(0.73, 0.65, 0.65),
  primary_upper  = c(0.87, 0.85, 0.86),
  primary_p      = c(0.001, 0.001, 0.001),
  secondary_hr   = c(0.84, 0.70, 0.69),
  secondary_p    = c(0.01, 0.001, 0.001)
)

tabviz(outcomes, label = "study",
  columns = list(
    viz_forest(point = "primary_hr", lower = "primary_lower",
               upper = "primary_upper", null_value = 1, scale = "log"),
    col_group("Primary",
      col_interval("primary_hr", "primary_lower", "primary_upper",
                   header = "HR (95% CI)"),
      col_pvalue("primary_p", "P")
    ),
    col_group("Secondary",
      col_numeric("secondary_hr", "HR", decimals = 2),
      col_pvalue("secondary_p", "P")
    )
  ),
  theme = web_theme_cochrane()
)
```

## Visualization columns

Four `viz_*()` columns render axis-backed plots inline: `viz_forest`, `viz_bar`, `viz_boxplot`, `viz_violin`. They share axis-control arguments (`axis_range`, `axis_ticks`, `axis_gridlines`, `axis_label`, `scale`) and are documented separately:

- [Forest plots](visualizations/forest-plots.qmd) — point + CI with annotations
- [Bar charts](visualizations/bar-charts.qmd) — horizontal bars with an axis
- [Boxplots](visualizations/boxplots.qmd) — quartiles, median, outliers
- [Violin plots](visualizations/violin-plots.qmd) — kernel-density distributions

Unlike `col_bar`, the viz columns share a consistent axis across rows.

All four viz columns honor the [marker-styling cascade](../concepts/styling-cascade.qmd#the-viz-cell-color-cascade): row semantic classes (`row_accent`/`row_emphasis`/`row_muted`) and per-row `marker_color` from `tabviz()` reach into the viz cell automatically. Single-effect viz columns get a fill replacement; multi-effect columns get an outline that preserves per-effect color identity. See [Row Styling → Cascading into viz cells](row-styling.qmd#cascading-into-viz-cells) for the full picture.

---

## Per-helper detail

### Text and identifiers

#### `col_text()` and `col_label()`

Plain text with optional truncation. `col_label()` is the convenience helper for the row-identifier column; it also prettifies the field name into a Title Case header by default.

```{r}
#| eval: false
col_text("drug")
col_text("notes", "Notes", max_chars = 50)      # truncate with ellipsis
col_text("description", wrap = TRUE, width = 220)   # wrap (up to 2 lines)
col_text("notes", wrap = 3, width = 200)            # up to 4 lines

col_label("study_id")                           # header auto: "Study Id"
col_label("study_id", header = "Study")          # explicit header
```

### Numbers

#### `col_numeric()`

General-purpose numeric column with `decimals` / `digits` / `thousands_sep` / `abbreviate`.

```{r}
#| eval: false
col_numeric("estimate")                          # 2 decimals
col_numeric("count", decimals = 0,
            thousands_sep = ",")                 # "12,345"
col_numeric("value", digits = 3)                 # 3 sig figs
col_numeric("population", abbreviate = TRUE)     # "1.2M"
```

#### `col_n()`

Convenience wrapper for integer counts: defaults to `decimals = 0`, `thousands_sep = ","`, header `"N"`.

```{r}
#| eval: false
col_n("n")
col_n("enrolled", header = "Enrolled", abbreviate = TRUE)
```

#### `col_percent()`

Formats proportions or percentages.

| Arg | Default | Effect |
|---|---|---|
| `multiply` | `TRUE` | `TRUE` when input is `[0, 1]` (multiply by 100); set `FALSE` if your data is already on the 0–100 scale |
| `symbol` | `TRUE` | Append `%` to the formatted number |

```{r}
#| eval: false
col_percent("rate")                           # 0.05 → "5.0%"
col_percent("accuracy", multiply = FALSE)     # data is already 0–100
col_percent("share", symbol = FALSE)          # drop the "%"
```

#### `col_currency()`

Money with currency symbol prefix/suffix.

| Arg | Default | Effect |
|---|---|---|
| `symbol` | `"$"` | Currency glyph (e.g. `"€"`, `"¥"`, `"R$"`) |
| `position` | `"prefix"` | `"suffix"` for post-fixed currencies |
| `thousands_sep` | `","` | String or `FALSE`; European users often pass `"."` |
| `abbreviate` | `FALSE` | `TRUE` for `"$1.2M"`-style short form |

```{r}
#| eval: false
col_currency("price")
col_currency("amount", symbol = "€", thousands_sep = ".", position = "suffix")
col_currency("revenue", abbreviate = TRUE)    # "$1.2M"
```

#### `col_pvalue()`

P-values with smart formatting and optional significance stars.

| Arg | Default | Effect |
|---|---|---|
| `stars` | `FALSE` | Append `*`/`**`/`***` based on `thresholds` |
| `thresholds` | `c(0.05, 0.01, 0.001)` | Star cutoffs |
| `format` | `"auto"` | `"auto"` / `"scientific"` / `"decimal"` |
| `digits` | `2` | Significant figures |
| `exp_threshold` | `0.001` | Below this, switch to scientific notation |
| `abbrev_threshold` | `NULL` | Below this, render as e.g. `"<0.0001"` |

```{r}
#| eval: false
col_pvalue("pval")
col_pvalue("pval", stars = TRUE)
col_pvalue("pval", abbrev_threshold = 0.0001)
```

### Paired values

#### `col_interval()`

Renders a point estimate with its confidence interval: `0.85 (0.70, 1.03)`.

| Arg | Default | Effect |
|---|---|---|
| `point`, `lower`, `upper` | required | Field names |
| `decimals` | `2` | Decimal places |
| `separator` | `" "` | Separator between point and CI block |
| `imprecise_threshold` | `NULL` | If `upper/lower > threshold`, render `"—"` instead |

```{r}
#| eval: false
col_interval("hr", "lower", "upper", header = "HR (95% CI)")
col_interval("or", "or_lo", "or_hi", decimals = 3, separator = ", ")
col_interval("hr", "lower", "upper", imprecise_threshold = 10)
```

#### `col_events()`

Events / N ratio, optionally with percentage.

```{r}
#| eval: false
col_events("events", "n")                           # "45/120"
col_events("events", "n", show_pct = TRUE)          # "45/120 (37.5%)"
col_events("events", "n", separator = " of ",
           abbreviate = TRUE)                       # "1.1K of 12K"
```

#### `col_range()`

Low – high range, optionally with an inline bar.

```{r}
#| eval: false
col_range("age_min", "age_max", "Age")
col_range("ci_lower", "ci_upper", decimals = 1)
col_range("low_price", "high_price", show_bar = TRUE)
```

### Dates

#### `col_date()`

Dates formatted with `strftime`-style tokens. R-side formatting, so the browser sees pre-formatted strings.

```{r}
#| eval: false
col_date("enrollment_date", format = "%Y-%m-%d")
col_date("published",       format = "%b %Y")          # "Jan 2024"
col_date("visit",           format = "%A, %d %B %Y")    # full
```

### Inline visuals

#### `col_bar()`

Small horizontal bar inside the cell — use for visual weight without a shared axis. For axis-backed bars across rows, reach for [`viz_bar()`](visualizations/bar-charts.qmd).

```{r}
#| eval: false
col_bar("weight")
col_bar("score", max_value = 100, color = "#2563eb")
col_bar("value", scale = "log")
```

#### `col_sparkline()`

Mini line/area/bar chart per row. Requires a list-column of numeric vectors.

```{r}
#| eval: false
col_sparkline("trend")                          # default: line
col_sparkline("trend", type = "area")
col_sparkline("trend", type = "bar", height = 30, color = "#16a34a")
```

:::{.callout-note}
## Preparing sparkline data
Create a list-column with `mutate(trend = list(c(1, 2, 3)))` or `tibble(trend = I(list(...)))`. Vectors do not need to be the same length across rows.
:::

#### `col_heatmap()`

Color-scale cell background with optional numeric overlay.

```{r}
#| eval: false
col_heatmap("score")                                   # default palette
col_heatmap("score", palette = c("#fef3c7", "#92400e"),
            min_value = 0, max_value = 100)
col_heatmap("magnitude", scale = "log", show_value = FALSE)
```

#### `col_progress()`

Horizontal progress bar with numeric label.

```{r}
#| eval: false
col_progress("completion", max_value = 100)
col_progress("share", max_value = 1, show_label = FALSE)
col_progress("fill", color = "#8b5cf6", height = 6)
```

### Status and categories

#### `col_badge()`

Colored chips keyed to row values.

| Arg | Effect |
|---|---|
| `variants` | Named list mapping values → semantic color (`"success"`, `"warning"`, `"error"`, `"info"`, `"muted"`) |
| `colors` | Named list mapping values → literal CSS colors (overrides variants) |
| `size` | `"base"` or `"sm"` |

```{r}
#| eval: false
col_badge("status")                              # plain chips
col_badge("status", variants = list(
  Published = "success", Preprint = "warning", Retracted = "error"
))
col_badge("priority", colors = list(
  high = "#dc2626", medium = "#f59e0b", low = "#22c55e"
))
```

#### `col_icon()`

Icons, emoji, or unicode glyphs.

```{r}
#| eval: false
col_icon("status_icon")                          # emoji already in data
col_icon("result", mapping = list(
  pass = "✔", fail = "✖", pending = "…"
))
col_icon("flag", size = "lg", color = "#16a34a")
```

#### `col_stars()`

Visual rating (filled/empty stars).

```{r}
#| eval: false
col_stars("quality")                              # 1–5 default
col_stars("quality", max_stars = 10, half_stars = TRUE)
col_stars("score_0_100", min_value = 0, max_value = 100)   # remap to star count
col_stars("rating", size = "lg")                            # larger glyphs
```

### Media and links

#### `col_img()`

Inline image cells. Images come from URL/path in the data column.

```{r}
#| eval: false
col_img("logo_url")
col_img("avatar", shape = "circle", width = 40)
col_img("thumbnail", fallback = "No image")
```

#### `col_reference()`

Truncated text (e.g. citation) with optional clickable link and external-link icon.

```{r}
#| eval: false
col_reference("citation")
col_reference("title", href_field = "doi_url",
              max_chars = 40)
col_reference("source", href_field = "url", show_icon = FALSE)
```

### Grouping

#### `col_group()`

Wraps child columns under a shared header row. See [Composing columns](#column-groups) above.

## See also

- [Cell Styling](cell-styling.qmd) — conditional formatting, formula pronouns
- [Row Styling](row-styling.qmd) — row-level bold/italic/color/bg/badge/icon
- [Visualizations](visualizations/index.qmd) — `viz_*()` axis-backed columns
- [`web_col()` reference](../reference/columns.html#web_col) — the low-level constructor
