---
title: "Package Datasets"
---
```{r}
#| include: false
library(webforest)
library(dplyr)
```

webforest includes four datasets for examples, testing, and learning. Each is designed to showcase different features and use cases.

## Available Datasets

| Dataset | Rows | Domain | Key Features |
|---------|------|--------|--------------|
| `glp1_trials` | 24 | Clinical trials | HR/CI, groups, subgroups, summary rows |
| `airline_delays` | 40 | Transportation | Sparkline trends, bar charts, grouped |
| `nba_efficiency` | 30 | Sports | Player stats, conference groups, badges |
| `climate_temps` | 20 | Science | Minimal structure, uncertainty ranges |

## glp1_trials

GLP-1 agonist cardiovascular outcomes trials - ideal for meta-analysis examples.

```{r}
data(glp1_trials)
glimpse(glp1_trials)
```

**Best for:** Meta-analyses, subgroup analyses, publication-style forest plots

```{r}
#| fig-height: 6
# Filter to main trials only
glp1_trials |>
  filter(group == "Main Trials") |>
  forest_plot(
    label = "study",
    point = "hr",
    lower = "lower",
    upper = "upper",
    columns = list(
      col_text("drug", "Drug"),
      col_n("n"),
      col_interval("HR (95% CI)")
    ),
    scale = "log",
    null_value = 1,
    theme = web_theme_lancet(),
    axis_label = "Hazard Ratio",
    title = "GLP-1 Agonist Cardiovascular Outcomes"
  )
```

## airline_delays

Airline carrier performance data with trends - ideal for sparklines and bar charts.

```{r}
data(airline_delays)
glimpse(airline_delays)
```

**Best for:** Sparkline columns, bar charts, grouped comparisons

```{r}
# Aggregate to carrier level for cleaner display
airline_summary <- airline_delays |>
  group_by(carrier) |>
  summarise(
    delay_vs_avg = mean(delay_vs_avg),
    delay_lower = mean(delay_lower),
    delay_upper = mean(delay_upper),
    on_time_pct = mean(on_time_pct),
    trend = list(unlist(trend[1]))
  )

forest_plot(
  airline_summary,
  label = "carrier",
  point = "delay_vs_avg",
  lower = "delay_lower",
  upper = "delay_upper",
  columns = list(
    col_percent("on_time_pct", "On-Time %"),
    col_sparkline("trend", "12M Trend"),
    col_interval("Delay (95% CI)")
  ),
  null_value = 0,
  theme = web_theme_modern(),
  axis_label = "Delay vs. Average (minutes)",
  title = "Airline Carrier Performance"
)
```

## nba_efficiency

NBA player statistics - a fun dataset for demonstrating grouped data and badges.
```{r}
data(nba_efficiency)
glimpse(nba_efficiency)
```

**Best for:** Grouped data, badges, non-medical examples

```{r}
#| fig-height: 7
forest_plot(
  nba_efficiency,
  label = "player",
  point = "per",
  lower = "per_lower",
  upper = "per_upper",
  group = "conference",
  columns = list(
    col_text("team", "Team", width = 60),
    col_numeric("ppg", "PPG"),
    col_interval("PER (95% CI)")
  ),
  row_badge = "award",
  null_value = 15,  # League average PER
  theme = web_theme_modern(),
  axis_label = "Player Efficiency Rating",
  title = "NBA Player Efficiency",
  subtitle = "2023-24 Season",
  caption = "PER: 15 = league average"
)
```

## climate_temps

Regional temperature anomalies - minimal dataset for clean examples.

```{r}
data(climate_temps)
glimpse(climate_temps)
```

**Best for:** Simple examples, minimal themes, uncertainty visualization

```{r}
climate_temps |>
  filter(category %in% c("Global", "Hemisphere", "Continental")) |>
  forest_plot(
    label = "region",
    point = "anomaly",
    lower = "lower",
    upper = "upper",
    group = "category",
    columns = list(
      col_text("certainty", "Confidence"),
      col_interval("Anomaly (95% CI)")
    ),
    null_value = 0,
    theme = web_theme_minimal(),
    axis_label = "Temperature Anomaly (°C)",
    title = "Regional Temperature Anomalies",
    subtitle = "2023 vs. Pre-Industrial Baseline"
  )
```

## Using Datasets in Your Code

All datasets are lazy-loaded. Simply reference them after loading the package:

```r
library(webforest)

# Datasets are available immediately
head(glp1_trials)

# Or explicitly load
data(glp1_trials)
```

See `?glp1_trials`, `?airline_delays`, `?nba_efficiency`, and `?climate_temps` for full documentation.
