---
title: "The Fluent API"
---

```{r}
#| include: false
library(tabviz)
library(dplyr)
```

tabviz exposes a single, unified fluent API for every kind of modification
you'd want to make — styling, column structure, row ordering, data, theme,
zoom, and cell edits. Every verb returns the modified object and dispatches
on input type, so the same call works on a static `WebSpec`, a rendered
htmlwidget, or a live Shiny proxy.

## Two ways to style

Styling can live on the `tabviz()` call or come from fluent modifiers chained after it. Both are fully supported and produce identical output. Pick based on intent:

| Use top-level args (`tabviz(...)`) | Use fluent modifiers (`|> set_*()`) |
|------------------------------------|-------------------------------------|
| One-shot construction | Post-hoc updates to an existing plot |
| Everything known upfront | Building variants from a single base |
| Concise static plots | Live updates inside a Shiny `tabviz_proxy()` |

```{r}
#| eval: false
# Top-level: everything in one call
tabviz(data, label = "study",
  columns = list(
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1)
  ),
  row_type = "rtype",
  row_bold = "rbold",
  marker_color = "effect_color",
  theme = web_theme_jama()
)

# Fluent: same result, modifier-driven
tabviz(data, label = "study",
  columns = list(
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1)
  )
) |>
  set_row_style(type = "rtype", bold = "rbold") |>
  set_marker_style(color = "effect_color") |>
  set_theme("jama")
```

For Shiny proxies the choice is made for you — incremental updates **must** go through fluent verbs.

## Static vs Proxy Dispatch

Every fluent verb accepts three kinds of input and behaves accordingly:

| Input type | What happens |
|------------|--------------|
| **`WebSpec`** (from `tabviz(..., .spec_only = TRUE)`) | Returns a new `WebSpec` with the change applied in R. |
| **htmlwidget** (from `tabviz(...)`) | Returns a new widget with the change baked into its spec. |
| **`tabviz_proxy`** (from `tabviz_proxy("id")`) | Sends a custom message to the running widget in Shiny — no re-render. |

```r
# Static WebSpec
spec <- tabviz(data, label = "study", .spec_only = TRUE) |>
  sort_rows("hr", "desc") |>
  add_column(col_numeric("n", "N"), after = "study")

# htmlwidget — same verbs
widget <- tabviz(data, ...) |>
  set_theme("jama") |>
  resize_column("hr", width = 140)

# Shiny proxy — same verbs
observeEvent(input$sort_hr, {
  tabviz_proxy("plot") |>
    sort_rows("hr", "desc")
})
```

A handful of verbs are **proxy-only** because they depend on row identity
post-render (`set_cell`, `set_row_label`, `clear_edits`). On a static
input they emit a warning and leave the object unchanged.

---

## The Fluent Pattern

Every verb returns the modified object, so calls chain:

```r
tabviz(data, ...) |>     # Create plot
  set_row_style(...) |>  # Modify rows
  add_column(...) |>     # Add a column
  sort_rows(...) |>      # Reorder rows
  set_theme(...)         # Apply theme
```

::: {.callout-note}
## Immutability

Each verb returns a **new** object; the input is unchanged. Safe to derive
multiple variants from a single base:

```r
base <- tabviz(data, ...)
jama   <- base |> set_theme("jama")
lancet <- base |> set_theme("lancet")
```
:::

---

## Styling Verbs

All styling verbs accept either **column names** (strings) or **formula
expressions** (`~ ...`):

```r
# Using column names (pre-computed)
plot |> set_row_style(bold = "is_significant")

# Using formula expressions (computed on the fly)
plot |> set_row_style(bold = ~ pvalue < 0.05)

# Mix both
plot |> set_marker_style(
  color = ~ ifelse(upper < 1, "#16a34a", "#64748b"),
  shape = "marker_shape_column"
)
```

See [Row Styling](row-styling.qmd#expression-based-styling) and
[Cell Styling](cell-styling.qmd#formula-expressions-for-cell-styling) for
formula details.

### set_row_style()

Row-level styling. Your primary tool for visual hierarchy — distinguishing
headers from data rows, highlighting summaries, showing nested structure.

```r
plot |> set_row_style(
  type    = "rtype",    # "header", "data", "summary", "spacer"
  bold    = "rbold",
  italic  = "ritalic",
  color   = "rcolor",   # CSS colors
  bg      = "rbg",
  badge   = "rbadge",   # text label anchored to first column
  icon    = "ricon",    # emoji/unicode anchored to first column
  indent  = "rindent"   # 0, 1, 2, ...
)
```

::: {.callout-warning}
`badge` and `icon` render into the **first visible text column's cell**.
If you reorder columns, they re-anchor to the new first column. See
[Binding rules](row-styling.qmd#binding-rules) for details.
:::

```{r}
styled_data <- data.frame(
  study = c("Primary", "  CV Death", "  MI", "Secondary", "  Stroke"),
  hr = c(NA, 0.82, 0.79, NA, 0.88),
  lower = c(NA, 0.72, 0.68, NA, 0.74),
  upper = c(NA, 0.94, 0.92, NA, 1.05),
  rtype = c("header", "data", "data", "header", "data"),
  rbold = c(TRUE, FALSE, FALSE, TRUE, FALSE),
  rindent = c(0, 1, 1, 0, 1)
)

forest_plot(styled_data,
  point = "hr", lower = "lower", upper = "upper", label = "study",
  scale = "log", null_value = 1
) |>
  set_row_style(type = "rtype", bold = "rbold", indent = "rindent") |>
  set_theme("modern")
```

### set_marker_style()

Interval marker styling (points and lines for effect estimates):

```r
plot |> set_marker_style(
  color   = "marker_color",  # CSS colors
  shape   = "marker_shape",  # "square", "circle", "diamond", "triangle"
  opacity = "marker_alpha",  # 0–1
  size    = "marker_size"    # size multiplier
)
```

```{r}
marker_data <- data.frame(
  study = c("Trial A", "Trial B", "Trial C", "Trial D"),
  hr = c(0.72, 0.95, 0.68, 1.15),
  lower = c(0.55, 0.78, 0.52, 0.90),
  upper = c(0.94, 1.15, 0.89, 1.47),
  direction = c("benefit", "neutral", "benefit", "harm")
) |>
  mutate(
    marker_color = case_when(
      upper < 1 ~ "#16a34a",
      lower > 1 ~ "#dc2626",
      TRUE ~ "#64748b"
    ),
    marker_shape = ifelse(direction == "benefit", "circle", "square")
  )

forest_plot(marker_data,
  point = "hr", lower = "lower", upper = "upper", label = "study",
  scale = "log", null_value = 1
) |>
  set_marker_style(color = "marker_color", shape = "marker_shape") |>
  set_theme("modern")
```

### set_column_style()

Per-cell styling for a specific column:

```r
plot |> set_column_style(column = "hr",
  bold  = "is_sig",
  color = "hr_color",
  bg    = "hr_bg"
)
```

::: {.callout-tip}
## Styling priority

When both row and column styles target the same cell, **column styles
win**. Row styles apply to cells without column overrides — so you can set
row-wide defaults with column-specific highlights.
:::

### set_theme() and theme modifiers

```r
# Apply by name
plot |> set_theme("jama")

# Apply by object
plot |> set_theme(web_theme_lancet())

# Modify theme properties
plot |> set_colors(primary = "#1a365d")
plot |> set_spacing(row_height = 24)
plot |> set_axis(gridlines = TRUE)
```

Theme modifiers (`set_colors`, `set_typography`, `set_spacing`, `set_shapes`,
`set_axis`, `set_layout`, `set_group_headers`, `set_effect_colors`,
`set_marker_shapes`) work on both plots and standalone `WebTheme` objects.
See [Styling reference](../reference/styling.qmd) for parameter tables.

---

## Column Operations

Restructure the columns of a running plot without re-rendering.

### add_column()

Insert a new column. Use `after` to position it; `NULL` appends.

```r
plot |> add_column(col_numeric("year", "Year"), after = "n")
plot |> add_column(viz_forest(point = "hr", lower = "lower", upper = "upper"))
```

**Static:**

```{r}
df <- data.frame(
  study = c("A", "B", "C"),
  hr = c(0.8, 1.1, 1.4),
  n = c(100, 200, 150)
)
tabviz(df, label = "study",
  columns = list(col_numeric("hr", "HR"))
) |>
  add_column(col_numeric("n", "N"), after = "hr")
```

**Proxy:**

```r
observeEvent(input$show_year, {
  tabviz_proxy("plot") |>
    add_column(col_numeric("year", "Year"), after = "n")
})
```

### remove_column()

```r
plot |> remove_column("notes")
```

On a proxy this hides the column without dropping data; it can be re-added
later.

### move_column()

```r
# Numeric index (1-based on R side; verb converts to 0-based for the store)
plot |> move_column("notes", to = 1)

# Or position relative to another column
plot |> move_column("notes", before = "hr")
```

### resize_column()

```r
plot |> resize_column("hr", width = 140)   # pixels
```

### update_column()

Change column properties — header, alignment, wrap, width, or any option
in the column's `options` list:

```r
plot |> update_column("hr", header = "Hazard Ratio", align = "right")
plot |> update_column("hr", options = list(decimals = 3))
```

Unknown keys are merged into `options`, matching R-side semantics.

---

## Row & Data Operations

### sort_rows()

```r
plot |> sort_rows("hr", direction = "asc")
plot |> sort_rows("hr", direction = "desc")
plot |> sort_rows("hr", direction = "none")   # restore original order
```

On a static `WebSpec`, `direction = "none"` is a no-op. On a proxy it
clears the sort.

### filter_rows() and clear_filters()

Operators: `"eq"`, `"neq"`, `"gt"`, `"lt"`, `"contains"`.

```r
plot |> filter_rows("n", operator = "gt", value = 100)
plot |> filter_rows("drug_class", operator = "eq", value = "SGLT2i")

plot |> clear_filters()
```

On a `WebSpec`, `filter_rows()` mutates `@data` in place. On a proxy it
applies a display-only filter — the original data is preserved.

### select_rows() and move_row()

```r
plot |> select_rows(c("study_A", "study_B"))       # by row id
plot |> move_row("study_C", to = 1)                # to top
```

`select_rows()` on a proxy forwards selection state to any Shiny input
bindings tied to the widget.

### update_data()

Replace the underlying data. On a proxy, this requires a full `WebSpec`
(not just a data frame) because column types must be re-serialized:

```r
new_spec <- tabviz(new_df, label = "study", columns = list(...),
                  .spec_only = TRUE)
tabviz_proxy("plot") |> update_data(new_spec)
```

---

## Groups & Zoom

### toggle_group()

Collapse or expand a group header:

```r
plot |> toggle_group("North America")               # flip current state
plot |> toggle_group("North America", collapsed = TRUE)
plot |> toggle_group("Europe",        collapsed = FALSE)
```

### set_zoom()

```r
plot |> set_zoom(zoom = 0.8)
plot |> set_zoom(zoom = 1.0, auto_fit = TRUE)
plot |> set_zoom(max_width = 900, max_height = 600)
plot |> set_zoom(show_controls = FALSE)
```

Any subset of arguments may be passed; unchanged knobs are left alone.

---

## Cell Edits (proxy-only)

These verbs mutate rendered state and require a live widget. On a static
`WebSpec` they warn and return unchanged.

```r
# Change a single cell value
tabviz_proxy("plot") |> set_cell("row_id", "hr", 1.23)

# Rename a row's label
tabviz_proxy("plot") |> set_row_label("row_id", "Renamed")

# Drop all in-browser edits
tabviz_proxy("plot") |> clear_edits()
```

---

## Conditional Styling with dplyr

Compute style columns with dplyr before plotting:

```{r}
analysis_data <- data.frame(
  study = c("EMPA-REG", "CANVAS", "DECLARE", "DAPA-HF"),
  hr = c(0.62, 0.86, 0.93, 0.74),
  lower = c(0.49, 0.75, 0.84, 0.65),
  upper = c(0.77, 0.97, 1.03, 0.85),
  pval = c(0.0001, 0.02, 0.17, 0.0001)
)

styled_analysis <- analysis_data |>
  mutate(
    is_sig = pval < 0.05,
    sig_color = ifelse(is_sig, "#16a34a", "#94a3b8")
  )

forest_plot(styled_analysis,
  point = "hr", lower = "lower", upper = "upper", label = "study",
  columns = list(col_pvalue("pval")),
  scale = "log", null_value = 1
) |>
  set_marker_style(color = "sig_color") |>
  set_row_style(bold = "is_sig") |>
  set_theme("modern")
```

::: {.callout-tip}
## Style column naming

Use consistent suffixes: `_bold`, `_color`, `_bg`, `_badge`. Makes it
clear which columns are data vs styling.
:::

---

## Reusable Pipelines

Encapsulate your organization's styling standards as a function:

```{r}
#| eval: false
apply_clinical_style <- function(plot) {
  plot |>
    set_marker_style(color = "sig_color") |>
    set_row_style(bold = "is_sig") |>
    set_theme("jama")
}

forest_plot(trial_a, ...) |> apply_clinical_style()
forest_plot(trial_b, ...) |> apply_clinical_style()
```

---

## Verb Reference

**Styling**

| Verb | Purpose |
|------|---------|
| `set_row_style()` | Row-level (bold, italic, color, type, indent, badge, icon) |
| `set_marker_style()` | Interval marker (color, shape, opacity, size) |
| `set_column_style()` | Per-cell styling for one column |
| `set_theme()` | Apply theme by name or object |
| `set_colors()`, `set_typography()`, `set_spacing()`, `set_shapes()`, `set_axis()`, `set_layout()`, `set_group_headers()`, `set_effect_colors()`, `set_marker_shapes()` | Theme-level modifiers |

**Columns**

| Verb | Purpose |
|------|---------|
| `add_column()` | Insert at a position |
| `remove_column()` | Drop / hide |
| `move_column()` | Reorder by index or relative position |
| `resize_column()` | Set width in pixels |
| `update_column()` | Modify header, alignment, options |

**Rows & data**

| Verb | Purpose |
|------|---------|
| `sort_rows()` | Sort by a column |
| `filter_rows()` | Apply a predicate |
| `clear_filters()` | Remove all filters |
| `select_rows()` | Set the selected-row set |
| `move_row()` | Reorder by index |
| `update_data()` | Replace underlying data (WebSpec on proxy) |

**Groups, cells, global**

| Verb | Purpose |
|------|---------|
| `toggle_group()` | Collapse/expand a group header |
| `set_cell()` | Change a cell value *(proxy-only)* |
| `set_row_label()` | Rename a row's label *(proxy-only)* |
| `clear_edits()` | Drop in-browser edits *(proxy-only)* |
| `set_zoom()` | Zoom / auto-fit / dimensions |

See [Styling reference](../reference/styling.qmd) and
[Data-ops reference](../reference/data-ops.qmd) for complete parameter
documentation.
