---
title: "Interactivity"
---

```{r}
#| include: false
library(tabviz)
library(dplyr)
```

Every tabviz widget is fully interactive by default. This page is the authoritative reference for `web_interaction()` — the spec object that toggles which UI chrome appears and which user capabilities are enabled.

## Default Interactions

Every tabviz widget includes these interactions out of the box:

| Interaction | Action |
|---|---|
| **Sort** | Click column headers to sort |
| **Select** | Click rows to select/deselect |
| **Hover** | Hover to highlight rows |
| **Collapse** | Click group headers to collapse/expand |
| **Resize** | Drag column borders to resize |
| **Filter** | Per-column filter popovers in headers |
| **Reorder** | Drag rows or columns to reorder |
| **Edit** | Double-click cells to edit — also works on existing title / subtitle / caption / footnote |
| **Paint** | Paint-brush button opens a popover for stamping `emphasis` / `muted` / `accent` onto rows or cells |
| **Export** | Download menu in header (hover to reveal) |
| **Themes** | Theme switcher in header |

```{r}
data(glp1_trials)

glp1_trials |>
  filter(row_type == "data", group == "Main Trials") |>
  head(5) |>
  tabviz(
    label = "study",
    columns = list(
      col_text("drug", "Drug"),
      viz_forest(point = "hr", lower = "lower", upper = "upper",
                 scale = "log", null_value = 1, width = 160),
      col_interval("hr", "lower", "upper", header = "HR (95% CI)")
    ),
    theme = web_theme_jama(),
    footnote = "Try: click headers to sort, hover over rows, drag column borders"
  )
```

## The `show_*` vs `enable_*` Rule

`web_interaction()` uses two prefixes deliberately:

- **`show_*`** controls **static UI visibility** — whether a piece of chrome appears at all (`show_legend`, `show_group_counts`). Nothing the user does at runtime brings it back.
- **`enable_*`** controls **user capabilities** — whether a behavior is available (`enable_sort`, `enable_filters`, `enable_edit`, …). The associated UI is rendered when the capability is on, hidden when off.

When in doubt: render-or-not is `show_`; can-the-user-do-it is `enable_`.

## `web_interaction()` Arguments

| Argument | Default | Purpose |
|---|---|---|
| `show_legend` | `TRUE` | Show legend for multi-effect plots |
| `show_group_counts` | `FALSE` | Show `(n)` after each row-group label |
| `tooltip_fields` | `NULL` | Character vector of column names to show in marker hover tooltip |
| `enable_sort` | `TRUE` | Click headers to sort |
| `enable_collapse` | `TRUE` | Click group headers to collapse/expand |
| `enable_select` | `TRUE` | Click rows to select |
| `enable_hover` | `TRUE` | Highlight rows on hover |
| `enable_resize` | `TRUE` | Drag column borders to resize |
| `enable_export` | `TRUE` | Show download button in header |
| `enable_filters` | `TRUE` | Per-column filter popovers in column headers |
| `enable_reorder_rows` | `TRUE` | Drag to reorder rows (within a group) |
| `enable_reorder_columns` | `TRUE` | Drag to reorder columns (within a column group) |
| `enable_edit` | `TRUE` | Double-click inline editing for text/numeric/label cells |
| `enable_themes` | `"default"` | Theme menu: `"default"`, `NULL`, or list of `WebTheme` |

## Customizing Interactions

```{r}
glp1_trials |>
  filter(row_type == "data") |>
  head(4) |>
  tabviz(
    label = "study",
    columns = list(
      col_text("drug", "Drug"),
      viz_forest(point = "hr", lower = "lower", upper = "upper",
                 scale = "log", null_value = 1, width = 160)
    ),
    interaction = web_interaction(
      enable_sort = FALSE,
      enable_select = FALSE
    ),
    theme = web_theme_jama()
  )
```

## Hover Tooltips

Add tooltips that appear when hovering over forest plot markers:

```{r}
glp1_trials |>
  filter(row_type == "data") |>
  head(4) |>
  tabviz(
    label = "study",
    columns = list(
      viz_forest(point = "hr", lower = "lower", upper = "upper",
                 scale = "log", null_value = 1, width = 180),
      col_interval("hr", "lower", "upper", header = "HR (95% CI)")
    ),
    interaction = web_interaction(
      tooltip_fields = c("drug", "n", "events", "pvalue")
    ),
    theme = web_theme_jama(),
    footnote = "Hover over markers to see details"
  )
```

Tooltips show the values from the specified columns for each row.

## Presets

tabviz ships three preset configurations.

| Preset | Description |
|---|---|
| `web_interaction()` | Full interactivity (the default) |
| `web_interaction_minimal()` | Hover only — no sorting, selection, resize, export, filtering, reorder, edit |
| `web_interaction_publication()` | Fully static — every interaction off, no legend |
| `web_interaction_full()` | Explicit "everything on" — same as `web_interaction()` defaults |

### Minimal Interaction

```{r}
glp1_trials |>
  filter(row_type == "data") |>
  head(4) |>
  tabviz(
    label = "study",
    columns = list(
      col_text("drug", "Drug"),
      viz_forest(point = "hr", lower = "lower", upper = "upper",
                 scale = "log", null_value = 1, width = 160)
    ),
    interaction = web_interaction_minimal(),
    theme = web_theme_jama()
  )
```

### Publication Mode

`web_interaction_publication()` removes all interactivity — suitable for print, static documents, or screenshots.

```{r}
#| eval: false
tabviz(data, label = "study",
  columns = list(...),
  interaction = web_interaction_publication()
)
```

## Theme Menu

The `enable_themes` argument controls the theme switcher. Three values are accepted:

| Value | Effect |
|---|---|
| `"default"` | Show all themes returned by `package_themes()` |
| `NULL` | Hide the theme menu entirely |
| `list(...)` | Show only the listed `WebTheme` objects (named entries override displayed names) |

```{r}
#| eval: false
# Hide the theme menu
tabviz(data, interaction = web_interaction(enable_themes = NULL), ...)

# Curate a specific shortlist
tabviz(data, interaction = web_interaction(
  enable_themes = list(
    web_theme_jama(),
    web_theme_nature(),
    web_theme_modern()
  )
), ...)
```

`selectable_themes()` is a fluent sibling that does the same thing post-construction:

```{r}
#| eval: false
tabviz(data, label = "study") |>
  set_theme("jama") |>
  selectable_themes(list(
    Classical = web_theme_jama(),
    Modern    = web_theme_modern()
  ))
```

The active theme is always auto-included in the menu so the user can revert.

## Pan & Zoom on Visualization Columns

Every viz column (`viz_forest`, `viz_bar`, `viz_boxplot`, `viz_violin`) supports direct-manipulation pan and zoom on its x-axis:

| Gesture | Action |
|---|---|
| **Mouse wheel** over the viz area | Zoom in/out, anchored on the cursor |
| **Click and drag** | Pan the axis |
| **Double-click** | Reset to the default domain |

State is per column — two forest columns side-by-side can be inspected independently. It is session-only (refreshing the page resets everything); there is no R-side argument to persist a zoom level.

Exports are WYSIWYG: zooming or panning in the browser changes what `save_plot()` writes to SVG/PNG and what the in-widget download button produces. Confidence intervals that fall outside the zoomed window are rendered with boundary arrows; marks on bar/boxplot/violin columns are clipped to the visible region.

On log-scale forests (`viz_forest(scale = "log")`), gestures that would push the lower bound to zero or below are ignored.

## Row Selection

Selected rows are tracked internally and persist during sorting and filtering.

```{r}
sample_data <- data.frame(
  item = c("Alpha", "Beta", "Gamma", "Delta"),
  value = c(10, 25, 15, 30),
  change = c(0.05, -0.12, 0.08, 0.22)
)

tabviz(sample_data, label = "item",
  columns = list(
    col_numeric("value", "Value"),
    col_percent("change", "Change")
  ),
  theme = web_theme_modern(),
  footnote = "Click rows to select (indicated by left border)"
)
```

## Paint Tool (v0.15+)

The paint-brush button on the widget toolbar opens a small popover
where authors can stamp semantic classes onto rows or cells
interactively. This is the in-widget equivalent of the R-side
`row_emphasis = <col>` / `row_muted = <col>` / `row_accent = <col>`
arguments, but applied by clicking rather than coded up front.

**Flow:**

1. Open the paint popover (brush icon in the top-right toolbar).
2. Pick a token: **Emphasis**, **Muted**, or **Accent**.
3. Flip the **Scope** switch between **Row** (whole row) and **Cell**
   (just one cell).
4. Click any row or cell to toggle the active flag on/off. The cursor
   turns into a crosshair while paint is active so you can spot where
   the next click lands.
5. Click the prominent **Exit paint mode** button at the top of the
   popover (or press **Esc**) when done. The brush icon stays lit in
   the toolbar's accent color as long as paint is armed.

**Persistence:** painted flags merge into `exportSpec`, so
`save_plot()` picks them up, and the flags round-trip through the
"View source" button in the settings panel. You can also clear all
paint with the "Clear all paint" button if you want to start over.

Painted flags coexist with R-side `row_emphasis` / etc. columns: the
R column sets the baseline, paint overrides per row.

## Configure Columns (v0.18+)

Every `col_*` and `viz_*` column has a **Configure…** entry in its
header context menu (right-click the column header, or use the
column-menu chevron). The popover exposes the column's relevant
authoring arguments so you can tweak them without rebuilding the
widget in R:

- **`col_numeric`** — decimals, format.
- **`col_interval`** — decimals, separator, imprecise threshold.
- **`col_text`** — max chars, wrap.
- **`viz_forest`** — scale, null value, axis label, and a full
  **Axis** section (range, tick count/values, gridlines, null tick,
  marker margin, symmetric).
- **`viz_bar` / `viz_boxplot` / `viz_violin`** — full CRUD on
  effects (add / remove / reorder / rename / recolor / opacity /
  data field), plus the same Axis section. Boxplot can toggle
  between array-shaped and pre-computed stats data.

Double-clicking a viz column's axis opens the configure popover
pre-scrolled to the Axis section.

Changes merge into `exportSpec` and round-trip through the "View
source" button — `save_plot()` picks them up. The interactive Axis
*theme* tab was retired in v0.18 since axis values are now per-column.

## Title / Subtitle / Caption / Footnote Editing (v0.15+)

When `enable_edit` is on:

- **Double-click** any existing title, subtitle, caption, or footnote
  to edit it inline (Enter to commit, Esc to cancel).
- **Add** a missing label via the **Basics** tab in the settings
  panel — it exposes four text fields (Title / Subtitle / Caption /
  Footnote). Populating any of them makes the matching slot appear.

Edits merge into `exportSpec.labels`, so `save_plot()` reflects them.

## View Source (v0.20+)

The **`</>` icon** in the widget toolbar opens a three-tab panel
showing the current widget as fluent R code:

- **Table ops** — your original `tabviz(...)` call (captured as
  written) plus every interactive structural change you've made
  (resize / add / remove / move columns, move rows, cell / title /
  caption / footnote edits, paint stamps, theme switches,
  watermark, split-forest align-widths toggle).
- **Theme** — the current theme state as a `web_theme_*()  |>
  set_*()` chain. Identical to the "View theme source" panel that
  used to live inside the Settings sidebar.
- **Combined** — both, stitched together:

```r
mytheme <- web_theme_default() |>
  set_inputs(brand = "#0891b2")

tbl <- tabviz(dt, label = "study", columns = list(...), theme = mytheme) |>
  resize_column("drug", 160) |>
  move_column("drug", to = 2) |>
  set_title("Updated")
```

The log is **append-only**: if you resize a column from 120 to
150 and back, both calls appear in order. No attempt is made to
dedupe or optimize — the raw sequence is the history.

The corresponding fluent R functions are all exported from the
package, so the emitted chain pastes into R as-is.

## See Also

- [Reference: Interaction Settings](../reference/interaction.qmd) — Full API reference
- [Export](export.qmd) — Exporting interactive tables
- [Shiny Integration](shiny.qmd) — Interactivity in Shiny apps
