---
title: "Interactivity"
---

```{r}
#| include: false
library(tabviz)
library(dplyr)
```

Every tabviz widget is fully interactive by default. This page is the authoritative reference for `web_interaction()` — the spec object that toggles which UI chrome appears and which user capabilities are enabled.

## Default Interactions

Every tabviz widget includes these interactions out of the box:

| Interaction | Action |
|---|---|
| **Sort** | Click column headers to sort |
| **Select** | Click rows to select/deselect |
| **Hover** | Hover to highlight rows |
| **Collapse** | Click group headers to collapse/expand |
| **Resize** | Drag column borders to resize |
| **Filter** | Per-column filter popovers in headers |
| **Reorder** | Drag rows or columns to reorder |
| **Edit** | Double-click cells to edit |
| **Export** | Download menu in header (hover to reveal) |
| **Themes** | Theme switcher in header |

```{r}
data(glp1_trials)

glp1_trials |>
  filter(row_type == "data", group == "Main Trials") |>
  head(5) |>
  tabviz(
    label = "study",
    columns = list(
      col_text("drug", "Drug"),
      viz_forest(point = "hr", lower = "lower", upper = "upper",
                 scale = "log", null_value = 1, width = 160),
      col_interval("hr", "lower", "upper", header = "HR (95% CI)")
    ),
    theme = web_theme_jama(),
    footnote = "Try: click headers to sort, hover over rows, drag column borders"
  )
```

## The `show_*` vs `enable_*` Rule

`web_interaction()` uses two prefixes deliberately:

- **`show_*`** controls **static UI visibility** — whether a piece of chrome appears at all (`show_legend`, `show_group_counts`). Nothing the user does at runtime brings it back.
- **`enable_*`** controls **user capabilities** — whether a behavior is available (`enable_sort`, `enable_filters`, `enable_edit`, …). The associated UI is rendered when the capability is on, hidden when off.

When in doubt: render-or-not is `show_`; can-the-user-do-it is `enable_`.

## `web_interaction()` Arguments

| Argument | Default | Purpose |
|---|---|---|
| `show_legend` | `TRUE` | Show legend for multi-effect plots |
| `show_group_counts` | `FALSE` | Show `(n)` after each row-group label |
| `tooltip_fields` | `NULL` | Character vector of column names to show in marker hover tooltip |
| `enable_sort` | `TRUE` | Click headers to sort |
| `enable_collapse` | `TRUE` | Click group headers to collapse/expand |
| `enable_select` | `TRUE` | Click rows to select |
| `enable_hover` | `TRUE` | Highlight rows on hover |
| `enable_resize` | `TRUE` | Drag column borders to resize |
| `enable_export` | `TRUE` | Show download button in header |
| `enable_filters` | `TRUE` | Per-column filter popovers in column headers |
| `enable_reorder_rows` | `TRUE` | Drag to reorder rows (within a group) |
| `enable_reorder_columns` | `TRUE` | Drag to reorder columns (within a column group) |
| `enable_edit` | `TRUE` | Double-click inline editing for text/numeric/label cells |
| `enable_themes` | `"default"` | Theme menu: `"default"`, `NULL`, or list of `WebTheme` |

## Customizing Interactions

```{r}
glp1_trials |>
  filter(row_type == "data") |>
  head(4) |>
  tabviz(
    label = "study",
    columns = list(
      col_text("drug", "Drug"),
      viz_forest(point = "hr", lower = "lower", upper = "upper",
                 scale = "log", null_value = 1, width = 160)
    ),
    interaction = web_interaction(
      enable_sort = FALSE,
      enable_select = FALSE
    ),
    theme = web_theme_jama()
  )
```

## Hover Tooltips

Add tooltips that appear when hovering over forest plot markers:

```{r}
glp1_trials |>
  filter(row_type == "data") |>
  head(4) |>
  tabviz(
    label = "study",
    columns = list(
      viz_forest(point = "hr", lower = "lower", upper = "upper",
                 scale = "log", null_value = 1, width = 180),
      col_interval("hr", "lower", "upper", header = "HR (95% CI)")
    ),
    interaction = web_interaction(
      tooltip_fields = c("drug", "n", "events", "pvalue")
    ),
    theme = web_theme_jama(),
    footnote = "Hover over markers to see details"
  )
```

Tooltips show the values from the specified columns for each row.

## Presets

tabviz ships three preset configurations.

| Preset | Description |
|---|---|
| `web_interaction()` | Full interactivity (the default) |
| `web_interaction_minimal()` | Hover only — no sorting, selection, resize, export, filtering, reorder, edit |
| `web_interaction_publication()` | Fully static — every interaction off, no legend |
| `web_interaction_full()` | Explicit "everything on" — same as `web_interaction()` defaults |

### Minimal Interaction

```{r}
glp1_trials |>
  filter(row_type == "data") |>
  head(4) |>
  tabviz(
    label = "study",
    columns = list(
      col_text("drug", "Drug"),
      viz_forest(point = "hr", lower = "lower", upper = "upper",
                 scale = "log", null_value = 1, width = 160)
    ),
    interaction = web_interaction_minimal(),
    theme = web_theme_jama()
  )
```

### Publication Mode

`web_interaction_publication()` removes all interactivity — suitable for print, static documents, or screenshots.

```{r}
#| eval: false
tabviz(data, label = "study",
  columns = list(...),
  interaction = web_interaction_publication()
)
```

## Theme Menu

The `enable_themes` argument controls the theme switcher. Three values are accepted:

| Value | Effect |
|---|---|
| `"default"` | Show all themes returned by `package_themes()` |
| `NULL` | Hide the theme menu entirely |
| `list(...)` | Show only the listed `WebTheme` objects (named entries override displayed names) |

```{r}
#| eval: false
# Hide the theme menu
tabviz(data, interaction = web_interaction(enable_themes = NULL), ...)

# Curate a specific shortlist
tabviz(data, interaction = web_interaction(
  enable_themes = list(
    web_theme_jama(),
    web_theme_nature(),
    web_theme_modern()
  )
), ...)
```

`selectable_themes()` is a fluent sibling that does the same thing post-construction:

```{r}
#| eval: false
tabviz(data, label = "study") |>
  set_theme("jama") |>
  selectable_themes(list(
    Classical = web_theme_jama(),
    Modern    = web_theme_modern()
  ))
```

The active theme is always auto-included in the menu so the user can revert.

## Pan & Zoom on Visualization Columns

Every viz column (`viz_forest`, `viz_bar`, `viz_boxplot`, `viz_violin`) supports direct-manipulation pan and zoom on its x-axis:

| Gesture | Action |
|---|---|
| **Mouse wheel** over the viz area | Zoom in/out, anchored on the cursor |
| **Click and drag** | Pan the axis |
| **Double-click** | Reset to the default domain |

State is per column — two forest columns side-by-side can be inspected independently. It is session-only (refreshing the page resets everything); there is no R-side argument to persist a zoom level.

Exports are WYSIWYG: zooming or panning in the browser changes what `save_plot()` writes to SVG/PNG and what the in-widget download button produces. Confidence intervals that fall outside the zoomed window are rendered with boundary arrows; marks on bar/boxplot/violin columns are clipped to the visible region.

On log-scale forests (`viz_forest(scale = "log")`), gestures that would push the lower bound to zero or below are ignored.

## Row Selection

Selected rows are tracked internally and persist during sorting and filtering.

```{r}
sample_data <- data.frame(
  item = c("Alpha", "Beta", "Gamma", "Delta"),
  value = c(10, 25, 15, 30),
  change = c(0.05, -0.12, 0.08, 0.22)
)

tabviz(sample_data, label = "item",
  columns = list(
    col_numeric("value", "Value"),
    col_percent("change", "Change")
  ),
  theme = web_theme_modern(),
  footnote = "Click rows to select (indicated by left border)"
)
```

## See Also

- [Reference: Interaction Settings](../reference/interaction.qmd) — Full API reference
- [Export](export.qmd) — Exporting interactive tables
- [Shiny Integration](shiny.qmd) — Interactivity in Shiny apps
