---
title: "Quick Start"
---

```{r}
#| include: false
library(tabviz)
library(dplyr)
```

Build your first interactive forest plot in about 30 lines, then layer in styling, themes, and conditional formatting.

## Installation

```r
# Install from GitHub using pak
install.packages("pak")
pak::pak("kaskarn/tabviz")
```

:::{.callout-note}
## Key concepts

- **Label column**: the row identifier (study name, category, item).
- **Column types**: `col_*()` for data columns, `viz_*()` for visualizations.
- **Row types**: `"data"`, `"header"`, `"summary"`, `"spacer"`.
- **Column mapping**: pass column names as strings; values are computed per row.
:::

## Your first table

```{r}
data(glp1_trials)

meta_data <- glp1_trials |>
  filter(row_type == "data", group == "Main Trials") |>
  head(5)

tabviz(meta_data,
  label = "study",
  columns = list(
    col_n("n"),
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1,
               axis_label = "Hazard Ratio (95% CI)"),
    col_interval("hr", "lower", "upper", header = "HR (95% CI)")
  )
)
```

:::{.callout-tip}
## Built-in datasets
This example uses `glp1_trials`, one of five datasets shipped with the package. See [Package Data](package-data.qmd) for all available datasets.
:::

## The core pattern

```r
tabviz(data,
  label = "label_column",    # row labels
  columns = list(...)        # what columns to show
)
```

Each `col_*()` function defines a column. The most common helpers:

| Function | Purpose | Example |
|----------|---------|---------|
| `col_text()` | Text | `col_text("category")` |
| `col_numeric()` | Numbers | `col_numeric("value", decimals = 1)` |
| `col_n()` | Sample sizes | `col_n("n")` |
| `viz_forest()` | Forest plot | `viz_forest(point = "hr", lower = "lo", upper = "hi")` |
| `col_interval()` | Point (CI) | `col_interval("hr", "lo", "hi", header = "HR")` |
| `col_pvalue()` | P-values | `col_pvalue("pval")` |

See [Columns](columns.qmd) for all 16+ column types.

## Adding more columns

```{r}
tabviz(meta_data,
  label = "study",
  columns = list(
    col_text("drug", "Drug"),
    col_n("n"),
    col_numeric("events", header = "Events"),
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1),
    col_interval("hr", "lower", "upper", header = "HR (95% CI)")
  )
)
```

## Applying themes

tabviz includes 9 publication-ready themes:

```{r}
#| layout-ncol: 2
#| fig-height: 3

tabviz(meta_data[1:3, ],
  label = "study",
  columns = list(
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1)
  ),
  theme = web_theme_jama(), title = "JAMA"
)

tabviz(meta_data[1:3, ],
  label = "study",
  columns = list(
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1)
  ),
  theme = web_theme_lancet(), title = "Lancet"
)
```

**Available themes**: `web_theme_cochrane()`, `web_theme_jama()`, `web_theme_lancet()`, `web_theme_cochrane()`, `web_theme_cochrane()`, `web_theme_cochrane()`, `web_theme_cochrane() |> set_variants(density = "spacious")`, `web_theme_dark()`, `web_theme_jama()`. See [Themes](themes.qmd).

## Row styling

Build visual hierarchy with row types and styling:

```{r}
styled_data <- data.frame(
  label = c("Primary Endpoint", "  CV Death", "  MI", "Summary"),
  hr = c(NA, 0.82, 0.79, 0.78),
  lower = c(NA, 0.72, 0.68, 0.65),
  upper = c(NA, 0.94, 0.92, 0.93),
  rtype = c("header", "data", "data", "summary"),
  rbold = c(TRUE, FALSE, FALSE, TRUE)
)

tabviz(styled_data,
  label = "label",
  columns = list(
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1),
    col_interval("hr", "lower", "upper", header = "HR (95% CI)")
  ),
  row_type = "rtype",
  row_bold = "rbold"
)
```

:::{.callout-note}
## Row types

| Type | Behavior |
|------|----------|
| `"data"` | Standard row with interval marker |
| `"header"` | Section header, no marker |
| `"summary"` | Summary row with diamond marker |
| `"spacer"` | Empty row for separation |

Header and spacer rows should have `NA` for point, lower, and upper.
:::

## Conditional styling

Style rows based on data values using formula expressions:

```{r}
tabviz(meta_data,
  label = "study",
  columns = list(
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1),
    col_interval("hr", "lower", "upper", header = "HR (95% CI)")
  ),
  row_bold = ~ upper < 1,
  marker_color = ~ ifelse(upper < 1, "#16a34a", "#64748b")
)
```

Or use pre-computed columns:

```{r}
direction_data <- meta_data |>
  mutate(
    marker_color = case_when(
      upper < 1 ~ "#16a34a",
      lower > 1 ~ "#dc2626",
      TRUE ~ "#64748b"
    )
  )

tabviz(direction_data,
  label = "study",
  columns = list(
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1)
  ),
  marker_color = "marker_color"
)
```

## Next steps

| Topic | What you'll learn |
|-------|-------------------|
| [Columns](columns.qmd) | All 16+ column types |
| [Row Styling](row-styling.qmd) | Headers, summaries, indentation, badges |
| [Cell Styling](cell-styling.qmd) | Conditional formatting for individual cells |
| [Row Groups](row-groups.qmd) | Collapsible groups |
| [Forest Plots](visualizations/forest-plots.qmd) | Advanced forest column options |
| [Themes](themes.qmd) | Preset themes and customization |
