---
title: "Recipes"
description: "Short 'how do I…' snippets for common tabviz patterns"
---

```{r}
#| include: false
library(tabviz)
library(dplyr)
library(shiny)
```

Short, self-contained snippets for specific patterns that aren't covered on a single dedicated page. Each recipe is a single drop-in pattern. For broader topics see the [Fluent API guide](fluent-api.qmd); for `col_*` / `viz_*` / `set_*` argument detail see the [reference section](../reference/index.qmd).

For recipes that *are* the bread and butter of one page (significance highlighting, p-value formatting, split-by with shared axis), follow the link to that page directly — those topics have full examples there.

## Composition

### Build a table-only view (no visualization column)

Drop `viz_forest()` (or any `viz_*`) — any table works without a viz column.

```{r}
#| eval: false
tabviz(data,
  label = "study",
  columns = list(
    col_text("drug", "Drug"),
    col_n("n"),
    col_events("events", "n", "Events"),
    col_interval("hr", "lower", "upper", header = "HR (95% CI)"),
    col_pvalue("pval", "P")
  )
)
```

### Show multiple effect sizes side-by-side

Stack `viz_forest()` columns with a shared `axis_range` so the eye can compare across them.

```{r}
#| eval: false
axis_range <- c(0.4, 2.0)

tabviz(data,
  label = "study",
  columns = list(
    viz_forest(point = "hr_primary", lower = "primary_lo", upper = "primary_hi",
               header = "Primary outcome",
               axis_range = axis_range, null_value = 1, scale = "log"),
    viz_forest(point = "hr_secondary", lower = "secondary_lo", upper = "secondary_hi",
               header = "Secondary outcome",
               axis_range = axis_range, null_value = 1, scale = "log")
  )
)
```

### Compose a reusable house-style function

Wrap your organization's standards in a function. Works on both `WebSpec` and htmlwidget inputs because every `set_*()` returns the same type it received.

```{r}
#| eval: false
house_style <- function(x) {
  x |>
    set_marker_style(color = "direction_color") |>
    set_row_style(bold = "is_primary", type = "row_type") |>
    set_theme(web_theme_jama() |>
                set_theme_field(c("axis", "gridlines"),     TRUE) |>
                set_theme_field(c("axis", "gridline_style"), "dotted"))
}

tabviz(trial_a, ...) |> house_style()
tabviz(trial_b, ...) |> house_style()
```

## Styling & annotations

### Badge the top-N rows

Rank by weight, tag the top 3 with a position badge.

```{r}
#| eval: false
data |>
  mutate(
    rank  = rank(-weight),
    badge = ifelse(rank <= 3, paste0("#", rank), NA_character_)
  ) |>
  tabviz(
    label = "study",
    columns = list(
      viz_forest(point = "hr", lower = "lower", upper = "upper",
                 null_value = 1, scale = "log")
    ),
    row_badge = "badge"
  )
```

### Pin a custom reference line on a forest

`null_value = 1` already paints the "no effect" line; add `refline()` entries to layer extra references such as a pooled estimate.

```{r}
#| eval: false
viz_forest(
  point = "hr", lower = "lower", upper = "upper",
  scale = "log", null_value = 1,
  annotations = list(
    refline(1,    label = "No effect",  style = "solid",  color = "#111"),
    refline(0.85, label = "Pooled HR",  style = "dashed", color = "#0066cc")
  )
)
```

## Shiny / proxy patterns

The proxy verbs below mutate the live widget without a re-render — scroll position, open/closed groups, and selection survive.

### Switch themes mid-session without a re-render

```{r}
#| eval: false
server <- function(input, output, session) {
  output$plot <- renderTabviz({
    tabviz(data, label = "study",
           columns = list(viz_forest(point = "hr", lower = "lower", upper = "upper",
                                     null_value = 1, scale = "log")))
  })

  observeEvent(input$theme, {
    tabviz_proxy("plot") |> set_theme(input$theme)
  })
}
```

### Filter rows from a slider

Proxy-dispatched `filter_rows()` updates in place. `clear_filters()` first to drop any earlier filter rather than stack on top of it.

```{r}
#| eval: false
ui <- fluidPage(
  sliderInput("max_p", "Max p-value", min = 0, max = 1, value = 0.1),
  tabvizOutput("plot")
)

server <- function(input, output, session) {
  output$plot <- renderTabviz({ tabviz(data, ...) })

  observeEvent(input$max_p, {
    tabviz_proxy("plot") |>
      clear_filters() |>
      filter_rows("pval", "lt", input$max_p)
  })
}
```

### Change a column's header or width mid-session

```{r}
#| eval: false
observeEvent(input$relabel, {
  tabviz_proxy("plot") |>
    update_column("hr", header = "Hazard Ratio", width = 160)
})
```

### Preserve scroll and selection across updates

Proxy verbs preserve UI state; `renderTabviz({ tabviz(...) })` re-renders the whole widget and resets it. Reach for the proxy whenever the change is incremental.

```{r}
#| eval: false
# GOOD: incremental update keeps scroll/selection
observeEvent(input$sort_col, {
  tabviz_proxy("plot") |> sort_rows(input$sort_col, "desc")
})

# BAD: full re-render loses scroll position
output$plot <- renderTabviz({
  data |>
    arrange(desc(.data[[input$sort_col]])) |>
    tabviz(...)
})
```

## Export

### Save at retina (2×) resolution

`scale` controls the pixel density of raster output.

```{r}
#| eval: false
plot <- tabviz(data, ...)

save_plot(plot, "forest.svg")                            # vector
save_plot(plot, "forest.png", scale = 2)                 # retina PNG
save_plot(plot, "forest.pdf", width = 800)               # vector PDF
```

For long tables that need to span multiple PDF pages, see [Export → Pagination](export.qmd#pagination).

## See also

These topics have full coverage on a dedicated page rather than a recipe here:

- **Conditional row styling by significance** → [Row Styling — Expression-Based Styling](row-styling.qmd#expression-based-styling)
- **P-value formatting with thresholds and stars** → [Columns — `col_pvalue()`](columns.qmd#col_pvalue)
- **Split by subgroup with shared axis** → [Split Plots — Shared Axis Range](split-plots.qmd#shared-axis-range)
- **All the proxy verbs** → [Fluent API guide](fluent-api.qmd) / [Data operations reference](../reference/data-ops.qmd)
