---
title: "Recipes"
description: "Short 'how do I…' snippets for common tabviz patterns"
---

```{r}
#| include: false
library(tabviz)
library(dplyr)
library(shiny)
```

Short, self-contained snippets for specific tasks. Each recipe is a single pattern you can lift into your own code. For broader patterns see the [Fluent API guide](fluent-api.qmd); for full parameter references see the [reference section](../reference/index.qmd).

---

## Color rows by significance threshold

Bold + green for `p < 0.05`, muted gray otherwise.

```{r}
#| eval: false
tabviz(data,
  label = "study",
  columns = list(viz_forest(point = "hr", lower = "lower", upper = "upper",
                            null_value = 1, scale = "log")),
  row_bold  = ~ pval < 0.05,
  row_color = ~ ifelse(pval < 0.05, "#166534", "#64748b")
)
```

---

## Badge the top-N rows

Rank by weight, tag the top 3.

```{r}
#| eval: false
data |>
  mutate(
    rank  = rank(-weight),
    badge = ifelse(rank <= 3, paste0("#", rank), NA_character_)
  ) |>
  tabviz(
    label = "study",
    columns = list(viz_forest(point = "hr", lower = "lower", upper = "upper",
                            null_value = 1, scale = "log")),
    row_badge = "badge"
  )
```

---

## Switch themes mid-Shiny-session

Use `set_theme()` on a proxy to re-skin without a full re-render.

```{r}
#| eval: false
server <- function(input, output, session) {
  output$plot <- renderTabviz({
    tabviz(data, label = "study",
           columns = list(viz_forest("hr", "lower", "upper")))
  })

  observeEvent(input$theme, {
    tabviz_proxy("plot") |> set_theme(input$theme)
  })
}
```

---

## Filter via a slider in Shiny

Proxy-dispatched `filter_rows()` updates in place.

```{r}
#| eval: false
ui <- fluidPage(
  sliderInput("max_p", "Max p-value", min = 0, max = 1, value = 0.1),
  tabvizOutput("plot")
)

server <- function(input, output, session) {
  output$plot <- renderTabviz({ tabviz(data, ...) })

  observeEvent(input$max_p, {
    proxy <- tabviz_proxy("plot")
    proxy |>
      clear_filters() |>
      filter_rows("pval", "lt", input$max_p)
  })
}
```

---

## Format p-values with thresholds

Switch to scientific notation for very small values, abbreviate below a cutoff, append significance stars.

```{r}
#| eval: false
tabviz(data,
  label = "study",
  columns = list(
    col_pvalue("pval", "P",
               exp_threshold    = 0.001,   # scientific notation below this
               abbrev_threshold = 1e-4,    # show "<0.0001" below this
               stars            = TRUE)    # append * / ** / *** by `thresholds`
  )
)
```

---

## Pin an annotation line at the null value

Use `refline()` in `viz_forest()`'s `annotations` list.

```{r}
#| eval: false
viz_forest(
  point = "hr", lower = "lower", upper = "upper",
  scale = "log", null_value = 1,
  annotations = list(
    refline(1,    label = "No effect",  style = "solid",  color = "#111"),
    refline(0.85, label = "Pooled HR", style = "dashed", color = "#0066cc")
  )
)
```

---

## Split by subgroup with a shared axis

Hierarchical navigation; every pane uses the same scale.

```{r}
#| eval: false
tabviz(data,
  label = "study",
  columns = list(viz_forest(point = "hr", lower = "lower", upper = "upper",
                            null_value = 1, scale = "log")),
  split_by     = c("region", "country"),
  shared_axis  = TRUE
)
```

---

## Export a static plot at 2× resolution

`scale` controls the pixel density of raster output.

```{r}
#| eval: false
plot <- tabviz(data, ...)

save_plot(plot, "forest.svg")                    # vector
save_plot(plot, "forest.png", scale = 2)         # retina PNG
save_plot(plot, "forest.pdf", width = 8, height = 6)
```

---

## Build a table-only view (no forest column)

Drop `viz_forest()` — any table works without a visualization column.

```{r}
#| eval: false
tabviz(data,
  label = "study",
  columns = list(
    col_text("drug", "Drug"),
    col_n("n"),
    col_events("events", "n", "Events"),
    col_interval("hr", "lower", "upper", header = "HR (95% CI)"),
    col_pvalue("pval", "P")
  )
)
```

---

## Show multiple effect sizes side-by-side

Stack `viz_forest()` columns with a shared axis range.

```{r}
#| eval: false
axis_range <- c(0.4, 2.0)

tabviz(data,
  label = "study",
  columns = list(
    viz_forest(point = "hr_primary", lower = "primary_lo", upper = "primary_hi",
               header = "Primary outcome",
               axis_range = axis_range, null_value = 1, scale = "log"),
    viz_forest(point = "hr_secondary", lower = "secondary_lo", upper = "secondary_hi",
               header = "Secondary outcome",
               axis_range = axis_range, null_value = 1, scale = "log")
  )
)
```

---

## Preserve scroll/selection across proxy updates

Proxy verbs (`sort_rows`, `filter_rows`, `update_column`, etc.) mutate state in place and preserve scroll position, open groups, and selection. Avoid `renderTabviz({ tabviz(...) })` for incremental changes — it re-renders the whole widget.

```{r}
#| eval: false
# GOOD: incremental update keeps scroll/selection
observeEvent(input$sort_col, {
  tabviz_proxy("plot") |> sort_rows(input$sort_col, "desc")
})

# BAD: full re-render loses scroll position
output$plot <- renderTabviz({
  data |>
    arrange(desc(.data[[input$sort_col]])) |>
    tabviz(...)
})
```

---

## Change a column's header and width mid-session

Use `update_column()` via a proxy — no re-render, no spec rebuild.

```{r}
#| eval: false
observeEvent(input$relabel, {
  tabviz_proxy("plot") |>
    update_column("hr", header = "Hazard Ratio", width = 160)
})
```

---

## Compose reusable styling pipelines

Wrap your organization's standards in a function. Works on both `WebSpec` and htmlwidget inputs because every `set_*()` returns the same type it received.

```{r}
#| eval: false
house_style <- function(x) {
  x |>
    set_marker_style(color = "direction_color") |>
    set_row_style(bold = "is_primary", type = "row_type") |>
    set_theme(web_theme_jama() |>
                set_theme_field(c("axis", "gridlines"), TRUE) |>
                set_theme_field(c("axis", "gridline_style"), "dotted"))
}

tabviz(trial_a, ...) |> house_style()
tabviz(trial_b, ...) |> house_style()
```

---

## See Also

- [Fluent API guide](fluent-api.qmd) — composing verbs, static-vs-proxy dispatch
- [Row styling](row-styling.qmd) — row-level parameters and binding rules
- [Data operations reference](../reference/data-ops.qmd) — every column/row/data/cell verb
- [Shiny integration](shiny.qmd) — proxy patterns end-to-end
