---
title: "Row Groups"
---

```{r}
#| include: false
library(tabviz)
```

Organize your forest plot into collapsible sections using the `group` parameter. Ideal for meta-analyses with natural categories like study design, region, or subpopulation.

## What you can pass to `group`

| Form | Effect |
|------|--------|
| `group = "region"` | Single flat grouping — one collapsible header per unique value |
| `group = c("region", "country")` | Hierarchical nesting — outer level collapses inner levels |
| `group = list(web_group("North", collapsed = FALSE), web_group("South", collapsed = TRUE))` | Explicit definitions; lets you set initial collapse state and custom labels |

## Basic Grouping

Pass a column name to `group` to create collapsible sections:

```{r}
grouped_data <- data.frame(
  study = c("Study A1", "Study A2", "Study B1", "Study B2", "Study B3"),
  region = c("North", "North", "South", "South", "South"),
  or = c(0.8, 0.9, 1.1, 1.3, 0.95),
  lower = c(0.6, 0.7, 0.85, 1.0, 0.7),
  upper = c(1.1, 1.15, 1.4, 1.7, 1.3)
)

forest_plot(grouped_data,
  point = "or", lower = "lower", upper = "upper",
  label = "study",
  group = "region",
  null_value = 1
)
```

Click the group headers to collapse or expand sections. Groups are rendered with:

- Distinct header rows with chevron indicators
- Automatic indentation of child rows
- Subtle background shading for visual hierarchy

---

## Hierarchical Grouping

For multi-level organization, pass multiple column names as a vector. tabviz creates nested, collapsible hierarchies:

```{r}
hierarchical_data <- data.frame(
  study = c(
    "Northeast Trial", "Southeast Trial", "Midwest Trial",
    "Ontario Study", "Quebec Study",
    "England RCT", "Scotland RCT",
    "Bavaria Trial", "Berlin Trial"
  ),
  region = c(
    rep("Americas", 5),
    rep("Europe", 4)
  ),
  country = c(
    rep("USA", 3), rep("Canada", 2),
    rep("UK", 2), rep("Germany", 2)
  ),
  or = c(0.82, 0.91, 0.78, 0.85, 0.88, 0.92, 0.87, 0.79, 0.84),
  lower = c(0.68, 0.75, 0.62, 0.70, 0.72, 0.76, 0.71, 0.64, 0.68),
  upper = c(0.99, 1.10, 0.98, 1.03, 1.07, 1.11, 1.07, 0.98, 1.04)
)

forest_plot(hierarchical_data,
  point = "or", lower = "lower", upper = "upper",
  label = "study",
  group = c("region", "country"),
  null_value = 1, scale = "log",
  theme = web_theme_modern()
)
```

Each level of the hierarchy:

- Can be collapsed independently
- Has progressively deeper indentation
- Uses subtle visual differentiation (font weight, background)

---

## Customizing Group Headers

Control the visual styling of group headers at each level using theme modifiers:

```{r}
#| eval: false
custom_theme <- web_theme_default() |>
  set_theme_field(c("row_group", "L1", "bg"),            "#E0F2FE") |>
  set_theme_field(c("row_group", "L1", "text", "size"),  "1.1rem") |>
  set_theme_field(c("row_group", "L1", "text", "weight"), 700) |>
  set_theme_field(c("row_group", "L2", "bg"),            "#F0F9FF") |>
  set_theme_field(c("row_group", "L2", "text", "italic"), TRUE) |>
  set_theme_field(c("row_group", "L2", "text", "size"),  "0.95rem") |>
  set_spacing(indent_per_level = 20)

forest_plot(hierarchical_data,
  point = "or", lower = "lower", upper = "upper",
  label = "study",
  group = c("region", "country"),
  null_value = 1, scale = "log",
  theme = custom_theme
)
```

The `row_group` cluster is one of the v2 theme's Tier 3 component
clusters. Its three tiers — `L1` (outermost), `L2`, `L3` — each carry
`bg`, `fg`, `rule`, `text` (a TextRole bundle: family / size / weight /
italic), and `border_bottom`. See [Themes guide](themes.qmd) for the
full structure and [Themes Reference](../reference/themes.qmd) for the
class definitions.

---

## Starting Collapsed

To have groups start in a collapsed state, you'll need to use explicit group definitions with `web_group()`:

```{r}
#| eval: false
forest_plot(grouped_data,
  point = "or", lower = "lower", upper = "upper",
  label = "study",
  group = list(
    web_group("North", collapsed = FALSE),
    web_group("South", collapsed = TRUE)
  ),
  null_value = 1
)
```

This is useful for large datasets where you want users to expand only the sections they're interested in.

---

## Grouping vs Row Styling

tabviz provides two complementary ways to organize data:

| Feature | Purpose | Best For |
|---------|---------|----------|
| **`group`** | Automatic collapsible sections from column values | Categorical data with natural groupings |
| **`row_type`** | Manual section headers, summaries, spacers | Structured layouts with subtotals |

Use `group` when your data has a grouping column. Use `row_type` when you need precise control over the layout structure, such as meta-analysis summary rows. See [Row Styling](row-styling.qmd) for the manual approach.

You can combine both approaches—use `group` for automatic organization and `row_type` for summary diamonds within each group.

---

## See Also

- [Row Styling](row-styling.qmd) — Manual row types, headers, summaries, and visual styling
- [Themes](themes.qmd) — Customize group header appearance
- [forest_plot()](../reference/tabviz.qmd) — Complete argument reference
