---
title: "Split Plots"
---

```{r}
#| include: false
library(tabviz)
library(dplyr)
```

Split plots produce **a separate plot for each value** of a categorical variable, with an interactive sidebar for navigation. Use them for subgroup analyses where each subgroup deserves its own focused view.

::: {.callout-tip}
## Split vs Group

| Feature | Use when |
|---|---|
| **`split_by`** / `split_table()` | Many subgroups (5+); each deserves focused attention |
| **`group`** | Few subgroups (2–4); want to see them all at once |

Split creates a sidebar with one plot per subgroup. Group creates collapsible sections within a single plot.
:::

## Two Ways to Split

| API | When to use |
|---|---|
| `tabviz(..., split_by = "region", shared_axis = TRUE)` | Quick: split inline as you build the widget |
| `tabviz(...) \|> split_table(by = ..., shared_axis = ...)` | Fluent: split after constructing a `WebSpec` |

Both produce a `SplitForest` object.

## `split_table()` Arguments

| Argument | Default | Purpose |
|---|---|---|
| `x` | required | A `WebSpec`, htmlwidget, or data frame |
| `by` | required | Column name (single string) or character vector for hierarchical splits |
| `shared_axis` | `FALSE` | Use the same axis range across all subgroups |
| `shared_column_widths` | `FALSE` | Stamp identical per-column widths across every sub-plot (v0.16+). Useful when stacking screenshots in slides / PowerPoint — sub-plot columns won't shift between tiles. Widths are computed once from the combined data. |

## Basic Split

```{r}
set.seed(42)
trial_data <- data.frame(
  study = paste0("Study ", 1:30),
  region = sample(c("North America", "Europe", "Asia Pacific"), 30, replace = TRUE),
  or = exp(rnorm(30, log(0.8), 0.3))
)
trial_data$lower <- trial_data$or * exp(-1.96 * 0.25)
trial_data$upper <- trial_data$or * exp(1.96 * 0.25)

tabviz(trial_data,
  label = "study",
  columns = list(
    viz_forest(point = "or", lower = "lower", upper = "upper",
               scale = "log", null_value = 1,
               axis_label = "Odds Ratio (95% CI)")
  ),
  split_by = "region",
  title = "Treatment Effect by Region"
)
```

The floating sidebar shows each region. Click to navigate between subgroups.

## Hierarchical Splits

Pass a character vector to `by` for nested navigation:

```{r}
set.seed(123)
n <- 50
subgroup_data <- data.frame(
  study = paste0("Study ", sprintf("%02d", 1:n)),
  sex = sample(c("Male", "Female"), n, replace = TRUE),
  age_group = sample(c("18-40", "41-65", "65+"), n, replace = TRUE),
  or = exp(rnorm(n, log(0.75), 0.25))
)
subgroup_data$lower <- subgroup_data$or * exp(-1.96 * 0.2)
subgroup_data$upper <- subgroup_data$or * exp(1.96 * 0.2)

tabviz(subgroup_data,
  label = "study",
  columns = list(
    viz_forest(point = "or", lower = "lower", upper = "upper",
               scale = "log", null_value = 1, axis_label = "Odds Ratio")
  ),
  theme = web_theme_cochrane()
) |>
  split_table(by = c("sex", "age_group"))
```

The sidebar shows "Sex" as the first level header, with "Age Group" nested under each sex.

## Adding Extra Columns

Columns in the underlying `tabviz()` carry over into every sub-plot:

```{r}
tabviz(subgroup_data,
  label = "study",
  columns = list(
    col_text("sex", header = "Sex"),
    col_text("age_group", header = "Age"),
    col_interval("or", "lower", "upper", header = "OR (95% CI)"),
    viz_forest(point = "or", lower = "lower", upper = "upper",
               scale = "log", null_value = 1)
  ),
  theme = web_theme_cochrane()
) |>
  split_table(by = c("sex", "age_group"))
```

## Shared Axis Range

By default each sub-plot auto-scales its axis. Pass `shared_axis = TRUE` for consistent visual scale across subgroups:

```{r}
tabviz(trial_data,
  label = "study",
  columns = list(
    viz_forest(point = "or", lower = "lower", upper = "upper",
               scale = "log", null_value = 1)
  ),
  theme = web_theme_cochrane()
) |>
  split_table(by = "region", shared_axis = TRUE)
```

## Shared Column Widths (v0.16+)

When you plan to stack sub-plot screenshots in slides, PowerPoint, or
side-by-side layouts, set `shared_column_widths = TRUE`. Each sub-plot
will render with identical per-column widths, computed once from the
combined data, so the columns line up tile-to-tile:

```{r}
tabviz(trial_data,
  label = "study",
  columns = list(
    viz_forest(point = "or", lower = "lower", upper = "upper",
               scale = "log", null_value = 1)
  ),
  theme = web_theme_cochrane()
) |>
  split_table(
    by = "region",
    shared_axis = TRUE,
    shared_column_widths = TRUE
  )
```

The default is `FALSE` to preserve the pre-v0.16 behavior where each
sub-plot sizes its columns to its own content.

Since v0.19 the sidebar also has an **Align column widths** checkbox
in its footer — users can toggle this behavior at runtime without
going back to R.

## Sidebar Features

The sidebar includes:

- **Section headers** — split-variable names displayed as uppercase labels
- **Collapsible** — click the chevron to collapse the sidebar
- **Search** — filter subgroups by typing
- **Keyboard navigation** — arrow keys move between subgroups
- **Bold active** — currently selected subgroup is highlighted

## Selecting a Subgroup in Shiny

Inside Shiny, switch the active subgroup programmatically with `split_tabviz_select()`:

```r
observeEvent(input$region, {
  split_tabviz_select(split_tabviz_proxy("split_plot"), input$region)
})
```

See [Shiny Integration](shiny.qmd) for the full pattern.

## Exporting Split Plots

Pass the `SplitForest` to `save_split_table()` (or `save_plot()`, which dispatches automatically). The directory structure mirrors the split hierarchy:

```r
p <- tabviz(data,
  label = "study",
  columns = list(
    viz_forest(point = "or", lower = "lower", upper = "upper",
               scale = "log", null_value = 1)
  )
) |>
  split_table(by = c("sex", "age_group"))

save_split_table(p, "output", format = "svg")
# Writes: output/Male/Male_Young.svg, output/Female/Female_Old.svg, ...
```

See [Export](export.qmd) for `save_plot()` / `save_split_table()` details.
