---
title: "Themes"
---

```{r}
#| include: false
library(tabviz)

meta_data <- data.frame(
  study = c("Smith 2020", "Jones 2021", "Lee 2022", "Chen 2023", "Park 2024"),
  hr    = c(0.72, 0.85, 0.91, 0.68, 0.79),
  lower = c(0.55, 0.70, 0.75, 0.52, 0.61),
  upper = c(0.95, 1.03, 1.10, 0.89, 1.02)
)
```

A theme decides every visual choice in your table that isn't data: surface colors, text tones, divider weights, spacing, fonts, marker fills. tabviz themes are designed so you can pick a preset, change one or two things, and never think about it again — or build a fully branded theme from a dozen inputs when you need to.

## Three steps from "pick a preset" to "fully bespoke"

Each step builds on the previous. Most users live at step 1.

```{mermaid}
flowchart LR
    A["1. Pick a preset"] --> B["2. Tweak inputs"]
    B --> C["3. Override component fields"]

    style A fill:#dbeafe
    style B fill:#bfdbfe
    style C fill:#93c5fd
```

* **Step 1** — `theme = web_theme_lancet()`. Done. Four presets cover the common publication and dashboard surfaces.
* **Step 2** — Swap a primary color, change density, switch to a serif font. The preset stays the foundation; you nudge a few inputs.
* **Step 3** — Reach into a specific component (e.g. the L1 row-group bar background) for surgical control. Rare; the preset cascade handles 95% of cases.

## Step 1: Pick a preset

```{r}
#| eval: false
forest_plot(
  meta_data, point = "hr", lower = "lower", upper = "upper",
  label = "study",
  theme = web_theme_lancet()
)
```

Four presets ship in the package — each leans into a clear identity rather than trying to be everything:

| Preset | Style | Designed for |
|--------|-------|--------------|
| `web_theme_cochrane()` | The package default. Cochrane teal + warm coral, Inter sans-serif, comfortable density | General purpose, systematic reviews |
| `web_theme_lancet()`   | Lancet navy + refined old-gold + warm cream, Georgia serif, comfortable | Lancet-style editorial layouts |
| `web_theme_jama()`     | Pure black-and-white, ultra-compact, strong dividers throughout, Arial | JAMA / print journal submissions |
| `web_theme_dark()`     | Catppuccin Mocha-inspired dark canvas with pastel marker palette | Dark-mode viewing |

Switch presets at any point with `set_theme()`:

```{r}
#| eval: false
my_table |> set_theme("jama")
```

The full registry is in `package_themes()`.

### Curating the in-widget switcher

The in-widget settings panel includes a theme dropdown that lets readers
swap between presets without reloading. By default it shows all four
package themes; pass `enable_themes = list(...)` to `web_interaction()`
or `selectable_themes()` to curate the menu:

```{r}
#| eval: false
# Flat list — no tabs, just a single column of choices.
selectable_themes(my_table, list(
  Cochrane = web_theme_cochrane(),
  Lancet   = web_theme_lancet()
))

# 2-level named list — each top-level key becomes a tab in the dropdown.
selectable_themes(my_table, list(
  Editorial = list(
    Cochrane = web_theme_cochrane(),
    Lancet   = web_theme_lancet(),
    JAMA     = web_theme_jama()
  ),
  Other = list(Dark = web_theme_dark())
))
```

The active theme is auto-included so readers can always revert. For
categorized lists, if the active theme isn't in any user-supplied
category, a `Current` tab is prepended.

## Step 2: Tweak inputs

Every theme — including all four presets — is determined by **a small set of customer-facing inputs**. There are twelve of them:

| Group | Input | What it controls |
|-------|-------|------------------|
| Chrome | `neutral` | 5-step ramp, lightest → darkest. Drives surfaces, dividers, content |
|  | `brand` + `primary_deep` | Brand fill (markers, accents) and dark companion (text/strokes, bold-mode header band) |
|  | `accent` + `accent_deep` | Chrome accent for hover, selection, links |
|  | `status_*` | Positive / negative / warning / info |
| Data | `series_anchors` | Per-effect anchor colors (variable length). The first anchor also drives the pooled-effect diamond. |
| Type | `font_body` | Primary face for cells, headers, labels |
|  | `font_display` | Optional title face (defaults to body) |
|  | `font_mono` | Optional monospace face (e.g. for IDs) |

Build a custom theme by overriding any of these:

```{r}
#| eval: false
my_theme <- web_theme(
  name = "my-brand",
  inputs = list(
    primary          = "#1F3A5F",   # navy
    accent         = "#B08938",   # warm gold
    series_anchors = c("#1F3A5F", "#B08938", "#6B8E7F"),
    font_body      = "Charter, Georgia, serif"
  )
)
```

`web_theme()` starts from `web_theme_cochrane()` unless you pass a different `base_theme`. Anything you *don't* set inherits from the base.

The deep companions mirror their primaries by default:

* `primary_deep` defaults to `brand` (override when you want a darker stroke than fill, or a richer bold-mode header band).
* `accent_deep` defaults to `accent`.
* `font_display` defaults to `font_body`.

The pooled-effect diamond reads from the first series slot bundle (`theme@series[[1]]`). Override `series_anchors[1]` to recolor it; reach into `theme@series[[1]]@stroke` directly for a journal-specific outline.

### What `primary_deep` actually drives

`primary_deep` is the dark identity tone — meant to land on text and rules
where you want the primary to read as serious, not decorative. The resolver
threads it into:

* Plot title color (`text@title@fg`)
* Forest-axis label and tick-label colors (`plot@axis_label@fg`, `plot@tick_label@fg`)
* Cell-hairline tint — `divider@subtle` is `neutral[3]` mixed with 8% `primary_deep`, so cell borders pick up a faint identity color without muddying out
* Bold-mode header band (`header@bold@bg`) and column-group bold band
* Bold-mode header rule (`header@bold@rule`) via the new `divider@strong_on_dark` helper — a 40%-deep / 60%-inverse mix that contrasts against the band instead of disappearing into it
* L1 group bar — but **only** when the active variant is `header_style = "bold"`. Under light-mode the L1 bar stays on `accent.tint_subtle` (chrome/data wall: primary only "leaks" into chrome when you've chosen brand-forward chrome).

To pin any of these, set the field directly via `set_theme_field()` and
your value survives re-resolution. The interactive panel surfaces the same
override semantics as a per-field reset icon.

### L1 group header and banding

When `theme@row@banding` cycles at the L1 level (the default for every
preset — `banding = "group"` with one nesting level), the banding cycle
already paints alternating fills across the L1 group rows. The renderer
detects this and **drops the explicit `theme@row_group@L1@bg` paint**
on the affected rows so the banded color reads as continuous and the
group-bar doesn't double-paint on top.

If you want a custom L1 bar paired with a banded background, switch
banding modes (e.g. `set_theme_field(c("row", "banding"), "row")` or
`"none"`) so the L1 paint isn't subsumed.

At deeper banding levels (e.g. `banding = "group-2"` with nested groups),
the L1 bar's explicit fill is preserved unchanged — only the banded
level loses its explicit fill.

### Semantic tokens — naming visual treatments

Tabviz ships with six **semantic tokens** that name visual treatments
you can apply to a row or cell. Each token is a `RowSemantic` bundle on
`theme.row`: `bg`, `fg`, `border`, `marker_fill`, `font_weight`,
`font_style`. The renderer picks one token per row/cell and paints its
bundle.

| Token       | Default treatment                                                    |
|-------------|----------------------------------------------------------------------|
| `mute`      | Lighter color, reduced prominence                                    |
| `bold`      | Just a weight bump — no color change                                 |
| `accent`    | Bold + accent-color text and marker (most common)                    |
| `highlight` | Bold + a pale "marker" background (defaults to a derived `theme.semantic.highlight`) |
| `fill`      | Bold + a strong row fill (defaults to a derived `theme.semantic.fill`); fg auto-contrasts |
| `emphasis`  | Legacy: bold + primary text and marker (kept for back-compat)        |

Two paths to apply a token to a row:

1. **Data column** — pass a logical column to `tabviz()`:

   ```{r}
   #| eval: false
   tabviz(data, label = "study",
     columns = list(viz_forest("hr", "lo", "hi")),
     row_accent = "is_primary",      # rows where is_primary == TRUE → accent token
     row_fill   = "is_overall"       # rows where is_overall == TRUE → fill token
   )
   ```

2. **Painter (interactive)** — click the paintbrush icon, pick a chip
   (`mute` / `bold` / `accent` / `highlight` / `fill`), then click rows
   or cells to stamp the token on. Hover before clicking shows a
   semi-transparent preview of the would-be visual. The painter
   defaults to `accent` + row-scope on first open. Toggle Row/Cell
   scope from the popover.

#### Customizing token visuals

Two ways to override:

* **Tier-2 named colors** for the highlight + fill backgrounds:

  ```{r}
  #| eval: false
  set_theme_field(t, c("semantic", "highlight"), "#FFE58F")  # custom marker yellow
  set_theme_field(t, c("semantic", "fill"),      "#1F3A5F")  # custom fill navy
  ```

  Or interactively: Settings panel → **Tokens** tab → Token colors.

* **Per-bundle override** for any field on any token:

  ```{r}
  #| eval: false
  set_theme_field(t, c("row", "fill", "fg"),          "#FFFFFF")  # force white text on fill rows
  set_theme_field(t, c("row", "highlight", "border"), "#C8553D")  # red border on highlight rows
  ```

  Or interactively: Settings panel → **Tokens** tab → expand a token →
  edit any field. All bundle fields are override-aware (reset reverts
  to the resolved default).

### Variants — three knobs every table can flip

Variants are per-table choices that don't change the theme's identity. They're orthogonal to inputs and freely composable:

```{r}
#| eval: false
web_theme(
  inputs = list(primary = "#1F3A5F"),
  variants = list(
    density            = "comfortable",   # compact | comfortable | spacious
    header_style       = "bold",          # light  | bold
    first_column_style = "bold"           # default | bold (Excel-style)
  )
)
```

* **`density`** drives every spacing token (row height, padding, gaps). The three presets cover almost every screen and print scenario.
* **`header_style`** picks between a light header on the muted surface and a bold header on `primary_deep` with inverse text. Both options are *always available* on every theme — you flip the variant per table.
* **`first_column_style`** gives the row-label column an Excel-style emphasized treatment (background tint + heavier weight). Useful when your first column is genuinely a label, not data.

You can also set variants directly on the `tabviz()` call as overrides (see [tabviz reference](../reference/tabviz.qmd)).

## How it cascades under the hood

A theme has three tiers. You only ever interact with Tier 1; the other tiers are computed.

```{mermaid}
flowchart TB
    subgraph T1["Tier 1 — Inputs (you edit these)"]
      neutral & primary & accent & series & fonts & density
    end

    subgraph T2C["Tier 2 — Chrome roles (derived)"]
      surface[surface.base / muted / raised]
      content[content.primary / secondary / muted / inverse]
      divider[divider.subtle / strong]
      tints[accent.default / muted / tint.subtle / tint.medium]
    end

    subgraph T2D["Tier 2 — Data roles (derived)"]
      slots[per-series SlotBundle: fill, stroke, fill.muted, ...]
    end

    subgraph T3["Tier 3 — Component bindings"]
      header & row & cell & rowGroup & plot & marks
    end

    T1 --> T2C
    T1 --> T2D
    T2C --> T3
    T2D --> T3

    style T1 fill:#dbeafe
    style T2C fill:#dcfce7
    style T2D fill:#fef3c7
    style T3 fill:#f3e8ff
```

**Two cascades, sharing only Tier 1.** Chrome derivation reads only chrome inputs (neutrals + primary + accent). Data derivation reads only data inputs (series anchors + status). Series anchors never affect surface tones; primary colors never bleed into row-group divider tints. This isolation makes themes predictable: changing `series_anchors` cannot accidentally shift the column-header background.

**OKLCH derivation.** All blends, lightenings, and darkenings happen in OKLCH (perceptually uniform). The result: tints stay readable across the brightness range, and high-chroma brands don't mud out when blended with surfaces.

**Slot bundles.** Each series anchor expands into a 7-field bundle: `fill`, `stroke`, `fill_muted`, `stroke_muted`, `fill_emphasis`, `stroke_emphasis`, `text_fg`. The renderer picks the right field based on row state (default / muted / emphasized) without you having to define every state by hand. Override individual fields if you need to.

If you want to inspect a resolved theme:

```{r}
#| eval: false
t <- web_theme_lancet()
t@surface@base       # "#FDFCFB"
t@content@primary    # "#1E3A5F"
t@accent@tint_subtle # tinted accent for L1 group bar
t@series[[1]]@fill   # first series anchor
t@series[[1]]@stroke # auto-darkened companion
t@row@hover@bg       # accent.muted blend
```

## Step 3: Customize without rewriting

Three pipe-friendly modifiers cover the common edits:

```{r}
#| eval: false
custom <- web_theme_jama() |>
  set_inputs(primary = "#1F3A5F", accent = "#B08938") |>
  set_variants(density = "comfortable", header_style = "bold") |>
  set_spacing(row_height = 28, axis_gap = 14)
```

* **`set_inputs(theme, ...)`** — update Tier 1 inputs. Cascade-aware: changing `brand` clears `primary_deep` so it re-mirrors; changing `series_anchors` rebuilds the slot bundles.
* **`set_variants(theme, ...)`** — flip per-table toggles (density / header_style / first_column_style). Density change resets spacing tokens so the new preset's values flow through.
* **`set_spacing(theme, ...)`** — per-token spacing override on top of the density preset. Set what you care about; the rest follow density.

For surgical overrides anywhere in the theme tree, use `set_theme_field()` with a path:

```{r}
#| eval: false
custom <- web_theme_cochrane() |>
  set_theme_field(c("row_group", "L1", "bg"), "#EEEEEE") |>
  set_theme_field(c("series", 1L, "fill"), "#FF0000") |>
  set_theme_field(c("plot", "gridline"), "#003300") |>
  set_theme_field(c("axis", "gridlines"), TRUE)
```

Path entries are property names (strings) or list indices (integers, currently only for `series`). The function re-resolves the theme before returning, so derived fields stay consistent with your override.

### Pinning chrome separately from data

Because the cascades share only Tier 1, you can override a Tier 2 chrome role without disturbing the data marks, or vice versa:

```{r}
#| eval: false
# Pin the surface to a specific cream, regardless of what neutrals say
web_theme_cochrane() |>
  set_theme_field(c("surface", "base"), "#FDFCFB")

# Adjust series 2's stroke without affecting any other slot or chrome
web_theme_cochrane() |>
  set_theme_field(c("series", 2L, "stroke"), "#444444")
```

Re-resolution preserves explicit overrides on subsequent edits — `set_inputs(primary = "...")` after the override above will refresh series colors but leave `surface.base` pinned to your cream.

## Going deeper: spacing and typography

### Density preset values

| Token | compact | comfortable | spacious |
|-------|--------:|------------:|---------:|
| `row_height` | 20 | 24 | 30 |
| `header_height` | 26 | 32 | 40 |
| `padding` | 8 | 12 | 16 |
| `axis_gap` | 8 | 12 | 16 |
| `cell_padding_x` | 8 | 10 | 14 |
| `column_group_padding` | 6 | 8 | 12 |
| `title_subtitle_gap` | 10 | 13 | 18 |
| `bottom_margin` | 12 | 16 | 22 |
| `indent_per_level` | 14 | 16 | 20 |

`set_spacing()` overrides any of these per-token; the rest fall back to the active density preset. Read [Density](../concepts/styling-cascade.qmd) for how spacing tokens flow into the renderer.

### Typography roles

A theme exposes eight named text roles:

| Role | Used for |
|------|----------|
| `title` | Plot title (display family, large, semibold) |
| `subtitle` | Subtitle (body family, medium) |
| `caption` | Italic caption above the plot |
| `footnote` | Footnote below the plot |
| `body` | Default body text (cells, headers when not bold) |
| `cell` | Data cell text (defaults to body) |
| `label` | Axis labels |
| `tick` | Axis tick labels (smaller) |

Edit individual fields with `set_theme_field()`:

```{r}
#| eval: false
web_theme_cochrane() |>
  set_theme_field(c("text", "title", "weight"), 700) |>
  set_theme_field(c("text", "title", "italic"), TRUE)
```

## The 9 presets — when to use which

```{r}
#| eval: false
# Black & white journal style — JAMA submissions, dense print
forest_plot(meta_data, point = "hr", lower = "lower", upper = "upper",
            label = "study", theme = web_theme_jama())

# Lancet navy with serif typography — Lancet submissions
forest_plot(meta_data, point = "hr", lower = "lower", upper = "upper",
            label = "study", theme = web_theme_lancet())

# Modern dashboard — vibrant, spacious, screen-first
forest_plot(meta_data, point = "hr", lower = "lower", upper = "upper",
            label = "study", theme = web_theme_cochrane())

# Slide deck — oversized markers, thick lines, large type
forest_plot(meta_data, point = "hr", lower = "lower", upper = "upper",
            label = "study", theme = web_theme_cochrane() |> set_variants(density = "spacious"))

# Dark mode — Catppuccin-inspired
forest_plot(meta_data, point = "hr", lower = "lower", upper = "upper",
            label = "study", theme = web_theme_dark())
```

## Editing themes interactively

Every rendered table carries a settings gear. The in-widget panel exposes the v2 theme model under five tabs:

| Tab | Edits |
|-----|-------|
| **Theme** | Tier 1 inputs — brand, accent, status, series anchors, fonts |
| **Layout** | Variants (density / header style / first column style), banding, container border |
| **Spacing** | Per-token overrides on top of the density preset (collapsed Advanced section) |
| **Marks** | Series slot bundles (anchor + expandable 7-field editor), summary slot, plot mark sizes |
| **Text** | Eight text-role bundles + L1/L2/L3 row-group accordions |

The "View source" button exports your edits as v2 R code (`web_theme(inputs = list(...))` or a chain of `set_*()` calls), so panel exploration translates back into reproducible theme construction.

## Saving a custom theme as a function

For team-wide use, wrap your custom theme in a named function:

```{r}
#| eval: false
web_theme_acme <- function() {
  web_theme(
    name = "acme",
    inputs = list(
      neutral        = c("#FFFFFF", "#FAFAFA", "#EEEEEE", "#888888", "#222222"),
      primary          = "#FF5500",
      primary_deep     = "#CC3300",
      accent         = "#0066CC",
      series_anchors = c("#FF5500", "#0066CC", "#00AA66", "#993399", "#666666"),
      font_body      = "Inter, system-ui, sans-serif",
      font_display   = "'Inter Display', Inter, system-ui, sans-serif"
    ),
    variants = list(density = "comfortable", header_style = "bold")
  )
}

# Now share it across the team
forest_plot(data, ..., theme = web_theme_acme())
```

Pass to `set_theme()` either by string (when registered as a preset) or by handing the constructed `WebTheme` object directly:

```{r}
#| eval: false
my_table |> set_theme(web_theme_acme())
```

## Reference summary

* `web_theme()` — custom-theme constructor (inputs + variants + base).
* `web_theme_*()` — 9 preset constructors.
* `package_themes()` — named list of all preset constructors.
* `set_theme(spec, theme)` — switch a `WebSpec` to a different theme.
* `set_inputs(theme, ...)` — update Tier 1 inputs.
* `set_variants(theme, ...)` — update per-table variants.
* `set_spacing(theme, ...)` — per-token spacing overrides.
* `set_theme_field(theme, path, value)` — generic deep setter.
* `selectable_themes(spec, themes)` — curate the in-widget theme switcher.

For complete argument lists and class definitions see the [Themes reference](../reference/themes.qmd).
