---
title: "Themes"
---

```{r}
#| include: false
library(tabviz)

meta_data <- data.frame(
  study = c("Smith 2020", "Jones 2021", "Lee 2022", "Chen 2023", "Park 2024"),
  hr    = c(0.72, 0.85, 0.91, 0.68, 0.79),
  lower = c(0.55, 0.70, 0.75, 0.52, 0.61),
  upper = c(0.95, 1.03, 1.10, 0.89, 1.02)
)
```

A theme decides every visual choice in your table that isn't data: surface colors, text tones, divider weights, spacing, fonts, marker fills. tabviz themes are designed so you can pick a preset, change one or two things, and never think about it again — or build a fully branded theme from a dozen inputs when you need to.

## Three steps from "pick a preset" to "fully bespoke"

Each step builds on the previous. Most users live at step 1.

```{mermaid}
flowchart LR
    A["1. Pick a preset"] --> B["2. Tweak inputs"]
    B --> C["3. Override component fields"]

    style A fill:#dbeafe
    style B fill:#bfdbfe
    style C fill:#93c5fd
```

* **Step 1** — `theme = web_theme_nejm()`. Done. Nine curated presets cover the common publication, dashboard, and illustrative surfaces.
* **Step 2** — Swap a primary color, change density, switch to a serif font. The preset stays the foundation; you nudge a few inputs.
* **Step 3** — Reach into a specific component (e.g. the L1 row-group bar background) for surgical control. Rare; the preset cascade handles 95% of cases.

## Step 1: Pick a preset

```{r}
#| eval: false
tabviz(
  meta_data, label = "study",
  columns = list(
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1)
  ),
  theme = web_theme_nejm()
)
```

Nine presets ship in the package across four registers — **clinical** (1),
**editorial** (2), **design** (2), and **expressive** (4). Each owns a
distinct expressive axis rather than trying to be everything.

| Preset | Register | Identity |
|--------|----------|----------|
| `web_theme_nejm()` | clinical | The package default — restraint-led: teal + warm-coral accent, Inter, comfortable density |
| `web_theme_newsprint()` | editorial | Texture: warm grain, serif body |
| `web_theme_dwarven()` | editorial | Fantasy: Cinzel display serif |
| `web_theme_ledger()` | design | Color: teal + oxblood, ruled rules |
| `web_theme_brutalist()` | design | Geometry: sharp corners, thick rules, no effects |
| `web_theme_aurora()` | expressive | Effects: glass + glow on a dark canvas |
| `web_theme_terminal()` | expressive | Mono phosphor CRT |
| `web_theme_blueprint()` | expressive | Draft / grid: cyanotype |
| `web_theme_synthwave()` | expressive | Neon |

Switch presets at any point with `set_theme()`:

```{r}
#| eval: false
my_table |> set_theme("nejm")
```

The full registry is in `package_themes()` (a 2-level named list keyed `clinical` / `editorial` / `design` / `expressive`).

### Curating the in-widget switcher

The in-widget settings panel includes a theme dropdown that lets readers
swap between presets without reloading. By default it shows every package
theme; pass `enable_themes = list(...)` to `web_interaction()` or
`selectable_themes()` to curate the menu:

```{r}
#| eval: false
# Flat list — no tabs, just a single column of choices.
selectable_themes(my_table, list(
  NEJM   = web_theme_nejm(),
  Ledger = web_theme_ledger()
))

# 2-level named list — each top-level key becomes a tab in the dropdown.
selectable_themes(my_table, list(
  Editorial = list(
    Newsprint = web_theme_newsprint(),
    Dwarven   = web_theme_dwarven()
  ),
  Expressive = list(Aurora = web_theme_aurora())
))
```

The active theme is auto-included so readers can always revert. For
categorized lists, if the active theme isn't in any user-supplied
category, a `Current` tab is prepended.

## Step 2: Tweak inputs

Every theme — including every preset — is determined by the **V4 Tier-1
inputs**: a handful of OKLCH color anchors plus orthogonal axes for
polarity, accessibility, typography, density, geometry, marks, and effects.

| Group | Input | What it controls |
|-------|-------|------------------|
| Anchors | `paper` | Light-end neutral anchor. The neutral ramp interpolates paper → ink; surfaces, dividers, and body text all derive from it. |
|  | `ink` | Dark-end neutral anchor (text, strong rules). |
|  | `brand` | The identity hue. Seeds the brand ramp: headers, gradients, series slot 1, title color. |
|  | `accent` | Optional engagement hue — hover, selected, callouts, highlight bar. Mirrors brand when unset. |
| Axes | `polarity` | `"light"` / `"dark"` — reflects every anchor's lightness around the midpoint. |
|  | `monochrome` | The neutral ramp rides the brand hue (phosphor / sepia / cyanotype themes in one toggle). |
|  | `mode` | Accessibility: `"standard"` / `"high-contrast"` / `"reduced-transparency"`. Orthogonal to polarity. |
| Status | `status_positive` … `status_info` | Semantic palette. Theme-scoped identity (synthwave's neon vs a journal's clinical green); reverts to a curated HC-safe palette under high-contrast mode. |
| Type | `fonts_body` / `fonts_display` / `fonts_mono` | Font stacks; `fonts_display` mirrors body when unset. `type_base_size`, `type_scale_ratio`, and `type_weights` shape the 9-role size/weight matrix. |
| Size | `density` (+ `density_factor`) | Spacing preset plus a continuous multiplier — identity-bearing for some presets (brutalist runs 0.88). |
|  | `geometry` | Radius scale (`sm/md/lg/pill`) and border-width scale (`hair/thin/regular/thick`). |
| Shell | `shell_mode` / `shell_texture` | The two-surface model: a shell band around the paper (flush / raised / float / transparent) and its texture (ruled / grid / dotted / grain). Textures fall through to the paper when the shell has no band. |
| Marks | `marks` | Viz-mark identity: `point_shape` (circle / square / diamond / triangle) and `interval_weight` (hair / regular / thick). |
| Effects | `effects` | Glow, gradient shell, elevation, glass (frosted / aurora), caption style (chip / stripe / both), header and title styles. |
| Data | `categorical` / `sequential` / `diverging` | Data palette schemes for multi-effect series. |

Build a custom theme by passing any of these to `web_theme()` — anchors
accept hex strings or `oklch()` triples:

```{r}
#| eval: false
my_theme <- web_theme(
  name   = "my-brand",
  brand  = "#1F3A5F",                       # navy identity
  accent = "#B08938",                       # warm gold engagement
  fonts_body = "Charter, Georgia, serif",
  density = "comfortable",
  effects = list(title_style = "bar")
) |>
  set_pin("--tv-ink2", "#862721")           # oxblood rubrication (pinnable token)
```

Defaults are predictable: `accent` mirrors `brand`; `fonts_display`
mirrors `fonts_body`; paper/ink default to a clinical near-white /
near-black pair carrying the brand hue at low chroma.

### What each anchor actually drives

* **`paper` → `ink`** define the neutral ramp — eleven OKLCH steps that
  every surface tone, divider, and text color binds to. Tint them
  (`paperC` / `inkC` in `derive_preset_anchors()`, or [tint_from_brand()])
  for warm-paper editorial looks.
* **`brand`** seeds the brand ramp: header bands, the brand gradient
  (caption stripe, title underline), series slot 1, and the pooled-effect
  diamond.
* **`accent`** owns the *engagement layer* — hover, selected, semantic
  callouts, the highlight bar. Keep it visually distinct from brand.

The `--tv-ink2` rubrication token survives as a pinnable token — set it
with `set_pin("--tv-ink2", "#hex")` for an editorial second color (e.g.
oxblood significance stars or caption chips) without disturbing the
cascade.

To pin any *derived* field directly, use `set_theme_field()` — explicit
pins survive re-resolution.

### L1 group header and banding

When `theme@row@banding` cycles at the L1 level (the default for every
preset — `banding = "group"` with one nesting level), the banding cycle
already paints alternating fills across the L1 group rows. The renderer
detects this and **drops the explicit `theme@row_group@L1@bg` paint**
on the affected rows so the banded color reads as continuous and the
group-bar doesn't double-paint on top.

If you want a custom L1 bar paired with a banded background, switch
banding modes (e.g. `set_theme_field(c("row", "banding"), "row")` or
`"none"`) so the L1 paint isn't subsumed.

At deeper banding levels (e.g. `banding = "group-2"` with nested groups),
the L1 bar's explicit fill is preserved unchanged — only the banded
level loses its explicit fill.

### Semantic tokens — naming visual treatments {#semantic-tokens}

Tabviz ships with **five semantic tokens** that name visual treatments
you can apply to a row or cell. Each token is a `RowSemantic` bundle on
`theme.row`: `bg`, `fg`, `border`, `marker_fill`, `font_weight`,
`font_style`. The renderer picks one token per row/cell and paints its
bundle.

| Token       | Default treatment                                                    |
|-------------|----------------------------------------------------------------------|
| `emphasis`  | Bold + primary text and marker                                       |
| `muted`     | Lighter color, reduced prominence                                    |
| `accent`    | Bold + accent-color text and marker (most common)                    |
| `bold`      | Just a weight bump — no color change                                 |
| `fill`      | Bold + a strong row fill (defaults to a derived `theme.semantic.fill`); `fg` auto-contrasts |

Two paths to apply a token to a row:

1. **Data column** — pass a logical column to `tabviz()`:

   ```{r}
   #| eval: false
   tabviz(data, label = "study",
     columns = list(viz_forest("hr", "lo", "hi")),
     row_accent = "is_primary",      # rows where is_primary == TRUE → accent token
     row_fill   = "is_overall"       # rows where is_overall == TRUE → fill token
   )
   ```

2. **Painter (interactive)** — click the paintbrush icon, pick a chip
   (`emphasis` / `muted` / `accent` / `bold` / `fill`), then click rows
   or cells to stamp the token on. Hover before clicking shows a
   semi-transparent preview of the would-be visual. The painter
   defaults to `accent` + row-scope on first open. Toggle Row/Cell
   scope from the popover.

#### Customizing token visuals

Two ways to override:

* **Tier-2 named color** for the fill background — pin a specific color
  for the `fill` token instead of letting it derive from `accent`:

  ```{r}
  #| eval: false
  set_theme_field(t, c("semantic", "fill"), "#1F3A5F")  # custom fill navy
  ```


* **Per-bundle override** for any field on any token:

  ```{r}
  #| eval: false
  set_theme_field(t, c("row", "fill",   "fg"),     "#FFFFFF")  # force white text on fill rows
  set_theme_field(t, c("row", "accent", "border"), "#C8553D")  # red border on accent rows
  ```

### Variants — three knobs every table can flip

Variants are per-table choices that don't change the theme's identity. They're orthogonal to inputs and freely composable:

```{r}
#| eval: false
web_theme(
  brand              = "#1F3A5F",
  density            = "comfortable",   # compact | comfortable | spacious
  header_style       = "tint",          # light | tint | bold (legacy variant)
  first_column_style = "bold",          # default | bold (Excel-style; opt-in, no preset uses it)
  effects = list(header_style = "fill") # Tier-1 form — supersedes the variant
)
```

* **`density`** drives every spacing token (row height, padding, gaps). The three presets cover almost every screen and print scenario.
* **`header_style`** picks between a light header, a brand-tinted band, and a solid-ink fill band with inverse text. Prefer the Tier-1 input (`effects = list(header_style = "fill")`), which supersedes the legacy variant; both are always available on every theme.
* **`first_column_style`** gives the row-label column an Excel-style emphasized treatment (background tint + heavier weight). Useful when your first column is genuinely a label, not data.

You can also set variants directly on the `tabviz()` call as overrides (see [tabviz reference](../reference/tabviz.qmd)).

## The three editing verbs — and what each edit travels with

Every theme-editing surface in tabviz — the R modifiers, the widget's
settings panel — performs exactly one of three verbs. Knowing
which verb you're using tells you how far the edit reaches and how it
behaves when the theme changes around it.

| Verb | What it does | Reach | Functions |
|---|---|---|---|
| **Re-tune a role** | Change what a named role *means* (`text-muted` → a darker grade; the `footnote` type role → a smaller step) | **Global** — everything bound to that role follows | `set_role()`, `set_type_role()`, `set_corners()`, `set_rules()` |
| **Re-route a component channel** | Point one component's channel at a *different* role (`title` color → `accent-text`; `axis-label` family → `numeric`) | **Local** to that component — but still cascade-coherent: the routed role keeps re-resolving under polarity flips, anchor edits, and high-contrast mode | `set_component()` (the interactive surface is being redesigned) |
| **Pin** | Hard-code one token to a raw value (`--tv-text-footnote-size: "0.7rem"`) | Local and **inert** — a pinned value ignores polarity, anchors, and accessibility modes. Last resort. | `set_pin()` |

```{r}
#| eval: false
web_theme_nejm() |>
  set_role("text-muted", "neutral", 9) |>            # re-tune (global)
  set_component("title", col = "accent-text") |>     # re-route (local, coherent)
  set_pin("--tv-text-footnote-size", "0.7rem")       # pin (local, inert)
```

Discover what each verb can address: [`list_roles()`] (every tunable
role across color, type, and geometry domains), [`list_components()`]
(every routable component, its region, states, and channels), and
[`inspect_token()`] (where any single token's value comes from).

### What travels with what

Edits land in one of three tiers, and the tier decides where the edit
follows you:

* **Theme** — roles, re-routes, pins, fonts, density. Travels with the
  portable theme artifact (`theme_to_wire()` / `write_theme()`, the
  interactive export). Send the JSON to a colleague and every edit arrives.
* **Figure** — column widths and order, row-kind height pins, banding
  override, watermark. Rides the spec (`spec.figureLayout`), survives
  data updates, round-trips through Shiny via `set_figure_layout()`.
* **View** — zoom, fit, contrast mode. Per-reader, per-document
  (browser local storage); never serialized.

The settings panel is organized around exactly this split: the theme
half edits what travels with the theme, the *This figure* half edits
figure state, and the toolbar's zoom controls are view state.

## How it cascades under the hood

A theme has three tiers. You only ever interact with Tier 1; the other tiers are computed.

```{mermaid}
flowchart TB
    subgraph T1["Tier 1 — Inputs (you edit these)"]
      anchors["anchors: paper · ink · brand · accent"] & axes["polarity · monochrome · mode"] & rest["fonts · density · geometry · marks · effects"]
    end

    subgraph T0["Ramps (generated)"]
      ramps["neutral / brand / accent — 11 OKLCH steps each + alpha companions"]
    end

    subgraph T2["Tier 2 — Roles (derived)"]
      roles["surface · text · border · accent-solid · highlight · status · series slots"]
    end

    subgraph T3["Tier 3 — Component tokens (the wire)"]
      tokens["~150 --tv-* CSS variables: row, header, cell, plot, shell, glass, caption …"]
    end

    T1 --> T0 --> T2 --> T3

    style T1 fill:#dbeafe
    style T0 fill:#fef3c7
    style T2 fill:#dcfce7
    style T3 fill:#f3e8ff
```

**One declarative pipeline.** Anchors expand into ramps; roles bind to
ramp grades; every component token declares which resolver realizes it.
Polarity reflection happens at the anchor level (one place); high-contrast
mode remaps roles and drops effects at resolve time. Changing `brand`
re-derives everything brand-touched and nothing else — explicit pins
survive re-resolution.

**OKLCH derivation.** All blends, lightenings, and darkenings happen in OKLCH (perceptually uniform). The result: tints stay readable across the brightness range, and high-chroma brands don't mud out when blended with surfaces.

**Slot bundles.** Each series anchor expands into a 7-field bundle: `fill`, `stroke`, `fill_muted`, `stroke_muted`, `fill_emphasis`, `stroke_emphasis`, `text_fg`. The renderer picks the right field based on row state (default / muted / emphasized) without you having to define every state by hand. Override individual fields if you need to.

If you want to inspect a resolved theme:

```{r}
#| eval: false
t <- web_theme_nejm()
# Resolved --tv-* tokens (the canonical inspection path):
theme_css_vars(t)[["--tv-surface-bg"]]  # paper / surface background
theme_css_vars(t)[["--tv-text"]]        # primary ink
theme_css_vars(t)[["--tv-accent"]]      # accent
# Or read S7 slots directly:
t@accent@default     # accent role
t@series[[1]]@fill   # first series anchor
t@row@hover@bg       # accent.muted blend
```

## Step 3: Customize without rewriting

Three pipe-friendly modifiers cover the common edits:

```{r}
#| eval: false
custom <- web_theme_nejm() |>
  set_brand("#1F3A5F") |>
  set_accent("#B08938") |>
  set_density("comfortable") |>
  set_effects(header_style = "fill", title_style = "bar") |>
  set_spacing(row_height = 28, axis_gap = 14)
```

* **Focused setters** — one verb per axis: `set_brand()`, `set_accent()`,
  `set_paper()`, `set_ink()`, `set_polarity()`,
  `set_monochrome()`, `set_mode()`, `set_density()`, `set_geometry()`,
  `set_curves()`, `set_effects()`, `set_status()` /
  `set_status_palette()`, and `tint_from_brand()` (nudges paper / ink /
  accent hues toward brand in one move).
* **`set_inputs(theme, ...)`** — multi-input escape hatch; accepts anchor
  names as hex/oklch plus any flat `ThemeInputs` slot. Cascade-aware:
  every change re-resolves with pins preserved.
* **`set_density(theme, x)`**, **`set_header_style(theme, x)`**, **`set_first_column_style(theme, x)`**  per-table variant setters. `set_density` re-resolves (drives spacing tokens); the style setters flip a post-resolution variant.
* **`set_spacing(theme, ...)`** — per-token spacing override on top of the density preset. Set what you care about; the rest follow density.

For surgical overrides anywhere in the theme tree, use `set_theme_field()` with a path:

```{r}
#| eval: false
custom <- web_theme_nejm() |>
  set_theme_field(c("row_group", "L1", "bg"), "#EEEEEE") |>
  set_theme_field(c("series", 1L, "fill"), "#FF0000") |>
  set_theme_field(c("plot", "gridline"), "#003300") |>
  set_theme_field(c("axis", "gridlines"), TRUE)
```

Path entries are property names (strings) or list indices (integers, currently only for `series`). The function re-resolves the theme before returning, so derived fields stay consistent with your override.

### Pinning chrome separately from data

Because the cascades share only Tier 1, you can override a Tier 2 chrome role without disturbing the data marks, or vice versa:

```{r}
#| eval: false
# Pin the surface to a specific cream, regardless of what neutrals say
web_theme_nejm() |>
  set_theme_field(c("surface", "base"), "#FDFCFB")

# Adjust series 2's stroke without affecting any other slot or chrome
web_theme_nejm() |>
  set_theme_field(c("series", 2L, "stroke"), "#444444")
```

Re-resolution preserves explicit overrides on subsequent edits — `set_inputs(brand = "...")` after the override above will refresh series colors but leave `surface.base` pinned to your cream.

## Column house styles — `column_defaults` {#column-defaults}

A theme can carry **per-column-TYPE defaults**, so picking a theme also
picks how its column types dress. The nejm preset turns on significance
stars and a pill for p-value columns; terminal sets numerals in mono and
boxes its badges square; synthwave oversizes glyph columns. Same data,
same `tabviz()` call — the house style rides the theme.

```{r}
#| eval: false
my_theme <- web_theme(
  brand = "#0B6E4F",
  column_defaults = list(
    pvalue = list(stars = TRUE, significantStyle = "pill"),
    badge  = list(shape = "square", outline = TRUE),
    numeric = list(fontClass = "mono")
  )
)
```

The same field exists for JS authors (`buildTheme({ column_defaults:
{...} })` — option keys are the wire camelCase). Three rules keep it
safe, enforced by the engine at every ingest:

1. **The author always wins.** A theme default fills an option only
   while it sits at its schema default; an explicit
   `col_pvalue(stars = FALSE)` is never clobbered. Switching themes
   re-bases cleanly — the old theme's house style never sticks.
2. **Kind gate.** Only options whose schema kind is `styling` or
   `presentation` are accepted; `core` options (the data semantics —
   fields, thresholds, precision) are dropped. Ask
   `column_schema("pvalue")` which options a theme may touch — the
   `themeable` column.
3. **Untrusted-wire safety.** `column_defaults` travels in the portable
   theme artifact, so every string value passes the same grammar gate
   as token pins before it can reach an exported SVG attribute.

Discover the vocabulary with `list_column_types()` (the `themeable`
count per type) and `column_schema(type)` (per-option detail).

## Going deeper: spacing and typography

### Density preset values

| Token | compact | comfortable | spacious |
|-------|--------:|------------:|---------:|
| `row_height` | 20 | 24 | 30 |
| `header_height` | 26 | 32 | 40 |
| `padding` | 8 | 12 | 16 |
| `axis_gap` | 8 | 12 | 16 |
| `cell_padding_x` | 8 | 10 | 14 |
| `column_group_padding` | 6 | 8 | 12 |
| `title_subtitle_gap` | 10 | 13 | 18 |
| `bottom_margin` | 12 | 16 | 22 |
| `indent_per_level` | 14 | 16 | 20 |

`set_spacing()` overrides any of these per-token; the rest fall back to the active density preset. Read [Density](../concepts/styling-cascade.qmd) for how spacing tokens flow into the renderer.

### Aspect-aware spacing {#aspect-aware-spacing}

When the in-widget [aspect-ratio slider](interactivity.qmd#aspect-ratio-slider) is engaged (or `set_aspect_ratio()` is pinned on the spec), the lever ladder scales `row_height` to absorb the height delta. The Spacing settings tab reflects this live:

- A small banner at the top of the tab explains: *"Row height is currently driven by the aspect-ratio slider. Editing it will clear the aspect target."*
- The `Row height` field reads its **effective** value (post-lever-ladder) rather than the raw spec value, and is labelled `Row height (aspect)` to flag its provenance.
- Editing the field clears `target_aspect` (so your typed value pins, the slider returns to natural). View-source emits the new `set_spacing(row_height = ...)` call cleanly, with no overlap with `set_aspect_ratio()`.

Other spacing tokens (`header_height`, `axis_gap`, `header_gap`, `footer_gap`, etc.) keep their authored values in the live widget, so the slider's effect is concentrated on `row_height`. Static export via `save_plot()` honours the full lever ladder including chrome scaling — when achieved aspect deviates from target by > 5 %, a `cli::cli_warn()` surfaces the residual.

### Typography roles

A theme exposes nine named text roles, each emitting
`--tv-text-<role>-{family,size,weight}` tokens from the modular scale
(`type_base_size` × `type_scale_ratio`^n):

| Role | Used for |
|------|----------|
| `title` | Plot title (display family, large, semibold) |
| `subtitle` | Subtitle (body family, medium) |
| `caption` | Caption above the plot |
| `footnote` | Footnote below the plot |
| `body` | Default body text (cells, headers when not bold) |
| `numeric` | Numeric cells (tabular figures) |
| `cell` | Data cell text (defaults to body) |
| `label` | Axis labels |
| `tick` | Axis tick labels (smaller) |

Edit individual fields with `set_theme_field()`:

```{r}
#| eval: false
web_theme_nejm() |>
  set_theme_field(c("text", "title", "weight"), 700) |>
  set_theme_field(c("text", "title", "italic"), TRUE)
```

## The nine presets — when to use which

The roster is nine committed identities (the 27→9 cull, 2026-06): the
**clinical** default `nejm` plus `ledger`, the **design movements**
`brutalist` and `newsprint`, the **moods** `aurora`, `terminal`,
`blueprint`, and `synthwave`, and the **storyteller** `dwarven`. Each
owns a distinct expressive axis; several also carry a column
[house style](#column-defaults).

Every preset is rendered side-by-side on one fixture in
[**the Preset Gallery**](../gallery/presets.qmd) — start there. Identity
is enforced: a CI distinctness gate guarantees no two presets share an
anchor tuple, and the roster spreads typography, density, radius, mark
shapes, and effects rather than just hue.

The everyday recipe is *preset + one brand swap*:

```{r}
#| eval: false
theme = set_brand(web_theme_ledger(), "#1F3A5F")
```

See the [Gallery → LOTR Easter Egg](../gallery/lotr.qmd) page for a worked LOTR example.

## Editing themes interactively

Two surfaces, deliberately split — and the settings panel itself is
**task-framed**: lean by default, with depth one click away.

* **The in-widget settings panel is being REDESIGNED** around five
  surfaces (Variations · Labels · Identity · Plots · Styling — the
  layered-engagement model in `docs/dev/settings-redesign.md`). Until
  it lands, treat the R verbs above as the stable editing surface; the
  current panel's sections are historical and not worth learning.
* **The studio is dormant** — it will be rebuilt after the core
  settings UX lands.

Discover the bindable roles from R with `list_roles()` (the `set_role()`
companion) and the pinnable tokens with `list_component_tokens()` (the
`set_pin()` companion).

## The theme JSON envelope

Both editing surfaces export and import **one** portable artifact:

```json
{
  "$schema": "tabviz-theme/v4",
  "name": "nejm",
  "inputs": { "anchors": { "...": "..." }, "polarity": "light", "...": "..." },
  "roleOverrides": { "text-muted": "brand.8" },
  "pins": { "--tv-text-footnote-size": "0.7rem" }
}
```

* `inputs` are the Tier-1 authoring values — everything the settings
  panel writes.
* `roleOverrides` (`set_role()`) and `pins` (`set_pin()`) are the optional
  total-control members; a settings-only export simply carries them
  empty. Role bindings serialize as stable **name aliases**
  (`"brand.8"` = ramp `brand`, grade 8) — DTCG-shaped and
  rename-migratable; the legacy `{ "ramp": "brand", "grade": 8 }`
  coordinate form is still read back.
* Every egress and ingress — the interactive surfaces and R's
  `read_theme()` — speaks this envelope; `theme_from_wire()` resolves
  one into a `WebTheme`. Files written by the resolved-blob
  `write_theme()` form are also read back transparently.

Pins and role overrides **survive Tier-1 edits**: change the brand
anchor on an imported pinned theme and the pins re-overlay after the
cascade re-resolves.

## Saving a custom theme as a function

For team-wide use, wrap your custom theme in a named function:

```{r}
#| eval: false
web_theme_acme <- function() {
  web_theme(
    name   = "acme",
    brand  = "#FF5500",
    accent = "#0066CC",
    fonts_body    = "'Source Sans 3', system-ui, sans-serif",
    fonts_display = "'Fraunces', Georgia, serif",
    density = "comfortable",
    effects = list(header_style = "tint")
  ) |>
    tint_from_brand("subtle") |>
    set_pin("--tv-ink2", "#CC3300")
}

# Now share it across the team
tabviz(data, label = "study", columns = list(...), theme = web_theme_acme())
```

Pass to `set_theme()` either by string (when registered as a preset) or by handing the constructed `WebTheme` object directly:

```{r}
#| eval: false
my_table |> set_theme(web_theme_acme())
```

## Reference summary

* `web_theme()` — custom-theme constructor: anchors (hex or `oklch()`)
  + polarity / monochrome / mode + fonts + density + geometry + marks +
  effects + status.
* `web_theme_*()` — 9 preset constructors across clinical / editorial /
  design / expressive ([Preset Gallery](../gallery/presets.qmd)).
* `package_themes()` — categorized named list of every preset.
* `set_theme(spec, theme)` — switch a `WebSpec` (or a live proxy) to a
  different theme.
* Focused setters — `set_brand()` `set_accent()`
  `set_paper()` `set_ink()` `set_polarity()` `set_monochrome()`
  `set_mode()` `set_density()` `set_geometry()` `set_curves()`
  `set_effects()` `set_status()` `set_status_palette()`
  `tint_from_brand()`.
* `set_inputs(theme, ...)` — multi-input escape hatch.
* `set_spacing(theme, ...)` — per-token spacing overrides.
* `set_theme_field(theme, path, value)` — generic deep setter; pins
  survive re-resolution.
* `selectable_themes(spec, themes)` — curate the in-widget theme switcher.

For complete argument lists and class definitions see the [Themes reference](../reference/themes.qmd).
