---
title: "Themes"
---

```{r}
#| include: false
library(tabviz)

meta_data <- data.frame(
  study = c("Smith 2020", "Jones 2021", "Lee 2022", "Chen 2023", "Park 2024"),
  hr    = c(0.72, 0.85, 0.91, 0.68, 0.79),
  lower = c(0.55, 0.70, 0.75, 0.52, 0.61),
  upper = c(0.95, 1.03, 1.10, 0.89, 1.02)
)
```

A theme decides every visual choice in your table that isn't data: surface colors, text tones, divider weights, spacing, fonts, marker fills. tabviz themes are designed so you can pick a preset, change one or two things, and never think about it again — or build a fully branded theme from a dozen inputs when you need to.

## Three steps from "pick a preset" to "fully bespoke"

Each step builds on the previous. Most users live at step 1.

```{mermaid}
flowchart LR
    A["1. Pick a preset"] --> B["2. Tweak inputs"]
    B --> C["3. Override component fields"]

    style A fill:#dbeafe
    style B fill:#bfdbfe
    style C fill:#93c5fd
```

* **Step 1** — `theme = web_theme_lancet()`. Done. Seven presets cover the common publication, dashboard, and illustrative surfaces.
* **Step 2** — Swap a primary color, change density, switch to a serif font. The preset stays the foundation; you nudge a few inputs.
* **Step 3** — Reach into a specific component (e.g. the L1 row-group bar background) for surgical control. Rare; the preset cascade handles 95% of cases.

## Step 1: Pick a preset

```{r}
#| eval: false
tabviz(
  meta_data, label = "study",
  columns = list(
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1)
  ),
  theme = web_theme_lancet()
)
```

Seven presets ship in the package — four **journal-style** themes designed for publication, and three **illustrative LOTR** themes that exercise the full surface for non-journal contexts (talks, blog posts, internal reports). Each leans into a clear identity rather than trying to be everything.

| Preset | Style | Designed for |
|--------|-------|--------------|
| `web_theme_cochrane()` | The package default. Cochrane teal + warm coral, Inter sans-serif, comfortable density | General purpose, systematic reviews |
| `web_theme_lancet()`   | Lancet navy + refined old-gold + warm cream, Georgia serif, comfortable | Lancet-style editorial layouts |
| `web_theme_jama()`     | Pure black-and-white, ultra-compact, strong dividers throughout, Arial | JAMA / print journal submissions |
| `web_theme_dark()`     | Catppuccin Mocha-inspired dark canvas with pastel marker palette | Dark-mode viewing |
| `web_theme_dwarven()`  | Stone neutrals, forge-glow accent, broad serif display | Talks, blog posts, themed reports |
| `web_theme_elvish()`   | Twilight surface, mithril silvers, calligraphic serif | Talks, blog posts, themed reports |
| `web_theme_hobbit()`   | Parchment + green, hand-set body face, hearth accent | Talks, blog posts, themed reports |

Switch presets at any point with `set_theme()`:

```{r}
#| eval: false
my_table |> set_theme("jama")
```

The full registry is in `package_themes()` (returns a 2-level list keyed `journals` / `lotr`).

### Curating the in-widget switcher

The in-widget settings panel includes a theme dropdown that lets readers
swap between presets without reloading. By default it shows every package
theme; pass `enable_themes = list(...)` to `web_interaction()` or
`selectable_themes()` to curate the menu:

```{r}
#| eval: false
# Flat list — no tabs, just a single column of choices.
selectable_themes(my_table, list(
  Cochrane = web_theme_cochrane(),
  Lancet   = web_theme_lancet()
))

# 2-level named list — each top-level key becomes a tab in the dropdown.
selectable_themes(my_table, list(
  Editorial = list(
    Cochrane = web_theme_cochrane(),
    Lancet   = web_theme_lancet(),
    JAMA     = web_theme_jama()
  ),
  Other = list(Dark = web_theme_dark())
))
```

The active theme is auto-included so readers can always revert. For
categorized lists, if the active theme isn't in any user-supplied
category, a `Current` tab is prepended.

## Step 2: Tweak inputs

Every theme — including every preset — is determined by the **V4 Tier-1
inputs**: a handful of OKLCH color anchors plus orthogonal axes for
polarity, accessibility, typography, density, geometry, marks, and effects.

| Group | Input | What it controls |
|-------|-------|------------------|
| Anchors | `paper` | Light-end neutral anchor. The neutral ramp interpolates paper → ink; surfaces, dividers, and body text all derive from it. |
|  | `ink` | Dark-end neutral anchor (text, strong rules). |
|  | `brand` | The identity hue. Seeds the brand ramp: headers, gradients, series slot 1, title color. |
|  | `accent` | Optional engagement hue — hover, selected, callouts, highlight bar. Mirrors brand when unset. |
|  | `ink2` | Optional **rubrication ink**. When set it seeds the accent ramp with precedence over `accent` — the editorial channel for significance stars and caption chips, without tinting interaction states by accident. |
| Axes | `polarity` | `"light"` / `"dark"` — reflects every anchor's lightness around the midpoint. |
|  | `monochrome` | The neutral ramp rides the brand hue (phosphor / sepia / cyanotype themes in one toggle). |
|  | `mode` | Accessibility: `"standard"` / `"high-contrast"` / `"reduced-transparency"`. Orthogonal to polarity. |
| Status | `status_positive` … `status_info` | Semantic palette. Theme-scoped identity (synthwave's neon vs a journal's clinical green); reverts to a curated HC-safe palette under high-contrast mode. |
| Type | `fonts_body` / `fonts_display` / `fonts_mono` | Font stacks; `fonts_display` mirrors body when unset. `type_base_size`, `type_scale_ratio`, and `type_weights` shape the 9-role size/weight matrix. |
| Size | `density` (+ `density_factor`) | Spacing preset plus a continuous multiplier — identity-bearing for some presets (brutalist runs 0.88). |
|  | `geometry` | Radius scale (`sm/md/lg/pill`) and border-width scale (`hair/thin/regular/thick`). |
| Shell | `shell_mode` / `shell_texture` | The two-surface model: a shell band around the paper (flush / raised / float / transparent) and its texture (ruled / grid / dotted / grain). Textures fall through to the paper when the shell has no band. |
| Marks | `marks` | Viz-mark identity: `point_shape` (circle / square / diamond / triangle) and `interval_weight` (hair / regular / thick). |
| Effects | `effects` | Glow, gradient shell, elevation, glass (frosted / aurora), caption style (chip / stripe / both), header and title styles. |
| Data | `categorical` / `sequential` / `diverging` | Data palette schemes for multi-effect series. |

Build a custom theme by passing any of these to `web_theme()` — anchors
accept hex strings or `oklch()` triples:

```{r}
#| eval: false
my_theme <- web_theme(
  name   = "my-brand",
  brand  = "#1F3A5F",                       # navy identity
  accent = "#B08938",                       # warm gold engagement
  ink2   = "#862721",                       # oxblood rubrication
  fonts_body = "Charter, Georgia, serif",
  density = "comfortable",
  effects = list(title_style = "bar")
)
```

Defaults are predictable: `accent` mirrors `brand`; `ink2` falls back to
`accent`; `fonts_display` mirrors `fonts_body`; paper/ink default to a
clinical near-white / near-black pair carrying the brand hue at low chroma.

### What each anchor actually drives

* **`paper` → `ink`** define the neutral ramp — eleven OKLCH steps that
  every surface tone, divider, and text color binds to. Tint them
  (`paperC` / `inkC` in `derive_preset_anchors()`, or [tint_from_brand()])
  for warm-paper editorial looks.
* **`brand`** seeds the brand ramp: header bands, the brand gradient
  (caption stripe, title underline), series slot 1, and the pooled-effect
  diamond.
* **`accent`** owns the *engagement layer* — hover, selected, semantic
  callouts, the highlight bar. Keep it visually distinct from brand.
* **`ink2`** seeds the accent ramp with precedence when set — so
  significance stars (`col_pvalue(stars = TRUE)`) and caption chips pick
  up the rubrication ink while the rest of the theme stays disciplined.

To pin any *derived* field directly, use `set_theme_field()` — explicit
pins survive re-resolution.

### L1 group header and banding

When `theme@row@banding` cycles at the L1 level (the default for every
preset — `banding = "group"` with one nesting level), the banding cycle
already paints alternating fills across the L1 group rows. The renderer
detects this and **drops the explicit `theme@row_group@L1@bg` paint**
on the affected rows so the banded color reads as continuous and the
group-bar doesn't double-paint on top.

If you want a custom L1 bar paired with a banded background, switch
banding modes (e.g. `set_theme_field(c("row", "banding"), "row")` or
`"none"`) so the L1 paint isn't subsumed.

At deeper banding levels (e.g. `banding = "group-2"` with nested groups),
the L1 bar's explicit fill is preserved unchanged — only the banded
level loses its explicit fill.

### Semantic tokens — naming visual treatments {#semantic-tokens}

Tabviz ships with **five semantic tokens** that name visual treatments
you can apply to a row or cell. Each token is a `RowSemantic` bundle on
`theme.row`: `bg`, `fg`, `border`, `marker_fill`, `font_weight`,
`font_style`. The renderer picks one token per row/cell and paints its
bundle.

| Token       | Default treatment                                                    |
|-------------|----------------------------------------------------------------------|
| `emphasis`  | Bold + primary text and marker                                       |
| `muted`     | Lighter color, reduced prominence                                    |
| `accent`    | Bold + accent-color text and marker (most common)                    |
| `bold`      | Just a weight bump — no color change                                 |
| `fill`      | Bold + a strong row fill (defaults to a derived `theme.semantic.fill`); `fg` auto-contrasts |

Two paths to apply a token to a row:

1. **Data column** — pass a logical column to `tabviz()`:

   ```{r}
   #| eval: false
   tabviz(data, label = "study",
     columns = list(viz_forest("hr", "lo", "hi")),
     row_accent = "is_primary",      # rows where is_primary == TRUE → accent token
     row_fill   = "is_overall"       # rows where is_overall == TRUE → fill token
   )
   ```

2. **Painter (interactive)** — click the paintbrush icon, pick a chip
   (`emphasis` / `muted` / `accent` / `bold` / `fill`), then click rows
   or cells to stamp the token on. Hover before clicking shows a
   semi-transparent preview of the would-be visual. The painter
   defaults to `accent` + row-scope on first open. Toggle Row/Cell
   scope from the popover.

#### Customizing token visuals

Two ways to override:

* **Tier-2 named color** for the fill background — pin a specific color
  for the `fill` token instead of letting it derive from `accent`:

  ```{r}
  #| eval: false
  set_theme_field(t, c("semantic", "fill"), "#1F3A5F")  # custom fill navy
  ```

  Or interactively: Settings panel → **Tokens** tab → Token colors.

* **Per-bundle override** for any field on any token:

  ```{r}
  #| eval: false
  set_theme_field(t, c("row", "fill",   "fg"),     "#FFFFFF")  # force white text on fill rows
  set_theme_field(t, c("row", "accent", "border"), "#C8553D")  # red border on accent rows
  ```

  Or interactively: Settings panel → **Tokens** tab → expand a token →
  edit any field. All bundle fields are override-aware (reset reverts
  to the resolved default).

### Variants — three knobs every table can flip

Variants are per-table choices that don't change the theme's identity. They're orthogonal to inputs and freely composable:

```{r}
#| eval: false
web_theme(
  brand              = "#1F3A5F",
  density            = "comfortable",   # compact | comfortable | spacious
  header_style       = "tint",          # light | tint | bold (legacy variant)
  first_column_style = "bold",          # default | tint | bold (Excel-style)
  effects = list(header_style = "fill") # Tier-1 form — supersedes the variant
)
```

* **`density`** drives every spacing token (row height, padding, gaps). The three presets cover almost every screen and print scenario.
* **`header_style`** picks between a light header, a brand-tinted band, and a solid-ink fill band with inverse text. Prefer the Tier-1 input (`effects = list(header_style = "fill")`), which supersedes the legacy variant; both are always available on every theme.
* **`first_column_style`** gives the row-label column an Excel-style emphasized treatment (background tint + heavier weight). Useful when your first column is genuinely a label, not data.

You can also set variants directly on the `tabviz()` call as overrides (see [tabviz reference](../reference/tabviz.qmd)).

## How it cascades under the hood

A theme has three tiers. You only ever interact with Tier 1; the other tiers are computed.

```{mermaid}
flowchart TB
    subgraph T1["Tier 1 — Inputs (you edit these)"]
      anchors["anchors: paper · ink · brand · accent · ink2"] & axes["polarity · monochrome · mode"] & rest["fonts · density · geometry · marks · effects"]
    end

    subgraph T0["Ramps (generated)"]
      ramps["neutral / brand / accent — 11 OKLCH steps each + alpha companions"]
    end

    subgraph T2["Tier 2 — Roles (derived)"]
      roles["surface · text · border · accent-solid · highlight · status · series slots"]
    end

    subgraph T3["Tier 3 — Component tokens (the wire)"]
      tokens["~150 --tv-* CSS variables: row, header, cell, plot, shell, glass, caption …"]
    end

    T1 --> T0 --> T2 --> T3

    style T1 fill:#dbeafe
    style T0 fill:#fef3c7
    style T2 fill:#dcfce7
    style T3 fill:#f3e8ff
```

**One declarative pipeline.** Anchors expand into ramps; roles bind to
ramp grades; every component token declares which resolver realizes it.
Polarity reflection happens at the anchor level (one place); high-contrast
mode remaps roles and drops effects at resolve time. Changing `brand`
re-derives everything brand-touched and nothing else — explicit pins
survive re-resolution.

**OKLCH derivation.** All blends, lightenings, and darkenings happen in OKLCH (perceptually uniform). The result: tints stay readable across the brightness range, and high-chroma brands don't mud out when blended with surfaces.

**Slot bundles.** Each series anchor expands into a 7-field bundle: `fill`, `stroke`, `fill_muted`, `stroke_muted`, `fill_emphasis`, `stroke_emphasis`, `text_fg`. The renderer picks the right field based on row state (default / muted / emphasized) without you having to define every state by hand. Override individual fields if you need to.

If you want to inspect a resolved theme:

```{r}
#| eval: false
t <- web_theme_lancet()
t@surface@base       # "#FDFCFB"
t@content@primary    # "#1E3A5F"
t@accent@tint_subtle # tinted accent for L1 group bar
t@series[[1]]@fill   # first series anchor
t@series[[1]]@stroke # auto-darkened companion
t@row@hover@bg       # accent.muted blend
```

## Step 3: Customize without rewriting

Three pipe-friendly modifiers cover the common edits:

```{r}
#| eval: false
custom <- web_theme_jama() |>
  set_brand("#1F3A5F") |>
  set_accent("#B08938") |>
  set_density("comfortable") |>
  set_effects(header_style = "fill", title_style = "bar") |>
  set_spacing(row_height = 28, axis_gap = 14)
```

* **Focused setters** — one verb per axis: `set_brand()`, `set_accent()`,
  `set_ink2()`, `set_paper()`, `set_ink()`, `set_polarity()`,
  `set_monochrome()`, `set_mode()`, `set_density()`, `set_geometry()`,
  `set_curves()`, `set_effects()`, `set_status()` /
  `set_status_palette()`, and `tint_from_brand()` (nudges paper / ink /
  accent / ink2 hues toward brand in one move).
* **`set_inputs(theme, ...)`** — multi-input escape hatch; accepts anchor
  names as hex/oklch plus any flat `ThemeInputs` slot. Cascade-aware:
  every change re-resolves with pins preserved.
* **`set_density(theme, x)`**, **`set_header_style(theme, x)`**, **`set_first_column_style(theme, x)`**  per-table variant setters. `set_density` re-resolves (drives spacing tokens); the style setters flip a post-resolution variant.
* **`set_spacing(theme, ...)`** — per-token spacing override on top of the density preset. Set what you care about; the rest follow density.

For surgical overrides anywhere in the theme tree, use `set_theme_field()` with a path:

```{r}
#| eval: false
custom <- web_theme_cochrane() |>
  set_theme_field(c("row_group", "L1", "bg"), "#EEEEEE") |>
  set_theme_field(c("series", 1L, "fill"), "#FF0000") |>
  set_theme_field(c("plot", "gridline"), "#003300") |>
  set_theme_field(c("axis", "gridlines"), TRUE)
```

Path entries are property names (strings) or list indices (integers, currently only for `series`). The function re-resolves the theme before returning, so derived fields stay consistent with your override.

### Pinning chrome separately from data

Because the cascades share only Tier 1, you can override a Tier 2 chrome role without disturbing the data marks, or vice versa:

```{r}
#| eval: false
# Pin the surface to a specific cream, regardless of what neutrals say
web_theme_cochrane() |>
  set_theme_field(c("surface", "base"), "#FDFCFB")

# Adjust series 2's stroke without affecting any other slot or chrome
web_theme_cochrane() |>
  set_theme_field(c("series", 2L, "stroke"), "#444444")
```

Re-resolution preserves explicit overrides on subsequent edits — `set_inputs(brand = "...")` after the override above will refresh series colors but leave `surface.base` pinned to your cream.

## Going deeper: spacing and typography

### Density preset values

| Token | compact | comfortable | spacious |
|-------|--------:|------------:|---------:|
| `row_height` | 20 | 24 | 30 |
| `header_height` | 26 | 32 | 40 |
| `padding` | 8 | 12 | 16 |
| `axis_gap` | 8 | 12 | 16 |
| `cell_padding_x` | 8 | 10 | 14 |
| `column_group_padding` | 6 | 8 | 12 |
| `title_subtitle_gap` | 10 | 13 | 18 |
| `bottom_margin` | 12 | 16 | 22 |
| `indent_per_level` | 14 | 16 | 20 |

`set_spacing()` overrides any of these per-token; the rest fall back to the active density preset. Read [Density](../concepts/styling-cascade.qmd) for how spacing tokens flow into the renderer.

### Aspect-aware spacing {#aspect-aware-spacing}

When the in-widget [aspect-ratio slider](interactivity.qmd#aspect-ratio-slider) is engaged (or `set_aspect_ratio()` is pinned on the spec), the lever ladder scales `row_height` to absorb the height delta. The Spacing settings tab reflects this live:

- A small banner at the top of the tab explains: *"Row height is currently driven by the aspect-ratio slider. Editing it will clear the aspect target."*
- The `Row height` field reads its **effective** value (post-lever-ladder) rather than the raw spec value, and is labelled `Row height (aspect)` to flag its provenance.
- Editing the field clears `target_aspect` (so your typed value pins, the slider returns to natural). View-source emits the new `set_spacing(row_height = ...)` call cleanly, with no overlap with `set_aspect_ratio()`.

Other spacing tokens (`header_height`, `axis_gap`, `header_gap`, `footer_gap`, etc.) keep their authored values in the live widget, so the slider's effect is concentrated on `row_height`. Static export via `save_plot()` honours the full lever ladder including chrome scaling — when achieved aspect deviates from target by > 5 %, a `cli::cli_warn()` surfaces the residual.

### Typography roles

A theme exposes nine named text roles, each emitting
`--tv-text-<role>-{family,size,weight}` tokens from the modular scale
(`type_base_size` × `type_scale_ratio`^n):

| Role | Used for |
|------|----------|
| `title` | Plot title (display family, large, semibold) |
| `subtitle` | Subtitle (body family, medium) |
| `caption` | Caption above the plot |
| `footnote` | Footnote below the plot |
| `body` | Default body text (cells, headers when not bold) |
| `numeric` | Numeric cells (tabular figures) |
| `cell` | Data cell text (defaults to body) |
| `label` | Axis labels |
| `tick` | Axis tick labels (smaller) |

Edit individual fields with `set_theme_field()`:

```{r}
#| eval: false
web_theme_cochrane() |>
  set_theme_field(c("text", "title", "weight"), 700) |>
  set_theme_field(c("text", "title", "italic"), TRUE)
```

## The twenty-seven presets — when to use which

The roster spans five categories — **journals** (Cochrane, Lancet, JAMA,
NEJM, Nature, BMJ), **modes** (dark), **design traditions** (Bauhaus,
Swiss, Tufte, Newsprint, Solarized ×2, Tonal ×2), **Middle-earth**
(Dwarven, Elvish, Hobbit), and the **showcase** set that exercises the
full effects surface (Synthwave, Brutalist, Atelier, Executive, Ledger,
Terminal, Aurora, Blueprint, Sunprint).

Every preset is rendered side-by-side on one fixture in
[**the Preset Gallery**](../gallery/presets.qmd) — start there. Identity
is enforced: a CI distinctness gate guarantees no two presets share an
anchor tuple, and the roster spreads typography, density, radius, mark
shapes, and effects rather than just hue.

The everyday recipe is *preset + one brand swap*:

```{r}
#| eval: false
theme = set_brand(web_theme_ledger(), "#1F3A5F")
```

See the [Gallery → LOTR Easter Egg](../gallery/lotr.qmd) page for a worked LOTR example.

## Editing themes interactively

Two surfaces, deliberately split:

* **The widget cog** is the *everyday* surface: theme switcher in the
  toolbar plus a curated drawer (polarity + density). Quick flips, not
  theme design.
* **The Studio** ([open it here](../studio.qmd)) is the *forge*:
  LCH-native anchor sliders (paper / ink / brand / accent / ink2), a
  five-tab rail (Color / Type / Size / Shell / Effects), a "Match brand"
  one-click hue alignment, 50-step undo, and a live cascade
  visualization. **Alt+click any element in the preview** to trace which
  theme token paints it. "Copy R code" exports your edits as a
  reproducible `web_theme()` / `set_*()` chain.

## Saving a custom theme as a function

For team-wide use, wrap your custom theme in a named function:

```{r}
#| eval: false
web_theme_acme <- function() {
  web_theme(
    name   = "acme",
    brand  = "#FF5500",
    accent = "#0066CC",
    ink2   = "#CC3300",
    fonts_body    = "'Source Sans 3', system-ui, sans-serif",
    fonts_display = "'Fraunces', Georgia, serif",
    density = "comfortable",
    effects = list(header_style = "tint")
  ) |>
    tint_from_brand("subtle")
}

# Now share it across the team
tabviz(data, label = "study", columns = list(...), theme = web_theme_acme())
```

Pass to `set_theme()` either by string (when registered as a preset) or by handing the constructed `WebTheme` object directly:

```{r}
#| eval: false
my_table |> set_theme(web_theme_acme())
```

## Reference summary

* `web_theme()` — custom-theme constructor: anchors (hex or `oklch()`)
  + polarity / monochrome / mode + fonts + density + geometry + marks +
  effects + status.
* `web_theme_*()` — 27 preset constructors across journals / modes /
  design / lotr / showcase ([Preset Gallery](../gallery/presets.qmd)).
* `package_themes()` — categorized named list of every preset.
* `set_theme(spec, theme)` — switch a `WebSpec` (or a live proxy) to a
  different theme.
* Focused setters — `set_brand()` `set_accent()` `set_ink2()`
  `set_paper()` `set_ink()` `set_polarity()` `set_monochrome()`
  `set_mode()` `set_density()` `set_geometry()` `set_curves()`
  `set_effects()` `set_status()` `set_status_palette()`
  `tint_from_brand()`.
* `set_inputs(theme, ...)` — multi-input escape hatch.
* `set_spacing(theme, ...)` — per-token spacing overrides.
* `set_theme_field(theme, path, value)` — generic deep setter; pins
  survive re-resolution.
* `selectable_themes(spec, themes)` — curate the in-widget theme switcher.

For complete argument lists and class definitions see the [Themes reference](../reference/themes.qmd).
