---
title: "Themes"
---

```{r}
#| include: false
library(tabviz)

# Sample data for examples
meta_data <- data.frame(
  study = c("Smith 2020", "Jones 2021", "Lee 2022", "Chen 2023", "Park 2024"),
  hr = c(0.72, 0.85, 0.91, 0.68, 0.79),
  lower = c(0.55, 0.70, 0.75, 0.52, 0.61),
  upper = c(0.95, 1.03, 1.10, 0.89, 1.02)
)
```

Themes control the visual appearance of forest plots—colors, typography, spacing, and shapes. You can use preset themes for common publication styles, or customize every aspect using the `set_*()` family of functions.

::: {.callout-tip}
## Theme Design Philosophy
Each theme is designed with a specific context in mind:

- **Publication themes** (JAMA, Lancet, Nature): Journal-specific styling for submissions
- **Presentation themes** (modern, presentation, dark): Optimized for screen viewing
- **Utility themes** (default, minimal, cochrane): Clean and adaptable
:::


## Component Summary

| Component | Function | Key Properties |
|-----------|----------|----------------|
| Colors | `set_colors()` | background, foreground, primary, accent, muted, ci_marker_fill, ci_line, summary_fill |
| Typography | `set_typography()` | font_family, font_size_*, font_weight_*, header_font_scale |
| Spacing | `set_spacing()` | row_height, header_height, padding, container_padding, cell_padding_x/y |
| Shapes | `set_shapes()` | point_size, summary_height, line_width, effect_colors, marker_shapes |
| Axis | `set_axis()` | range_min/max, tick_count, tick_values, gridlines, symmetric |
| Layout | `set_layout()` | plot_position, row_border, container_border |
| Group Headers | `set_group_headers()` | level*_font_size, level*_background, indent_per_level |
| Effects | `set_effect_colors()`, `set_marker_shapes()` | Multi-effect default colors/shapes |


## Color Architecture

Colors serve different purposes in tabviz:

| Category | Purpose | Colors |
|----------|---------|--------|
| **Structural** | Define the canvas | `background`, `foreground`, `border` |
| **Semantic** | Convey meaning through styling | `primary`, `accent`, `muted` |
| **Data** | Visualize data elements | `ci_marker_fill`, `ci_line`, `summary_fill` |

### How Palette Colors Connect to Styling

The palette defines **what colors look like**. Semantic styles **apply those colors** to specific elements:

```
┌─────────────────────────────────────────────────────────────────┐
│  PALETTE (set_colors)              STYLES (row_*/column)       │
│  ─────────────────────             ────────────────────        │
│  accent: "#8b5cf6"      ───────►   row_accent, cell accent     │
│  muted: "#94a3b8"       ───────►   row_muted, cell muted       │
│  foreground: "#333"     ───────►   row_emphasis (+ bold)       │
│  primary: "#0891b2"     ───────►   group headers, hover        │
└─────────────────────────────────────────────────────────────────┘
```

**Key insight**: Changing `accent` in the theme automatically updates all `row_accent` and cell `accent` styling.

::: {.callout-note}
## Semantic Styles at Row and Cell Level

The same semantic vocabulary works at both levels:

| Style | Effect | Row Parameter | Cell Parameter |
|-------|--------|---------------|----------------|
| **Emphasis** | Bold + foreground color | `row_emphasis` | `emphasis` in `web_col()` |
| **Muted** | Muted color (de-emphasized) | `row_muted` | `muted` in `web_col()` |
| **Accent** | Accent color (highlighted) | `row_accent` | `accent` in `web_col()` |

Row-level parameters accept column names with TRUE/FALSE values. Cell-level parameters in `web_col()` also accept column names.
:::

## Preset Themes

tabviz includes 9 preset themes designed for different use cases. All themes can be customized using a fluent API.

| Builder | Style | Best for |
|---------|-------|----------|
| `web_theme_default()` | Clean modern with subtle cyan accents, system fonts | General purpose |
| `web_theme_jama()` | Maximum density, pure black & white, Arial, 18px rows | JAMA submissions, print |
| `web_theme_lancet()` | Lancet navy (#00407A), warm off-white, Georgia serif | Lancet submissions |
| `web_theme_nature()` | Nature blue (#1976D2), Helvetica Neue, tight typography | Nature family journals |
| `web_theme_cochrane()` | Teal accents (#0099CC), Arial, 20px rows | Systematic reviews |
| `web_theme_modern()` | Inter font, bright blue (#3B82F6), 36px rows, rounded corners | Web/screen viewing |
| `web_theme_presentation()` | Extra-large fonts, thick lines, 44px rows | Slides and posters |
| `web_theme_dark()` | Catppuccin Mocha-inspired low-contrast palette | Dark-mode viewing |
| `web_theme_minimal()` | Pure black & white, Georgia serif | Classic academic look |

:::{.panel-tabset}

### Default

```{r}
forest_plot(meta_data, point = "hr", lower = "lower", upper = "upper",
  label = "study", null_value = 1, scale = "log",
  theme = web_theme_default())
```

### JAMA

```{r}
forest_plot(meta_data, point = "hr", lower = "lower", upper = "upper",
  label = "study", null_value = 1, scale = "log",
  theme = web_theme_jama())
```

### Lancet

```{r}
forest_plot(meta_data, point = "hr", lower = "lower", upper = "upper",
  label = "study", null_value = 1, scale = "log",
  theme = web_theme_lancet())
```

### Modern

```{r}
forest_plot(meta_data, point = "hr", lower = "lower", upper = "upper",
  label = "study", null_value = 1, scale = "log",
  theme = web_theme_modern())
```

### Presentation

```{r}
forest_plot(meta_data, point = "hr", lower = "lower", upper = "upper",
  label = "study", null_value = 1, scale = "log",
  theme = web_theme_presentation())
```

### Dark

```{r}
forest_plot(meta_data, point = "hr", lower = "lower", upper = "upper",
  label = "study", null_value = 1, scale = "log",
  theme = web_theme_dark())
```

### Minimal

```{r}
forest_plot(meta_data, point = "hr", lower = "lower", upper = "upper",
  label = "study", null_value = 1, scale = "log",
  theme = web_theme_minimal())
```

### Cochrane

```{r}
forest_plot(meta_data, point = "hr", lower = "lower", upper = "upper",
  label = "study", null_value = 1, scale = "log",
  theme = web_theme_cochrane())
```

### Nature

```{r}
forest_plot(meta_data, point = "hr", lower = "lower", upper = "upper",
  label = "study", null_value = 1, scale = "log",
  theme = web_theme_nature())
```

:::

## Customizing Themes

Use the `set_*()` functions to override individual aspects of a preset theme. Each one takes the theme as its first argument and returns a modified copy:

```{r}
custom_theme <- web_theme_jama()
custom_theme <- set_colors(custom_theme,
  primary = "#0066cc", ci_marker_fill = "#0066cc", ci_line = "#0066cc")
custom_theme <- set_spacing(custom_theme, row_height = 24)
custom_theme <- set_axis(custom_theme, gridlines = TRUE, gridline_style = "dotted")

forest_plot(meta_data, point = "hr", lower = "lower", upper = "upper",
  label = "study", null_value = 1, scale = "log",
  theme = custom_theme)
```

::: {.callout-tip}
## Pipe-friendly equivalents

The same `set_*()` calls compose naturally with the pipe (`web_theme_jama() |> set_colors(...) |> set_spacing(...)`). See [The Fluent API](fluent-api.qmd) for that style.
:::

## Theme Components

### Colors

```r
set_colors(
  background        = "#ffffff",   # Background color
  foreground        = "#1a1a1a",   # Primary text color
  primary           = "#2563eb",   # Primary accent for highlights
  secondary         = "#64748b",   # Secondary text/UI color
  accent            = "#8b5cf6",   # Accent color for emphasis
  muted             = "#94a3b8",   # Muted/disabled text color
  border            = "#e2e8f0",   # Border color for dividers
  ci_marker_fill    = "#2563eb",   # Default CI marker fill
  ci_marker_positive = NULL,       # Optional override when upper < null
  ci_marker_negative = NULL,       # Optional override when lower > null
  ci_marker_neutral  = NULL,       # Optional override when CI crosses null
  ci_line           = "#475569",   # CI whisker line color
  summary_fill      = "#2563eb",   # Summary diamond fill
  summary_border    = "#1d4ed8"    # Summary diamond border
)
```

:::{.callout-note}
## Color usage reference

| Color | Where it's used |
|-------|-----------------|
| `background` | Plot container background, table cells |
| `foreground` | Primary text: labels, data values, column headers; `row_emphasis` text |
| `primary` | Row group header backgrounds (reduced opacity), alternating row stripes, hover/selection |
| `secondary` | Subtitles, less prominent UI text |
| `accent` | `row_accent` text styling; annotation labels |
| `muted` | Footnotes, captions; `row_muted` text styling |
| `border` | Row dividers, container borders, table grid lines |
| `ci_marker_fill` | Default CI marker fill (squares, circles, diamonds) |
| `ci_marker_positive` / `_negative` / `_neutral` | Direction-aware fill overrides |
| `ci_line` | CI whisker lines connecting lower to upper bounds |
| `summary_fill` | Summary row diamond fill color |
| `summary_border` | Summary diamond outline/stroke |

**Row styling colors**: use `row_emphasis`, `row_muted`, or `row_accent` to apply these colors to specific rows. See [Row Styling](row-styling.qmd).
:::

:::{.callout-tip}
## Semantic color tips
- Set `ci_marker_fill` and `primary` to the same value for a cohesive look.
- Use a slightly darker `ci_line` than `ci_marker_fill` for subtle depth.
- For conditional coloring (significant vs not), add a color column to your data and map it with `marker_color` in `viz_forest()`.
:::

### Typography

```r
set_typography(
  font_family = "system-ui, sans-serif",  # CSS font-family string
  font_size_sm = "0.75rem",               # Small text (badges, counts)
  font_size_base = "0.875rem",            # Default text size
  font_size_lg = "1rem",                  # Large text (headers)
  font_weight_normal = 400,               # Normal weight
  font_weight_medium = 500,               # Medium weight
  font_weight_bold = 600,                 # Bold weight
  header_font_scale = 1.05                # Header cell font scale relative to base
)
```

### Spacing

```r
set_spacing(
  row_height = 28,          # Data row height in pixels
  header_height = 36,       # Header row height in pixels
  padding = 12,             # Padding around forest plot SVG
  container_padding = 0,    # Left/right padding for outer container
  cell_padding_x = 10,      # Horizontal cell padding
  cell_padding_y = 4,       # Vertical cell padding
  axis_gap = 12             # Gap between table and x-axis
)
```

::: {.callout-note}
## Row Height Guidelines
| Context | Recommended Height |
|---------|-------------------|
| Compact/print | 18-22px |
| Standard web | 28-32px |
| Touch/mobile | 36-40px |
| Presentation | 40-48px |
:::

### Shapes

```r
set_shapes(
  point_size = 6,         # Point marker radius in pixels
  summary_height = 10,    # Summary diamond height
  line_width = 1.5,       # Confidence interval line width
  border_radius = 4,      # Container border radius
  effect_colors = NULL,   # Default colors for multi-effect visualizations
  marker_shapes = c("square", "circle", "diamond", "triangle")
)
```

### Multi-Effect Color Defaults

For visualizations with multiple effects per row (forest plots, bar charts, boxplots, violin plots), colors cycle through theme-defined `effect_colors`:

```{r}
#| eval: false
# Set custom effect colors and shapes for multi-effect visualizations
custom_theme <- web_theme_default()
custom_theme <- set_effect_colors(custom_theme, c("#2563eb", "#dc2626", "#16a34a"))
custom_theme <- set_marker_shapes(custom_theme, c("square", "circle", "diamond"))

forest_plot(data, ..., theme = custom_theme)
```

Effects without explicit `color` in their specification use these theme defaults in order, cycling if there are more effects than colors.

#### Built-in Theme Palettes

Each theme includes a curated 5-color `effect_colors` palette. The swatches below come from each theme's actual `shapes@effect_colors` slot — change a theme and this table re-renders.

```{r}
#| echo: false
#| output: asis
themes_to_show <- list(
  default      = web_theme_default(),
  modern       = web_theme_modern(),
  jama         = web_theme_jama(),
  lancet       = web_theme_lancet(),
  nature       = web_theme_nature(),
  cochrane     = web_theme_cochrane(),
  dark         = web_theme_dark(),
  minimal      = web_theme_minimal(),
  presentation = web_theme_presentation()
)
styles <- c(
  "Balanced, accessible", "Vibrant blue-first",
  "Grayscale for print", "Journal-inspired reds/blues",
  "Warm coral palette", "Strong primaries",
  "Pastel for dark mode", "Subtle slate tones",
  "High-contrast for slides"
)

swatch <- function(hex) {
  sprintf(
    "<span style=\"display:inline-block;width:14px;height:14px;background:%s;border:1px solid #ddd;border-radius:2px;vertical-align:middle;margin-right:2px\"></span>",
    hex
  )
}

rows <- vapply(seq_along(themes_to_show), function(i) {
  th <- themes_to_show[[i]]
  cols <- th@shapes@effect_colors
  swatches <- paste0(vapply(cols, swatch, character(1)), collapse = "")
  hexes <- paste0("<code style=\"font-size:11px\">", cols, "</code>", collapse = " ")
  sprintf(
    "<tr><td><code>web_theme_%s()</code></td><td>%s<br><span style=\"font-size:11px\">%s</span></td><td>%s</td></tr>",
    names(themes_to_show)[i], swatches, hexes, styles[i]
  )
}, character(1))

cat(
  "<table><thead><tr><th>Theme</th><th>Effect Colors</th><th>Style</th></tr></thead><tbody>",
  paste(rows, collapse = ""),
  "</tbody></table>",
  sep = ""
)
```

### Axis

```r
set_axis(
  range_min = NULL,        # Minimum axis value (NULL = auto)
  range_max = NULL,        # Maximum axis value (NULL = auto)
  tick_count = 5,          # Target number of ticks
  tick_values = NULL,      # Custom tick positions (overrides tick_count)
  gridlines = FALSE,       # Show vertical gridlines
  gridline_style = "solid",# "solid", "dashed", or "dotted"
  symmetric = NULL,        # Force symmetric axis around null (NULL = auto)
  include_null = TRUE,     # Always include null value in range
  null_tick = TRUE         # Always show tick at null value
)
```

::: {.callout-tip}
## When to Use Symmetric Axes
Use `symmetric = TRUE` when comparing benefit vs harm magnitudes. This makes a HR of 0.5 and 2.0 appear equidistant from 1.0 on log scale.
:::

### Layout

```r
set_layout(
  plot_position = "right",    # "left" or "right"
  row_border = TRUE,          # Show row borders
  row_border_style = "solid", # "solid", "dashed", "dotted"
  container_border = FALSE,   # Show border around container
  container_border_radius = 8 # Corner radius in pixels
)
```

### Group Headers

Control the hierarchical styling for nested row groups (h1/h2/h3-style visual hierarchy):

```r
set_group_headers(
  level1_font_size = "0.9375rem",
  level1_font_weight = 600,
  level1_background = NULL,      # NULL = 15% primary opacity
  level1_border_bottom = FALSE,

  level2_font_size = "0.875rem",
  level2_font_weight = 500,
  level2_italic = FALSE,
  level2_background = NULL,      # NULL = 10% primary opacity
  level2_border_bottom = FALSE,

  level3_font_size = "0.875rem",
  level3_font_weight = 400,
  level3_background = NULL,      # NULL = 6% primary opacity
  level3_border_bottom = FALSE,

  indent_per_level = 16          # Indentation per nesting level (px)
)
```

::: {.callout-note}
## Automatic Background Colors
When `levelN_background = NULL`, the theme automatically computes a subtle background from the primary color at reduced opacity. Override with explicit hex colors for full control.
:::

## Creating Custom Themes

For extensive customization, use `web_theme()` to extend a base theme:

```{r}
#| eval: false
publication_theme <- web_theme(
  name = "my_journal",
  colors = list(
    primary = "#1a365d",
    ci_marker_fill = "#1a365d",
    ci_line = "#2d3748"
  ),
  typography = list(
    font_family = "Georgia, serif",
    font_size_base = "0.8rem"
  ),
  spacing = list(
    row_height = 22
  ),
  base_theme = web_theme_minimal()  # Start from minimal
)
```

See [Themes Reference](../reference/themes.qmd) for complete parameter documentation.
