---
title: "Bar Charts"
---

```{r}
#| include: false
library(tabviz)
library(dplyr)
```

`viz_bar()` creates horizontal bar charts for comparing magnitude across rows. Bars share a common axis so the eye can compare values directly.

::: {.callout-note}
## `viz_bar()` vs `col_bar()`

`viz_bar()` is a **focal column** with its own axis — designed for cross-row comparisons. If you only need a simple inline weight indicator inside a cell (no axis, no scale), reach for [`col_bar()`](../columns.qmd#visualizations) instead.
:::

## Basic Usage

```{r}
bar_data <- data.frame(
  metric = c("Response Rate", "Completion Rate", "Retention Rate", "Satisfaction"),
  value = c(78, 92, 85, 88),
  category = c("Primary", "Primary", "Secondary", "Secondary")
)

tabviz(bar_data, label = "metric",
  columns = list(
    col_text("category", "Category"),
    viz_bar(
      effect_bar("value"),
      header = "Rate (%)",
      width = 200,
      axis_range = c(0, 100)
    ),
    col_numeric("value", "%", width = 60)
  ),
  title = "Performance Metrics"
)
```

## Multiple Effects

Wrap several `effect_bar()` items to compare values per row with grouped bars:

```{r}
comparison_data <- data.frame(
  region = c("North America", "Europe", "Asia Pacific", "Latin America"),
  q1_sales = c(245, 189, 312, 98),
  q2_sales = c(278, 215, 345, 125),
  q3_sales = c(312, 242, 398, 156)
)

tabviz(comparison_data, label = "region",
  columns = list(
    viz_bar(
      effect_bar("q1_sales", label = "Q1", color = "#3b82f6"),
      effect_bar("q2_sales", label = "Q2", color = "#22c55e"),
      effect_bar("q3_sales", label = "Q3", color = "#f59e0b"),
      header = "Quarterly Sales ($K)",
      width = 250
    )
  ),
  title = "Regional Sales Comparison"
)
```

## `viz_bar()` Arguments

| Argument | Default | Purpose |
|---|---|---|
| `header` | `NULL` | Column header text |
| `width` | `150` | Column width in pixels |
| `scale` | `"linear"` | `"linear"` or `"log"` |
| `axis_range` | auto | Fixed axis range `c(min, max)` |
| `axis_ticks` | auto | Custom tick positions |
| `axis_gridlines` | `FALSE` | Show vertical gridlines |
| `axis_label` | `"Value"` | Label text below the axis |
| `show_axis` | `TRUE` | Show the axis |

## `effect_bar()` Arguments

| Argument | Purpose |
|---|---|
| `value` | Column name containing the bar value |
| `label` | Display label (used in legends with multiple effects) |
| `color` | Bar fill color |
| `opacity` | Bar opacity 0–1 (default `0.85`) |

## See Also

- [Visualizations Overview](index.qmd) — Tooltips, export, common features
- [Boxplots](boxplots.qmd) — Distribution summaries with `viz_boxplot()`
- [Violin Plots](violin-plots.qmd) — Full distributions with `viz_violin()`
- [Columns: `col_bar()`](../columns.qmd#visualizations) — Inline weight bars without an axis
