---
title: "Bar Charts"
---

```{r}
#| include: false
library(tabviz)
library(dplyr)
```

`viz_bar()` creates horizontal bar charts that share a common axis across rows so the eye can compare magnitudes directly. Use it for counts, scores, percentages, or any single-value-per-row metric.

::: {.callout-note}
## `viz_bar()` vs `col_bar()`

`viz_bar()` is a **focal column** with its own axis — designed for cross-row comparisons. If you only need a quick inline weight indicator inside a cell (no axis, no scale), reach for [`col_bar()`](../columns.qmd#col_bar) instead.
:::

## Quick example

```{r}
bar_data <- data.frame(
  metric = c("Response Rate", "Completion Rate", "Retention Rate", "Satisfaction"),
  value  = c(78, 92, 85, 88),
  category = c("Primary", "Primary", "Secondary", "Secondary")
)

tabviz(bar_data, label = "metric",
  columns = list(
    col_text("category", "Category"),
    viz_bar(
      effect_bar("value"),
      header = "Rate (%)",
      width  = 200,
      axis_range = c(0, 100)
    ),
    col_numeric("value", "%", width = 60)
  ),
  title = "Performance Metrics"
)
```

## Multiple effects (grouped bars)

Pass several `effect_bar()` items to compare values per row with grouped bars:

```{r}
comparison_data <- data.frame(
  region   = c("North America", "Europe", "Asia Pacific", "Latin America"),
  q1_sales = c(245, 189, 312, 98),
  q2_sales = c(278, 215, 345, 125),
  q3_sales = c(312, 242, 398, 156)
)

tabviz(comparison_data, label = "region",
  columns = list(
    viz_bar(
      effect_bar("q1_sales", label = "Q1", color = "#3b82f6"),
      effect_bar("q2_sales", label = "Q2", color = "#22c55e"),
      effect_bar("q3_sales", label = "Q3", color = "#f59e0b"),
      header = "Quarterly Sales ($K)",
      width  = 250
    )
  ),
  title = "Regional Sales Comparison"
)
```

When you don't pass per-effect colors, the theme's `series_anchors` fill them automatically — switching themes restyles every bar. See [Visualization Styling — Layer 1](styling.qmd#layer-1-theme-palette) for the full palette flow.

## `viz_bar()` arguments

| Argument | Default | Purpose |
|---|---|---|
| `...` | required | One or more `effect_bar()` items, optionally followed by named cell-styling args (see below) |
| `header` | `NULL` | Column header. `NULL` resolves to the primary effect's label / field name |
| `show_header` | `TRUE` | Render the header cell above the chart |
| `header_align` | `"center"` | `"left"` / `"center"` / `"right"` |
| `width` | `NULL` (auto) | Column width in pixels |
| `scale` | `"linear"` | `"linear"` or `"log"` (log requires all-positive values) |
| `axis_range` | auto | Numeric `c(min, max)`; auto-computed when `NULL` |
| `axis_ticks` | auto | Numeric vector of explicit tick positions |
| `axis_gridlines` | `FALSE` | Show vertical gridlines at tick positions |
| `axis_label` | `"Value"` | Label rendered below the axis |
| `show_axis` | `TRUE` | Hide the axis with `FALSE` for compact dashboard tiles |
| `null_value` | `NULL` | Optional reference value — paints a dashed line at this position (a synthetic `refline()`) |
| `annotations` | `NULL` | List of [`refline()`](../recipes.qmd#pin-a-custom-reference-line-on-a-forest) items for additional reference lines (`forest_annotation()` is forest-specific and ignored here) |

::: {.callout-note}
## Cell-styling passthrough

Named arguments that follow the `effect_bar()` items are forwarded to `web_col()`:
`bold`, `italic`, `color`, `bg`, `badge`, `icon`, `emphasis`, `muted`, `accent`. Cell-level args that don't make sense for a graphical column (`na_text`, `tooltip`, `formatter`) are silently ignored — rows with no valid data are skipped, not text-filled.
:::

## `effect_bar()` arguments

| Argument | Default | Purpose |
|---|---|---|
| `value` | required | Column name containing the bar value |
| `label` | `NULL` | Display label (used in legends and as the resolved column header for single-effect viz) |
| `color` | `NULL` | Bar fill color (overrides the theme palette for this effect) |
| `opacity` | `NULL` | Bar opacity in `[0, 1]` |

## Axis control

Every axis option lives on `viz_bar()` directly. Set them per column.

### Fixed range

```{r}
tabviz(bar_data, label = "metric",
  columns = list(
    viz_bar(effect_bar("value"),
            header = "Rate (%)",
            axis_range = c(0, 100))
  ),
  title = "Axis pinned to 0–100"
)
```

### Custom ticks and gridlines

```{r}
tabviz(bar_data, label = "metric",
  columns = list(
    viz_bar(effect_bar("value"),
            header = "Rate (%)",
            axis_range = c(0, 100),
            axis_ticks = c(0, 25, 50, 75, 100),
            axis_gridlines = TRUE)
  ),
  title = "Custom ticks + gridlines"
)
```

### Log scale

`scale = "log"` requires every bar value to be positive. Useful for population counts, dollar magnitudes, or any quantity that spans orders of magnitude.

```{r}
log_data <- data.frame(
  city = c("Tokyo", "Delhi", "Shanghai", "São Paulo", "Manila"),
  pop  = c(37e6, 32e6, 28e6, 22e6, 14e6)
)

tabviz(log_data, label = "city",
  columns = list(
    viz_bar(effect_bar("pop"), header = "Population",
            scale = "log",
            axis_label = "Metro population (log scale)")
  )
)
```

### Hide the axis for dense dashboards

```{r}
mini <- bar_data |> head(3)

tabviz(mini, label = "metric",
  columns = list(
    viz_bar(effect_bar("value"),
            header = "%",
            show_axis = FALSE,
            width = 120,
            axis_range = c(0, 100))
  )
)
```

## Reference lines

Use `null_value` for a single dashed line at a baseline / target / threshold. For multiple lines, pass a list of `refline()` items via `annotations`.

```{r}
target_data <- data.frame(
  team = c("Alpha", "Bravo", "Charlie", "Delta"),
  score = c(72, 88, 65, 91)
)

tabviz(target_data, label = "team",
  columns = list(
    viz_bar(effect_bar("score"),
            header = "Score",
            axis_range = c(0, 100),
            null_value = 80)
  ),
  title = "null_value = 80 — target line"
)
```

```{r}
tabviz(target_data, label = "team",
  columns = list(
    viz_bar(effect_bar("score"),
            header = "Score",
            axis_range = c(0, 100),
            annotations = list(
              refline(60, label = "Floor",  style = "dashed", color = "#dc2626"),
              refline(80, label = "Target", style = "solid",  color = "#16a34a"),
              refline(95, label = "Stretch", style = "dotted", color = "#0369a1")
            ))
  ),
  title = "Custom annotation set"
)
```

## Styling cascade

Bar colors resolve through the same four-layer cascade as every other `viz_*()` column — the theme palette is the default, per-effect `color = "..."` pins a series identity, semantic row tokens (`row_accent`, `row_emphasis`, `row_muted`, `row_bold`, `row_fill`) reach into bars on tagged rows, and per-row `marker_color` wins for exact pixel control. See [Visualization Styling](styling.qmd) for the full cascade with rendered examples.

```{r}
tabviz(target_data, label = "team",
  columns = list(
    viz_bar(effect_bar("score"),
            header = "Score",
            axis_range = c(0, 100),
            null_value = 80)
  ),
  row_accent = ~ score >= 80,
  title = "row_accent reaches into bars",
  subtitle = "Teams ≥ 80 take the theme accent fill"
)
```

## When to reach for what

- **`viz_bar()`** when you need a focal axis-backed comparison across rows (counts, scores, percentages).
- **`col_bar()`** ([Columns](../columns.qmd#col_bar)) when you only need an inline weight indicator inside a cell — no shared axis, no scale, fits next to a label.
- **`viz_boxplot()` / `viz_violin()`** when you have a *distribution* per row, not a single value.
- **`viz_forest()`** when you have point + lower + upper (confidence intervals).

## See also

- [Visualization Styling](styling.qmd) — The 4-layer color cascade across all viz types
- [Visualizations Overview](index.qmd) — Tooltips, export, shared features
- [Columns — `col_bar()`](../columns.qmd#col_bar) — Inline cell-level bars
- [Forest Plots](forest-plots.qmd) — Confidence intervals
- [Boxplots](boxplots.qmd) / [Violin Plots](violin-plots.qmd) — Distribution viz
