---
title: "Boxplots"
---

```{r}
#| include: false
library(tabviz)
library(dplyr)
```

`viz_boxplot()` summarizes distributions with median, quartiles, and outliers. It accepts either raw arrays of values or pre-computed summary statistics.

::: {.callout-tip}
## When to Use Boxplots

Best for comparing distribution summaries when you care about median, spread, quartile ranges, and outliers. If you need to see distribution **shape** (bimodality, skewness), use [violin plots](violin-plots.qmd) instead.
:::

::: {.callout-important}
## Array Columns Need `I()`

When passing a list-of-vectors as a column, wrap it with `I()` so `data.frame()` preserves the list structure:

```r
data <- data.frame(
  group = c("A", "B", "C"),
  values = I(list(c(1, 2, 3), c(2, 4, 6), c(3, 5, 7)))
)
```

Without `I()`, R will try to unlist the vectors and fail.
:::

## Mode 1 — From Array Data

When each row contains an array of values, boxplot statistics are computed automatically:

```{r}
set.seed(42)
boxplot_data <- data.frame(
  treatment = c("Control", "Low Dose", "Medium Dose", "High Dose"),
  response = I(list(
    rnorm(50, mean = 100, sd = 15),
    rnorm(50, mean = 110, sd = 18),
    rnorm(50, mean = 125, sd = 20),
    rnorm(50, mean = 140, sd = 25)
  )),
  n = c(50, 50, 50, 50)
)

tabviz(boxplot_data, label = "treatment",
  columns = list(
    col_n("n"),
    viz_boxplot(
      effect_boxplot(data = "response", color = "#8b5cf6"),
      header = "Response Distribution",
      width = 250,
      show_outliers = TRUE
    )
  ),
  title = "Dose-Response Distribution"
)
```

## Mode 2 — From Pre-computed Statistics

If you already have quartile statistics, map them directly:

```{r}
summary_data <- data.frame(
  study = c("Study A", "Study B", "Study C", "Study D"),
  min = c(45, 52, 38, 60),
  q1 = c(62, 68, 55, 72),
  median = c(78, 82, 71, 85),
  q3 = c(89, 94, 86, 95),
  max = c(105, 112, 98, 108)
)

tabviz(summary_data, label = "study",
  columns = list(
    viz_boxplot(
      effect_boxplot(
        min = "min", q1 = "q1", median = "median", q3 = "q3", max = "max",
        color = "#22c55e"
      ),
      header = "Score Distribution",
      width = 220
    ),
    col_numeric("median", "Median")
  ),
  title = "Study Score Distributions"
)
```

## Multiple Effects

Compare distributions across conditions with multiple `effect_boxplot()` items:

```{r}
set.seed(123)
multi_box_data <- data.frame(
  timepoint = c("Baseline", "Week 4", "Week 8", "Week 12"),
  placebo = I(list(
    rnorm(40, 50, 12), rnorm(40, 52, 13), rnorm(40, 51, 14), rnorm(40, 53, 12)
  )),
  treatment = I(list(
    rnorm(40, 50, 12), rnorm(40, 62, 14), rnorm(40, 70, 15), rnorm(40, 75, 13)
  ))
)

tabviz(multi_box_data, label = "timepoint",
  columns = list(
    viz_boxplot(
      effect_boxplot(data = "placebo", label = "Placebo", color = "#94a3b8"),
      effect_boxplot(data = "treatment", label = "Treatment", color = "#3b82f6"),
      header = "Score by Treatment Arm",
      width = 280,
      show_outliers = TRUE
    )
  ),
  title = "Treatment Effect Over Time"
)
```

## `viz_boxplot()` Arguments

| Argument | Default | Purpose |
|---|---|---|
| `header` | `NULL` | Column header text |
| `width` | `150` | Column width in pixels |
| `scale` | `"linear"` | `"linear"` or `"log"` |
| `axis_range` | auto | Fixed axis range `c(min, max)` |
| `axis_ticks` | auto | Custom tick positions |
| `axis_gridlines` | `FALSE` | Show vertical gridlines |
| `show_outliers` | `TRUE` | Display outlier points |
| `whisker_type` | `"iqr"` | `"iqr"` (1.5×IQR) or `"minmax"` |
| `axis_label` | `"Value"` | Label text below the axis |
| `show_axis` | `TRUE` | Show the axis |

## `effect_boxplot()` Arguments

| Argument | Purpose |
|---|---|
| `data` | Column containing array of values (mode 1) |
| `min`, `q1`, `median`, `q3`, `max` | Column names for pre-computed stats (mode 2) |
| `outliers` | Column containing outlier values (optional, mode 2) |
| `label` | Display label for legend |
| `color` | Box fill color |
| `opacity` | Box fill opacity (default `0.7`). Was `fill_opacity` before 0.9.0. |

## See Also

- [Visualization Styling](styling.qmd) — How `row_accent` / `marker_color` cascade into box fills
- [Visualizations Overview](index.qmd) — Tooltips, export, common features
- [Bar Charts](bar-charts.qmd) — Single values with `viz_bar()`
- [Violin Plots](violin-plots.qmd) — Full distributions with `viz_violin()`
