---
title: "Distribution Plots"
---

```{r}
#| include: false
library(tabviz)
library(dplyr)
```

A side-by-side guide to the three distribution-oriented `viz_*()` columns: `viz_bar()`, `viz_boxplot()`, and `viz_violin()`. Use this page to pick the right column for your data; follow the per-type pages for the full surface area.

## Choosing the Right Visualization

| Visualization | Use when… | Data shape |
|---|---|---|
| [`viz_bar()`](bar-charts.qmd) | Comparing single values across rows | One numeric value per row |
| [`viz_boxplot()`](boxplots.qmd) | You need median, quartiles, outliers | Array per row **or** pre-computed quartile stats |
| [`viz_violin()`](violin-plots.qmd) | Distribution shape matters (bimodal, skewed) | Array per row |

## Quick Examples

### Bar Charts

```{r}
bar_data <- data.frame(
  metric = c("Response Rate", "Completion Rate", "Retention Rate"),
  value = c(78, 92, 85)
)

tabviz(bar_data, label = "metric",
  columns = list(
    viz_bar(effect_bar("value"), header = "Rate (%)", axis_range = c(0, 100)),
    col_numeric("value", "%")
  )
)
```

### Boxplots

```{r}
set.seed(42)
box_data <- data.frame(
  group = c("Control", "Treatment"),
  values = I(list(rnorm(50, 100, 15), rnorm(50, 120, 20)))
)

tabviz(box_data, label = "group",
  columns = list(
    viz_boxplot(effect_boxplot(data = "values", color = "#8b5cf6"),
                header = "Distribution")
  )
)
```

### Violin Plots

```{r}
set.seed(456)
violin_data <- data.frame(
  group = c("A", "B"),
  scores = I(list(c(rnorm(80, 72, 8), rnorm(20, 90, 5)), rnorm(100, 68, 12)))
)

tabviz(violin_data, label = "group",
  columns = list(
    viz_violin(effect_violin(data = "scores", color = "#f59e0b"),
               header = "Scores", show_median = TRUE)
  )
)
```

## Combining Visualizations

Mix viz columns for multi-faceted dashboards:

```{r}
set.seed(321)
combined_data <- data.frame(
  site = c("Site A", "Site B", "Site C", "Site D"),
  enrollment = c(245, 189, 312, 156),
  completion_rate = c(0.87, 0.92, 0.78, 0.95),
  response_times = I(list(
    rnorm(50, 5.2, 1.1),
    rnorm(50, 4.8, 0.9),
    rnorm(50, 6.1, 1.5),
    rnorm(50, 4.5, 0.8)
  ))
)

tabviz(combined_data, label = "site",
  columns = list(
    viz_bar(
      effect_bar("enrollment", color = "#3b82f6"),
      header = "Enrollment",
      width = 140,
      axis_range = c(0, 350)
    ),
    viz_bar(
      effect_bar("completion_rate", color = "#22c55e"),
      header = "Completion",
      width = 120,
      axis_range = c(0, 1)
    ),
    viz_violin(
      effect_violin(data = "response_times", color = "#f59e0b"),
      header = "Response Time (days)",
      width = 180,
      show_median = TRUE
    )
  ),
  title = "Clinical Site Performance Dashboard"
)
```

## Common Options

All `viz_*()` columns share these arguments. See the per-type pages for the full set.

| Option | Default | Purpose |
|---|---|---|
| `header` | `NULL` | Column header text |
| `width` | `150` | Column width in pixels |
| `scale` | `"linear"` | `"linear"` or `"log"` |
| `axis_range` | auto | Fixed axis range `c(min, max)` |
| `axis_ticks` | auto | Custom tick positions |
| `axis_gridlines` | `FALSE` | Show vertical gridlines |
| `axis_label` | `"Value"` | Label text below the axis |

## See Also

- [Bar Charts](bar-charts.qmd) — Full `viz_bar()` documentation
- [Boxplots](boxplots.qmd) — Full `viz_boxplot()` documentation
- [Violin Plots](violin-plots.qmd) — Full `viz_violin()` documentation
- [Forest Plots](forest-plots.qmd) — Confidence intervals with `viz_forest()`
