---
title: "Forest Plots"
---

```{r}
#| include: false
library(tabviz)
library(dplyr)
library(tidyr)
```

Forest plots display point estimates with confidence intervals — the standard visualization for meta-analyses and clinical trial results. Build them with the `viz_forest()` column inside `tabviz()`, or use the `forest_plot()` shortcut for quick one-liners.

::: {.callout-tip}
## Two Ways to Add a Forest Column

```r
# tabviz() + viz_forest() (full control)
tabviz(data, label = "study", columns = list(
  viz_forest(point = "hr", lower = "lo", upper = "hi",
             scale = "log", null_value = 1)
))

# forest_plot() (quick shortcut)
forest_plot(data, point = "hr", lower = "lo", upper = "hi",
            label = "study", scale = "log", null_value = 1)
```
:::

## Data Mapping

`viz_forest()` uses a column-mapping pattern: name the columns that hold your data, not the values themselves.

```{r}
data(effect_sizes)

effect_sizes |>
  head(4) |>
  tabviz(label = "study", columns = list(
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1)
  ))
```

### Required Mappings

| Argument | What it is | Example |
|---|---|---|
| `point` | Point estimate column | Hazard ratio, odds ratio, mean difference |
| `lower` | Lower CI bound column | 95% CI lower |
| `upper` | Upper CI bound column | 95% CI upper |

### Handling Missing Values

tabviz uses `NA` strategically for structured layouts:

- **Header rows**: `row_type = "header"` with `NA` for effect estimates
- **Spacer rows**: `row_type = "spacer"` for visual separation
- **Missing estimates**: data rows where the effect couldn't be calculated

```{r}
data(glp1_trials)

glp1_trials |>
  filter(group == "Main Trials") |>
  head(5) |>
  tabviz(label = "study", row_type = "row_type", row_bold = "row_bold",
    columns = list(
      viz_forest(point = "hr", lower = "lower", upper = "upper",
                 scale = "log", null_value = 1)
    )
  )
```

## `viz_forest()` Arguments

| Argument | Default | Purpose |
|---|---|---|
| `point` / `lower` / `upper` | required (single-effect mode) | Column names for the point estimate and CI bounds |
| `effects` | `NULL` | List of `effect_forest()` items for multi-effect mode |
| `header` | `NULL` | Column header text |
| `width` | auto | Column width in pixels |
| `scale` | `"linear"` | `"linear"` or `"log"` |
| `null_value` | `NULL` | Position of the dashed reference line (e.g. `1` for ratios, `0` for differences) |
| `axis_range` | auto | `c(min, max)` overriding the auto-calculated range |
| `axis_ticks` | auto | Custom tick positions |
| `axis_gridlines` | `FALSE` | Show vertical gridlines at tick positions |
| `axis_label` | `NULL` | X-axis label |
| `annotations` | `NULL` | List of `refline()` / `forest_annotation()` items |
| `shared_axis` | `FALSE` | Share axis range with sibling forest columns |
| `show_axis` | `TRUE` | Show the axis |

## Scale and Null Value

### Log Scale

Use for **ratios** (hazard ratio, odds ratio, risk ratio) — effects are multiplicative around `null_value = 1`:

```{r}
effect_sizes |>
  filter(outcome == "Primary") |>
  head(4) |>
  tabviz(label = "study", columns = list(
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1,
               axis_label = "Hazard Ratio (95% CI)")
  ))
```

::: {.callout-warning}
## Log Scale Requires Positive Values

All values in `point`, `lower`, and `upper` must be **positive**. Zero or negative values will error.
:::

### Linear Scale

Use for **differences** (mean difference, risk difference, SMD) — effects are additive around `null_value = 0`:

```{r}
diff_data <- data.frame(
  comparison = c("Treatment A", "Treatment B", "Treatment C"),
  mean_diff = c(-2.5, 1.3, -0.8),
  lower = c(-4.1, -0.2, -2.1),
  upper = c(-0.9, 2.8, 0.5)
)

tabviz(diff_data, label = "comparison", columns = list(
  viz_forest(point = "mean_diff", lower = "lower", upper = "upper",
             scale = "linear", null_value = 0,
             axis_label = "Mean Difference (95% CI)")
))
```

## Multiple Effects

Display multiple effect estimates per row using `effects = list(effect_forest(...), ...)`. Useful for comparing analyses or sensitivity scenarios:

```{r}
multi_data <- data.frame(
  study = c("PIONEER", "SUMMIT", "HORIZON"),
  n = c(2450, 1890, 3200),
  itt_or = c(0.72, 0.78, 0.65),
  itt_lo = c(0.58, 0.64, 0.52),
  itt_hi = c(0.89, 0.95, 0.81),
  pp_or = c(0.68, 0.73, 0.61),
  pp_lo = c(0.53, 0.58, 0.47),
  pp_hi = c(0.87, 0.92, 0.79)
)

tabviz(multi_data, label = "study", title = "Sensitivity Analysis Comparison",
  columns = list(
    col_n("n"),
    col_interval("itt_or", "itt_lo", "itt_hi", header = "Primary OR"),
    viz_forest(
      scale = "log", null_value = 1,
      effects = list(
        effect_forest("itt_or", "itt_lo", "itt_hi", label = "ITT", color = "#2563eb"),
        effect_forest("pp_or", "pp_lo", "pp_hi", label = "Per-Protocol", color = "#16a34a")
      )
    )
  )
)
```

### `effect_forest()` Arguments

| Argument | Purpose |
|---|---|
| `point` | Column name for point estimates |
| `lower` | Column name for lower bounds |
| `upper` | Column name for upper bounds |
| `label` | Display label for legend |
| `color` | Override color for this effect |
| `shape` | Override shape: `"square"`, `"circle"`, `"diamond"`, `"triangle"` |

### Side-by-Side Forest Columns

When comparing entirely different outcomes, separate `viz_forest()` columns are often clearer than overlaid effects:

```{r}
comparison_data <- data.frame(
  study = c("Cardiovascular", "Mortality", "Hospitalization", "Composite"),
  na_hr = c(0.72, 0.85, 0.78, 0.76),
  na_lo = c(0.58, 0.71, 0.64, 0.65),
  na_hi = c(0.89, 1.02, 0.95, 0.89),
  eu_hr = c(0.68, 0.79, 0.82, 0.74),
  eu_lo = c(0.52, 0.62, 0.67, 0.61),
  eu_hi = c(0.89, 1.01, 1.00, 0.90)
)

tabviz(comparison_data, label = "study",
  title = "Regional Comparison: North America vs Europe",
  columns = list(
    viz_forest(point = "na_hr", lower = "na_lo", upper = "na_hi",
               header = "North America", scale = "log", null_value = 1, width = 180),
    viz_forest(point = "eu_hr", lower = "eu_lo", upper = "eu_hi",
               header = "Europe", scale = "log", null_value = 1, width = 180)
  )
)
```

::: {.callout-tip}
## Side-by-Side vs Multi-Effect

- **Side-by-side `viz_forest()` columns** — comparing different outcomes or populations on potentially different scales.
- **Multi-effect overlays via `effect_forest()`** — comparing the same outcome across analyses (sensitivity, ITT vs PP) on a single shared scale.
:::

### `shared_axis`

Set `shared_axis = TRUE` on multiple `viz_forest()` columns to lock them to a common axis range so the eye can compare across columns directly.

## Marker Customization

Customize marker appearance based on data values using formula expressions on `tabviz()`:

```{r}
effect_sizes |>
  filter(outcome == "Primary") |>
  head(6) |>
  tabviz(label = "study", title = "Marker Styling with Formulas",
    marker_shape = ~ ifelse(phase == "Phase 3", "circle", "square"),
    marker_color = ~ ifelse(significant, "#16a34a", "#94a3b8"),
    marker_opacity = ~ ifelse(n > 200, 1, 0.7),
    columns = list(
      col_text("phase", "Phase"),
      viz_forest(point = "hr", lower = "lower", upper = "upper",
                 scale = "log", null_value = 1)
    )
  )
```

| Argument | Type | Effect |
|---|---|---|
| `marker_color` | string / formula | CSS color for marker fill |
| `marker_shape` | string / formula | `"square"`, `"circle"`, `"diamond"`, `"triangle"` |
| `marker_opacity` | numeric / formula | Transparency (0–1) |
| `marker_size` | numeric / formula / `"col"` | Size multiplier (1 = default), or column name for proportional sizing |

All marker arguments accept either a column name or a formula expression (`~ ...`). Pass `marker_size = "col_name"` to scale markers by a column (e.g. study weight in a meta-analysis).

## Axis Control

Axis arguments live on `viz_forest()` itself — they're per-column.

```{r}
#| include: false
axis_demo <- effect_sizes |>
  filter(outcome == "Primary") |>
  head(4)
```

### Axis Range

```{r}
tabviz(axis_demo, label = "study", columns = list(
  viz_forest(point = "hr", lower = "lower", upper = "upper",
             scale = "log", null_value = 1, axis_range = c(0.3, 2.0))
))
```

### Custom Tick Values

```{r}
tabviz(axis_demo, label = "study", columns = list(
  viz_forest(point = "hr", lower = "lower", upper = "upper",
             scale = "log", null_value = 1,
             axis_range = c(0.3, 2.0),
             axis_ticks = c(0.5, 0.7, 1.0, 1.5, 2.0))
))
```

### Gridlines

```{r}
tabviz(axis_demo, label = "study", columns = list(
  viz_forest(point = "hr", lower = "lower", upper = "upper",
             scale = "log", null_value = 1, axis_gridlines = TRUE)
))
```

## Reference Lines and Annotations

`null_value` draws a single dashed reference line. For more, pass `annotations = list(...)` of `refline()` / `forest_annotation()` items.

```{r}
tabviz(axis_demo, label = "study", title = "With Custom Reference Line",
  columns = list(
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1,
               annotations = list(
                 refline(0.8, label = "Threshold", style = "dashed", color = "#e11d48")
               ))
  )
)
```

```{r}
tabviz(axis_demo, label = "study", title = "Clinical Thresholds",
  columns = list(
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1,
               annotations = list(
                 refline(0.5, label = "Strong", style = "solid", color = "#16a34a"),
                 refline(0.8, label = "Moderate", style = "dashed", color = "#f59e0b")
               ))
  )
)
```

### `refline()` Arguments

| Argument | Purpose |
|---|---|
| `value` | X-axis position |
| `label` | Optional label text |
| `style` | `"solid"`, `"dashed"`, `"dotted"` |
| `color` | Line color |
| `width` | Line width (default `1.5`) |
| `opacity` | Line opacity 0–1 (default `0.6`) |

`forest_annotation()` adds richer annotations (regions, text callouts) — see the reference for the full surface.

## Working with Meta-Analysis Results

```{r}
#| eval: false
library(metafor)

res <- rma(yi = log_or, sei = se, data = studies, method = "REML")

forest_data <- studies |>
  mutate(
    or = exp(log_or),
    lower = exp(log_or - 1.96 * se),
    upper = exp(log_or + 1.96 * se)
  ) |>
  bind_rows(
    tibble(
      study = "Pooled Estimate",
      or = exp(res$b),
      lower = exp(res$ci.lb),
      upper = exp(res$ci.ub),
      row_type = "summary",
      row_bold = TRUE
    )
  )

tabviz(forest_data,
  label = "study", row_type = "row_type", row_bold = "row_bold",
  columns = list(
    viz_forest(point = "or", lower = "lower", upper = "upper",
               scale = "log", null_value = 1)
  )
)
```

## See Also

- [Visualizations Overview](index.qmd) — Tooltips, export, common features
- [Titles & Labels](../titles.qmd) — Adding context to your plots
- [Row Styling](../row-styling.qmd) — Headers, summaries, indentation
- [Themes](../themes.qmd) — Visual styling
