---
title: "Visualizations"
---

```{r}
#| include: false
library(tabviz)
library(dplyr)
```

tabviz embeds visualizations directly inside table columns. Four `viz_*()` column types cover the common patterns; each gets its own focused page.

## The Four `viz_*` Columns

| Column | Use this when… | Guide |
|---|---|---|
| `viz_forest()` | You have point estimates with confidence intervals (meta-analyses, treatment effects) | [Forest Plots](forest-plots.qmd) |
| `viz_bar()` | You're comparing single values across rows (counts, scores, percentages) | [Bar Charts](bar-charts.qmd) |
| `viz_boxplot()` | You care about median, spread, and outliers across distributions | [Boxplots](boxplots.qmd) |
| `viz_violin()` | You need to see full distribution shapes (bimodality, skewness) | [Violin Plots](violin-plots.qmd) |

::: {.callout-note}
## `viz_*` vs `col_*`

`viz_*()` columns are **focal visualizations** with their own axes and scales — designed for direct visual comparison across rows. The simpler `col_*()` family (`col_bar()`, `col_sparkline()`) is for **inline cell content** that fits in a cell without a dedicated axis.

If you want a quick weight indicator inside a row, reach for `col_bar()`. If you want a true comparable bar chart, reach for `viz_bar()`.
:::

## Common Features

All `viz_*()` columns share:

- **Axis configuration** — `axis_range`, `axis_ticks`, `axis_gridlines`, `axis_label`
- **Pan and zoom** — mouse wheel, drag, double-click reset (per column)
- **Theming** — colors adapt to the active theme; multi-effect series use the theme's `effect_colors` palette
- **Marker-styling cascade** — row-level `row_accent` / `row_emphasis` / `row_muted` and per-row `marker_color` reach into viz cells. Single-effect viz: fill replacement. Multi-effect viz: outline augment. See [Visualization Styling](styling.qmd) for the full 4-layer cascade with rendered examples across viz types.
- **Annotations** — `refline()` for reference lines (and `forest_annotation()` on `viz_forest`). Use `null_value` as a shortcut for a single themed reference line.
- **Tooltips** — opt in via `web_interaction(tooltip_fields = ...)`
- **Export** — SVG / PNG / PDF via `save_plot()`

::: {.callout-tip}
## Theme Effect Colors

When you have multiple effects per row (multiple bar series, boxplot groups, or forest effects), colors come from the theme's `series_anchors` input. Each anchor expands into a 7-field slot bundle (fill, stroke, fill_muted, etc.) at resolve time:

```r
web_theme_cochrane() |>
  set_inputs(series_anchors = c("#2563eb", "#dc2626", "#16a34a"))
```

Individual effects can still override the theme via their own `color` parameter.
:::

## Tooltips

Tooltips are opt-in. Specify the column names you want to surface on marker hover:

```{r}
data(glp1_trials)

glp1_trials |>
  filter(row_type == "data", group == "Main Trials") |>
  head(4) |>
  tabviz(label = "study",
    interaction = web_interaction(tooltip_fields = c("n", "events", "pvalue")),
    footnote = "Hover over markers to see tooltip",
    columns = list(
      viz_forest(point = "hr", lower = "lower", upper = "upper",
                 scale = "log", null_value = 1)
    )
  )
```

See [Interactivity](../interactivity.qmd) for the full set of `web_interaction()` controls.

## Exporting Visualizations

```r
p <- tabviz(data, label = "study", columns = list(
  viz_forest(point = "hr", lower = "lower", upper = "upper",
             scale = "log", null_value = 1)
))

save_plot(p, "figure.svg")                           # vector, no deps
save_plot(p, "figure.png", width = 1200, scale = 2)  # raster, needs rsvg
save_plot(p, "figure.pdf", width = 800)              # vector, needs rsvg
```

See [Export](../export.qmd) for details.

## See Also

- [Forest Plots](forest-plots.qmd) — Confidence intervals with `viz_forest()`
- [Bar Charts](bar-charts.qmd) — Magnitude comparisons with `viz_bar()`
- [Boxplots](boxplots.qmd) — Distribution summaries with `viz_boxplot()`
- [Violin Plots](violin-plots.qmd) — Full distributions with `viz_violin()`
- [Visualization Styling](styling.qmd) — The 4-layer color cascade across all viz types
- [Distribution Plots](distribution-plots.qmd) — Choosing among bar / box / violin
- [Columns](../columns.qmd) — Inline columns including `col_bar()` and `col_sparkline()`
