---
title: "Violin Plots"
---

```{r}
#| include: false
library(tabviz)
library(dplyr)
```

`viz_violin()` shows the full distribution shape using kernel density estimation. Use it when distribution shape matters — bimodality, skewness, asymmetry — rather than just summary statistics.

::: {.callout-tip}
## When to Use Violin Plots

Best for revealing distribution shapes (bimodal, multimodal, skewed). If you only need to compare medians, quartiles, and outliers, use [boxplots](boxplots.qmd).
:::

::: {.callout-important}
## Array Columns Need `I()`

`viz_violin()` always works from raw arrays of values. Wrap list columns with `I()` so `data.frame()` preserves their structure:

```r
data <- data.frame(
  group = c("A", "B", "C"),
  values = I(list(
    rnorm(100, mean = 5, sd = 2),
    rnorm(100, mean = 8, sd = 3),
    c(rnorm(50, 4, 1), rnorm(50, 10, 1))  # bimodal
  ))
)
```

Without `I()`, R will try to unlist the vectors and fail.
:::

## Basic Usage

```{r}
set.seed(456)
violin_data <- data.frame(
  group = c("Young Adults", "Middle Age", "Seniors"),
  scores = I(list(
    c(rnorm(80, 72, 8), rnorm(20, 90, 5)),     # bimodal
    rnorm(100, 68, 12),                          # normal
    c(rnorm(70, 60, 10), runif(30, 40, 80))     # skewed
  )),
  n = c(100, 100, 100)
)

tabviz(violin_data, label = "group",
  columns = list(
    col_n("n"),
    viz_violin(
      effect_violin(data = "scores", color = "#f59e0b"),
      header = "Score Distribution",
      width = 250,
      show_median = TRUE,
      show_quartiles = TRUE
    )
  ),
  title = "Score Distributions by Age Group"
)
```

## Comparing Distributions

Use multiple `effect_violin()` items to overlay distributions:

```{r}
set.seed(789)
compare_violin <- data.frame(
  condition = c("Healthy", "Mild", "Moderate", "Severe"),
  biomarker_a = I(list(
    rnorm(60, 25, 5),
    rnorm(60, 35, 8),
    rnorm(60, 50, 12),
    rnorm(60, 70, 15)
  )),
  biomarker_b = I(list(
    rnorm(60, 30, 6),
    rnorm(60, 40, 7),
    rnorm(60, 55, 10),
    rnorm(60, 65, 18)
  ))
)

tabviz(compare_violin, label = "condition",
  columns = list(
    viz_violin(
      effect_violin(data = "biomarker_a", label = "Biomarker A", color = "#3b82f6"),
      effect_violin(data = "biomarker_b", label = "Biomarker B", color = "#ef4444"),
      header = "Biomarker Distributions",
      width = 280,
      show_median = TRUE
    )
  ),
  title = "Biomarker Profiles by Disease Severity"
)
```

## `viz_violin()` Arguments

| Argument | Default | Purpose |
|---|---|---|
| `header` | `NULL` | Column header text |
| `width` | `150` | Column width in pixels |
| `scale` | `"linear"` | `"linear"` or `"log"` |
| `axis_range` | auto | Fixed axis range `c(min, max)` |
| `axis_ticks` | auto | Custom tick positions |
| `axis_gridlines` | `FALSE` | Show vertical gridlines |
| `bandwidth` | `NULL` | KDE bandwidth (`NULL` uses Silverman's rule of thumb) |
| `show_median` | `TRUE` | Show the median line |
| `show_quartiles` | `FALSE` | Show Q1/Q3 dashed lines |
| `axis_label` | `"Value"` | Label text below the axis |
| `show_axis` | `TRUE` | Show the axis |

## `effect_violin()` Arguments

| Argument | Purpose |
|---|---|
| `data` | Column containing array of values (required) |
| `label` | Display label for legend |
| `color` | Violin fill color |
| `opacity` | Fill opacity (default `0.5`). Was `fill_opacity` before 0.9.0. |

## See Also

- [Visualization Styling](styling.qmd) — How `row_accent` / `marker_color` cascade into violin fills
- [Visualizations Overview](index.qmd) — Tooltips, export, common features
- [Bar Charts](bar-charts.qmd) — Single values with `viz_bar()`
- [Boxplots](boxplots.qmd) — Distribution summaries with `viz_boxplot()`
