---
title: "Styling Overview"
---

```{r}
#| include: false
library(tabviz)
```

tabviz styling is a four-layer cascade вҖ” each layer targets a more specific scope and overrides the layer above. This page is the map; follow the links for the details.

## The Cascade

```
в”Ңв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”җ
в”Ӯ  Theme        вҖ” base colors, fonts, spacing         в”Ӯ
в”Ӯ  в”Ңв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”җ  в”Ӯ
в”Ӯ  в”Ӯ  Row         вҖ” bold, color, background per row в”Ӯ  в”Ӯ
в”Ӯ  в”Ӯ  в”Ңв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”җ  в”Ӯ  в”Ӯ
в”Ӯ  в”Ӯ  в”Ӯ  Cell      вҖ” per-column formatting       в”Ӯ  в”Ӯ  в”Ӯ
в”Ӯ  в”Ӯ  в”Ӯ  в”Ңв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”җ  в”Ӯ  в”Ӯ  в”Ӯ
в”Ӯ  в”Ӯ  в”Ӯ  в”Ӯ  Marker  вҖ” forest-plot points      в”Ӯ  в”Ӯ  в”Ӯ  в”Ӯ
в”Ӯ  в”Ӯ  в”Ӯ  в””в”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”ҳ  в”Ӯ  в”Ӯ  в”Ӯ
в”Ӯ  в”Ӯ  в””в”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”ҳ  в”Ӯ  в”Ӯ
в”Ӯ  в””в”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”ҳ  в”Ӯ
в””в”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”ҳ
```

| Layer | Where you set it | Use it for |
|---|---|---|
| **Theme** | `theme = web_theme_*()` on `tabviz()` | Overall look and feel |
| **Row** | `row_bold`, `row_color`, `row_bg`, вҖҰ on `tabviz()` | Whole-row conditional formatting |
| **Cell** | `bold`, `color`, `bg`, вҖҰ inside `col_*()` | Per-column overrides |
| **Marker** | `marker_color`, `marker_shape`, `marker_size`, `marker_opacity` on `tabviz()` | Forest-plot point appearance |

More specific layers win: marker beats cell, cell beats row, row beats theme. Formula expressions and column-name strings are equivalent at every layer вҖ” pick whichever reads better at the call site.

::: {.callout-tip}
## Semantic shortcuts

Theme-aware semantic tokens вҖ” `row_emphasis`, `row_muted`, `row_accent`, `row_bold`, `row_fill` (and the matching `emphasis` / `muted` / `accent` / `bold` / `fill` cell-level args) вҖ” let you say "this row/cell is special" without picking literal colors. They restyle automatically when you swap themes. See [Themes вҖ” Semantic tokens](../themes.qmd#semantic-tokens) for the full list.
:::

## 1. Theme Level

Themes set the foundational palette, typography, spacing, and shapes.

```{r}
data(glp1_trials)
sample_data <- head(glp1_trials[glp1_trials$row_type == "data", ], 4)

tabviz(sample_data, label = "study",
  columns = list(
    col_text("drug", "Drug"),
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1, width = 160),
    col_interval("hr", "lower", "upper", header = "HR (95% CI)")
  ),
  theme = web_theme_nejm()
)
```

## 2. Row Level

Row styling applies formatting to entire rows based on data conditions.

```{r}
tabviz(sample_data, label = "study",
  columns = list(
    col_text("drug", "Drug"),
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1, width = 160),
    col_interval("hr", "lower", "upper", header = "HR (95% CI)")
  ),
  row_bold = ~ upper < 1,
  theme = web_theme_nejm()
)
```

## 3. Cell Level

Cell styling targets individual columns. Use `.x` inside formulas to refer to the column's own value.

```{r}
tabviz(sample_data, label = "study",
  columns = list(
    col_text("drug", "Drug"),
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1, width = 160),
    col_numeric("hr", "HR", decimals = 2,
                bold = ~ .x < 1,
                color = ~ ifelse(.x < 1, "#16a34a", "#64748b"))
  ),
  theme = web_theme_nejm()
)
```

## 4. Marker Level

Marker styling controls forest-plot point appearance: color, shape, size, opacity.

```{r}
tabviz(sample_data, label = "study",
  columns = list(
    col_text("drug", "Drug"),
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1, width = 160),
    col_interval("hr", "lower", "upper", header = "HR (95% CI)")
  ),
  marker_color = ~ ifelse(upper < 1, "#16a34a", "#64748b"),
  marker_shape = ~ ifelse(upper < 1, "diamond", "square"),
  theme = web_theme_nejm()
)
```

## Formula Expressions

tabviz uses R's formula syntax (`~ expression`) for conditional styling.

- **Row and marker formulas** can reference any column in the data:
  ```r
  row_bold = ~ pvalue < 0.05
  marker_color = ~ ifelse(group == "Treatment", "blue", "gray")
  ```
- **Cell formulas** use `.x` for the current column's value:
  ```r
  col_numeric("hr", "HR", bold = ~ .x < 1)
  col_pvalue("pvalue", "P", color = ~ ifelse(.x < 0.05, "red", "black"))
  ```

## Next Steps

- [Row Styling](../row-styling.qmd) вҖ” Row-level conditional formatting
- [Cell Styling](../cell-styling.qmd) вҖ” Per-column formatting
- [Themes](../themes.qmd) вҖ” Built-in themes and customization
- [Fluent API](../fluent-api.qmd) вҖ” Pipe-based modifiers for complex cases
