---
title: "Bilingual Dashboard (R + TS)"
description: "A KPI dashboard with bars, sparklines, badges, and rings — rendered live from both R and TypeScript"
---

```{r}
#| include: false
library(tabviz)
```

The [Genetic Association example](genetic-association.qmd) showcased the bilingual API on a forest plot. This page proves the same R ↔ TypeScript parity for the *visual* column family: `col_bar` / `col_sparkline` / `col_badge` / `col_ring` / `col_stars`. The R chunk renders the live widget; the OJS chunk loads `@tabviz/core` from esm.sh and mounts the equivalent TS-built spec — both shown on the same page.

## Data: regional sales KPIs

```{r}
#| cache: false
set.seed(2026)

regions <- c("North", "South", "East", "West", "Central")
revenue <- c(2400000, 1900000, 3200000, 1650000, 2750000)
growth  <- c(0.12, -0.04, 0.18, 0.03, 0.09)

# Per-region 12-month revenue sparkline (slight growth ± monthly noise).
sales_trend <- lapply(seq_along(regions), function(i) {
  base <- revenue[i] / 12
  trend <- seq(0, growth[i], length.out = 12)
  jitter <- rnorm(12, sd = 0.08)
  round(base * (1 + trend + jitter))
})

kpi <- data.frame(
  region        = regions,
  revenue       = revenue,
  growth        = growth,
  health        = c("green", "amber", "green", "red", "green"),
  satisfaction  = c(4.5, 3.2, 4.8, 2.9, 4.1),
  pipeline_pct  = c(0.78, 0.45, 0.92, 0.38, 0.71)
)
kpi$trend <- I(sales_trend)

# Expose to OJS (plus the locked Svelte version + @tabviz/core version,
# same pattern as the genetic-association page).
svelte_version <- jsonlite::fromJSON(
  "../../srcjs/package-lock.json"
)$packages[["node_modules/svelte"]]$version
tabviz_core_version <- jsonlite::fromJSON(
  "../../srcjs/package.json"
)$version

ojs_define(kpi_in = kpi,
           svelte_version = svelte_version,
           tabviz_core_version = tabviz_core_version)
```

## R: `tabviz(...)` rendering the live widget

```{r}
#| fig-width: 11
#| fig-height: 4
tabviz(
  kpi,
  label = "region",
  label_header = "Region",
  theme = web_theme_cochrane(),
  columns = list(
    col_currency("revenue", header = "Revenue", symbol = "$", decimals = 0,
                 thousands_sep = ",", abbreviate = TRUE),
    col_bar("growth", header = "YoY growth", scale = "linear"),
    col_sparkline("trend", header = "12-month trend", type = "area"),
    col_badge("health", header = "Health"),
    col_stars("satisfaction", header = "CSAT", max_stars = 5),
    col_ring("pipeline_pct", header = "Pipeline", min_value = 0, max_value = 1)
  ),
  title = "Regional sales KPIs",
  caption = "Quarter-end snapshot. Revenue abbreviated; sparkline shows last 12 months."
)
```

## TypeScript via `@tabviz/core`, rendered in the browser

```{ojs}
//| echo: false
core = import(`https://esm.sh/@tabviz/core@${tabviz_core_version}?deps=svelte@${svelte_version}`)

// Transpose the column-oriented kpi data + ensure trend is a per-row array.
kpi_rows = {
  const cols = transpose(kpi_in);
  // `kpi.trend` was set R-side via `kpi$trend <- I(sales_trend)` — a
  // list-column with one 12-element numeric vector per row. Quarto's
  // `ojs_define()` + jsonlite serializes that as a nested array, so
  // `kpi_in.trend[i]` is already the per-row series; no flat slicing
  // needed.
  return cols.map((row, i) => ({
    ...row,
    trend: kpi_in.trend[i],
  }));
}

styleHref = `https://esm.sh/@tabviz/core@${tabviz_core_version}/dist/style.css`
_styleInjected = {
  if (!document.querySelector(`link[href="${styleHref}"]`)) {
    const link = document.createElement("link");
    link.rel = "stylesheet";
    link.href = styleHref;
    document.head.appendChild(link);
  }
  return true;
}

tsWidget = {
  const { tabviz, createTabviz,
          colCurrency, colBar, colSparkline, colBadge, colStars, colRing } = await core;
  _styleInjected;

  const spec = tabviz({
    data: kpi_rows,
    label: "region",
    labelHeader: "Region",
    theme: "cochrane",
    columns: [
      colCurrency({ field: "revenue", header: "Revenue", symbol: "$",
                    decimals: 0, thousandsSep: ",", abbreviate: true }),
      colBar({ field: "growth", header: "YoY growth", scale: "linear" }),
      colSparkline({ field: "trend", header: "12-month trend", type: "area" }),
      colBadge({ field: "health", header: "Health" }),
      colStars({ field: "satisfaction", header: "CSAT", maxStars: 5 }),
      colRing({ field: "pipeline_pct", header: "Pipeline", minValue: 0, maxValue: 1 }),
    ],
    title: "Regional sales KPIs",
    caption: "Quarter-end snapshot. Revenue abbreviated; sparkline shows last 12 months.",
  });

  const host = document.createElement("div");
  host.style.minHeight = "240px";
  createTabviz(host, spec);
  return host;
}
```

::: {.callout-note collapse="true"}
## The pure-TypeScript snippet (for a Vite / Next.js project)

```ts
import {
  tabviz, createTabviz,
  colCurrency, colBar, colSparkline, colBadge, colStars, colRing,
} from "@tabviz/core";
import "@tabviz/core/style.css";

declare const kpiData: Array<{
  region: string;
  revenue: number;
  growth: number;
  trend: number[];          // 12-month sparkline series
  health: "green" | "amber" | "red";
  satisfaction: number;     // 0-5 stars
  pipeline_pct: number;     // 0-1
}>;

const spec = tabviz({
  data: kpiData,
  label: "region",
  labelHeader: "Region",
  theme: "cochrane",
  columns: [
    colCurrency({ field: "revenue", header: "Revenue", symbol: "$",
                  decimals: 0, thousandsSep: ",", abbreviate: true }),
    colBar({ field: "growth", header: "YoY growth", scale: "linear" }),
    colSparkline({ field: "trend", header: "12-month trend", type: "area" }),
    colBadge({ field: "health", header: "Health" }),
    colStars({ field: "satisfaction", header: "CSAT", maxStars: 5 }),
    colRing({ field: "pipeline_pct", header: "Pipeline",
              minValue: 0, maxValue: 1 }),
  ],
  title: "Regional sales KPIs",
});

createTabviz(document.querySelector("#plot")!, spec);
```
:::

## What's the same, what's different?

Same as the [genetic-association page](genetic-association.qmd) — argument names + defaults map snake_case ↔ camelCase, positional R args become named TS args, and `theme = web_theme_cochrane()` becomes `theme: "cochrane"`. The visual column family (bars, sparklines, badges, rings, stars) has the same mechanical translation as the forest-plot family.
