---
title: "Compose Your Own Brand"
description: "Building a branded theme from scratch — neutrals, identity, accent, fonts, variants, pins"
---

```{r}
#| include: false
library(tabviz)
library(dplyr)
```

The V4 theme system is built so authoring a branded theme is a sequence of **single-input edits** — each one obviously about one thing — rather than a big monolithic `web_theme_acme()` wrapper to bisect when something goes wrong. This page walks the build end-to-end for a fictional brand, *Acme Cardio Research*, with the rendered table at every step so you can see exactly which input controls what.

Each step is one named input on `web_theme()` (or one `set_*()` call). At the end, we wrap the result as a one-shot constructor.

## The shared dataset

Every step renders the same eight-row meta-analysis so the only thing changing across the renders is the theme:

```{r}
acme_data <- data.frame(
  study  = c("ACME-101", "PROVE-IT", "CADENCE", "VANGUARD-2",
             "STELLAR", "ENDEAVOUR", "SUMMIT", "Pooled"),
  cohort = c("Phase III", "Phase III", "Phase III",
             "Real-world", "Real-world", "Real-world",
             "Registry",   "Pooled"),
  n      = c(2400, 1850, 3120, 980,  4200, 2180, 1640,  16370),
  hr     = c(0.78, 0.82, 0.74, 0.85, 0.79, 0.81, 0.88,  0.79),
  lower  = c(0.65, 0.71, 0.62, 0.69, 0.69, 0.70, 0.74,  0.74),
  upper  = c(0.93, 0.95, 0.88, 1.05, 0.91, 0.94, 1.05,  0.85),
  pvalue = c(0.005, 0.008, 0.0001, 0.13, 0.0002, 0.005, 0.18, NA),
  row_type = c(rep("data", 7), "summary"),
  is_pooled = c(rep(FALSE, 7), TRUE)
)

render_acme <- function(theme, title = NULL, subtitle = NULL) {
  tabviz(acme_data,
    label = "study",
    group = "cohort",
    columns = list(
      col_n("n"),
      col_pvalue("pvalue", "P", stars = TRUE),
      viz_forest(point = "hr", lower = "lower", upper = "upper",
                 scale = "log", null_value = 1,
                 axis_label = "Hazard ratio (95% CI)"),
      col_interval("hr", "lower", "upper", header = "HR (95% CI)")
    ),
    row_type = "row_type",
    row_fill = "is_pooled",
    theme = theme,
    title = title %||% "Acme Cardio Research — pooled meta-analysis",
    subtitle = subtitle
  )
}
```

## Step 0 — start from `web_theme_nejm()`

The package default is the cleanest, broadest-use preset; everything that follows is a delta from this baseline.

```{r}
render_acme(
  theme    = web_theme_nejm(),
  subtitle = "Step 0 — package default"
)
```

## Step 1 — pick the brand identity

Acme's brand seed is `#1E3A5F` — a deep editorial navy. Everything identity
(marker fills, title color, the brand ramp behind headers and gradients)
derives from this one V4 anchor.

```{r}
acme_t1 <- web_theme(brand = "#1E3A5F")

render_acme(acme_t1, subtitle = "Step 1 — navy brand")
```

## Step 2 — tint the neutrals toward the brand

The neutral ramp (paper → ink) drives every surface tone, divider, and the
chrome body-text color. Rather than hand-picking greys, nudge the paper/ink
anchors' *hues* toward the brand in one move with [`tint_from_brand()`] —
`"subtle"` travels 30% of the angular hue distance, leaving lightness
untouched. Acme wants a quietly cool, editorial chrome.

```{r}
acme_t2 <- tint_from_brand(acme_t1, "subtle")

render_acme(acme_t2, subtitle = "Step 2 — brand-tinted neutrals")
```

Notice the marker fill, the title color, and the cell-hairline tint all picked up the new identity at once. The chrome cascade does this automatically — there's no per-component knob to flip.

## Step 3 — pick the accent (engagement layer)

`accent` is reserved for engagement: hover, selected rows, semantic row callouts (`row_accent`), the selection edge, and the `status.info` fallback. **Never identity** — keep it visually distinct from `brand`. Acme picks a warm amber to counterpoint the cool navy.

```{r}
acme_t3 <- web_theme(
  brand  = "#1E3A5F",
  accent = "#C8893F"           # amber engagement
) |>
  tint_from_brand("subtle")

render_acme(acme_t3, subtitle = "Step 3 — amber accent")
```

## Step 4 — series colors for multi-effect plots

Per-effect colors (for multi-arm forest, grouped bars, multiple boxplots per
row) are *derived* — the resolved `series` slot bundles come from the `brand`
identity plus the `categorical` data scheme; the first slot also drives the
pooled-effect diamond. You don't pin them as an input. To choose the scheme,
set `categorical`; to override a single resolved slot fill, reach for
`set_theme_field(c("series", i, "fill"), …)`.

```{r}
acme_t4 <- web_theme(
  brand  = "#1E3A5F",
  accent = "#C8893F",
  categorical = "okabe_ito"    # the per-effect data palette
) |>
  tint_from_brand("subtle") |>
  # pin the first series slot to the brand navy (drives the pooled diamond)
  set_theme_field(c("series", 1L, "fill"), "#1E3A5F")

render_acme(acme_t4, subtitle = "Step 4 — derived series palette (single-effect here)")
```

The single-effect forest in this example only consumes the first slot; switch to a multi-effect `effect_forest()` list and the rest cycle in from the categorical scheme.

## Step 5 — fonts

Acme picks an editorial serif for display copy and a humanist sans for body. `fonts_display` falls back to `fonts_body` when unset, so we name both explicitly.

```{r}
acme_t5 <- web_theme(
  brand        = "#1E3A5F",
  accent       = "#C8893F",
  categorical  = "okabe_ito",
  fonts_body    = "'IBM Plex Sans', system-ui, sans-serif",
  fonts_display = "'IBM Plex Serif', Georgia, serif",
  web_fonts = list(
    web_font("IBM Plex Sans",
             "https://fonts.googleapis.com/css2?family=IBM+Plex+Sans:wght@400;500;600&display=swap"),
    web_font("IBM Plex Serif",
             "https://fonts.googleapis.com/css2?family=IBM+Plex+Serif:wght@500;600&display=swap")
  )
) |>
  tint_from_brand("subtle")

render_acme(acme_t5, subtitle = "Step 5 — IBM Plex Sans + Serif via web_fonts")
```

The widget injects `<link rel="stylesheet">` tags on mount for any `web_fonts` declared on the theme; SVG / PNG / PDF export embeds the `@font-face` rules into the SVG so the static output carries its glyphs (when a fallback is needed, the browser substitutes — see [Themes guide](../guide/themes.qmd) for details).

## Step 6 — pick variants

Variants are the per-table toggles that don't change identity. Acme picks **comfortable** density and a **tint** header — a brand-tinted band that sits halfway between the default light header and the maximum-chrome bold header.

```{r}
acme_t6 <- web_theme(
  brand        = "#1E3A5F",
  accent       = "#C8893F",
  categorical  = "okabe_ito",
  fonts_body    = "'IBM Plex Sans', system-ui, sans-serif",
  fonts_display = "'IBM Plex Serif', Georgia, serif",
  density            = "comfortable",
  header_style       = "tint",
  first_column_style = "default",
  web_fonts = list(
    web_font("IBM Plex Sans",
             "https://fonts.googleapis.com/css2?family=IBM+Plex+Sans:wght@400;500;600&display=swap"),
    web_font("IBM Plex Serif",
             "https://fonts.googleapis.com/css2?family=IBM+Plex+Serif:wght@500;600&display=swap")
  )
) |>
  tint_from_brand("subtle") |>
  set_theme_field(c("series", 1L, "fill"), "#1E3A5F")

render_acme(acme_t6, subtitle = "Step 6 — comfortable density, tint header")
```

## Step 7 — pin a single field with `set_theme_field()`

Almost every theme can be expressed entirely through Tier 1 inputs. The one place you reach for `set_theme_field()` is when you want to override a single derived field without disturbing the rest of the cascade — e.g., turning gridlines on, forcing a specific axis tick color, or pinning the row-group L1 background to a brand-specific shade.

Here, Acme wants visible axis gridlines (default off) and a slightly stronger tick label:

```{r}
acme_t7 <- acme_t6 |>
  set_theme_field(c("axis", "gridlines"), TRUE) |>
  set_theme_field(c("plot", "tick_label", "weight"), 500)

render_acme(acme_t7, subtitle = "Step 7 — gridlines on, slightly stronger tick label")
```

These overrides survive subsequent `set_inputs()` calls — the resolver fills NA-valued fields and leaves explicit pins alone.

## Step 8 — wrap as a function

Once you're happy, wrap the whole theme in a named function for team-wide use. Anyone on your team gets a one-line theme without learning the input model.

```{r}
web_theme_acme <- function() {
  web_theme(
    name         = "acme",
    brand        = "#1E3A5F",
    accent       = "#C8893F",
    categorical  = "okabe_ito",
    fonts_body    = "'IBM Plex Sans', system-ui, sans-serif",
    fonts_display = "'IBM Plex Serif', Georgia, serif",
    density            = "comfortable",
    header_style       = "tint",
    first_column_style = "default",
    web_fonts = list(
      web_font("IBM Plex Sans",
               "https://fonts.googleapis.com/css2?family=IBM+Plex+Sans:wght@400;500;600&display=swap"),
      web_font("IBM Plex Serif",
               "https://fonts.googleapis.com/css2?family=IBM+Plex+Serif:wght@500;600&display=swap")
    )
  ) |>
    tint_from_brand("subtle") |>
    set_theme_field(c("series", 1L, "fill"), "#1E3A5F") |>
    set_theme_field(c("axis", "gridlines"), TRUE) |>
    set_theme_field(c("plot", "tick_label", "weight"), 500)
}

# Now the whole team writes:
render_acme(web_theme_acme(), subtitle = "Step 8 — wrapped as web_theme_acme()")
```

## Putting it on a different table

Same theme, different data — column types and content shift, the brand stays consistent:

```{r}
sales <- data.frame(
  region   = c("Americas", "Europe", "Asia-Pac", "MENA"),
  q1       = c(2.4, 1.9, 2.1, 0.7),
  q2       = c(2.7, 2.0, 2.3, 0.8),
  q3       = c(3.1, 2.2, 2.5, 0.9),
  q4       = c(3.4, 2.4, 2.8, 1.0),
  yoy      = c(0.18, 0.11, 0.16, 0.22)
)

tabviz(sales, label = "region",
  columns = list(
    col_currency("q1", "Q1", decimals = 1),
    col_currency("q2", "Q2", decimals = 1),
    col_currency("q3", "Q3", decimals = 1),
    col_currency("q4", "Q4", decimals = 1),
    col_percent("yoy", "YoY"),
    viz_bar(effect_bar("q4"), header = "Q4 ($B)", axis_range = c(0, 4))
  ),
  theme = web_theme_acme(),
  title = "Acme Cardio — regional revenue",
  subtitle = "Same theme function, different schema."
)
```

## See also

- [Themes — Guide](../guide/themes.qmd) — the conceptual model of inputs / variants / overrides.
- [Themes — Reference](../reference/themes.qmd) — every `web_theme()` argument and `set_*()` modifier.
- [Theming Manifesto](../concepts/theming-manifesto.qmd) — *why* the cascade is shaped this way (two-knob principle, three-axis contract, anti-patterns).
- [LOTR (Easter Egg)](lotr.qmd) — three real shipped examples of theme functions wrapped this same way.
