---
title: "Clinical Examples"
---

```{r}
#| include: false
library(tabviz)
library(dplyr)
```

Meta-analyses, forest plots, and clinical trial visualizations.

## JAMA-Style Forest Plot

Classic medical journal style with clean typography:

```{r}
data(glp1_trials)

tabviz(
  glp1_trials,
  label = "study",
  group = "group",
  columns = list(
    col_text("drug", "Drug"),
    col_n("n"),
    col_events("events", "n", "Events"),
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1,
               axis_label = "Hazard Ratio (95% CI)"),
    col_interval("HR (95% CI)", point = "hr", lower = "lower", upper = "upper"),
    col_pvalue("pvalue", "P")
  ),
  row_type = "row_type",
  row_bold = "row_bold",
  theme = web_theme_jama(),
  title = "GLP-1 Agonist Cardiovascular Outcomes",
  subtitle = "Major adverse cardiovascular events (MACE)"
)
```

::: {.callout-tip}
## Theme Variations
The same data structure works with any theme. Try `web_theme_lancet()`, `web_theme_nature()`, or `web_theme_cochrane()` for different journal styles. See [Feature Showcases](showcases.qmd#all-theme-variants) for a visual comparison of all 9 built-in themes.
:::

---

## Subgroup Analysis

Forest plot with subgroup headers and indentation:

```{r}
subgroup_data <- tibble(
  label = c(
    "Overall Population", "",
    "Age",
    "  < 65 years", "  >= 65 years", "",
    "Sex",
    "  Male", "  Female", "",
    "Prior CV Disease",
    "  Yes", "  No"
  ),
  hr = c(0.78, NA, NA, 0.72, 0.85, NA, NA, 0.76, 0.82, NA, NA, 0.74, 0.81),
  lower = c(0.68, NA, NA, 0.58, 0.72, NA, NA, 0.64, 0.68, NA, NA, 0.61, 0.66),
  upper = c(0.89, NA, NA, 0.89, 1.00, NA, NA, 0.90, 0.99, NA, NA, 0.90, 0.99),
  n = c(15234, NA, NA, 7821, 7413, NA, NA, 9140, 6094, NA, NA, 8542, 6692),
  rtype = c("summary", "spacer", "header", "data", "data", "spacer",
            "header", "data", "data", "spacer", "header", "data", "data"),
  rbold = c(TRUE, FALSE, TRUE, FALSE, FALSE, FALSE,
            TRUE, FALSE, FALSE, FALSE, TRUE, FALSE, FALSE)
)

tabviz(subgroup_data,
  label = "label",
  columns = list(
    col_n("n"),
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1,
               axis_label = "Hazard Ratio"),
    col_interval("HR (95% CI)", point = "hr", lower = "lower", upper = "upper")
  ),
  row_type = "rtype",
  row_bold = "rbold",
  theme = web_theme_jama(),
  title = "Subgroup Analysis"
)
```

---

## Multi-Arm Dose-Response

Compare three treatment arms across multiple endpoints using `col_group` to organize results:

```{r}
dose_response <- data.frame(
  endpoint = c("Overall Survival", "Progression-Free Survival",
               "Objective Response", "Quality of Life"),
  low_hr  = c(0.88, 0.82, 0.85, 0.92),
  low_lo  = c(0.74, 0.68, 0.70, 0.78),
  low_hi  = c(1.05, 0.99, 1.03, 1.08),
  mid_hr  = c(0.76, 0.71, 0.74, 0.83),
  mid_lo  = c(0.62, 0.58, 0.60, 0.69),
  mid_hi  = c(0.93, 0.87, 0.91, 1.00),
  high_hr = c(0.65, 0.60, 0.62, 0.75),
  high_lo = c(0.51, 0.46, 0.48, 0.61),
  high_hi = c(0.83, 0.78, 0.80, 0.92)
)

tabviz(dose_response, label = "endpoint",
  columns = list(
    viz_forest(
      effects = list(
        effect_forest("low_hr", "low_lo", "low_hi", label = "Low Dose"),
        effect_forest("mid_hr", "mid_lo", "mid_hi", label = "Standard"),
        effect_forest("high_hr", "high_lo", "high_hi", label = "High Dose")
      ),
      scale = "log", null_value = 1,
      axis_label = "Hazard Ratio"
    ),
    col_group("Low Dose",
      col_interval("HR (95% CI)", point = "low_hr", lower = "low_lo", upper = "low_hi")
    ),
    col_group("Standard",
      col_interval("HR (95% CI)", point = "mid_hr", lower = "mid_lo", upper = "mid_hi")
    ),
    col_group("High Dose",
      col_interval("HR (95% CI)", point = "high_hr", lower = "high_lo", upper = "high_hi")
    )
  ),
  theme = web_theme_cochrane(),
  title = "Dose-Response Analysis",
  subtitle = "Three-arm randomized trial across key endpoints"
)
```

---

## Annotated Forest Plot

Combine reference lines with annotations, direction-coded marker colors, and significance-based styling in a single comprehensive example:

```{r}
annotated_data <- glp1_trials |>
  filter(row_type == "data") |>
  mutate(
    sig_color = case_when(
      upper < 1 ~ "#16a34a",
      lower > 1 ~ "#dc2626",
      TRUE ~ "#64748b"
    ),
    sig_shape = case_when(
      upper < 1 ~ "circle",
      lower > 1 ~ "diamond",
      TRUE ~ "square"
    )
  )

tabviz(annotated_data,
  label = "study",
  columns = list(
    col_text("drug", "Drug"),
    col_n("n", "N"),
    col_events("events", "n", "Events"),
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1,
               axis_gridlines = TRUE,
               annotations = list(
                 refline(1, label = "No effect", style = "solid"),
                 refline(0.85, label = "Pooled estimate", style = "dashed", color = "#2563eb")
               )),
    col_interval("HR (95% CI)", point = "hr", lower = "lower", upper = "upper"),
    col_pvalue("pvalue", "P", stars = TRUE)
  ),
  marker_color = "sig_color",
  marker_shape = "sig_shape",
  row_bold = ~ pvalue < 0.05,
  theme = web_theme_nature(),
  title = "Cardiovascular Outcomes with Annotations",
  subtitle = "Direction-coded markers: green = significant benefit, gray = non-significant"
)
```

---

## Large-Scale Meta-Analysis

A stress test: 150 synthetic studies split into four regional panels via `split_by`. With `shared_axis = TRUE`, all panels share a single x-axis range, and the sidebar becomes a navigable tree — no endless scroll.

```{r}
#| fig-height: 9
set.seed(42)
n_studies <- 150
regions <- c("North America", "Europe", "Asia-Pacific", "Latin America")
region_pool <- sample(regions, n_studies,
                      replace = TRUE, prob = c(0.32, 0.30, 0.25, 0.13))

large_meta <- data.frame(
  study = sprintf("Study-%03d", seq_len(n_studies)),
  region = region_pool,
  n = round(pmax(60, rlnorm(n_studies, meanlog = 6.2, sdlog = 0.7))),
  hr = round(exp(rnorm(n_studies, mean = -0.18, sd = 0.22)), 3)
) |>
  mutate(
    se = 0.08 + 0.18 / sqrt(n / 200),
    lower = round(exp(log(hr) - 1.96 * se), 3),
    upper = round(exp(log(hr) + 1.96 * se), 3),
    weight = round(1 / se^2, 1),
    pvalue = round(2 * pnorm(-abs(log(hr) / se)), 4)
  ) |>
  select(-se)

tabviz(large_meta,
  label = "study",
  split_by = "region",
  shared_axis = TRUE,
  columns = list(
    col_n("n"),
    col_bar("weight", "Weight"),
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1,
               axis_range = c(0.3, 2.5),
               axis_label = "Hazard Ratio (95% CI)",
               annotations = list(
                 refline(1, label = "No effect", style = "solid"),
                 refline(0.83, label = "Pooled", style = "dashed", color = "#2563eb")
               )),
    col_interval("HR (95% CI)", point = "hr", lower = "lower", upper = "upper"),
    col_pvalue("pvalue", "P", stars = TRUE)
  ),
  marker_color = ~ case_when(
    upper < 1 ~ "#16a34a",
    lower > 1 ~ "#dc2626",
    TRUE ~ "#64748b"
  ),
  theme = web_theme_jama(),
  title = "Large-Scale Regional Meta-Analysis",
  subtitle = "150 synthetic studies across 4 regions, shared axis for cross-panel comparison",
  footnote = "Use the sidebar tree to navigate regions. Pooled estimate shown as blue dashed reference line."
)
```

---

## Nested Hierarchical Groups

Three levels of nesting (Region > Country > Site) give multi-center trial programs a compact, collapsible structure. Click any group header to collapse an entire branch and all its children.

```{r}
nested_data <- tibble(
  site = c(
    "Boston General", "Mass Eye & Ear", "Johns Hopkins", "Cleveland Clinic",
    "UCL Hospital", "Imperial College", "Charite Berlin", "LMU Munich",
    "Tokyo University", "Osaka Medical", "Peking Union", "Shanghai Ruijin"
  ),
  region = c(
    rep("americas", 4),
    rep("europe", 4),
    rep("asia_pacific", 4)
  ),
  country = c(
    "usa", "usa", "usa", "usa",
    "uk", "uk", "germany", "germany",
    "japan", "japan", "china", "china"
  ),
  hr = c(0.72, 0.68, 0.75, 0.71, 0.78, 0.82, 0.69, 0.74, 0.65, 0.70, 0.67, 0.72),
  lower = c(0.58, 0.52, 0.61, 0.56, 0.64, 0.68, 0.54, 0.59, 0.50, 0.55, 0.52, 0.57),
  upper = c(0.89, 0.89, 0.92, 0.90, 0.95, 0.99, 0.88, 0.93, 0.85, 0.89, 0.86, 0.91),
  n = c(245, 189, 312, 278, 156, 134, 298, 267, 445, 389, 512, 478)
)

tabviz(
  nested_data,
  label = "site",
  group = c("region", "country"),  # Hierarchical: region > country
  columns = list(
    col_n("n"),
    col_interval("HR (95% CI)", point = "hr", lower = "lower", upper = "upper"),
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1, axis_label = "Hazard Ratio")
  ),
  theme = web_theme_modern(),
  title = "Nested Hierarchical Groups",
  subtitle = "Region > Country > Site (3 levels)",
  caption = "Click any group header to collapse that branch and all children"
)
```
