---
title: "Domain Dashboards"
description: "Information-dense dashboards across sports, survey, finance, ops, and climate"
---

```{r}
#| include: false
library(tabviz)
library(dplyr)
```

`tabviz()` works as well outside the journal as inside. This page collects dense, information-rich dashboards from six domains — sports analytics, customer feedback, portfolio performance, product metrics, airline ops, climate science — built with the same `columns = list(...)` API. Each example leans into the column types that fit the domain (rings for percentages, pictograms for counts and ratings, sparklines for trends, viz_forest for confidence intervals on any interval-shaped metric).

## Sports analytics — NBA player efficiency

Three dimensions in one row: shooting efficiency (rings — `FG%` / `3P%` / `FT%`), accumulated honors (pictograms in rating mode for All-Star appearances), and a focal forest plot for Player Efficiency Rating with 95% CIs. Row groups by conference; row accent on All-Stars.

```{r}
#| fig-height: 9
data(nba_efficiency)

# Synthesize a few realistic-feeling secondary stats so we can exercise
# col_ring / col_pictogram / viz_bar on top of the per-CI forest.
set.seed(2024)
nba_dash <- nba_efficiency |>
  mutate(
    fg_pct      = round(runif(n(), 0.42, 0.58), 3),
    three_pct   = round(runif(n(), 0.30, 0.42), 3),
    ft_pct      = round(runif(n(), 0.70, 0.92), 3),
    all_star_n  = pmin(15L, sample(0:9, n(), replace = TRUE) +
                            ifelse(all_star, 1L, 0L)),
    is_mvp      = award == "MVP"
  )

tabviz(nba_dash,
  label = "player",
  group = "conference",
  columns = list(
    col_group("Player",
      col_text("team", "Team", width = 50),
      col_text("position", "Pos", width = 40)
    ),
    col_group("Volume",
      col_numeric("games", "GP", decimals = 0),
      col_numeric("ppg", "PPG", decimals = 1)
    ),
    col_group("Shooting",
      col_ring("fg_pct",    header = "FG%",
               thresholds = c(0.45, 0.52)),
      col_ring("three_pct", header = "3P%",
               thresholds = c(0.34, 0.39)),
      col_ring("ft_pct",    header = "FT%",
               thresholds = c(0.75, 0.85))
    ),
    col_pictogram("all_star_n", header = "All-Star apps",
                  glyph = "star", max_glyphs = 15,
                  value_label = "trailing", label_format = "integer"),
    viz_forest(point = "per", lower = "per_lower", upper = "per_upper",
               null_value = 15,
               axis_label = "Player Efficiency Rating (95% CI)"),
    col_badge("award", "Award",
      variants = list(
        MVP                = "success",
        `All-NBA 1st`      = "info",
        `All-NBA 2nd`      = "warning",
        `All-NBA 3rd`      = "muted",
        `Defensive POY`    = "info",
        `6th Man`          = "muted",
        `Most Improved`    = "muted",
        `Rookie of Year`   = "muted",
        `All-Star`         = "muted"
      ))
  ),
  row_accent = "all_star",       # bold + theme accent on All-Stars
  row_fill   = "is_mvp",         # strong fill on MVP row
  theme = web_theme_cochrane() |>
    set_inputs(brand = "#1d428a", accent = "#c8102e"),  # NBA blue + red
  title    = "NBA Player Efficiency Dashboard",
  subtitle = "PER = 15 is league average. Rings show shooting efficiency by typical-range thresholds.",
  caption  = "All-Star rows accented; MVP row gets the row.fill token."
)
```

::: {.callout-tip}
## What this exercises

- **`col_ring()` with `thresholds`**: each shooting-percentage column flips color at a typical-range cutoff (e.g. FG% green ≥ 0.52, amber 0.45–0.52, red < 0.45). No hex literals — the cascade resolves to the theme's status palette.
- **`col_pictogram()` rating mode**: `max_glyphs = 15` with `value_label = "trailing"` shows accumulated All-Star appearances visually + numerically.
- **`row_accent` and `row_fill` together**: the 26 All-Stars take the accent token (bold + accent color), and the single MVP row layers the fill token on top — strongest visual hierarchy reserved for the rarest case.
- **Custom-tinted theme** via `set_inputs()` — NBA blue + red without rebuilding from scratch.
:::

---

## Customer satisfaction survey

Likert-style satisfaction scores as rings (with auto-defaulted positive / warning / negative palette), detractor counts as pictograms, response progress as bars. Row fill flags the segment with the strongest result; row accent flags any with concerning scores.

```{r}
#| fig-height: 6
survey <- data.frame(
  segment      = c("Enterprise — strategic", "Enterprise — mid-market",
                   "SMB — paid",             "SMB — self-serve",
                   "Education",              "Public sector"),
  responses    = c(412, 287, 1820, 4560, 198, 142),
  response_rate= c(0.71, 0.58, 0.32, 0.18, 0.62, 0.49),
  csat         = c(0.87, 0.82, 0.74, 0.66, 0.89, 0.78),         # 0–1
  nps          = c( 62,  48,  31,   8,  74,  44),                # -100 to 100
  ces          = c(0.82, 0.76, 0.68, 0.55, 0.85, 0.71),
  detractors   = c(  2,   3,   8,  14,   1,   4),
  promoters    = c(  9,   8,   6,   3,  11,   7),
  trend        = I(list(
    c(78, 80, 82, 85, 87), c(72, 75, 78, 80, 82),
    c(65, 68, 70, 72, 74), c(60, 62, 64, 65, 66),
    c(82, 85, 87, 88, 89), c(70, 73, 75, 77, 78)
  )),
  is_top       = c(FALSE, FALSE, FALSE, FALSE, TRUE, FALSE),     # row_fill
  is_concern   = c(FALSE, FALSE, FALSE, TRUE,  FALSE, FALSE)     # row_accent
)

tabviz(survey,
  label = "segment",
  columns = list(
    col_n("responses", "N"),
    col_progress("response_rate", "Response rate", max_value = 1,
                 show_label = TRUE),
    col_group("Satisfaction",
      col_ring("csat", header = "CSAT", thresholds = c(0.70, 0.85)),
      col_ring("ces",  header = "CES",  thresholds = c(0.65, 0.80))
    ),
    col_numeric("nps", "NPS", decimals = 0),
    col_pictogram("detractors", header = "Detractors",
                  glyph = "person", layout = "row"),
    col_pictogram("promoters",  header = "Promoters",
                  glyph = "person", layout = "row"),
    col_sparkline("trend", "5q trend")
  ),
  row_fill   = "is_top",
  row_accent = "is_concern",
  theme = web_theme_cochrane(),
  title    = "Customer satisfaction by segment",
  subtitle = "CSAT and CES bands set per the program's typical-range thresholds; trend on the right.",
  caption  = "SMB self-serve flagged for response-rate concern; Education leads on every metric."
)
```

::: {.callout-note}
## When to reach for what

- **Use `col_ring()`** when the metric has a natural percentage / fraction shape *and* falls into typical "good / OK / concern" bands. The threshold-driven palette tells the story at a glance.
- **Use `col_pictogram()`** when the metric is a **count** that benefits from visual heft (one icon per detractor, one star per appearance) — it reads faster than a bare number for small counts.
- **Use `col_sparkline()`** when you need direction without precision; for axis-comparable trends across rows reach for `viz_bar()` or `viz_forest()` instead.
:::

---

## Portfolio performance

Twelve-month return with a 95% CI as the focal forest column, allocation share as a ring, position counts as a pictogram, and a Sharpe-ratio bar. Row-level accent on assets that beat their benchmark; row fill on the overall portfolio summary.

```{r}
#| fig-height: 7
portfolio <- data.frame(
  asset        = c("US Large Cap", "US Small Cap", "International Dev",
                   "Emerging Mkts", "US IG Bonds",  "TIPS",
                   "Gold",         "Real Estate",   "Cash",
                   "PORTFOLIO"),
  category     = c("Equity", "Equity", "Equity", "Equity",
                   "Fixed Income", "Fixed Income",
                   "Alternative", "Alternative", "Cash",
                   "Total"),
  allocation   = c(0.32, 0.10, 0.12, 0.06, 0.22, 0.05,
                   0.04, 0.06, 0.03, 1.00),
  positions    = c(  4,    3,    5,    3,    6,    1,
                     1,    2,    0,   25L),
  return_pct   = c(0.183, 0.122, 0.094, 0.072, 0.041, 0.025,
                   0.158, 0.108, 0.043, 0.118),
  return_lo    = c(0.142, 0.062, 0.038, 0.005, 0.020, 0.008,
                   0.090, 0.058, 0.041, 0.092),
  return_hi    = c(0.224, 0.182, 0.150, 0.139, 0.062, 0.042,
                   0.226, 0.158, 0.045, 0.144),
  sharpe       = c(1.45, 0.86, 0.72, 0.41, 1.02, 0.62,
                   0.94, 0.81, NA,   1.18),
  beats_bench  = c(TRUE, FALSE, FALSE, FALSE, TRUE, TRUE,
                   FALSE, TRUE, FALSE, TRUE),
  row_type     = c(rep("data", 9), "summary"),
  is_total     = c(rep(FALSE, 9), TRUE)
)

tabviz(portfolio,
  label = "asset",
  group = "category",
  columns = list(
    col_ring("allocation", header = "Alloc",
             thresholds = c(0.05, 0.20)),     # tiny / typical / heavy
    col_pictogram("positions", header = "Positions",
                  glyph = "person",
                  value_label = "trailing", label_format = "integer",
                  max_glyphs = 6),
    viz_forest(point = "return_pct", lower = "return_lo", upper = "return_hi",
               null_value = 0,
               axis_label = "12-month return (95% CI)"),
    col_percent("return_pct", "Return", multiply = TRUE),
    col_numeric("sharpe",     "Sharpe", decimals = 2),
    col_badge("beats_bench", "vs Benchmark",
      variants = list(`TRUE` = "success", `FALSE` = "muted"))
  ),
  row_type   = "row_type",
  row_accent = "beats_bench",     # bold + accent on outperformers
  row_fill   = "is_total",        # strong fill on overall portfolio row
  theme = web_theme_lancet(),
  title    = "Portfolio performance — trailing twelve months",
  subtitle = "Allocation share, position count, return CI, and Sharpe ratio per holding category.",
  caption  = "Outperforming holdings highlighted with the accent token; overall portfolio row uses the fill token."
)
```

---

## Product metrics

Five product variants tracked across revenue, growth, margin, units sold, and a 12-month trend:

```{r}
product_metrics <- data.frame(
  product = c("Widget Pro", "Widget Lite", "Widget Plus", "Widget Max", "Widget Mini"),
  revenue = c(2450000, 1890000, 3200000, 1560000, 980000),
  growth = c(0.12, -0.05, 0.28, 0.08, 0.15),
  margin = c(0.35, 0.42, 0.28, 0.38, 0.45),
  units_sold = c(12500, 9800, 18200, 7800, 15600),
  satisfaction = c(4.2, 3.8, 4.5, 4.1, 4.3),
  trend = I(list(
    c(180, 195, 210, 205, 225, 240, 235, 250, 260, 275, 290, 305),
    c(220, 215, 205, 210, 200, 195, 190, 185, 188, 182, 180, 178),
    c(200, 220, 240, 260, 290, 320, 340, 360, 380, 400, 420, 440),
    c(120, 125, 130, 128, 135, 140, 138, 145, 150, 148, 155, 160),
    c(60, 68, 75, 82, 88, 95, 100, 108, 115, 122, 128, 135)
  ))
)

tabviz(product_metrics, label = "product",
  columns = list(
    col_currency("revenue", "Revenue", decimals = 0),
    col_percent("growth", "YoY"),
    col_percent("margin", "Margin"),
    viz_bar(
      effect_bar("units_sold"),
      header = "Units Sold",
      axis_range = c(0, 20000)
    ),
    col_sparkline("trend", "12M Trend"),
    col_stars("satisfaction", "Rating", max_stars = 5)
  ),
  row_color = ~ ifelse(growth < 0, "#dc2626", "#64748b"),
  theme = web_theme_cochrane(),
  title = "Product Performance Dashboard"
)
```

---

## A/B test results

Conversion lift with a 95% CI per variant, p-value, and a verdict badge. The control row is a `row_type = "summary"` baseline.

```{r}
ab_test_results <- data.frame(
  variant = c("Control", "Variant A", "Variant B", "Variant C"),
  visitors = c(50000, 48500, 51200, 47800),
  conversions = c(2500, 2910, 2816, 2629),
  rate = c(0.050, 0.060, 0.055, 0.055),
  lift = c(0, 0.20, 0.10, 0.10),
  lower = c(NA, 0.08, -0.02, -0.03),
  upper = c(NA, 0.32, 0.22, 0.23),
  pvalue = c(NA, 0.002, 0.12, 0.15),
  verdict = c("Baseline", "Winner", "Inconclusive", "Inconclusive"),
  row_type = c("summary", "data", "data", "data"),
  row_bold = c(TRUE, FALSE, FALSE, FALSE)
)

tabviz(ab_test_results, label = "variant",
  columns = list(
    col_n("visitors", "Visitors", abbreviate = TRUE),
    col_n("conversions", "Conv"),
    col_percent("rate", "CVR"),
    viz_forest(
      point = "lift", lower = "lower", upper = "upper",
      null_value = 0,
      axis_label = "Lift vs Control"
    ),
    col_percent("lift", "Lift"),
    col_pvalue("pvalue", "P", stars = TRUE),
    col_badge("verdict", "Verdict",
      variants = list(
        Baseline = "info",
        Winner = "success",
        Inconclusive = "warning"
      ))
  ),
  row_type = "row_type",
  row_bold = "row_bold",
  marker_color = ~ ifelse(is.na(lower), "#64748b",
                          ifelse(lower > 0, "#16a34a", "#64748b")),
  theme = web_theme_cochrane(),
  title = "Homepage Redesign A/B Test"
)
```

---

## Airline performance

Carrier delays with a heatmap on on-time percentage, a progress bar for satisfaction, and a 12-month delay trend:

```{r}
data(airline_delays)

airline_summary <- airline_delays |>
  filter(month == "Apr") |>
  select(carrier, carrier_type, on_time_pct, delay_vs_avg, satisfaction, flights, trend)

tabviz(airline_summary, label = "carrier",
  columns = list(
    col_badge("carrier_type", "Type",
      variants = list(Legacy = "info", `Low-Cost` = "success",
                      `Ultra Low-Cost` = "warning")),
    col_n("flights", "Flights", abbreviate = TRUE),
    col_heatmap("on_time_pct", "On-Time",
      palette = c("#fee2e2", "#dcfce7"),
      min_value = 0.65, max_value = 0.95, decimals = 2, show_value = TRUE),
    col_progress("satisfaction", "Satisfaction", max_value = 5),
    col_numeric("delay_vs_avg", "Delay vs Avg", decimals = 1),
    col_sparkline("trend", "12-Month Trend")
  ),
  theme = web_theme_cochrane(),
  title = "Airline Performance — April"
)
```

---

## Climate — regional temperature anomalies

The same interval-estimate structure that runs clinical forest plots maps cleanly onto temperature anomalies vs. a pre-industrial baseline. Marker shape encodes confidence; marker color signals severity:

```{r}
data(climate_temps)

tabviz(climate_temps,
  label = "region",
  group = "category",
  columns = list(
    col_text("period", "Period"),
    col_text("certainty", "Confidence"),
    col_interval("anomaly", "lower", "upper", header = "Anomaly °C (95% CI)"),
    viz_forest(point = "anomaly", lower = "lower", upper = "upper",
               null_value = 0, axis_label = "Temperature Anomaly (°C vs. pre-industrial)")
  ),
  marker_shape = ~ ifelse(certainty == "High", "circle", "square"),
  marker_color = ~ ifelse(anomaly > 1.5, "#dc2626", ifelse(anomaly > 1.0, "#f59e0b", "#3b82f6")),
  theme = web_theme_cochrane() |>
    set_theme_field(c("axis", "gridlines"), TRUE),
  title    = "Regional Temperature Anomalies",
  subtitle = "Grouped by geographic category, shaped by confidence level",
  caption  = "Red ≥ 1.5 °C, amber 1.0–1.5 °C, blue < 1.0 °C; circle = high confidence."
)
```

---

## See also

- [Long Meta-Analysis](meta-analysis.qmd) — pooled summaries + pagination across 70 studies.
- [Pictogram & Ring Deep Dive](viz-glyphs.qmd) — every mode of `col_pictogram()` and `col_ring()`.
- [Compose Your Own Brand](brand-design.qmd) — building the NBA / Lancet / journal themes from scratch.
- [Visualization Styling](../guide/visualizations/styling.qmd) — the four-layer marker-color cascade across `viz_*` columns.
- [Reference → Columns](../reference/columns.qmd) — every `col_*` and `viz_*` argument.
