---
title: "Dashboard Examples"
---

```{r}
#| include: false
library(tabviz)
library(dplyr)
```

Information-dense dashboards built from `tabviz()` column primitives — heatmaps, progress bars, sparklines, badges, currency, and the occasional in-line forest plot. Each example targets a different domain (product, ops, climate, sports) but uses the same data-frame-plus-columns API.

## Product Metrics Dashboard

Track product KPIs with sparklines, bars, stars, currency, and trend indicators:

```{r}
product_metrics <- data.frame(
  product = c("Widget Pro", "Widget Lite", "Widget Plus", "Widget Max", "Widget Mini"),
  revenue = c(2450000, 1890000, 3200000, 1560000, 980000),
  growth = c(0.12, -0.05, 0.28, 0.08, 0.15),
  margin = c(0.35, 0.42, 0.28, 0.38, 0.45),
  units_sold = c(12500, 9800, 18200, 7800, 15600),
  satisfaction = c(4.2, 3.8, 4.5, 4.1, 4.3),
  trend = I(list(
    c(180, 195, 210, 205, 225, 240, 235, 250, 260, 275, 290, 305),
    c(220, 215, 205, 210, 200, 195, 190, 185, 188, 182, 180, 178),
    c(200, 220, 240, 260, 290, 320, 340, 360, 380, 400, 420, 440),
    c(120, 125, 130, 128, 135, 140, 138, 145, 150, 148, 155, 160),
    c(60, 68, 75, 82, 88, 95, 100, 108, 115, 122, 128, 135)
  ))
)

tabviz(product_metrics, label = "product",
  columns = list(
    col_currency("revenue", "Revenue", decimals = 0),
    col_percent("growth", "YoY"),
    col_percent("margin", "Margin"),
    viz_bar(
      effect_bar("units_sold"),
      header = "Units Sold",
      axis_range = c(0, 20000)
    ),
    col_sparkline("trend", "12M Trend"),
    col_stars("satisfaction", "Rating", max_stars = 5)
  ),
  row_color = ~ ifelse(growth < 0, "#dc2626", "#64748b"),
  theme = web_theme_modern(),
  title = "Product Performance Dashboard"
)
```

---

## A/B Test Results

Compare conversion rates across experiment variants with confidence intervals and winner badges:

```{r}
ab_test_results <- data.frame(
  variant = c("Control", "Variant A", "Variant B", "Variant C"),
  visitors = c(50000, 48500, 51200, 47800),
  conversions = c(2500, 2910, 2816, 2629),
  rate = c(0.050, 0.060, 0.055, 0.055),
  lift = c(0, 0.20, 0.10, 0.10),
  lower = c(NA, 0.08, -0.02, -0.03),
  upper = c(NA, 0.32, 0.22, 0.23),
  pvalue = c(NA, 0.002, 0.12, 0.15),
  verdict = c("Baseline", "Winner", "Inconclusive", "Inconclusive"),
  row_type = c("summary", "data", "data", "data"),
  row_bold = c(TRUE, FALSE, FALSE, FALSE)
)

tabviz(ab_test_results, label = "variant",
  columns = list(
    col_n("visitors", "Visitors", abbreviate = TRUE),
    col_n("conversions", "Conv"),
    col_percent("rate", "CVR"),
    viz_forest(
      point = "lift", lower = "lower", upper = "upper",
      null_value = 0,
      axis_label = "Lift vs Control"
    ),
    col_percent("lift", "Lift"),
    col_pvalue("pvalue", "P", stars = TRUE),
    col_badge("verdict", "Verdict",
      variants = list(
        Baseline = "info",
        Winner = "success",
        Inconclusive = "warning"
      ))
  ),
  row_type = "row_type",
  row_bold = "row_bold",
  marker_color = ~ ifelse(is.na(lower), "#64748b",
                          ifelse(lower > 0, "#16a34a", "#64748b")),
  theme = web_theme_nature(),
  title = "Homepage Redesign A/B Test"
)
```

---

## Airline Performance

Airline delay metrics with heatmaps for on-time percentage and progress bars for satisfaction:

```{r}
data(airline_delays)

airline_summary <- airline_delays |>
  filter(month == "Apr") |>
  select(carrier, carrier_type, on_time_pct, delay_vs_avg, satisfaction, flights, trend)

tabviz(
  airline_summary,
  label = "carrier",
  columns = list(
    col_badge("carrier_type", "Type",
      variants = list(
        Legacy = "info",
        `Low-Cost` = "success",
        `Ultra Low-Cost` = "warning"
      )),
    col_n("flights", "Flights", abbreviate = TRUE),
    col_heatmap("on_time_pct", "On-Time %",
      palette = c("#fee2e2", "#dcfce7"),
      min_value = 0.65, max_value = 0.95,
      decimals = 2, show_value = TRUE),
    col_progress("satisfaction", "Satisfaction", max_value = 5),
    col_numeric("delay_vs_avg", "Delay vs Avg", decimals = 1),
    col_sparkline("trend", "12-Month Trend")
  ),
  row_bg = ~ case_when(
    on_time_pct > 0.85 ~ "#dcfce7",
    on_time_pct < 0.75 ~ "#fee2e2",
    TRUE ~ NA_character_
  ),
  theme = web_theme_modern(),
  title = "Airline Performance Dashboard"
)
```

---

## Executive Summary

A visually rich C-suite dashboard using progress bars, heatmaps, sparklines, and badges:

```{r}
exec_data <- data.frame(
  metric = c("Revenue", "Profit Margin", "Customer Count", "Retention Rate",
             "Net Promoter Score", "Employee Satisfaction"),
  current_val = c(142000000, 0.197, 45200, 0.92, 72, 4.1),
  target_val = c(150000000, 0.200, 50000, 0.95, 75, 4.5),
  pct_to_target = c(0.947, 0.985, 0.904, 0.968, 0.960, 0.911),
  health = c(88, 92, 78, 95, 85, 72),
  status = c("On Track", "On Track", "Behind", "On Track", "On Track", "At Risk"),
  trend = I(list(
    c(120, 125, 130, 128, 135, 138, 140, 142),
    c(0.18, 0.185, 0.19, 0.188, 0.192, 0.195, 0.196, 0.197),
    c(38, 40, 41, 42, 43, 44, 44.5, 45.2),
    c(0.88, 0.89, 0.90, 0.91, 0.91, 0.92, 0.92, 0.92),
    c(65, 67, 68, 70, 71, 72, 72, 72),
    c(3.8, 3.9, 3.9, 4.0, 4.0, 4.1, 4.1, 4.1)
  ))
)

tabviz(exec_data,
  label = "metric",
  columns = list(
    col_badge("status", "Status",
      variants = list(
        `On Track` = "success",
        Behind = "warning",
        `At Risk` = "error"
      )),
    col_progress("pct_to_target", "Progress to Target", max_value = 1),
    col_heatmap("health", "Health",
      palette = c("#fef3c7", "#dcfce7"),
      min_value = 60, max_value = 100, decimals = 0),
    col_sparkline("trend", "8-Quarter Trend")
  ),
  theme = web_theme_presentation(),
  title = "Q4 Executive Summary"
)
```

---

## Dark Mode Dashboard

A server monitoring dashboard using the dark theme with heatmaps, progress bars, badges, and sparklines:

```{r}
server_data <- data.frame(
  server = c("prod-web-01", "prod-web-02", "prod-api-01",
             "prod-db-01", "prod-cache-01", "staging-01"),
  status = c("Healthy", "Healthy", "Warning", "Healthy", "Critical", "Maintenance"),
  cpu_load = c(42, 67, 88, 35, 95, 12),
  memory_pct = c(0.58, 0.72, 0.91, 0.45, 0.97, 0.22),
  disk_pct = c(0.45, 0.62, 0.78, 0.88, 0.93, 0.31),
  uptime_days = c(142, 142, 89, 365, 23, 3),
  requests = I(list(
    c(1200, 1350, 1500, 1420, 1600, 1550, 1700, 1650, 1800, 1750, 1900, 1850),
    c(1100, 1250, 1400, 1350, 1500, 1450, 1600, 1550, 1700, 1650, 1800, 1750),
    c(800, 900, 950, 1100, 1300, 1500, 1800, 2100, 2400, 2600, 2800, 3000),
    c(500, 520, 510, 530, 540, 535, 550, 545, 560, 555, 570, 565),
    c(400, 420, 450, 480, 520, 580, 650, 750, 900, 1100, 1400, 1800),
    c(50, 60, 55, 45, 40, 35, 30, 25, 20, 15, 10, 8)
  ))
)

tabviz(server_data, label = "server",
  columns = list(
    col_badge("status", "Status",
      variants = list(
        Healthy = "success",
        Warning = "warning",
        Critical = "error",
        Maintenance = "info"
      )),
    col_heatmap("cpu_load", "CPU %",
      palette = c("#dcfce7", "#fee2e2"),
      min_value = 0, max_value = 100, decimals = 0),
    col_progress("memory_pct", "Memory", max_value = 1),
    col_progress("disk_pct", "Disk", max_value = 1),
    col_n("uptime_days", "Uptime (d)"),
    col_sparkline("requests", "Traffic (12h)")
  ),
  theme = web_theme_dark(),
  title = "Infrastructure Monitor"
)
```

---

## Org Chart Executive Dashboard

A table-only view (no forest plot) of company org structure — row styling carries the hierarchy: summary row at the top, bolded department headers, indented sub-teams, and a sparkline showing 6-month headcount trend.

```{r}
exec_dashboard <- tibble(
  department = c(
    "COMPANY TOTAL", "",
    "Engineering", "  Platform", "  Infrastructure", "  Mobile",
    "",
    "Product", "  Design", "  Research",
    "",
    "Sales", "  Enterprise", "  SMB"
  ),
  headcount = c(1250, NA, 480, 180, 150, 150, NA, 220, 85, 135, NA, 550, 320, 230),
  revenue_m = c(185, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, 185, 142, 43),
  growth = c(24, NA, 32, 45, 28, 18, NA, 28, 35, 22, NA, 18, 22, 12),
  satisfaction = c(78, NA, 82, 85, 80, 81, NA, 76, 79, 74, NA, 75, 77, 72),
  trend = list(
    c(1050, 1100, 1150, 1180, 1220, 1250), NULL,
    c(380, 400, 420, 440, 460, 480), c(140, 150, 160, 170, 175, 180),
    c(120, 130, 140, 145, 148, 150), c(120, 120, 125, 130, 140, 150), NULL,
    c(180, 190, 200, 208, 215, 220), c(65, 70, 75, 80, 82, 85),
    c(115, 120, 125, 128, 132, 135), NULL,
    c(490, 505, 520, 532, 542, 550), c(280, 290, 300, 308, 315, 320),
    c(210, 215, 220, 224, 227, 230)
  ),
  effect = rep(1, 14), lower = rep(0.9, 14), upper = rep(1.1, 14),
  rtype = c("summary", "spacer", "header", rep("data", 3), "spacer",
            "header", "data", "data", "spacer", "header", "data", "data"),
  rbold = c(TRUE, FALSE, TRUE, rep(FALSE, 3), FALSE, TRUE, FALSE, FALSE, FALSE, TRUE, FALSE, FALSE),
  rindent = c(0, 0, 0, 1, 1, 1, 0, 0, 1, 1, 0, 0, 1, 1),
  rcolor = c("#16a34a", NA, "#2563eb", NA, NA, NA, NA, "#2563eb", NA, NA, NA, "#2563eb", NA, NA)
)

tabviz(
  exec_dashboard,
  label = "department",
  columns = list(
    col_numeric("headcount", "HC"),
    col_numeric("revenue_m", "Rev $M"),
    col_bar("growth", "Growth %"),
    col_numeric("satisfaction", "eNPS"),
    col_sparkline("trend", "6M Trend")
  ),
  row_type = "rtype", row_bold = "rbold", row_indent = "rindent", row_color = "rcolor",
  theme = web_theme_modern(),
  title = "Executive Dashboard",
  subtitle = "Table-only with hierarchical org structure",
  caption = "Using tabviz() with row styling for org charts"
)
```

---

## Financial Benchmarks

A SaaS peer-group benchmark: revenue in millions USD, YoY growth, gross margin, quartile ranking, and a quarterly ARR trend sparkline. Row background is colored by quartile.

```{r}
set.seed(2026)
saas_peers <- data.frame(
  company = c("Aurora Cloud", "Beacon Labs", "Cirrus IO", "Delta Forge",
              "Ember Systems", "Fathom AI", "Helios Data", "Ion Works",
              "Juno Metrics", "Kestrel Security"),
  revenue_m = c(842, 612, 1204, 318, 186, 2450, 945, 428, 276, 1567),
  yoy_growth = c(0.28, 0.42, 0.18, 0.65, 0.88, 0.22, 0.35, 0.52, 0.71, 0.15),
  margin = c(0.74, 0.68, 0.81, 0.59, 0.52, 0.85, 0.77, 0.63, 0.55, 0.79),
  arr_trend = I(lapply(1:10, function(i) {
    base <- runif(1, 80, 220)
    cumsum(c(base, rnorm(11, mean = base * 0.08, sd = base * 0.04)))
  })),
  quartile = c("Q2", "Q1", "Q2", "Q1", "Q1", "Q3", "Q2", "Q1", "Q1", "Q3")
)

tabviz(saas_peers, label = "company",
  columns = list(
    col_currency("revenue_m", "Revenue ($M)", decimals = 0),
    col_bar("yoy_growth", "YoY Growth", max_value = 1),
    col_bar("margin", "Gross Margin", max_value = 1),
    col_sparkline("arr_trend", "ARR (12Q)"),
    col_badge("quartile", "Quartile",
      variants = list(Q1 = "success", Q2 = "info", Q3 = "warning", Q4 = "error"))
  ),
  row_bg = ~ case_when(
    quartile == "Q1" ~ "#ecfdf5",
    quartile == "Q4" ~ "#fef2f2",
    TRUE ~ NA_character_
  ),
  theme = web_theme_modern(),
  title = "SaaS Peer Benchmark",
  subtitle = "Revenue, growth, margin, and ARR trajectory across 10 companies"
)
```

---

## Academic / Education

Meta-analysis of test-score interventions grouped by grade band. Effect sizes are Cohen's *d* with 95% CIs; forest plot uses `null_value = 0` with a linear scale.

```{r}
set.seed(7)
n_int <- 14
interventions <- data.frame(
  intervention = c(
    "Phonics Tutoring", "Small-Group Reading", "1:1 Math Coach",
    "Peer Tutoring", "Summer Bridge", "After-School Math",
    "Writing Workshop", "STEM Lab Rotation", "Flipped Classroom",
    "Project-Based Learning", "Socratic Seminars", "AP Prep Bootcamp",
    "Career Mentorship", "Digital Literacy"
  ),
  grade_level = c("K-5", "K-5", "K-5", "K-5",
                  "6-8", "6-8", "6-8", "6-8", "6-8",
                  "9-12", "9-12", "9-12", "9-12", "9-12"),
  n = c(842, 1204, 416, 988, 612, 524, 368, 892, 1056, 742, 488, 934, 612, 1120),
  effect_size = c(0.42, 0.58, 0.71, 0.33, 0.28, 0.12, 0.45, 0.38,
                  0.52, 0.22, 0.61, 0.35, 0.48, -0.04),
  ci_lo = c(0.28, 0.41, 0.52, 0.18, 0.11, -0.05, 0.22, 0.22,
            0.38, 0.05, 0.42, 0.19, 0.28, -0.20),
  ci_hi = c(0.56, 0.75, 0.90, 0.48, 0.45, 0.29, 0.68, 0.54,
            0.66, 0.39, 0.80, 0.51, 0.68, 0.12)
) |>
  mutate(
    status = case_when(
      ci_lo > 0.2 ~ "Pass",
      ci_hi < 0.1 ~ "Fail",
      TRUE ~ "Improvement"
    )
  )

tabviz(interventions,
  label = "intervention",
  group = "grade_level",
  columns = list(
    col_n("n"),
    viz_forest(point = "effect_size", lower = "ci_lo", upper = "ci_hi",
               scale = "linear", null_value = 0,
               axis_label = "Effect size (Cohen's d)"),
    col_interval("effect_size", "ci_lo", "ci_hi", header = "d (95% CI)")
  ),
  row_badge = "status",
  theme = web_theme_lancet(),
  title = "Test-Score Interventions by Grade Band",
  subtitle = "Row groups: K-5 / 6-8 / 9-12. Badge shows pre-registered pass criterion (d > 0.2, CI excludes 0).",
  footnote = "Synthetic data modeled on typical education meta-analysis effect sizes."
)
```

---

## Operations / QC

Manufacturing-line process monitoring: capacity utilization bars, yield sparklines, a defect-rate heatmap, and a status badge. `row_bg` flags critical lines.

```{r}
set.seed(99)
qc_lines <- data.frame(
  line = c("Line-01", "Line-02", "Line-03", "Line-04", "Line-05",
           "Line-06", "Line-07", "Line-08", "Line-09"),
  defect_rate = c(0.018, 0.042, 0.009, 0.068, 0.021, 0.015, 0.055, 0.033, 0.012),
  target = rep(0.025, 9),
  capacity_used = c(0.82, 0.74, 0.91, 0.58, 0.88, 0.79, 0.45, 0.85, 0.93),
  yield_trend = I(lapply(1:9, function(i) {
    base <- runif(1, 0.88, 0.98)
    pmin(1, pmax(0.75, base + cumsum(rnorm(24, 0, 0.01))))
  })),
  status = c("In control", "Warning", "In control", "Critical", "In control",
             "In control", "Critical", "Warning", "In control")
)

tabviz(qc_lines, label = "line",
  columns = list(
    col_heatmap("defect_rate", "Defect Rate",
      palette = c("#dcfce7", "#fee2e2"),
      min_value = 0, max_value = 0.08, decimals = 3),
    col_numeric("target", "Target", decimals = 3),
    col_bar("capacity_used", "Capacity", max_value = 1),
    col_sparkline("yield_trend", "Yield (24h)"),
    col_badge("status", "Status",
      variants = list(
        `In control` = "success",
        Warning = "warning",
        Critical = "error"
      ))
  ),
  row_bg = ~ case_when(
    status == "Critical" ~ "#fef2f2",
    status == "Warning" ~ "#fffbeb",
    TRUE ~ NA_character_
  ),
  theme = web_theme_modern(),
  title = "Production-Line QC Dashboard",
  subtitle = "Defect rate vs target, capacity utilization, and 24-hour yield trend"
)
```

---

## Sports Analytics: NBA Player Efficiency

Non-clinical use case — forest plots work for any interval data. Here player efficiency ratings with 95% CIs are grouped by conference, with team/position column groups and award badges.

```{r}
#| fig-height: 8
data(nba_efficiency)

tabviz(
  nba_efficiency,
  label = "player",
  group = "conference",
  columns = list(
    col_group("Player Info",
      col_text("team", "Team", width = 50),
      col_text("position", "Pos", width = 40)
    ),
    col_group("Stats",
      col_numeric("ppg", "PPG"),
      col_numeric("games", "GP", decimals = 0)
    ),
    viz_forest(point = "per", lower = "per_lower", upper = "per_upper",
               null_value = 15,
               axis_label = "Player Efficiency Rating"),
    col_interval("per", "per_lower", "per_upper", header = "PER (95% CI)")
  ),
  row_badge = "award",
  theme = web_theme_modern() |>
    set_inputs(brand = "#C9082A"),  # NBA red
  title = "NBA Player Efficiency Ratings",
  subtitle = "Row groups by conference + column groups for Player Info / Stats",
  caption = "PER = 15 is league average. Badge shows major awards.",
  footnote = "Data simulated based on typical NBA statistics"
)
```

---

## Climate Science: Regional Temperature Anomalies

Environmental data example using the `climate_temps` dataset — the same interval-estimate structure used in clinical forest plots maps cleanly onto temperature anomalies vs. a pre-industrial baseline.

```{r}
data(climate_temps)

tabviz(
  climate_temps,
  label = "region",
  group = "category",
  columns = list(
    col_text("period", "Period"),
    col_text("certainty", "Confidence"),
    col_interval("anomaly", "lower", "upper", header = "Anomaly °C (95% CI)"),
    viz_forest(point = "anomaly", lower = "lower", upper = "upper",
               null_value = 0, axis_label = "Temperature Anomaly (°C vs. pre-industrial)")
  ),
  marker_shape = ~ ifelse(certainty == "High", "circle", "square"),
  marker_color = ~ ifelse(anomaly > 1.5, "#dc2626", ifelse(anomaly > 1.0, "#f59e0b", "#3b82f6")),
  theme = web_theme_modern() |>
    set_theme_field(c("axis", "gridlines"), TRUE),
  title = "Regional Temperature Anomalies",
  subtitle = "Grouped by geographic category, shaped by confidence level",
  caption = "Red = significant warming, Blue = cooling or below baseline"
)
```
