---
title: "Dashboard Examples"
---

```{r}
#| include: false
library(tabviz)
library(dplyr)
```

Business dashboards, KPI displays, and data tables with embedded visualizations.

## NBA Player Efficiency

Performance dashboard using Player Efficiency Rating (PER):

```{r}
data(nba_efficiency)

tabviz(
  nba_efficiency,
  label = "player",
  columns = list(
    col_text("team", "Team"),
    col_text("position", "Pos"),
    col_numeric("ppg", "PPG", decimals = 1),
    col_numeric("games", "GP"),
    viz_forest(point = "per", lower = "per_lower", upper = "per_upper",
               scale = "linear", null_value = 15,
               axis_label = "PER"),
    col_interval("PER (95% CI)", point = "per", lower = "per_lower", upper = "per_upper"),
    col_badge("award", "Award")
  ),
  row_bold = ~ per > 25,
  theme = web_theme_modern(),
  title = "Player Efficiency Rating Dashboard"
)
```

## Climate Temperature Anomalies

Scientific data showing temperature anomaly trends:

```{r}
data(climate_temps)

tabviz(
  climate_temps,
  label = "region",
  columns = list(
    col_text("period", "Period"),
    col_numeric("anomaly", "Anomaly (°C)", decimals = 2),
    viz_forest(point = "anomaly", lower = "lower", upper = "upper",
               scale = "linear", null_value = 0,
               axis_label = "Temperature Anomaly"),
    col_interval("95% CI", point = "anomaly", lower = "lower", upper = "upper"),
    col_text("certainty", "Confidence"),
    col_badge("category", "Type")
  ),
  row_color = ~ ifelse(anomaly > 1.5, "#dc2626", "#64748b"),
  theme = web_theme_minimal(),
  title = "Global Temperature Anomalies"
)
```

## Airline Performance

Airline delay metrics analysis:

```{r}
data(airline_delays)

# Show one row per carrier (first month)
airline_summary <- airline_delays |>
  filter(month == "Apr") |>
  select(carrier, carrier_type, on_time_pct, delay_vs_avg, satisfaction, flights, trend)

tabviz(
  airline_summary,
  label = "carrier",
  columns = list(
    col_text("carrier_type", "Type"),
    col_numeric("flights", "Flights", abbreviate = TRUE),
    col_percent("on_time_pct", "On-Time"),
    col_numeric("delay_vs_avg", "Delay vs Avg", decimals = 1),
    col_sparkline("trend", "12-Month Trend"),
    col_numeric("satisfaction", "Rating", decimals = 1)
  ),
  row_bg = ~ case_when(
    on_time_pct > 85 ~ "#dcfce7",
    on_time_pct < 75 ~ "#fee2e2",
    TRUE ~ NA_character_
  ),
  theme = web_theme_modern(),
  title = "Airline Performance Dashboard"
)
```

## Executive Summary Table

Minimal table for C-suite presentations:

```{r}
exec_data <- data.frame(
  metric = c("Revenue", "Profit", "Customers", "Retention", "NPS"),
  current = c("$142M", "$28M", "45.2K", "92%", "72"),
  change = c(12.5, 8.3, 15.2, 2.1, 5),
  trend = I(list(
    c(120, 125, 130, 128, 135, 142),
    c(22, 24, 23, 26, 27, 28),
    c(38, 40, 41, 43, 44, 45.2),
    c(88, 89, 90, 91, 91, 92),
    c(65, 67, 68, 70, 71, 72)
  )),
  status = c("above", "above", "above", "on_track", "above")
)

tabviz(exec_data,
  label = "metric",
  columns = list(
    col_text("current", "Current"),
    col_percent("change", "YoY Change"),
    col_sparkline("trend", "6-Month Trend"),
    col_badge("status", "Status")
  ),
  row_color = ~ ifelse(change > 10, "#16a34a", "#64748b"),
  theme = web_theme_presentation(),
  title = "Q4 Executive Summary"
)
```

## Dark Mode Dashboard

Modern dark theme for monitoring displays:

```{r}
tabviz(
  nba_efficiency[1:8, ],
  label = "player",
  columns = list(
    col_text("team", "Team"),
    col_numeric("ppg", "PPG"),
    viz_forest(point = "per", lower = "per_lower", upper = "per_upper",
               scale = "linear", null_value = 15),
    col_interval("PER", point = "per", lower = "per_lower", upper = "per_upper"),
    col_badge("award", "Award")
  ),
  theme = web_theme_dark(),
  title = "Live Performance Monitor"
)
```

## Comparison Table

Side-by-side product/feature comparison:

```{r}
comparison_data <- data.frame(
  feature = c("Price", "Storage", "CPU Cores", "Memory", "Support", "Rating"),
  plan_a = c("$99/mo", "100 GB", "4", "16 GB", "Email", "4.2"),
  plan_b = c("$199/mo", "500 GB", "8", "32 GB", "24/7 Phone", "4.7"),
  plan_c = c("$399/mo", "2 TB", "16", "64 GB", "Dedicated", "4.9"),
  recommendation = c(NA, NA, "Best Value", NA, NA, "Highest Rated")
)

tabviz(comparison_data,
  label = "feature",
  columns = list(
    col_text("plan_a", "Basic"),
    col_text("plan_b", "Pro"),
    col_text("plan_c", "Enterprise"),
    col_badge("recommendation", "Note")
  ),
  theme = web_theme_minimal(),
  title = "Plan Comparison"
)
```
