---
title: "Genetic Association (R + TS)"
description: "The same forest plot, two ways — both rendered live, from R and from TypeScript"
---

```{r}
#| include: false
library(tabviz)
```

When `tabviz` was just an R package, "the JS code" meant "the bundled htmlwidget runtime" — opaque to consumers, only useful as a black box behind the R API. As of 0.2.0, the JS side is a first-class authoring target: the same `tabviz()` / `col_*()` / `viz_*()` / `theme_*()` API mirrored in TypeScript, available from npm as [`@tabviz/core`](https://www.npmjs.com/package/@tabviz/core).

This page renders the same forest plot **twice on the same page** — once via R, once via TypeScript loaded over the wire from [esm.sh](https://esm.sh). The pixels should match.

## Data: genetic associations with cardiovascular events

A small GWAS-style result table — SNP-level effect estimates from a polygenic-risk-score study. Rows are SNPs; columns carry the locus name, allele frequency, effect size + CI, sample size, and a p-value. A forest plot on the right shows the effect estimates aligned with the table.

```{r}
#| cache: false
snps <- data.frame(
  snp         = c("rs7412",   "rs429358",  "rs2230806", "rs5174",   "rs1801133",
                  "rs5882",   "rs7041",    "rs662",     "rs6025",   "rs1799963"),
  gene        = c("APOE",     "APOE",      "ABCA1",     "LRP1",     "MTHFR",
                  "CETP",     "GC",        "PON1",      "F5",       "F2"),
  maf         = c(0.07,       0.14,        0.22,        0.41,       0.30,
                  0.35,       0.40,        0.31,        0.04,       0.02),
  effect      = c(-0.18,      0.28,        -0.11,       0.07,       0.09,
                  -0.13,      0.05,        0.10,        0.34,       0.41),
  lower       = c(-0.27,      0.19,        -0.18,       0.02,       0.02,
                  -0.20,     -0.01,       0.03,        0.18,        0.21),
  upper       = c(-0.09,      0.37,        -0.04,       0.12,       0.16,
                  -0.06,      0.11,       0.17,        0.50,        0.61),
  n_studies   = c(34,         38,          12,          9,          21,
                  17,         8,           14,          22,          11),
  p_value     = c(3.2e-5,     1.4e-9,      4.7e-3,      0.018,      0.014,
                  2.6e-4,     0.067,       7.3e-3,      1.1e-5,      2.1e-5)
)

# Expose the data to the OJS runtime below.
ojs_define(snps_in = snps)
```

## R: `tabviz(...)` rendering the live widget

```{r}
#| fig-width: 11
#| fig-height: 5
tabviz(
  snps,
  label = "snp",
  label_header = "SNP",
  theme = web_theme_lancet(),
  columns = list(
    col_text("gene", header = "Gene"),
    col_numeric("maf", header = "MAF", decimals = 2),
    col_interval("effect", "lower", "upper", header = "Effect (95% CI)"),
    col_n("n_studies", header = "Studies"),
    col_pvalue("p_value"),
    viz_forest(
      point = "effect", lower = "lower", upper = "upper",
      axis_label = "Effect estimate",
      scale = "linear",
      null_value = 0
    )
  ),
  title = "Genetic associations with cardiovascular events",
  caption = "Effect = log odds ratio per allele; CI = 95%."
)
```

## TypeScript via `@tabviz/core`, rendered in the browser

Below is the same plot built from the TypeScript-mirror of the R API and mounted by `createTabviz` — executed live in this page (no R involved). The package is loaded from esm.sh; Svelte's optional peer dep is resolved transitively.

```{ojs}
//| echo: false
// Pin Svelte to the version @tabviz/core@0.2.1 was compiled against —
// esm.sh's default resolution of the optional peer picks the latest 5.x,
// which can differ in internal helpers (`validate_each_keys` etc.) and
// throws at mount. Lockstep version pin is the simplest workaround.
core = import("https://esm.sh/@tabviz/core@0.2.1?deps=svelte@5.46.1")

// Quarto exposes `snps_in` as a column-oriented table; transpose to rows.
rows = transpose(snps_in)

styleHref = "https://esm.sh/@tabviz/core@0.2.1/dist/style.css"
// Inject the @tabviz/core stylesheet once per page.
_styleInjected = {
  if (!document.querySelector(`link[href="${styleHref}"]`)) {
    const link = document.createElement("link");
    link.rel = "stylesheet";
    link.href = styleHref;
    document.head.appendChild(link);
  }
  return true;
}

tsWidget = {
  const { tabviz, createTabviz,
          colText, colNumeric, colInterval, colN, colPvalue,
          vizForest } = await core;
  _styleInjected;

  const spec = tabviz({
    data: rows,
    label: "snp",
    labelHeader: "SNP",
    theme: "lancet",
    columns: [
      colText({ field: "gene", header: "Gene" }),
      colNumeric({ field: "maf", header: "MAF", decimals: 2 }),
      colInterval({
        point: "effect", lower: "lower", upper: "upper",
        header: "Effect (95% CI)",
      }),
      colN({ field: "n_studies", header: "Studies" }),
      colPvalue({ field: "p_value" }),
      vizForest({
        point: "effect", lower: "lower", upper: "upper",
        axisLabel: "Effect estimate",
        scale: "linear",
        nullValue: 0,
      }),
    ],
    title: "Genetic associations with cardiovascular events",
    caption: "Effect = log odds ratio per allele; CI = 95%.",
  });

  const host = document.createElement("div");
  host.style.minHeight = "320px";
  createTabviz(host, spec);
  return host;
}
```

The source of that `ojs` block, in plain TypeScript:

::: {.callout-note collapse="true"}
## The TypeScript snippet (for use in a Vite/Next.js project)

```ts
import {
  tabviz, createTabviz,
  colText, colNumeric, colInterval, colN, colPvalue,
  vizForest,
} from "@tabviz/core";
import "@tabviz/core/style.css";

// `snpsData` is your row array — same shape as the R `snps` data.frame.
declare const snpsData: Array<{
  snp: string; gene: string; maf: number;
  effect: number; lower: number; upper: number;
  n_studies: number; p_value: number;
}>;

const spec = tabviz({
  data: snpsData,
  label: "snp",
  labelHeader: "SNP",
  theme: "lancet",
  columns: [
    colText({ field: "gene", header: "Gene" }),
    colNumeric({ field: "maf", header: "MAF", decimals: 2 }),
    colInterval({
      point: "effect", lower: "lower", upper: "upper",
      header: "Effect (95% CI)",
    }),
    colN({ field: "n_studies", header: "Studies" }),
    colPvalue({ field: "p_value" }),
    vizForest({
      point: "effect", lower: "lower", upper: "upper",
      axisLabel: "Effect estimate",
      scale: "linear",
      nullValue: 0,
    }),
  ],
  title: "Genetic associations with cardiovascular events",
  caption: "Effect = log odds ratio per allele; CI = 95%.",
});

const instance = createTabviz(document.querySelector("#plot")!, spec);
```
:::

## What's the same, what's different?

The argument names, defaults, and shape of the returned spec are identical:

- R: `col_interval("effect", "lower", "upper")` ↔ TS: `colInterval({ point: "effect", lower: "lower", upper: "upper" })`
- R: snake_case kwargs ↔ TS: camelCase keys
- R: `theme = web_theme_lancet()` ↔ TS: `theme: "lancet"` (or `theme: themeLancet()`)

R is positional-friendly (`col_*` helpers take `field` as the first positional arg); TS is purely named-argument because it's a single object literal. Beyond that, the helpers are mechanical 1:1 translations — change one to fit the other by remembering snake-case ↔ camelCase and keyword-vs-positional.
