---
title: "Gallery"
---

Explore example visualizations organized by domain. Each example includes live interactive output and source code.

## By Domain

::: {.grid}
::: {.g-col-4}
### [Clinical Examples](clinical.qmd)

Meta-analyses, forest plots, and clinical trial visualizations.

- JAMA-style forest plots
- Subgroup analyses
- Multi-effect comparisons
:::

::: {.g-col-4}
### [Dashboard Examples](dashboards.qmd)

Business dashboards and KPI displays.

- Executive dashboards
- Sparklines and bars
- Status indicators
:::

::: {.g-col-4}
### [Feature Showcases](showcases.qmd)

Demonstrations of specific tabviz features.

- Row grouping
- Themes
- Split views
:::
:::

## Quick Examples

### Clinical Trial Meta-Analysis

```{r}
#| echo: false
library(tabviz)
data(glp1_trials)

tabviz(
  glp1_trials,
  label = "study",
  group = "group",
  columns = list(
    col_text("drug", "Drug"),
    col_n("n"),
    col_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1,
               axis_label = "Hazard Ratio"),
    col_interval("HR (95% CI)", point = "hr", lower = "lower", upper = "upper"),
    col_pvalue("pvalue", "P")
  ),
  row_type = "row_type",
  row_bold = "row_bold",
  theme = web_theme_nature()
)
```

### Dashboard Table

```{r}
#| echo: false
library(tabviz)
data(nba_efficiency)

tabviz(
  nba_efficiency,
  label = "player",
  columns = list(
    col_text("team", "Team"),
    col_numeric("ppg", "PPG", decimals = 1),
    col_percent("fg_pct", "FG%"),
    col_sparkline("last_10", "Last 10"),
    col_bar("efficiency", "Efficiency")
  ),
  theme = web_theme_modern()
)
```

## Example Sources

All examples are available as R scripts in the package:

```r
# List available examples
list.files(system.file("examples", package = "tabviz"))

# Run an example
source(system.file("examples", "jama_style.R", package = "tabviz"))
```
