---
title: "Gallery"
---

```{r}
#| include: false
library(tabviz)
library(dplyr)
```

Explore example visualizations organized by use case. Each example below is a live widget — sort columns, click filter funnels, drag rows or columns where enabled, double-click to edit.

## By Category

::: {.grid}
::: {.g-col-6 .g-col-md-3}
### [Interactivity](interactive.qmd)

Sort, filter, drag-and-drop, and double-click editing — with WYSIWYG export.

- Full-interactivity tour
- Filter + sort dashboard
- In-place forest cell editing
:::

::: {.g-col-6 .g-col-md-3}
### [Clinical Examples](clinical.qmd)

Meta-analyses, forest plots, and clinical trial visualizations.

- JAMA-style forest plots
- Subgroup analyses
- Multi-arm dose-response
:::

::: {.g-col-6 .g-col-md-3}
### [Dashboard Examples](dashboards.qmd)

Business dashboards, KPIs, and data tables.

- Product metrics
- A/B test analysis
- Executive summaries
:::

::: {.g-col-6 .g-col-md-3}
### [Feature Showcases](showcases.qmd)

Demonstrations of specific tabviz features.

- Rich column types
- Distribution plots
- Theme customization
:::
:::

---

## Row Groups × Column Groups

One widget, two axes of grouping: rows grouped by treatment arm (collapsible), columns grouped by purpose ("Demographics", "Efficacy"). Drag rows inside a treatment group to reorder; drag columns inside a column group without disturbing the other group.

```{r}
data(effect_sizes)

tabviz(effect_sizes,
  label = "study",
  group = "treatment",
  columns = list(
    col_group("Demographics",
      col_text("region", "Region"),
      col_text("phase", "Phase"),
      col_n("n", "N")
    ),
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1,
               axis_label = "Hazard Ratio"),
    col_group("Efficacy",
      col_interval("HR (95% CI)", point = "hr", lower = "lower", upper = "upper"),
      col_badge("direction", "Direction",
        variants = list(
          `Favors Treatment` = "success",
          `Favors Control` = "error",
          `No Effect` = "muted"
        )),
      col_badge("significant", "Signif",
        variants = list(`TRUE` = "success", `FALSE` = "muted"))
    )
  ),
  interaction = web_interaction_full(),
  theme = web_theme_default(),
  title = "Row groups and column groups on the same plot"
)
```

---

## Project Portfolio Dashboard

A data-rich dashboard combining heatmaps, progress bars, currency, dates, badges, and sparklines — no forest plot needed. Full interactivity is on; try filtering the "Status" column.

```{r}
portfolio <- data.frame(
  project = c("Cloud Migration", "Mobile App v3", "Data Platform",
               "Security Audit", "API Gateway", "ML Pipeline"),
  status = c("On Track", "At Risk", "Completed", "On Track", "Delayed", "On Track"),
  budget = c(2400000, 890000, 1250000, 340000, 720000, 1800000),
  spent_pct = c(0.62, 0.88, 1.00, 0.45, 0.71, 0.33),
  health = c(92, 58, 100, 85, 42, 78),
  due_date = as.Date(c("2025-12-31", "2025-09-30", "2025-06-15",
                        "2025-08-31", "2025-11-15", "2026-03-31")),
  team_size = c(12, 8, 6, 4, 9, 7),
  velocity = I(list(
    c(34, 38, 42, 45, 48, 52, 55, 58, 60, 62, 65, 68),
    c(28, 32, 30, 25, 22, 18, 20, 15, 12, 14, 16, 18),
    c(40, 45, 50, 55, 60, 65, 70, 75, 80, 85, 90, 95),
    c(15, 18, 20, 22, 25, 28, 30, 32, 35, 38, 40, 42),
    c(30, 28, 25, 22, 20, 18, 22, 24, 20, 18, 15, 12),
    c(10, 15, 20, 25, 28, 32, 35, 38, 42, 45, 48, 50)
  ))
)

tabviz(portfolio, label = "project",
  columns = list(
    col_badge("status", "Status",
      variants = list(
        `On Track` = "success",
        `At Risk` = "warning",
        Completed = "info",
        Delayed = "error"
      )),
    col_currency("budget", "Budget", decimals = 0),
    col_progress("spent_pct", "Budget Used", max_value = 1),
    col_heatmap("health", "Health Score",
      palette = c("#fee2e2", "#dcfce7"),
      min_value = 0, max_value = 100, decimals = 0),
    col_date("due_date", "Due Date", format = "%b %d, %Y"),
    col_n("team_size", "Team"),
    col_sparkline("velocity", "Velocity (12w)")
  ),
  interaction = web_interaction_full(),
  theme = web_theme_modern(),
  title = "Project Portfolio Dashboard"
)
```

---

## Example Sources

All examples are available as R scripts in the package:

```r
# List available examples
list.files(system.file("examples", package = "tabviz"))

# Run an example
source(system.file("examples", "jama_style.R", package = "tabviz"))
```
