---
title: "Interactivity Showcase"
---

```{r}
#| include: false
library(tabviz)
library(dplyr)
```

Every widget below is live. Sort, filter, drag, edit — any change flows through to the
SVG/PNG export at the click of the download button (WYSIWYG).

```{r}
#| include: false
# Convenience: a one-liner that enables everything
full <- web_interaction_full
```

## The whole kitchen sink

Sort, filter, column reorder, row reorder, double-click edit, hover tooltips — one plot, every feature on.

```{r}
data(glp1_trials)

tabviz(
  glp1_trials,
  label = "study",
  group = "group",
  columns = list(
    col_group("Study Info",
      col_text("drug", "Drug"),
      col_n("year", "Year"),
      col_n("n"),
      col_events("events", "n", "Events")
    ),
    viz_forest(
      point = "hr", lower = "lower", upper = "upper",
      scale = "log", null_value = 1,
      axis_label = "Hazard Ratio (95% CI)",
      axis_range = c(0.4, 1.5),
      axis_ticks = c(0.5, 0.75, 1.0, 1.25),
      axis_gridlines = TRUE,
      annotations = list(
        refline(0.85, label = "Pooled HR", style = "dashed", color = "#00407a")
      )
    ),
    col_group("Results",
      col_interval("HR (95% CI)", point = "hr", lower = "lower", upper = "upper"),
      col_pvalue("pvalue", "P", stars = TRUE)
    )
  ),
  row_type = "row_type",
  row_bold = "row_bold",
  interaction = web_interaction_full(),
  theme = web_theme_nature(),
  title = "GLP-1 Agonist Cardiovascular Outcomes",
  subtitle = "All interactivity on — try everything"
)
```

::: {.callout-tip}
## How to try each feature
- **Sort** → click a column header. Click again for descending, again to clear.
- **Filter** → click the funnel icon next to a header. You'll get a text / numeric / categorical UI depending on the column.
- **Reorder rows** → hover a row, grab the drag handle on the left, drop within the same group.
- **Reorder columns** → hover a header, grab its drag handle. Columns inside "Study Info" stay inside "Study Info".
- **Edit** → double-click an editable cell (text, numeric, percent, p-value, date, interval) to edit inline, or double-click a column-group header to rename the span.
- **Export** → download SVG or PNG; the file reflects your current state exactly.
:::

---

## Filter + sort on a dashboard

Large-ish dashboard where filters really pay off — apply any combination of filters and the export matches the filtered view.

```{r}
data(airline_delays)

tabviz(airline_delays, label = "carrier",
  group = "month",
  columns = list(
    col_group("Carrier",
      col_badge("carrier_type", "Type",
        variants = list(
          Legacy = "info",
          `Low-Cost` = "success",
          `Ultra Low-Cost` = "warning"
        )),
      col_n("flights", "Flights", abbreviate = TRUE)
    ),
    col_group("Performance",
      col_heatmap("on_time_pct", "On-Time",
        palette = c("#fee2e2", "#dcfce7"),
        min_value = 0.65, max_value = 0.95,
        decimals = 2, show_value = TRUE),
      col_progress("satisfaction", "Satisfaction", max_value = 5),
      col_numeric("delay_vs_avg", "Delay", decimals = 1)
    ),
    viz_forest(point = "delay_vs_avg", lower = "delay_lower", upper = "delay_upper",
               null_value = 0,
               axis_label = "Delay vs avg (min)"),
    col_sparkline("trend", "12-month trend")
  ),
  interaction = web_interaction_full(),
  theme = web_theme_modern(),
  title = "Airline Performance by Month",
  subtitle = "Try filtering by carrier_type or on_time_pct; export matches the view"
)
```

---

## Drag to reorder within column groups

Columns inside "Baseline" and "Efficacy" swap cleanly; the forest column stays where it is. Whole column groups can also be dragged among top-level siblings.

```{r}
set.seed(42)
n_studies <- 8
trial_data <- data.frame(
  study = paste0("Trial-", LETTERS[1:n_studies]),
  arm = rep(c("Active", "Control"), length.out = n_studies),
  n_baseline = sample(200:800, n_studies),
  age_mean = round(runif(n_studies, 55, 72), 1),
  female_pct = round(runif(n_studies, 0.35, 0.65), 2),
  response_rate = round(runif(n_studies, 0.3, 0.8), 2),
  median_duration_months = round(runif(n_studies, 8, 24), 1),
  hr = round(runif(n_studies, 0.55, 1.05), 2)
) |>
  mutate(
    lo = pmax(0.3, hr - 0.2),
    hi = pmin(1.4, hr + 0.2)
  )

tabviz(trial_data, label = "study",
  columns = list(
    col_badge("arm", "Arm",
      variants = list(Active = "success", Control = "muted")),
    col_group("Baseline",
      col_n("n_baseline", "N"),
      col_numeric("age_mean", "Age"),
      col_percent("female_pct", "Female")
    ),
    col_group("Efficacy",
      col_percent("response_rate", "Response"),
      col_numeric("median_duration_months", "Dur (mo)", decimals = 1)
    ),
    viz_forest(point = "hr", lower = "lo", upper = "hi",
               scale = "log", null_value = 1,
               axis_label = "HR"),
    col_interval("HR (95% CI)", point = "hr", lower = "lo", upper = "hi")
  ),
  interaction = web_interaction_full(),
  theme = web_theme_cochrane(),
  title = "Reorder within column groups",
  subtitle = "Try dragging columns inside \"Baseline\" — they stay inside"
)
```

---

## Edit forest cell values in place

Double-click any marker to pop up the estimate / lower / upper editor. The marker moves, and the export picks up the edit.

```{r}
set.seed(11)
effects <- data.frame(
  metric = c("Mortality", "MI", "Stroke", "HF Hospitalization",
             "CV Death", "All-cause Death"),
  hr = c(0.74, 0.81, 0.92, 0.68, 0.71, 0.85),
  lo = c(0.62, 0.68, 0.77, 0.54, 0.58, 0.72),
  hi = c(0.88, 0.95, 1.10, 0.85, 0.87, 0.99),
  events = c(412, 289, 174, 356, 298, 512),
  n = c(12800, 12800, 12800, 12800, 12800, 12800),
  pvalue = c(0.001, 0.02, 0.22, 0.0004, 0.002, 0.04)
)

tabviz(effects, label = "metric",
  columns = list(
    col_n("n", "N"),
    col_events("events", "n", "Events"),
    viz_forest(point = "hr", lower = "lo", upper = "hi",
               scale = "log", null_value = 1,
               axis_range = c(0.4, 1.4),
               axis_label = "Hazard Ratio"),
    col_interval("HR (95% CI)", point = "hr", lower = "lo", upper = "hi"),
    col_pvalue("pvalue", "P", stars = TRUE)
  ),
  interaction = web_interaction_full(),
  theme = web_theme_jama(),
  title = "Double-click a forest marker to edit",
  subtitle = "Estimate / Lower / Upper popover; export reflects edits"
)
```

---

## Proxy-Driven Updates

Inside a Shiny app, `tabviz_proxy()` exposes a fluent verb API — `add_column()`, `sort_rows()`, `filter_rows()`, `set_theme()`, `set_zoom()`, and more — that mutates the live widget without triggering a full re-render. The snippet below is excerpted from [`inst/examples/fluent_shiny_demo.R`](https://github.com/antoine-sachet/forest/blob/main/inst/examples/fluent_shiny_demo.R); Shiny apps can't execute inside Quarto, so the code is shown for reference only.

```{r}
#| eval: false
library(shiny)
library(tabviz)

trial_data <- data.frame(
  study = c("EMPA-REG", "CANVAS", "DECLARE", "CREDENCE", "LEADER", "SUSTAIN-6"),
  hr    = c(0.86, 0.86, 0.93, 0.80, 0.87, 0.74),
  lower = c(0.74, 0.75, 0.84, 0.67, 0.78, 0.58),
  upper = c(0.99, 0.97, 1.03, 0.95, 0.97, 0.95),
  n     = c(7020, 10142, 17160, 4401, 9340, 3297),
  year  = c(2015, 2017, 2019, 2019, 2016, 2016)
)

ui <- fluidPage(
  fluidRow(
    column(3,
      actionButton("sort_desc",    "Sort HR desc"),
      actionButton("filter_hr",    "Filter HR < 0.9"),
      actionButton("clear_flt",    "Clear filter"),
      actionButton("theme_jama",   "Theme: JAMA"),
      actionButton("theme_lancet", "Theme: Lancet"),
      actionButton("add_col",      "Add 'year' column"),
      actionButton("rename_hr",    "Rename HR to 'Hazard'")
    ),
    column(9, tabvizOutput("plot", height = "500px"))
  )
)

server <- function(input, output, session) {
  output$plot <- renderTabviz({
    tabviz(trial_data, label = "study",
      columns = list(
        col_numeric("hr", "HR", decimals = 2),
        col_numeric("n",  "N"),
        viz_forest(point = "hr", lower = "lower", upper = "upper",
                   header = "HR (95% CI)")
      ),
      null_value = 1, scale = "log"
    )
  })

  proxy <- tabviz_proxy("plot")

  # Each verb sends a custom message to the live widget — no re-render.
  observeEvent(input$sort_desc,    sort_rows(proxy, "hr", direction = "desc"))
  observeEvent(input$filter_hr,    filter_rows(proxy, "hr", operator = "lt", value = 0.9))
  observeEvent(input$clear_flt,    clear_filters(proxy))
  observeEvent(input$theme_jama,   set_theme(proxy, "jama"))
  observeEvent(input$theme_lancet, set_theme(proxy, "lancet"))
  observeEvent(input$add_col,      add_column(proxy, col_numeric("year", "Year"), after = "n"))
  observeEvent(input$rename_hr,    update_column(proxy, "hr", header = "Hazard"))
}

shinyApp(ui, server)
```

What the demo shows:

- **Sort / filter / clear** → row operations applied client-side via the proxy; the `${id}_filtered_rows` input updates so the server knows which rows are visible.
- **Theme switch** → swaps colors, typography, and spacing in place without losing scroll or edit state.
- **Add / rename / resize columns** → structural mutations with no full render.
- **Zoom** → live scale change (0.8, 1.0, ...) on the whole widget.

The same verbs (`add_column`, `sort_rows`, `set_theme`, ...) work on static `WebSpec` and htmlwidget objects too; on a proxy they dispatch a custom message, otherwise they mutate the spec in place.

To run locally:

```r
shiny::runApp(system.file("examples/fluent_shiny_demo.R", package = "tabviz"))
```

---

## Enable only what you need

`web_interaction()` takes any combination of flags. Use `web_interaction_full()` for everything on, `web_interaction_minimal()` for a read-only widget, `web_interaction_publication()` for no UI chrome at all, or dial it in yourself:

```r
web_interaction(
  enable_sort          = TRUE,
  enable_filters       = TRUE,
  enable_reorder_rows  = TRUE,   # drag rows within their group
  enable_reorder_columns = FALSE,
  enable_edit          = TRUE,   # double-click to edit
  enable_select        = TRUE,
  enable_hover         = TRUE,
  enable_resize        = TRUE,
  enable_export        = TRUE
)
```

All reorder / edit state lives in the browser session. SVG/PNG exports capture the current view, but a page reload returns to the server-supplied data. For Shiny round-trips, the `${id}_selected` and `${id}_hover` inputs fire on selection/hover changes, and the `tabviz-proxy` message handler accepts `applyFilter`, `clearFilter`, `sortBy`, and `toggleGroup` methods.
