---
title: "Long Meta-Analysis"
description: "Many studies, three cohorts, pooled summaries, and a paginated PDF"
---

```{r}
#| include: false
library(tabviz)
library(dplyr)
```

A real meta-analysis often has 50–100+ studies, several cohorts, pooled estimates per cohort, and an overall summary. That's the case `tabviz()` was built for: declarative columns, semantic row tokens, and pagination that lets the same spec render as a paginated HTML viewer **and** as a multi-page PDF without rewriting anything.

This page builds the canonical meta-analysis display step by step. We end up with a 70-row paginated forest plot that breaks across three pages by cohort.

## A 70-study cardiovascular meta-analysis

Three cohorts, each with its own pooled summary plus an overall pooled diamond:

```{r}
set.seed(2026)

# Per-cohort row counts; each cohort gets one pooled summary at the end.
cohorts <- c("Phase III RCT", "Real-world evidence", "Registry")
counts  <- c(22, 26, 18)
n_data  <- sum(counts)

# Synthetic data — designed so each cohort lands in the budget when we
# paginate at 30 rows per page later.
trials <- data.frame(
  study  = c(sprintf("Trial %02d", seq_len(counts[1])),
             sprintf("RWE %02d",   seq_len(counts[2])),
             sprintf("Reg %02d",   seq_len(counts[3]))),
  cohort = rep(cohorts, times = counts),
  year   = sample(2010:2024, n_data, replace = TRUE),
  n      = sample(c(120, 240, 480, 960, 1900, 3800), n_data, replace = TRUE),
  hr     = round(rnorm(n_data, 0.92, 0.18), 2),
  se     = round(runif(n_data, 0.06, 0.20), 2)
)
trials$lower <- pmax(0.30, round(trials$hr - 1.96 * trials$se, 2))
trials$upper <- round(trials$hr + 1.96 * trials$se, 2)
trials$row_type <- "data"

# Pooled summary per cohort (diamond marker via row_type = "summary").
pooled <- trials |>
  group_by(cohort) |>
  summarize(
    study  = paste0("Pooled — ", dplyr::first(cohort)),
    year   = NA_integer_,
    n      = sum(n),
    hr     = round(exp(mean(log(hr))), 2),
    se     = NA_real_,
    lower  = round(exp(mean(log(hr)) - 1.96 * 0.04), 2),
    upper  = round(exp(mean(log(hr)) + 1.96 * 0.04), 2),
    row_type = "summary",
    .groups = "drop"
  )

# Overall pooled diamond (one row at the very end of each cohort? overall here)
overall <- data.frame(
  study  = "Overall",
  cohort = "Registry",                       # last cohort in display order
  year   = NA_integer_,
  n      = sum(trials$n),
  hr     = round(exp(mean(log(trials$hr))), 2),
  se     = NA_real_,
  lower  = round(exp(mean(log(trials$hr)) - 1.96 * 0.03), 2),
  upper  = round(exp(mean(log(trials$hr)) + 1.96 * 0.03), 2),
  row_type = "summary"
)

# Assemble in display order: each cohort's data → its pooled → next cohort,
# with the overall pooled at the very bottom.
meta <- bind_rows(
  trials  |> filter(cohort == cohorts[1]),
  pooled  |> filter(cohort == cohorts[1]),
  trials  |> filter(cohort == cohorts[2]),
  pooled  |> filter(cohort == cohorts[2]),
  trials  |> filter(cohort == cohorts[3]),
  pooled  |> filter(cohort == cohorts[3]),
  overall
) |>
  mutate(
    is_significant = !is.na(upper) & upper < 1,   # entire CI below 1
    is_overall     = study == "Overall"
  )
```

```{r}
tabviz(
  meta,
  label = "study",
  group = "cohort",
  columns = list(
    col_numeric("year", "Year", decimals = 0),
    col_n("n"),
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1,
               axis_label = "Hazard ratio (95% CI)"),
    col_interval("hr", "lower", "upper", header = "HR (95% CI)")
  ),
  row_type   = "row_type",
  row_accent = "is_significant",      # bold + theme accent on significant rows
  row_fill   = "is_overall",          # strong fill on the overall pooled row
  theme = web_theme_jama(),
  title    = "GLP-1 receptor agonists — pooled cardiovascular meta-analysis",
  subtitle = "70 studies across three cohorts, with cohort-level and overall pooled estimates",
  caption  = "Significant studies (CI strictly below 1) are highlighted with the accent token.",
  paginate = paginate_letter()
)
```

::: {.callout-tip}
## What's happening on this plot

- **Three collapsible cohorts** via `group = "cohort"` — click any header to fold a section.
- **Pooled diamonds** at the bottom of each cohort and at the very end of the table, driven by `row_type = "summary"` (no diamond on data rows; pooled rows render as diamonds via the marker dispatch).
- **`row_accent`** on the rows whose CI is strictly below 1 — bold + theme accent color. Themable: changing the theme restyles the highlight automatically.
- **`row_fill`** on the single Overall pooled row — strong row-fill background with auto-contrasting foreground. The fifth semantic token in the `RowCluster` set; use it when a row needs to *be* the conclusion, not just stand out.
- **Pagination at 30 rows/page** via `paginate_letter()`. Page count appears below the plot ("Page 1 of 3"); use the prev / next buttons or toggle continuous mode for scrolling.
- **Group integrity** keeps each cohort intact across page breaks — `keep_groups = TRUE` is the default, so a cohort either fits on the current page or moves to the next.
:::

## Saving as a multi-page PDF

The same spec emits a multi-page PDF where each logical page becomes one PDF page:

```r
save_plot(meta_table, "meta-analysis.pdf", width = 800)
# Multi-page PDF merged via qpdf::pdf_combine()
```

For the rendering pipeline (V8 → SVG → rsvg → qpdf), the qpdf-missing fallback (numbered series), and the per-render override surface, see [Export → Pagination](../guide/export.qmd#pagination).

## Adjusting the pagination for slides

The same spec at slide density (16 rows per page) — useful when each cohort needs its own slide deck:

```{r}
#| eval: false
tabviz(meta, ..., paginate = paginate_slide())   # 16 rows / page
```

Or fold pagination off for a single-image PNG without warning:

```{r}
#| eval: false
save_plot(meta_table, "wall.png", paginate = NULL)
```

## See also

- [Export → Pagination](../guide/export.qmd#pagination) — full surface and PDF export details
- [Concepts → Data Flow](../concepts/data-flow.qmd) — how `paginate.pages` is computed once R-side and shipped on the wire
- [Reference → `paginate_spec()`](../reference/columns.qmd#paginate) — every knob (`rows`, `break_on`, `keep_groups`, `orphan_min`, `repeat_*`, `footnotes_on`, `page_label`, `oversized_group_policy`)
- [Clinical examples](clinical.qmd) — shorter clinical forest patterns
- [LOTR](lotr.qmd) — illustrative themes that exercise the full surface
