---
title: "The Preset Gallery"
description: "All nine theme presets on one table — pick one, then make it yours with set_brand()"
---

```{r}
#| include: false
library(tabviz)
```

The fastest road to a publication-ready table is **not** building a theme
from inputs — it's picking the preset whose *identity* matches your
document, then swapping in your brand color:

```{r}
#| eval: false
tabviz(df, ..., theme = set_brand(web_theme_ledger(), "#1F3A5F"))
```

Every preset below renders the same five-row meta-analysis, so the only
thing changing is the theme. Identity is more than color: presets pin
typography, density, corner radius, shell texture, mark shapes, effects
(glow / gradient / glass), and status palettes — and a CI gate enforces
that no two presets converge.

```{r}
demo <- data.frame(
  study = c("Alpha 2024", "Beta 2023", "Gamma 2022", "Delta 2021", "Epsilon 2024"),
  n     = c(245, 189, 312, 278, 478),
  hr    = c(0.72, 0.81, 0.66, 0.74, 0.91),
  lo    = c(0.58, 0.65, 0.50, 0.59, 0.78),
  hi    = c(0.89, 1.01, 0.86, 0.93, 1.06),
  p     = c(0.004, 0.062, 0.0008, 0.011, 0.23)
)

show <- function(theme, tag = NULL) {
  tabviz(demo,
    label = "study",
    columns = list(
      col_n("n"),
      col_interval("hr", "lo", "hi", header = "HR (95% CI)"),
      col_pvalue("p", "P", stars = TRUE),
      viz_forest(point = "hr", lower = "lo", upper = "hi",
                 scale = "log", null_value = 1, axis_label = "Hazard ratio")
    ),
    theme = theme,
    title = theme@name,
    tag = tag
  )
}
```

The roster is the four registers `package_themes()` groups by: one
**clinical** default, two **editorial** voices, two **design** movements,
and four **expressive** moods. Each preset owns a distinct expressive axis.

## Clinical

Restrained clinical authority — the package default.

```{r}
show(web_theme_nejm())
```

## Editorial

Typographic register: a warm-grain newsprint and a Cinzel-display fantasy
serif.

```{r}
show(web_theme_newsprint())
```

```{r}
show(web_theme_dwarven(), tag = "Table 1")
```

## Design

Design-movement identities — color discipline and hard geometry.

```{r}
show(web_theme_ledger())
```

```{r}
show(web_theme_brutalist())
```

## Expressive

The expressivity ceiling: glow, glass, neon, phosphor aliasing, and
cyanotype grid — the moods that exercise the full surface.

```{r}
show(web_theme_aurora())
```

```{r}
show(web_theme_terminal())
```

```{r}
show(web_theme_blueprint())
```

```{r}
show(web_theme_synthwave())
```

## Make it yours

Three moves cover most customization, in order of reach:

```{r}
#| eval: false
# 1. Swap the brand — everything identity re-derives from one seed.
my_theme <- set_brand(web_theme_ledger(), "#1F3A5F")

# 2. Pull the supporting hues toward the brand (subtle / medium / vivid).
my_theme <- tint_from_brand(my_theme, "subtle")

# 3. Reach for a focused setter when one axis needs attention:
my_theme <- my_theme |>
  set_density("compact", factor = 0.95) |>
  set_effects(glow_intensity = "subtle", title_style = "bar") |>
  set_status_palette(list(negative = "#D42320")) |>
  set_pin("--tv-ink2", "#862721")             # rubrication ink (pinnable token)
```

The full input model — anchors, polarity, monochrome, geometry, marks,
effects — is the [Themes guide](../guide/themes.qmd); every argument is
in the [Themes reference](../reference/themes.qmd). For an end-to-end
worked example, see [Compose Your Own Brand](brand-design.qmd).
