---
title: "Feature Showcases"
---

```{r}
#| include: false
library(tabviz)
library(dplyr)
```

Spotlights for individual tabviz features. Each example highlights one feature family — column types, statistical models, distribution viz, theming, conditional styling, hierarchical splits — and combines them in the closing "Full Monty" smoke test.

## Rich Column Types

A comprehensive showcase of column types — heatmaps, progress bars, currency, dates, badges, stars, sparklines, bars, ranges, and icons — using a project management dataset:

```{r}
project_data <- data.frame(
  task = c("Backend API", "Frontend UI", "Database Migration",
           "Auth System", "Deployment Pipeline", "Documentation"),
  status = c("Active", "Active", "Completed", "At Risk", "On Hold", "Active"),
  priority = c(5, 4, 3, 5, 2, 3),
  budget = c(85000, 62000, 41000, 55000, 28000, 15000),
  spent_pct = c(0.65, 0.42, 1.00, 0.88, 0.15, 0.30),
  health = c(88, 75, 100, 45, 60, 82),
  start_date = as.Date(c("2025-01-15", "2025-02-01", "2024-11-01",
                          "2025-03-10", "2025-04-01", "2025-03-15")),
  est_low = c(30, 20, 15, 25, 10, 8),
  est_high = c(45, 35, 20, 40, 18, 12),
  velocity = I(list(
    c(8, 10, 12, 14, 15, 18, 20, 22),
    c(5, 8, 10, 12, 10, 8, 12, 14),
    c(15, 18, 20, 22, 25, 28, 30, 32),
    c(12, 10, 8, 6, 5, 4, 3, 2),
    c(2, 3, 4, 3, 2, 1, 1, 0),
    c(3, 5, 6, 8, 10, 12, 14, 15)
  )),
  flag = c("ok", "ok", "done", "blocked", "paused", "ok")
)

tabviz(project_data, label = "task",
  columns = list(
    col_badge("status", "Status",
      variants = list(
        Active = "success", Completed = "info",
        `At Risk` = "error", `On Hold` = "warning"
      )),
    col_stars("priority", "Priority", max_stars = 5),
    col_currency("budget", "Budget", decimals = 0),
    col_progress("spent_pct", "Spent", max_value = 1),
    col_heatmap("health", "Health",
      palette = c("#fee2e2", "#dcfce7"),
      min_value = 0, max_value = 100, decimals = 0),
    col_date("start_date", "Started", format = "%b %Y"),
    col_range("est_low", "est_high", header = "Est. Days"),
    col_sparkline("velocity", "Velocity"),
    col_icon("flag", "Flag",
      mapping = list(ok = "\u2714", done = "\u2605", blocked = "\u2716", paused = "\u23f8"))
  ),
  theme = web_theme_cochrane(),
  title = "Rich Column Types Showcase"
)
```

---

## Statistical Models

Forest plots aren't just for hazard ratios — any point estimate with a confidence interval fits. Here we visualize regression coefficients from a linear model of fuel economy on `mtcars`, with `null_value = 0` and a linear scale:

```{r}
model <- lm(mpg ~ cyl + disp + hp + wt + am, data = mtcars)
coefs <- summary(model)$coefficients
ci <- confint(model)

reg_data <- data.frame(
  term = rownames(coefs),
  estimate = coefs[, "Estimate"],
  std_error = coefs[, "Std. Error"],
  conf.low = ci[, 1],
  conf.high = ci[, 2],
  p.value = coefs[, "Pr(>|t|)"],
  row.names = NULL
) |>
  filter(term != "(Intercept)")

tabviz(reg_data,
  label = "term",
  columns = list(
    col_numeric("estimate", "Estimate", decimals = 3),
    col_numeric("std_error", "SE", decimals = 3),
    viz_forest(point = "estimate", lower = "conf.low", upper = "conf.high",
               scale = "linear", null_value = 0,
               axis_label = "Coefficient (95% CI)"),
    col_interval("estimate", "conf.low", "conf.high", header = "95% CI"),
    col_pvalue("p.value", "P", stars = TRUE)
  ),
  marker_color = ~ case_when(
    conf.high < 0 ~ "#dc2626",
    conf.low > 0 ~ "#16a34a",
    TRUE ~ "#64748b"
  ),
  theme = web_theme_cochrane(),
  title = "Linear Regression Coefficients",
  subtitle = "lm(mpg ~ cyl + disp + hp + wt + am, data = mtcars)",
  footnote = "Red = negative effect, green = positive effect, gray = not significant at alpha = 0.05."
)
```

---

## Distribution Visualizations

All three visualization types — `viz_bar`, `viz_boxplot`, and `viz_violin` — in one combined clinical site dashboard:

```{r}
set.seed(654)
clinical_dashboard <- data.frame(
  site = c("Boston", "London", "Tokyo", "Sydney", "Toronto"),
  enrollment = c(245, 189, 312, 156, 198),
  completion_rate = c(0.87, 0.92, 0.78, 0.95, 0.89),
  response_times = I(list(
    rnorm(50, 5.2, 1.1),
    rnorm(50, 4.8, 0.9),
    rnorm(50, 6.1, 1.5),
    rnorm(50, 4.5, 0.8),
    rnorm(50, 5.5, 1.2)
  )),
  satisfaction = I(list(
    c(rnorm(60, 8.0, 1.0), rnorm(20, 5, 1.5)),
    rnorm(80, 8.5, 0.7),
    c(rnorm(50, 7.0, 1.5), runif(20, 3, 9)),
    rnorm(75, 9.0, 0.5),
    c(rnorm(40, 7.5, 1.2), rnorm(30, 6, 1.8))
  ))
)

tabviz(clinical_dashboard, label = "site",
  columns = list(
    viz_bar(
      effect_bar("enrollment"),
      header = "Enrolled",
      axis_range = c(0, 350)
    ),
    viz_bar(
      effect_bar("completion_rate"),
      header = "Completion",
      axis_range = c(0, 1)
    ),
    viz_boxplot(
      effect_boxplot(data = "response_times"),
      header = "Response Time (days)",
      show_outliers = TRUE
    ),
    viz_violin(
      effect_violin(data = "satisfaction"),
      header = "Satisfaction (1-10)",
      show_median = TRUE,
      show_quartiles = TRUE
    )
  ),
  theme = web_theme_cochrane(),
  title = "Clinical Site Performance Dashboard"
)
```

---

## Column Groups & Split Plots

Combine `col_group()` headers with `split_by` for faceted views with organized columns:

```{r}
regional_data <- data.frame(
  study = rep(paste0("Study ", 1:4), 3),
  region = rep(c("North America", "Europe", "Asia Pacific"), each = 4),
  n = rep(c(1250, 980, 1560, 1100), 3),
  hr = c(0.72, 0.85, 0.68, 0.91,
         0.78, 0.82, 0.75, 0.88,
         0.65, 0.71, 0.62, 0.79),
  lo = c(0.55, 0.70, 0.52, 0.75,
         0.62, 0.68, 0.58, 0.72,
         0.48, 0.55, 0.45, 0.62),
  hi = c(0.95, 1.03, 0.89, 1.10,
         0.98, 0.99, 0.97, 1.08,
         0.88, 0.92, 0.86, 1.01)
)

tabviz(regional_data, label = "study",
  columns = list(
    col_group("Study Info",
      col_n("n")
    ),
    viz_forest(point = "hr", lower = "lo", upper = "hi",
               scale = "log", null_value = 1),
    col_group("Results",
      col_interval("hr", "lo", "hi", header = "HR (95% CI)")
    )
  ),
  split_by = "region",
  shared_axis = TRUE,
  theme = web_theme_jama(),
  title = "Treatment Effect by Region"
)
```

---

## Pagination — long tables across multiple pages

Long meta-analyses or multi-cohort registries often need to span more rows than fit on a printed page. `paginate = paginate_spec()` (or one of the presets — `paginate_letter()`, `paginate_a4()`, `paginate_slide()`) breaks the table into pages: the HTML viewer renders one page at a time with prev/next controls and a continuous-mode toggle, and `save_plot(*.pdf)` emits a true multi-page PDF.

```{r}
set.seed(2026)
n_studies <- 72
cohorts <- c("Phase III RCT", "Real-world evidence", "Registry")
counts  <- c(24, 30, 18)

meta <- data.frame(
  study = c(
    sprintf("Trial %02d", 1:counts[1]),
    sprintf("RWE %02d",  1:counts[2]),
    sprintf("Reg %02d",  1:counts[3])
  ),
  cohort = rep(cohorts, times = counts),
  year   = sample(2010:2024, n_studies, replace = TRUE),
  n      = sample(c(120, 240, 480, 960, 1900, 3800), n_studies, replace = TRUE),
  hr     = round(rnorm(n_studies, 0.92, 0.18), 2),
  se     = round(runif(n_studies, 0.06, 0.20), 2)
)
meta$lower <- pmax(0.20, round(meta$hr - 1.96 * meta$se, 2))
meta$upper <- round(meta$hr + 1.96 * meta$se, 2)

tabviz(
  meta,
  label = "study",
  group = "cohort",
  columns = list(
    col_numeric("year", "Year", decimals = 0),
    col_n("n"),
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               header = "HR (95% CI)", scale = "log", null_value = 1,
               axis_label = "Hazard ratio"),
    col_interval("hr", "lower", "upper", header = "HR")
  ),
  theme = web_theme_jama(),
  title = "GLP-1 receptor agonists — pooled cardiovascular meta-analysis",
  subtitle = "72 studies across three cohorts; paginate at 30 rows per page",
  caption = "Use the prev/next controls below the plot to walk through pages, or switch to continuous mode to scroll.",
  paginate = paginate_letter()
)
```

`paginate` cascades into every `split_by` subview — each subview owns its own page set, paginating independently:

```{r}
set.seed(404)
regions <- c("North America", "Europe", "Asia-Pacific", "Latin America")
sites_per_region <- 18
n <- length(regions) * sites_per_region

split_data <- data.frame(
  site = unlist(lapply(seq_along(regions), function(i) {
    sprintf("%s site %02d", substr(regions[i], 1, 4), seq_len(sites_per_region))
  })),
  region = rep(regions, each = sites_per_region),
  arm    = rep(c("Treatment", "Placebo"), length.out = n),
  n      = sample(c(80, 120, 240, 400), n, replace = TRUE),
  rr     = round(rnorm(n, 0.85, 0.14), 2)
)
split_data$lower <- pmax(0.30, round(split_data$rr - 0.18, 2))
split_data$upper <- round(split_data$rr + 0.18, 2)

tabviz(
  split_data,
  label = "site",
  group = "arm",
  split_by = "region",
  columns = list(
    col_n("n"),
    viz_forest(point = "rr", lower = "lower", upper = "upper",
               header = "Risk ratio (95% CI)",
               scale = "log", null_value = 1,
               axis_label = "Risk ratio"),
    col_interval("rr", "lower", "upper", header = "RR")
  ),
  theme = web_theme_cochrane(),
  title = "Multi-region efficacy",
  subtitle = "Each region paginates independently at 10 rows per page",
  paginate = paginate_spec(rows = 10)
)
```

See [Export → Pagination](../guide/export.qmd#pagination) for the full configuration surface and PDF export details.

---

## All Theme Variants {#all-theme-variants}

Compare the seven built-in themes — four journal presets and three LOTR-flavored editorial themes:

```{r}
#| layout-ncol: 3
#| fig-height: 2.5

sample_data <- data.frame(
  study = c("A", "B", "C"),
  hr = c(0.72, 0.85, 0.91),
  lo = c(0.55, 0.70, 0.75),
  hi = c(0.95, 1.03, 1.10)
)

make_plot <- function(theme_fn, name) {
  tabviz(sample_data, label = "study",
    columns = list(viz_forest(point = "hr", lower = "lo", upper = "hi",
                             scale = "log", null_value = 1)),
    theme = theme_fn(), title = name
  )
}

# Journal presets
make_plot(web_theme_cochrane, "Cochrane")
make_plot(web_theme_lancet,   "Lancet")
make_plot(web_theme_jama,     "JAMA")
make_plot(web_theme_dark,     "Dark")

# LOTR presets
make_plot(web_theme_dwarven,  "Dwarven")
make_plot(web_theme_elvish,   "Elvish")
make_plot(web_theme_hobbit,   "Hobbit")
```

---

## Conditional Styling

One comprehensive example showing formulas for `row_bold`, `row_color`, `marker_shape`, `marker_color`, and cell-level `color`/`bold` on columns:

```{r}
styling_data <- data.frame(
  study = c("RCT Alpha", "RCT Beta", "Cohort Gamma", "Cohort Delta",
            "Registry Epsilon", "RCT Zeta", "Cohort Eta"),
  design = c("RCT", "RCT", "Cohort", "Cohort", "Registry", "RCT", "Cohort"),
  hr = c(0.65, 0.78, 0.95, 1.12, 0.88, 0.72, 1.05),
  lo = c(0.50, 0.62, 0.78, 0.95, 0.72, 0.58, 0.88),
  hi = c(0.84, 0.98, 1.15, 1.32, 1.08, 0.89, 1.25),
  pval = c(0.001, 0.02, 0.15, 0.03, 0.08, 0.005, 0.22),
  quality = c(1.0, 0.9, 0.7, 0.6, 0.5, 0.95, 0.65)
)

tabviz(styling_data,
  label = "study",
  columns = list(
    col_text("design", "Design"),
    viz_forest(point = "hr", lower = "lo", upper = "hi",
               scale = "log", null_value = 1),
    col_numeric("hr", "HR",
      color = ~ ifelse(.x < 1, "#16a34a", "#dc2626"),
      bold = ~ .x < 0.8 | .x > 1.1
    ),
    col_pvalue("pval", "P",
      stars = TRUE,
      bold = ~ .x < 0.05
    ),
    col_badge("design", "Type")
  ),
  # Row-level styling
  row_bold = ~ pval < 0.01,
  row_color = ~ case_when(
    hi < 1 ~ "#16a34a",
    lo > 1 ~ "#dc2626",
    TRUE ~ "#64748b"
  ),
  # Marker styling by study design
  marker_shape = ~ case_when(
    design == "RCT" ~ "circle",
    design == "Cohort" ~ "square",
    TRUE ~ "diamond"
  ),
  marker_color = ~ case_when(
    hi < 1 ~ "#16a34a",
    lo > 1 ~ "#dc2626",
    TRUE ~ "#64748b"
  ),
  marker_opacity = "quality",
  marker_size = ~ ifelse(design == "RCT", 1.2, 1),
  theme = web_theme_cochrane(),
  title = "Conditional Styling Showcase",
  subtitle = "Row, cell, and marker styling with formulas"
)
```

---

## Fluent Theme Customization

Build a fully customized theme by chaining v2 modifiers off a base preset:

```{r}
custom_theme <- web_theme(
  name         = "warm",
  brand        = "#CA8A04",
  accent       = "#CA8A04",
  fonts_body   = "Georgia, serif",
  density      = "spacious"
) |>
  set_spacing(row_height = 32, cell_padding_x = 12) |>
  set_theme_field(c("plot", "point_size"), 8) |>
  set_theme_field(c("plot", "line_width"), 2) |>
  set_theme_field(c("layout", "container_border_radius"), 8) |>
  set_theme_field(c("row", "banding"), "row")

theme_data <- data.frame(
  study = c("PIONEER-1", "SUMMIT-2", "HORIZON-3", "ELEVATE-4", "APEX-5"),
  n = c(2450, 1890, 3200, 1560, 2100),
  hr = c(0.72, 0.85, 0.68, 0.91, 0.78),
  lo = c(0.58, 0.70, 0.52, 0.75, 0.62),
  hi = c(0.89, 1.03, 0.89, 1.10, 0.98),
  status = c("Significant", "Non-significant", "Significant",
             "Non-significant", "Significant")
)

tabviz(theme_data,
  label = "study",
  columns = list(
    col_n("n"),
    viz_forest(point = "hr", lower = "lo", upper = "hi",
               scale = "log", null_value = 1,
               axis_label = "Hazard Ratio"),
    col_interval("hr", "lo", "hi", header = "HR (95% CI)"),
    col_badge("status", "Result",
      variants = list(
        Significant = "success",
        `Non-significant` = "warning"
      ))
  ),
  marker_color = ~ ifelse(hi < 1, "#16a34a", "#64748b"),
  theme = custom_theme,
  title = "Custom Warm Theme via the v2 Inputs API",
  subtitle = "web_theme(brand = ...) + set_spacing() + set_theme_field()"
)
```

::: {.callout-tip}
## Theme modifiers (v2)
Four modifiers cover the surface: `set_inputs()` (Tier 1 inputs),
`set_density()` / `set_header_style()` (per-table variants), `set_spacing()` (per-token spacing),
and `set_theme_field(theme, path, value)` (generic deep setter).
See [Fluent API](../guide/fluent-api.qmd) and [Themes](../guide/themes.qmd).
:::

---

## Publication Templates

A single drop-in journal-style template demonstrating a row-grouped, column-grouped efficacy figure for a Lancet-style submission. To produce a JAMA / Cochrane / dark / LOTR variant, swap the `theme = web_theme_lancet()` line — every other element survives unchanged.

```{r}
lancet_data <- tibble(
  outcome = c(
    "CV death or HF hospitalization",
    "Cardiovascular death",
    "HF hospitalization",
    "All-cause mortality",
    "Change in KCCQ score"
  ),
  category = c("Primary", "Components", "Components", "Secondary", "Secondary"),
  hr = c(0.74, 0.82, 0.70, 0.88, 0.85),
  lower = c(0.66, 0.72, 0.61, 0.76, 0.74),
  upper = c(0.83, 0.94, 0.81, 1.02, 0.98),
  events_tx = c(447, 156, 291, 222, NA),
  events_ctrl = c(606, 199, 407, 253, NA),
  n_tx = c(2373, 2373, 2373, 2373, 2200),
  n_ctrl = c(2371, 2371, 2371, 2371, 2180)
)

tabviz(
  lancet_data,
  label = "outcome",
  group = "category",
  columns = list(
    col_group("Treatment",
      col_n("events_tx", "Events"),
      col_n("n_tx", "N")
    ),
    col_group("Control",
      col_n("events_ctrl", "Events"),
      col_n("n_ctrl", "N")
    ),
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1,
               axis_label = "Hazard Ratio (95% CI)"),
    col_interval("hr", "lower", "upper", header = "HR (95% CI)")
  ),
  theme = web_theme_lancet(),
  title    = "Figure 3: Efficacy Outcomes",
  subtitle = "Row groups by endpoint type + column groups by treatment arm",
  caption  = "HR<1 favours treatment.",
  footnote = "Cox proportional hazards model stratified by region."
)
```

For a worked walkthrough of building a custom branded theme from scratch — neutrals → primary → accent → series anchors → fonts → variants → field pins → wrapped function — see [Compose Your Own Brand](brand-design.qmd). For a full long-table treatment with pagination and pooled summaries see [Long Meta-Analysis](meta-analysis.qmd).

---

## Hierarchical Split Navigation

`split_by` with multiple variables turns the sidebar into an expandable tree (Outcome > Treatment here), letting you drill into nested subgroups without cluttering the main view.

```{r}
data(effect_sizes)

tabviz(
  effect_sizes,
  label = "study",
  split_by = c("outcome", "treatment"),
  columns = list(
    col_text("region", "Region"),
    col_text("phase", "Phase"),
    col_n("n"),
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1,
               axis_label = "Hazard Ratio"),
    col_interval("hr", "lower", "upper", header = "HR (95% CI)")
  ),
  theme = web_theme_cochrane(),
  title = "Hierarchical Subgroup Analysis",
  subtitle = "Outcome > Treatment navigation tree",
  caption = "Expand tree nodes to drill into subgroups"
)
```

---

## The Full Monty

Everything at once — nested row groups, three stacked effects, sparklines, weight bars, annotations, badges, and a custom dark theme. Useful as a smoke test and as a reference for how features compose.

```{r}
full_monty <- tibble(
  study = c(
    "ALPHA-01", "ALPHA-02",
    "BETA-01", "BETA-02", "BETA-03",
    "GAMMA-01", "GAMMA-02"
  ),
  program = c(rep("program_a", 2), rep("program_b", 3), rep("program_c", 2)),
  phase = c("Phase_II", "Phase_II", "Phase_III", "Phase_III", "Phase_III", "Phase_II", "Phase_III"),
  # Three effects
  primary_hr = c(0.68, 0.72, 0.65, 0.71, 0.74, 0.78, 0.69),
  primary_lo = c(0.52, 0.56, 0.50, 0.55, 0.58, 0.62, 0.53),
  primary_hi = c(0.89, 0.93, 0.85, 0.92, 0.95, 0.98, 0.90),
  secondary_hr = c(0.72, 0.76, 0.69, 0.75, 0.78, 0.82, 0.73),
  secondary_lo = c(0.56, 0.60, 0.53, 0.59, 0.62, 0.66, 0.57),
  secondary_hi = c(0.93, 0.96, 0.90, 0.95, 0.98, 1.02, 0.94),
  safety_hr = c(0.85, 0.88, 0.82, 0.86, 0.89, 0.92, 0.84),
  safety_lo = c(0.68, 0.72, 0.65, 0.69, 0.72, 0.75, 0.67),
  safety_hi = c(1.06, 1.08, 1.04, 1.07, 1.10, 1.13, 1.05),
  n = c(420, 380, 1250, 980, 1100, 560, 890),
  weight = c(8, 7, 22, 18, 20, 10, 15),
  pvalue = c(0.008, 0.015, 0.001, 0.004, 0.012, 0.042, 0.003),
  trend = list(
    c(0.85, 0.78, 0.72, 0.68), c(0.88, 0.82, 0.76, 0.72),
    c(0.82, 0.75, 0.69, 0.65), c(0.86, 0.80, 0.75, 0.71),
    c(0.88, 0.82, 0.78, 0.74), c(0.92, 0.88, 0.84, 0.78),
    c(0.84, 0.78, 0.73, 0.69)
  ),
  badge = c("Lead", NA, "Pivotal", NA, NA, NA, "Fast Track")
)

# Custom theme
monty_theme <- web_theme_dark() |>
  set_inputs(brand = "#f59e0b") |>
  set_spacing(row_height = 38) |>
  set_theme_field(c("axis", "gridlines"), TRUE) |>
  set_theme_field(c("axis", "gridline_style"), "dotted")

tabviz(
  full_monty,
  label = "study", group = c("program", "phase"),
  columns = list(
    col_n("n"),
    col_bar("weight"),
    viz_forest(
      effects = list(
        effect_forest("primary_hr", "primary_lo", "primary_hi", label = "Primary", color = "#22c55e"),
        effect_forest("secondary_hr", "secondary_lo", "secondary_hi", label = "Secondary", color = "#3b82f6"),
        effect_forest("safety_hr", "safety_lo", "safety_hi", label = "Safety", color = "#f59e0b")
      ),
      scale = "log", null_value = 1,
      axis_range = c(0.4, 1.2),
      axis_label = "Hazard Ratio",
      annotations = list(
        refline(0.75, label = "Target", style = "dashed", color = "#a855f7")
      )
    ),
    col_group("Results",
      col_interval("primary_hr", "primary_lo", "primary_hi", header = "HR (95% CI)"),
      col_pvalue("pvalue", "P")
    ),
    col_sparkline("trend", "Trend")
  ),
  row_badge = "badge",
  theme = monty_theme,
  title = "The Full Monty",
  subtitle = "Nested groups + 3 effects + sparklines + weights + annotations + custom theme",
  caption = "Every major feature combined in one visualization",
  footnote = "Green=Primary, Blue=Secondary, Orange=Safety. Purple line=Target."
)
```
