---
title: "webforest"
subtitle: "Interactive Forest Plots for R"
---

```{r}
#| include: false
library(webforest)
```

## Overview

**webforest** is an R htmlwidgets package for creating interactive, web-native forest plots. While traditionally used for meta-analysis, it supports any data with point estimates and confidence intervals.

```{r}
#| echo: false
library(dplyr)

styled_data <- tibble(
  label = c(
    "Primary Endpoint",
    "  Composite MACE", "  CV Death", "  MI", "  Stroke",
    "",
    "Secondary Endpoints",
    "  All-cause mortality", "  HF hospitalization",
    "",
    "Overall Summary"
  ),
  hr = c(NA, 0.82, 0.88, 0.79, 0.76, NA, NA, 0.91, 0.72, NA, 0.80),
  lower = c(NA, 0.74, 0.76, 0.68, 0.62, NA, NA, 0.79, 0.61, NA, 0.73),
  upper = c(NA, 0.91, 1.02, 0.92, 0.93, NA, NA, 1.05, 0.85, NA, 0.88),
  events = c(NA, 856, 312, 298, 246, NA, NA, 445, 412, NA, 1268),
  rtype = c("header", rep("data", 4), "spacer", "header", "data", "data", "spacer", "summary"),
  rbold = c(TRUE, FALSE, FALSE, FALSE, FALSE, FALSE, TRUE, FALSE, FALSE, FALSE, TRUE),
  rindent = c(0, 1, 2, 2, 2, 0, 0, 1, 1, 0, 0),
  rcolor = c("#2563eb", NA, NA, NA, NA, NA, "#2563eb", NA, NA, NA, "#16a34a"),
  rbadge = c(NA, "Primary", NA, NA, NA, NA, NA, NA, "Key", NA, "Pooled")
)

forest_plot(
  styled_data,
  point = "hr", lower = "lower", upper = "upper",
  label = "label",
  columns = list(
    col_numeric("events", "Events"),
    col_interval("HR (95% CI)")
  ),
  row_type = "rtype", row_bold = "rbold", row_indent = "rindent",
  row_color = "rcolor", row_badge = "rbadge",
  theme = web_theme_modern(),
  scale = "log", null_value = 1,
  axis_label = "Hazard Ratio",
  caption = "Headers in blue, summary in green, with badges and indentation"
)
```

## Features

- **Modern web rendering** using Svelte 5 and D3.js
- **Multiple preset themes** (JAMA, Lancet, presentation, dark mode)
- **Flexible column system** with special renderers (bars, p-values, sparklines)
- **Hierarchical grouping** with collapsible sections
- **Interactive features** (hover tooltips, row selection)
- **Static export** to SVG, PNG, and PDF
- **Fluent API** for theme customization

## Installation

```r
# install.packages("pak")
pak::pak("kaskarn/webforest")
```

## Quick Example

```r
library(webforest)

forest_plot(data,
  point = "estimate",
  lower = "ci_lower",
  upper = "ci_upper",
  label = "study",
  null_value = 1,
  scale = "log"
)
```

## Guide

Learn how to use webforest:

| Chapter | Description |
|---------|-------------|
| [Quick Start](guide/quick-start.qmd) | Installation and basic usage |
| [Themes](guide/themes.qmd) | Preset themes and customization |
| [Columns](guide/columns.qmd) | Data columns and special formatting |
| [Grouping & Rows](guide/grouping-rows.qmd) | Hierarchical data and row styling |
| [Exporting](guide/exporting.qmd) | Save plots as SVG, PNG, or PDF |
| [Axis & Annotations](guide/axis-annotations.qmd) | Axis control and reference lines |

## More Resources

- [Cookbook](cookbook.qmd) - Common tasks and patterns
- [Gallery](gallery.qmd) - Interactive examples
- [Reference](reference.qmd) - API documentation
