---
title: "tabviz"
subtitle: "Interactive Tables with Embedded Visualizations for R"
---

```{r}
#| include: false
library(tabviz)
library(dplyr)
```

## Overview

**tabviz** is an R htmlwidgets package for building interactive tables with embedded visualizations. It turns a data frame into a publication-quality, fully interactive widget — sortable, filterable, draggable, and editable — with SVG/PNG export that captures exactly what you see on screen.

```{r}
#| echo: false
# Full-interactivity hero: drag rows within groups, drag columns within groups,
# click headers to sort, click the funnel to filter, double-click cells to edit.
data(glp1_trials)

tabviz(
  glp1_trials,
  label = "study",
  group = "group",
  columns = list(
    col_group("Study Info",
      col_text("drug", "Drug"),
      col_n("n"),
      col_events("events", "n", "Events")
    ),
    viz_forest(
      point = "hr", lower = "lower", upper = "upper",
      scale = "log", null_value = 1,
      axis_label = "Hazard Ratio (95% CI)",
      axis_range = c(0.4, 1.5),
      axis_ticks = c(0.5, 0.75, 1.0, 1.25),
      axis_gridlines = TRUE,
      annotations = list(
        refline(0.85, label = "Pooled HR", style = "dashed", color = "#00407a")
      )
    ),
    col_group("Results",
      col_interval("HR (95% CI)", point = "hr", lower = "lower", upper = "upper"),
      col_pvalue("pvalue", "P", stars = TRUE)
    )
  ),
  row_type = "row_type", row_bold = "row_bold",
  interaction = web_interaction_full(),
  theme = web_theme_nature(),
  title = "GLP-1 Agonist Cardiovascular Outcomes",
  subtitle = "Major adverse cardiovascular events (MACE)"
)
```

::: {.callout-tip}
## Try the interactivity
- **Sort** — click any column header (chevron appears)
- **Filter** — click the funnel icon in a header and pick a text / numeric / categorical filter
- **Reorder rows** — hover a row, grab the drag handle, drop inside the same group
- **Reorder columns** — hover a column header, grab its handle; columns stay within their group ("Study Info" / "Results")
- **Edit values** — double-click any label, text, or numeric cell; double-click a forest marker for an estimate / lower / upper popover
- **Export** — the download button's SVG & PNG capture exactly your current view
:::

## Build any table in a few lines

Use `tabviz()` with explicit column definitions — the declarative API mirrors the output:

```r
tabviz(data,
  label = "study",
  columns = list(
    col_group("Demographics",
      col_text("site"),
      col_n("n")
    ),
    viz_forest(point = "hr", lower = "lo", upper = "hi",
               scale = "log", null_value = 1),
    col_group("Results",
      col_interval("HR (95% CI)", point = "hr", lower = "lo", upper = "hi"),
      col_pvalue("pvalue", "P")
    )
  ),
  interaction = web_interaction_full()
)
```

## Feature highlights

A few of tabviz's strengths, each in a compact example:

### Row groups + column groups together

```{r}
#| echo: false
data(effect_sizes)

tabviz(effect_sizes,
  label = "study",
  group = "treatment",       # collapsible row groups (Drug A, Drug B, ...)
  columns = list(
    col_group("Trial",
      col_text("region", "Region"),
      col_text("phase", "Phase"),
      col_n("n")
    ),
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1,
               axis_label = "Hazard Ratio"),
    col_group("Effect",
      col_interval("HR (95% CI)", point = "hr", lower = "lower", upper = "upper"),
      col_badge("direction", "Direction",
        variants = list(
          `Favors Treatment` = "success",
          `Favors Control` = "error",
          `No Effect` = "muted"
        ))
    )
  ),
  interaction = web_interaction_full(),
  theme = web_theme_default(),
  title = "Row Groups × Column Groups",
  subtitle = "Click a group header to collapse; drag within any group to reorder"
)
```

### Multiple effects per row

```{r}
#| echo: false
dose <- data.frame(
  endpoint = c("Overall Survival", "Progression-Free Survival",
               "Objective Response", "Quality of Life"),
  low_hr  = c(0.88, 0.82, 0.85, 0.92),
  low_lo  = c(0.74, 0.68, 0.70, 0.78),
  low_hi  = c(1.05, 0.99, 1.03, 1.08),
  mid_hr  = c(0.76, 0.71, 0.74, 0.83),
  mid_lo  = c(0.62, 0.58, 0.60, 0.69),
  mid_hi  = c(0.93, 0.87, 0.91, 1.00),
  high_hr = c(0.65, 0.60, 0.62, 0.75),
  high_lo = c(0.51, 0.46, 0.48, 0.61),
  high_hi = c(0.83, 0.78, 0.80, 0.92)
)

tabviz(dose, label = "endpoint",
  columns = list(
    viz_forest(
      effects = list(
        effect_forest("low_hr", "low_lo", "low_hi", label = "Low"),
        effect_forest("mid_hr", "mid_lo", "mid_hi", label = "Standard"),
        effect_forest("high_hr", "high_lo", "high_hi", label = "High")
      ),
      scale = "log", null_value = 1,
      axis_label = "Hazard Ratio"
    )
  ),
  interaction = web_interaction(enable_sort = TRUE, enable_filters = TRUE,
                                enable_reorder_rows = TRUE),
  theme = web_theme_cochrane(),
  title = "Dose-Response (three arms per row)"
)
```

### Rich dashboards — heatmaps, progress, sparklines, badges

```{r}
#| echo: false
data(airline_delays)
april <- airline_delays |>
  filter(month == "Apr") |>
  select(carrier, carrier_type, on_time_pct, satisfaction, flights, delay_vs_avg, trend)

tabviz(april, label = "carrier",
  columns = list(
    col_badge("carrier_type", "Type",
      variants = list(Legacy = "info", `Low-Cost` = "success",
                      `Ultra Low-Cost` = "warning")),
    col_n("flights", "Flights", abbreviate = TRUE),
    col_heatmap("on_time_pct", "On-Time",
      palette = c("#fee2e2", "#dcfce7"),
      min_value = 0.65, max_value = 0.95, decimals = 2, show_value = TRUE),
    col_progress("satisfaction", "Satisfaction", max_value = 5),
    col_numeric("delay_vs_avg", "Delay vs Avg", decimals = 1),
    col_sparkline("trend", "12-Month Trend")
  ),
  interaction = web_interaction_full(),
  theme = web_theme_modern(),
  title = "Airline Performance (April)"
)
```

## Features

::: {.grid}
::: {.g-col-6}
**Visualization**

- 9 preset themes (JAMA, Lancet, Nature, Cochrane, …)
- 20+ column types — forest plots, bars, p-values, sparklines, heatmaps, progress bars, currency, dates, badges, icons, stars
- Hierarchical row groups and column groups with collapsible sections
- Multiple effects per row for dose-response and multi-arm designs
:::

::: {.g-col-6}
**Interactivity**

- Sort columns by header click (grouped tables preserve group structure)
- Per-column filter popovers (text / numeric range / categorical)
- Drag-and-drop reordering of rows, row groups, columns, and column groups
- Double-click to edit any label, text, or numeric cell; forest cells get an estimate / lower / upper popover
- Hover tooltips with custom content
:::
:::

::: {.grid}
::: {.g-col-6}
**Export**

- SVG, PNG, PDF output at publication quality
- **WYSIWYG**: exports capture your current sort / filter / reorder / edits
- Same layout engine drives both browser view and headless export
:::

::: {.g-col-6}
**Developer Experience**

- Fluent API with R pipes
- Declarative column system
- Type-safe styling via NSE column mapping (`row_bold = ~ pvalue < 0.05`)
- Shiny integration with proxy methods for reactive updates
:::
:::

## Installation

```r
# install.packages("pak")
pak::pak("kaskarn/tabviz")
```

## Quick Example

```r
library(tabviz)

tabviz(data,
  label = "study",
  columns = list(
    col_text("category"),
    viz_forest(point = "estimate", lower = "ci_lower", upper = "ci_upper",
               scale = "log", null_value = 1),
    col_interval("Est (95% CI)", point = "estimate",
                 lower = "ci_lower", upper = "ci_upper"),
    col_numeric("n")
  ),
  interaction = web_interaction_full()
)
```

## Guide

Learn how to use tabviz:

| Chapter | Description |
|---------|-------------|
| [Quick Start](guide/quick-start.qmd) | Installation and basic usage |
| [Columns](guide/columns.qmd) | Column types, usage, and column groups |
| [Row Styling](guide/row-styling.qmd) | Headers, summaries, indentation, and badges |
| [Cell Styling](guide/cell-styling.qmd) | Per-cell conditional formatting |
| [Row Groups](guide/row-groups.qmd) | Collapsible sections from column values |
| [Forest Plots](guide/visualizations/forest-plots.qmd) | Adding forest columns with viz_forest() |
| [Split Plots](guide/split-plots.qmd) | Sidebar navigation for subgroups |
| [Themes](guide/themes.qmd) | Preset themes and customization |
| [Shiny Integration](guide/shiny.qmd) | Reactive tables |
| [Export](guide/export.qmd) | SVG, PNG, PDF output |

## More Resources

- [Concepts](concepts/index.qmd) — architecture and mental models
- [Cheatsheet](cheatsheet.qmd) — quick reference
- [Gallery](gallery/index.qmd) — examples by domain, including a dedicated interactivity tour
- [Reference](reference/index.qmd) — API documentation
