---
title: "Annotations"
---

```{r}
#| include: false
library(tabviz)
```

Annotations decorate a `viz_forest()` column with reference lines and custom shapes. They are passed via the `annotations = list(...)` argument of `viz_forest()` (and the equivalent shorthand in `forest_plot()`).

---

## `refline()` {#refline}

Vertical reference line at a specific x-axis position. Most often used to mark the null effect (`x = 0` linear, `x = 1` log) or a clinically meaningful threshold.

```r
refline(x,
        label = NULL,
        style = c("dashed", "solid", "dotted"),
        color = NULL,
        width = 1,
        opacity = 0.6)
```

| Argument | Type | Default | Description |
|---|---|---|---|
| `x` | `numeric(1)` | required | X-axis position. |
| `label` | `character(1)` | `NULL` | Text rendered near the line. |
| `style` | choice | `"dashed"` | `"dashed"`, `"solid"`, or `"dotted"`. |
| `color` | `character(1)` | `NULL` | CSS color. `NULL` uses the theme default. |
| `width` | `numeric(1)` | `1` | Line width in px. |
| `opacity` | `numeric(1)` | `0.6` | 0ŌĆō1. |

**Returns:** a `ReferenceLine` object.

```{r}
#| eval: false
data <- data.frame(
  study = c("Study A", "Study B", "Study C"),
  hr    = c(0.72, 0.85, 1.15),
  lower = c(0.55, 0.70, 0.90),
  upper = c(0.95, 1.03, 1.45)
)

tabviz(data, label = "study",
  columns = list(
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1,
               annotations = list(
                 refline(1,    label = "No effect",            style = "solid"),
                 refline(0.75, label = "Clinically meaningful", style = "dashed")
               ))
  )
)
```

---

## `forest_annotation()` {#forest_annotation}

Custom shape attached to a specific study row.

```r
forest_annotation(
  study_id,
  shape = c("circle", "square", "triangle", "star"),
  position = c("after", "before", "overlay"),
  color = "#8b5cf6",
  size = 1.0
)
```

| Argument | Type | Default | Description |
|---|---|---|---|
| `study_id` | `character(1)` | required | Identifier of the study to annotate. |
| `shape` | choice | `"circle"` | One of `"circle"`, `"square"`, `"triangle"`, `"star"`. |
| `position` | choice | `"after"` | Position relative to the point: `"before"`, `"after"`, or `"overlay"`. |
| `color` | `character(1)` | `"#8b5cf6"` | CSS color. |
| `size` | `numeric(1)` | `1.0` | Size multiplier. |

**Returns:** a `CustomAnnotation` object.

:::{.callout-warning}
## Frontend rendering not implemented

`forest_annotation()` is currently a placeholder ŌĆö the spec is built but the frontend has no renderer for it yet. Use `refline()` for visual annotations until this lands.
:::

## Related

- [`viz_forest()`](columns.qmd#viz_forest) ŌĆö the `annotations` argument
- [`forest_plot()`](forest_plot.qmd) ŌĆö quick wrapper that forwards an `annotations` list
