---
title: "Columns"
---

```{r}
#| include: false
library(tabviz)
```

Reference for every column-defining helper: data columns (`col_*()`), per-row visualization columns (`viz_*()`), the low-level `web_col()` constructor, column groups, and the per-effect helpers used by multi-effect viz columns.

:::{.callout-important}
## `field` is now a required argument

`col_n()`, `col_bar()`, `col_pvalue()`, and `col_sparkline()` previously had implicit defaults for `field` (`"n"`, `"weight"`, `"pvalue"`, `"trend"`). Those defaults are soft-deprecated since 0.9.0 — pass the column name explicitly. Other `col_*()` helpers have always required `field`.
:::

All `col_*()` helpers also accept the cell-styling arguments documented in [Cell-level styling](#cell-level-styling) below: `bold`, `italic`, `color`, `bg`, `badge`, `icon`, `emphasis`, `muted`, `accent`, plus `align`, `header_align`, `wrap`, `sortable`, `na_text`, `tooltip`, and `formatter`.

::: {.callout-note}
## `tooltip` and `formatter` (added in 0.9.0)

- `tooltip = "<col>"` — column whose values are the per-cell hover text
- `formatter = function(x) ...` — escape hatch for custom value rendering. The function receives the raw column vector and must return a character vector of the same length; the output is rendered as text. Use this when the built-in numeric/interval/percent formatting can't express the desired output.
:::

---

## Basic data columns

### `col_text()` {#col_text}

Plain text column.

```r
col_text(field, header = NULL, width = NULL, max_chars = NULL, ...)
```

| Argument | Type | Default | Description |
|---|---|---|---|
| `field` | `character(1)` | required | Data column name. |
| `header` | `character(1)` | `NULL` (uses `field`) | Display header. |
| `width` | `numeric(1)` | `NULL` (auto) | Column width in pixels. |
| `max_chars` | `integer(1)` | `NULL` | Truncate with ellipsis past this length. |
| `...` | — | — | Forwarded to `web_col()` (cell styling, alignment). |

### `col_label()` {#col_label}

Same as `col_text()` but auto-prettifies the default header (e.g. `"study_id"` → `"Study Id"`). Useful for the leftmost row-identifier column when you want explicit control.

```r
col_label(field, header = NULL, width = NULL, max_chars = NULL, ...)
```

### `col_numeric()` {#col_numeric}

Numeric column with formatting, sig-fig support, large-number abbreviation.

```r
col_numeric(field, header = NULL, width = NULL,
            decimals = 2, digits = NULL,
            thousands_sep = FALSE, abbreviate = FALSE, ...)
```

| Argument | Type | Default | Description |
|---|---|---|---|
| `field` | `character(1)` | required | Data column. |
| `header` | `character(1)` | `NULL` | Header text. |
| `width` | `numeric(1)` | `NULL` | Width in px. |
| `decimals` | `integer(1)` | `2` | Fixed decimal places. Mutually exclusive with `digits`. |
| `digits` | `integer(1)` | `NULL` | Significant figures. |
| `thousands_sep` | `character(1)`/`FALSE` | `FALSE` | Pass `","` (or other) to enable. |
| `abbreviate` | `logical(1)` | `FALSE` | Shorten ≥1000 (e.g. `1100` → `"1.1K"`). |

```r
col_numeric("estimate")
col_numeric("count", decimals = 0, thousands_sep = ",")
col_numeric("population", abbreviate = TRUE)
```

### `col_n()` {#col_n}

Sample-size / count column. Defaults to integers with `,` thousands separator.

```r
col_n(field, header = "N", width = NULL,
      decimals = 0, digits = NULL,
      thousands_sep = ",", abbreviate = FALSE, ...)
```

| Argument | Type | Default | Description |
|---|---|---|---|
| `field` | `character(1)` | required | Sample-size column. |
| `header` | `character(1)` | `"N"` | |
| `decimals` | `integer(1)` | `0` | |
| `thousands_sep` | character/`FALSE` | `","` | |
| `abbreviate` | `logical(1)` | `FALSE` | |

### `col_percent()` {#col_percent}

```r
col_percent(field, header = NULL, width = NULL,
            decimals = 1, digits = NULL,
            multiply = TRUE, symbol = TRUE, ...)
```

| Argument | Type | Default | Description |
|---|---|---|---|
| `field` | `character(1)` | required | Numeric column. |
| `decimals` | `integer(1)` | `1` | |
| `multiply` | `logical(1)` | `TRUE` | When `TRUE`, expects 0–1 proportions. Set `FALSE` if data is already 0–100. |
| `symbol` | `logical(1)` | `TRUE` | Show `%`. |

### `col_pvalue()` {#col_pvalue}

```r
col_pvalue(field,
           header = "P-value",
           width = NULL,
           stars = FALSE,
           thresholds = c(0.05, 0.01, 0.001),
           format = c("auto", "scientific", "decimal"),
           digits = 2,
           exp_threshold = 0.001,
           abbrev_threshold = NULL, ...)
```

| Argument | Type | Default | Description |
|---|---|---|---|
| `field` | `character(1)` | required | P-value column. |
| `stars` | `logical(1)` | `FALSE` | Append `*` / `**` / `***`. |
| `thresholds` | `numeric(3)` | `c(0.05, 0.01, 0.001)` | Star thresholds. |
| `format` | choice | `"auto"` | `"auto"`, `"scientific"`, or `"decimal"`. |
| `digits` | `integer(1)` | `2` | Significant figures. |
| `exp_threshold` | `numeric(1)` | `0.001` | Below this, use exponential. |
| `abbrev_threshold` | `numeric(1)` | `NULL` | Below this, render `"<x"`. |

### `col_interval()` {#col_interval}

Formatted interval text (e.g. `"1.20 (0.90, 1.50)"`).

```r
col_interval(point, lower, upper,
             header = "95% CI", width = NULL,
             decimals = 2, digits = NULL,
             thousands_sep = FALSE, abbreviate = FALSE,
             separator = " ",
             imprecise_threshold = NULL,
             na_text = NULL, ...)
```

| Argument | Type | Default | Description |
|---|---|---|---|
| `point` | `character(1)` | required | Point estimate column. |
| `lower` | `character(1)` | required | Lower bound column. |
| `upper` | `character(1)` | required | Upper bound column. |
| `header` | `character(1)` | `"95% CI"` | |
| `decimals` | `integer(1)` | `2` | Decimal places. Mutually exclusive with `digits`. |
| `digits` | `integer(1)` | `NULL` | Significant figures. Mutually exclusive with `decimals`. |
| `thousands_sep` | `character(1)` or `FALSE` | `FALSE` | Thousands separator string or `FALSE`. |
| `abbreviate` | `logical(1)` | `FALSE` | Shorten large values (`"1.2K"`). |
| `separator` | `character(1)` | `" "` | Between point and `(lower, upper)`. |
| `imprecise_threshold` | `numeric(1)` | `NULL` | When `upper / lower` exceeds this, render `"--"`. |
| `na_text` | `character(1)` | `NULL` | Text for NA point values. |

:::{.callout-warning}
## Updated in 0.9.0
- `point` / `lower` / `upper` are now required. The implicit fallback that read these from a sibling `viz_forest()` is soft-deprecated and emits `lifecycle::deprecate_warn()`.
- `digits`, `thousands_sep`, `abbreviate` added to match the rest of the numeric formatting vocabulary.
- `sep` was renamed to `separator`; the old name still works with a deprecation warning.
:::

### `col_events()` {#col_events}

```r
col_events(events, n,
           header = "Events", width = NULL,
           separator = "/", show_pct = FALSE,
           thousands_sep = ",", abbreviate = FALSE,
           na_text = NULL, ...)
```

| Argument | Type | Default | Description |
|---|---|---|---|
| `events` | `character(1)` | required | Event-count column. |
| `n` | `character(1)` | required | Total-N column. |
| `separator` | `character(1)` | `"/"` | |
| `show_pct` | `logical(1)` | `FALSE` | Append `(xx.x%)`. |
| `thousands_sep` | `character(1)` or `FALSE` | `","` | Thousands separator. |
| `abbreviate` | `logical(1)` | `FALSE` | Shorten large values (`"1.1K/12K"`). |
| `na_text` | `character(1)` | `NULL` | Text for NA values. |

::: {.callout-warning}
## Renamed in 0.9.0
The arguments were called `events_field` and `n_field`. Both are still accepted with a deprecation warning; new code should use `events` and `n`.
:::

### `col_currency()` {#col_currency}

```r
col_currency(field, header = NULL, width = NULL,
             symbol = "$", decimals = 2, digits = NULL,
             thousands_sep = ",", abbreviate = FALSE,
             position = c("prefix", "suffix"),
             na_text = NULL, ...)
```

| Argument | Type | Default | Description |
|---|---|---|---|
| `symbol` | `character(1)` | `"$"` | Currency glyph. |
| `decimals` | `numeric(1)` | `2` | Decimal places. Mutually exclusive with `digits`. |
| `digits` | `numeric(1)` | `NULL` | Significant figures. Mutually exclusive with `decimals`. |
| `thousands_sep` | `character(1)` or `FALSE` | `","` | Separator string (`","`, `" "`, `"."`, …) or `FALSE`. The legacy `TRUE` value is still accepted. |
| `abbreviate` | `logical(1)` | `FALSE` | Shorten large values (`"$1.2M"`). |
| `position` | `"prefix"` or `"suffix"` | `"prefix"` | Symbol placement. |
| `na_text` | `character(1)` | `NULL` | Text for NA values. |

::: {.callout-warning}
## Updated in 0.9.0
`thousands_sep` was a `TRUE/FALSE` flag; now accepts a string. `digits` and `abbreviate` are new.
:::

### `col_date()` {#col_date}

```r
col_date(field, header = NULL, width = NULL, format = "%Y-%m-%d", ...)
```

`format` follows `?strftime`. Defaults: body alignment `"left"`.

### `col_heatmap()` {#col_heatmap}

```r
col_heatmap(field, header = NULL, width = NULL,
            palette = c("#f7fbff", "#08306b"),
            min_value = NULL, max_value = NULL,
            decimals = 2, show_value = TRUE,
            scale = c("linear", "log", "sqrt"), ...)
```

### `col_progress()` {#col_progress}

```r
col_progress(field, header = NULL, width = NULL,
             max_value = 100, color = NULL,
             show_label = TRUE,
             scale = c("linear", "log", "sqrt"), ...)
```

### `col_range()` {#col_range}

```r
col_range(low, high,
          header = "Range", width = NULL,
          separator = " - ",
          decimals = NULL, digits = NULL,
          thousands_sep = FALSE, abbreviate = FALSE,
          show_bar = FALSE,
          na_text = NULL, ...)
```

| Argument | Type | Default | Description |
|---|---|---|---|
| `low` | `character(1)` | required | Field name for the lower bound. |
| `high` | `character(1)` | required | Field name for the upper bound. |
| `separator` | `character(1)` | `" - "` | Glue between the two values. |
| `decimals` | `numeric(1)` | `NULL` (auto) | Decimal places. Mutually exclusive with `digits`. |
| `digits` | `integer(1)` | `NULL` | Significant figures. Mutually exclusive with `decimals`. |
| `thousands_sep` | `character(1)` or `FALSE` | `FALSE` | Thousands separator string or `FALSE`. |
| `abbreviate` | `logical(1)` | `FALSE` | Shorten large values (`"1.2K - 3.4M"`). |
| `show_bar` | `logical(1)` | `FALSE` | Render a visual bar between low and high. |
| `na_text` | `character(1)` | `NULL` | Text when either bound is NA. |

Body cells default to `align = "right"` (numeric convention).

::: {.callout-warning}
## Updated in 0.9.0
- `min_field` and `max_field` are still accepted with a deprecation warning; new code should use `low` and `high`.
- `digits`, `thousands_sep`, `abbreviate` added to match the rest of the numeric formatting vocabulary.
:::

### `col_reference()` {#col_reference}

Truncated text with optional outbound link and tooltip.

```r
col_reference(field, header = "Reference", width = NULL,
              href_field = NULL, max_chars = 30,
              show_icon = TRUE, ...)
```

| Argument | Type | Default | Description |
|---|---|---|---|
| `field` | `character(1)` | required | Reference text. |
| `href_field` | `character(1)` | `NULL` | URL column (optional). |
| `max_chars` | `integer(1)` | `30` | Truncation length. |
| `show_icon` | `logical(1)` | `TRUE` | Show external-link icon when `href_field` is set. |

:::{.callout-note}
## `icon` is deprecated
Use `show_icon` instead. The old `icon` argument still works for one cycle and emits `lifecycle::deprecate_warn()`.
:::

---

## Inline visualization columns

### `col_bar()` {#col_bar}

```r
col_bar(field, header = NULL, width = NULL,
        max_value = NULL, show_label = TRUE, color = NULL,
        scale = c("linear", "log", "sqrt"), ...)
```

| Argument | Type | Default | Description |
|---|---|---|---|
| `field` | `character(1)` | required | Bar value column. |
| `max_value` | `numeric(1)` | `NULL` (auto) | Bar normalization. |
| `show_label` | `logical(1)` | `TRUE` | |
| `color` | `character(1)` | `NULL` | CSS color. |

### `col_sparkline()` {#col_sparkline}

```r
col_sparkline(field, header = "Trend", width = NULL,
              type = c("line", "bar", "area"),
              height = 20, color = NULL, ...)
```

The `field` column should be a list-column of numeric vectors (one vector per row).

### `col_icon()` {#col_icon}

```r
col_icon(field, header = NULL, width = NULL,
         mapping = NULL,
         size = c("base", "sm", "lg"),
         color = NULL, ...)
```

`mapping` is a named character vector mapping data values to icons/emoji. Body cells default to `align = "center"`.

### `col_badge()` {#col_badge}

```r
col_badge(field, header = NULL, width = NULL,
          variants = NULL, colors = NULL,
          size = c("base", "sm"), ...)
```

| Argument | Type | Default | Description |
|---|---|---|---|
| `field` | `character(1)` | required | Badge text column. |
| `variants` | named `character` | `NULL` | Map values to semantic variants: `"default"`, `"success"`, `"warning"`, `"error"`, `"info"`, `"muted"`. |
| `colors` | named `character` | `NULL` | Map values to custom hex colors (override `variants`). |

```r
col_badge("status", variants = c(
  published = "success", draft = "warning", rejected = "error"
))
```

### `col_stars()` {#col_stars}

```r
col_stars(field, header = NULL, width = NULL,
          max_stars = 5, color = "#f59e0b",
          empty_color = "#d1d5db",
          half_stars = FALSE, domain = NULL, ...)
```

`domain = c(min, max)` remaps arbitrary input ranges into `[0, max_stars]`.

### `col_img()` {#col_img}

```r
col_img(field, header = NULL, width = NULL,
        height = NULL, max_width = NULL,
        fallback = "[img]",
        shape = c("square", "circle", "rounded"), ...)
```

---

## Per-row visualization columns (`viz_*`)

These columns plot one image per row with a shared axis across rows. They define their own effects and are not sortable.

### `viz_forest()` {#viz_forest}

Forest plot column (point + CI). Either pass `point` / `lower` / `upper` for a single effect, or `effects = list(effect_forest(...), ...)` for several effects overlaid in the same column.

```r
viz_forest(
  header = NULL,
  show_header = TRUE,
  header_align = "center",
  width = NULL,
  point = NULL, lower = NULL, upper = NULL,
  effects = NULL,
  scale = c("linear", "log"),
  null_value = NULL,
  axis_label = "Effect",
  axis_range = NULL,
  axis_ticks = NULL,
  axis_gridlines = FALSE,
  show_axis = TRUE,
  annotations = NULL,
  shared_axis = NULL,
  ...
)
```

| Argument | Type | Default | Description |
|---|---|---|---|
| `header` | `character(1)` | `NULL` | Falls back to single-effect label, then blank for multi-effect. |
| `show_header` | `logical(1)` | `TRUE` | |
| `header_align` | choice | `"center"` | |
| `width` | `numeric(1)` | `NULL` (auto) | |
| `point` / `lower` / `upper` | `character(1)` | `NULL` | Single-effect column refs (mutually exclusive with `effects`). |
| `effects` | `list` of `effect_forest()` | `NULL` | Multi-effect spec. |
| `scale` | choice | `"linear"` | `"linear"` or `"log"`. |
| `null_value` | `numeric(1)` | `0` (linear) / `1` (log) | Reference line. |
| `axis_label` | `character(1)` | `"Effect"` | |
| `axis_range` | `numeric(2)` | `NULL` | |
| `axis_ticks` | `numeric` | `NULL` | |
| `axis_gridlines` | `logical(1)` | `FALSE` | |
| `show_axis` | `logical(1)` | `TRUE` | |
| `annotations` | `list` of `refline()` | `NULL` | |
| `shared_axis` | `logical(1)` | `NULL` | In split forests, opt this column in/out of the split-level shared axis. |

### `effect_forest()` {#effect_forest}

```r
effect_forest(point, lower, upper,
              label = NULL, color = NULL,
              shape = NULL, opacity = NULL)
```

| Argument | Type | Default | Description |
|---|---|---|---|
| `point` | `character(1)` | required | Point column. |
| `lower` | `character(1)` | required | Lower bound column. |
| `upper` | `character(1)` | required | Upper bound column. |
| `label` | `character(1)` | `point` | Legend label. |
| `color` | `character(1)` | `NULL` (theme cycle) | |
| `shape` | choice | `NULL` (theme cycle) | One of `"square"`, `"circle"`, `"diamond"`, `"triangle"`. |
| `opacity` | `numeric(1)` | `NULL` | 0–1. |

:::{.callout-note}
## No per-effect `size`

In multi-effect plots, marker size is row-driven (via `set_marker_style(size = ...)` or `marker_size` on `tabviz()`) so weights remain comparable across effects.
:::

### `viz_bar()` {#viz_bar}

```r
viz_bar(...,                  # one or more effect_bar() objects
        header = NULL, show_header = TRUE, header_align = "center",
        width = NULL,
        scale = c("linear", "log"),
        axis_range = NULL, axis_ticks = NULL,
        axis_gridlines = FALSE,
        axis_label = "Value", show_axis = TRUE)
```

#### `effect_bar()`

```r
effect_bar(value, label = NULL, color = NULL, opacity = NULL)
```

### `viz_boxplot()` {#viz_boxplot}

```r
viz_boxplot(...,              # one or more effect_boxplot() objects
            header = NULL, show_header = TRUE, header_align = "center",
            width = NULL,
            scale = c("linear", "log"),
            axis_range = NULL, axis_ticks = NULL,
            axis_gridlines = FALSE,
            show_outliers = TRUE,
            whisker_type = c("iqr", "minmax"),
            axis_label = "Value", show_axis = TRUE)
```

#### `effect_boxplot()`

Two modes — provide either `data` (raw array column) or all five summary stats.

```r
# From raw array data (auto-compute quartiles)
effect_boxplot(data,
               outliers = NULL, label = NULL,
               color = NULL, opacity = 0.7)

# From pre-computed summary stats
effect_boxplot(min, q1, median, q3, max,
               outliers = NULL, label = NULL,
               color = NULL, opacity = 0.7)
```

::: {.callout-warning}
## Renamed in 0.9.0
`fill_opacity` is now `opacity` (matching `effect_forest()` / `effect_bar()`). The old name still works with a deprecation warning.
:::

### `viz_violin()` {#viz_violin}

```r
viz_violin(...,                # one or more effect_violin() objects
           header = NULL, show_header = TRUE, header_align = "center",
           width = NULL,
           scale = c("linear", "log"),
           axis_range = NULL, axis_ticks = NULL,
           axis_gridlines = FALSE,
           bandwidth = NULL,
           show_median = TRUE, show_quartiles = FALSE,
           axis_label = "Value", show_axis = TRUE)
```

#### `effect_violin()`

```r
effect_violin(data, label = NULL, color = NULL, opacity = 0.5)
```

::: {.callout-warning}
## Renamed in 0.9.0
`fill_opacity` is now `opacity`. The old name still works with a deprecation warning.
:::

---

## Column groups

### `col_group()` {#col_group}

Wrap several columns under a shared header.

```r
col_group(header, ...)
```

| Argument | Type | Description |
|---|---|---|
| `header` | `character(1)` | Group header label. |
| `...` | `ColumnSpec` | One or more `col_*()` / `viz_*()` columns. |

```r
col_group("Treatment Effect",
  col_numeric("hr", "HR"),
  col_interval("hr", "lower", "upper", header = "95% CI")
)
```

---

## Low-level constructor

### `web_col()` {#web_col}

The constructor every `col_*()` helper delegates to. Use this directly only when you need a column type the helpers don't cover.

```r
web_col(field,
        header = NULL,
        type = c("text", "numeric", "interval", "bar", "pvalue",
                 "sparkline", "icon", "badge", "stars", "img",
                 "reference", "range", "forest", "heatmap",
                 "progress", "viz_bar", "viz_boxplot",
                 "viz_violin", "custom"),
        width = NULL,
        align = NULL, header_align = NULL,
        show_header = NULL,
        wrap = FALSE, sortable = TRUE,
        options = list(),
        na_text = NULL,
        # Cell styling
        bold = NULL, italic = NULL,
        color = NULL, bg = NULL,
        badge = NULL, icon = NULL,
        emphasis = NULL, muted = NULL, accent = NULL)
```

Returns a `ColumnSpec`.

---

## Cell-level styling {#cell-level-styling}

Every `col_*()` helper accepts the same set of cell-styling arguments via `...`. Each takes either a **column name** (resolved at render time) or a **formula** evaluated against the row (use `.x` to refer to the column's own value).

| Argument | Resolved value | Effect |
|---|---|---|
| `bold` | logical | Bold text. |
| `italic` | logical | Italic text. |
| `color` | CSS color | Text color. |
| `bg` | CSS color | Cell background. |
| `badge` | character | Inline badge in the cell. |
| `icon` | emoji/unicode | Icon next to value. |
| `emphasis` | logical | Theme-aware emphasis (bold + foreground). |
| `muted` | logical | Theme-aware muted (lighter). |
| `accent` | logical | Theme accent color. |

```r
# Bold significant p-values, color them
col_pvalue("pval",
           bold  = ~ .x < 0.05,
           color = ~ ifelse(.x < 0.05, "#16a34a", "#71717a"))

# Reference another column
col_numeric("hr", "HR",
            color = ~ ifelse(pval < 0.05, "#16a34a", "#71717a"))
```

---

## Alignment

Two slots: `align` for body cells, `header_align` for the header. Defaults are type-aware:

| Helper | Body `align` default |
|---|---|
| `col_text`, `col_numeric`, `col_pvalue`, `col_bar`, `col_interval`, `col_sparkline`, `col_percent`, `col_reference`, `col_heatmap`, `col_progress`, `col_date` | `"left"` |
| `col_icon`, `col_badge`, `col_stars`, `col_img` | `"center"` |
| `col_range`, `col_currency` | `"right"` |
| `viz_forest` | header `"center"` |
| `viz_bar` | header `"left"` |
| `viz_boxplot`, `viz_violin` | header `"center"` |

`header_align` defaults to "follow the body" — set it explicitly to decouple.

---

## Examples

```{r}
data <- data.frame(
  study  = c("Trial A", "Trial B", "Trial C"),
  hr     = c(0.72, 0.85, 0.68),
  lower  = c(0.55, 0.70, 0.52),
  upper  = c(0.95, 1.03, 0.89),
  n      = c(245, 180, 320),
  weight = c(35, 28, 37),
  pval   = c(0.018, 0.092, 0.003)
)

tabviz(data,
  label = "study",
  columns = list(
    col_n("n"),
    col_bar("weight"),
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1),
    col_interval("hr", "lower", "upper", header = "HR (95% CI)"),
    col_pvalue("pval", stars = TRUE)
  )
)
```

```r
# Column groups
columns = list(
  col_n("n"),
  col_group("Treatment Effect",
    col_numeric("hr", "HR"),
    col_interval("hr", "lower", "upper", header = "95% CI")
  ),
  col_pvalue("pval")
)
```

## Related

- [`tabviz()`](tabviz.qmd) — passes a `columns` list of these helpers
- [Styling modifiers](styling.qmd) — `set_column_style()` for after-the-fact per-cell styling
- [Annotations](annotations.qmd) — `refline()` for `viz_forest(annotations = ...)`
