---
title: "Data Operations"
---

```{r}
#| include: false
library(tabviz)
library(dplyr)
```

The data-operation verbs provide a fluent, pipe-friendly interface for adding, removing, rearranging, filtering, sorting, and editing the columns, rows, and cells of a table. Every verb dispatches on input type:

- **`WebSpec`** — returns a new `WebSpec` with the change applied.
- **htmlwidget** (returned by `tabviz()`) — extracts the spec, mutates it, re-wraps as a widget.
- **`tabviz_proxy`** (Shiny only) — sends a targeted `sendCustomMessage()` to the running widget; no full re-render.

This single surface means the same code can compose a static plot or drive a live Shiny widget.

::: {.callout-tip}
## See Also

- [Styling functions](styling.qmd) for row/column/marker visual styling.
- [Fluent API guide](../guide/fluent-api.qmd) for end-to-end workflows.
- [Shiny integration](shiny.qmd) for proxy usage patterns.
:::

---

## Column Operations

### add_column() {#add_column}

Insert a new column into the table.

```r
add_column(x, col, after = NULL)
```

| Argument | Description |
|----------|-------------|
| `x` | A `WebSpec`, htmlwidget, or `tabviz_proxy` |
| `col` | A `ColumnSpec` from `col_text()`, `col_numeric()`, `col_bar()`, etc. |
| `after` | Field/id of an existing column to insert after. `NULL` (default) appends; `"__start__"` prepends. |

**Returns**: Modified input (same type).

```{r}
#| eval: false
spec <- web_spec(data, columns = list(col_text("study"), col_numeric("hr")))

# Append a column
spec |> add_column(col_numeric("se", "SE"))

# Insert after a specific field
spec |> add_column(col_badge("sig"), after = "hr")

# Prepend
spec |> add_column(col_icon("flag"), after = "__start__")
```

### remove_column() {#remove_column}

Remove a column by field or id.

```r
remove_column(x, field)
```

| Argument | Description |
|----------|-------------|
| `x` | A `WebSpec`, htmlwidget, or `tabviz_proxy` |
| `field` | Field or id of the column to remove |

On a proxy this sends `hideColumn`, which hides the column client-side without destroying its spec — `add_column()` with the same field will restore it.

```{r}
#| eval: false
spec |> remove_column("notes")
```

### move_column() {#move_column}

Change a column's position.

```r
move_column(x, field, to)
```

| Argument | Description |
|----------|-------------|
| `x` | A `WebSpec`, htmlwidget, or `tabviz_proxy` |
| `field` | Field or id of the column to move |
| `to` | 1-based integer index, or the field name of another column |

On the R side, indices are 1-based. The proxy payload converts to 0-based before sending to JS — you do not need to adjust.

```{r}
#| eval: false
# Move to position 3
spec |> move_column("hr", to = 3L)

# Move before the "pval" column
spec |> move_column("hr", to = "pval")
```

::: {.callout-note}
Within-group moves via static dispatch are currently unsupported — `move_column()` on a column that lives inside a `col_group()` will error. Use proxy dispatch or restructure groups.
:::

### resize_column() {#resize_column}

Set a column's pixel width.

```r
resize_column(x, field, width)
```

| Argument | Description |
|----------|-------------|
| `x` | A `WebSpec`, htmlwidget, or `tabviz_proxy` |
| `field` | Field or id of the column |
| `width` | Width in pixels (≥ 40) |

```{r}
#| eval: false
spec |> resize_column("study", width = 240)
```

### update_column() {#update_column}

Edit properties of an existing column in place.

```r
update_column(x, field, ...)
```

| Argument | Description |
|----------|-------------|
| `x` | A `WebSpec`, htmlwidget, or `tabviz_proxy` |
| `field` | Field or id of the column |
| `...` | Named properties to update |

Top-level slots written directly: `header`, `align`, `header_align`, `wrap`, `sortable`, `width`.
Any other name is merged into the column's `options` list (useful for type-specific knobs such as `decimals`, `format`, `colors`).

```{r}
#| eval: false
spec |>
  update_column("hr",
    header = "Hazard Ratio",
    align = "right",
    decimals = 3
  )
```

---

## Row & Data Operations

### sort_rows() {#sort_rows}

Order rows by a column's values.

```r
sort_rows(x, by, direction = c("asc", "desc", "none"))
```

| Argument | Description |
|----------|-------------|
| `x` | A `WebSpec`, htmlwidget, or `tabviz_proxy` |
| `by` | Field or id of the column to sort by |
| `direction` | `"asc"`, `"desc"`, or `"none"` (clears sort on a proxy) |

On static inputs, `direction = "none"` is a no-op; the spec data frame keeps its existing order.

```{r}
#| eval: false
spec |> sort_rows("hr", direction = "desc")
```

### filter_rows() {#filter_rows}

Keep rows that match a predicate.

```r
filter_rows(x, field, operator = c("eq", "neq", "gt", "lt", "contains"), value)
```

| Argument | Description |
|----------|-------------|
| `x` | A `WebSpec`, htmlwidget, or `tabviz_proxy` |
| `field` | Field to filter on |
| `operator` | `"eq"`, `"neq"`, `"gt"`, `"lt"`, or `"contains"` |
| `value` | Scalar value to compare against |

```{r}
#| eval: false
spec |> filter_rows("region", "eq", "North America")
spec |> filter_rows("hr", "lt", 1)
```

On static inputs, rows where the predicate is `NA` are dropped.

### clear_filters() {#clear_filters}

Remove any runtime filters.

```r
clear_filters(x)
```

No-op for static inputs (there is no persisted filter state). On a proxy, sends `clearFilter` to reset the widget's filter stack.

### select_rows() {#select_rows}

Replace the current row selection (proxy only).

```r
select_rows(x, row_ids)
```

| Argument | Description |
|----------|-------------|
| `x` | A `tabviz_proxy` |
| `row_ids` | Character vector of row ids to select |

On static inputs, warns and returns the input unchanged — selection is a runtime concept with no spec-level representation.

```{r}
#| eval: false
observeEvent(input$highlight, {
  proxy <- tabviz_proxy("plot")
  select_rows(proxy, c("Trial A", "Trial C"))
})
```

### move_row() {#move_row}

Reposition a row within the data frame.

```r
move_row(x, row_id, to)
```

| Argument | Description |
|----------|-------------|
| `x` | A `WebSpec`, htmlwidget, or `tabviz_proxy` |
| `row_id` | Row id (the value in the leftmost label/id column) |
| `to` | 1-based integer index, or another row id |

```{r}
#| eval: false
spec |> move_row("Summary", to = 1L)
```

### update_data() {#update_data}

Replace the entire underlying data frame.

```r
update_data(x, data)
```

| Argument | Description |
|----------|-------------|
| `x` | A `WebSpec`, htmlwidget, or `tabviz_proxy` |
| `data` | A data frame (static inputs) or `WebSpec` (proxy) |

For a proxy you must pass a complete `WebSpec`, not a raw data frame — column specs, themes, and annotations need to be re-serialized together. Build a fresh spec with `tabviz()` or `web_spec()` and pass it in.

```{r}
#| eval: false
# Static: swap the data frame
spec |> update_data(new_df)

# Proxy: full re-specification
observeEvent(input$refresh, {
  proxy <- tabviz_proxy("plot")
  fresh <- tabviz(reactive_data(), label = "study",
                  columns = list(viz_forest("hr", "lo", "hi")))
  update_data(proxy, attr(fresh, "webspec"))
})
```

---

## Group Operations

### toggle_group() {#toggle_group}

Collapse or expand a named group.

```r
toggle_group(x, group_id, collapsed = NULL)
```

| Argument | Description |
|----------|-------------|
| `x` | A `WebSpec`, htmlwidget, or `tabviz_proxy` |
| `group_id` | Id of the group to toggle |
| `collapsed` | `TRUE` to collapse, `FALSE` to expand, `NULL` to flip current state |

Groups are defined via the `group` argument of `tabviz()`; if a spec has none, `toggle_group()` warns for static inputs.

```{r}
#| eval: false
spec |> toggle_group("secondary_endpoints", collapsed = TRUE)
```

---

## Cell Edits (Proxy Only)

These verbs carry changes directly into a running widget. They have no spec-level counterpart — on static inputs they warn and return unchanged.

### set_cell() {#set_cell}

Overwrite a single cell.

```r
set_cell(x, row_id, field, value)
```

```{r}
#| eval: false
set_cell(proxy, row_id = "Trial A", field = "hr", value = 0.84)
```

### set_row_label() {#set_row_label}

Rename a row's label.

```r
set_row_label(x, row_id, label)
```

```{r}
#| eval: false
set_row_label(proxy, "Trial A", "Trial A (updated)")
```

### clear_edits() {#clear_edits}

Discard every in-widget cell and label edit applied since the last full render.

```r
clear_edits(x)
```

---

## Global controls

`set_theme()` and `set_zoom()` are documented in the [styling reference](styling.qmd#set_theme) — they are global presentation modifiers rather than data verbs, but they accept the same input shapes (spec / widget / proxy where applicable).

---

## Verb Reference

### Columns

| Function | Purpose |
|----------|---------|
| `add_column()` | Insert a new column |
| `remove_column()` | Remove a column (proxy: hide) |
| `move_column()` | Change column position |
| `resize_column()` | Set pixel width |
| `update_column()` | Edit header, alignment, options |

### Rows & data

| Function | Purpose |
|----------|---------|
| `sort_rows()` | Order rows by a column |
| `filter_rows()` | Keep rows matching a predicate |
| `clear_filters()` | Clear runtime filters (proxy) |
| `select_rows()` | Select rows (proxy) |
| `move_row()` | Reposition a row |
| `update_data()` | Replace the data frame |

### Groups & cells

| Function | Purpose |
|----------|---------|
| `toggle_group()` | Collapse/expand a group |
| `set_cell()` | Overwrite a cell (proxy) |
| `set_row_label()` | Rename a row (proxy) |
| `clear_edits()` | Discard in-widget edits (proxy) |

## See Also

- [Styling functions](styling.qmd)
- [Fluent API guide](../guide/fluent-api.qmd)
- [Shiny integration](shiny.qmd)
