---
title: "Export"
---

```{r}
#| include: false
library(tabviz)
```

Export tabviz widgets to static file formats (SVG, PDF, PNG).

The two exported entry points are:

| Function | Use case |
|---|---|
| [`save_plot()`](save_plot.qmd#save_plot) | Export a single widget / `WebSpec` to SVG, PDF, or PNG. |
| [`save_split_table()`](save_plot.qmd#save_split_table) | Export every sub-plot of a `SplitForest` to a directory tree. |

Both delegate to a shared SVG generator (the same one the browser uses), then convert to PDF / PNG via `rsvg`. See [`save_plot()`](save_plot.qmd) for the full reference, dependencies, and resolution guidance.

```{r}
#| eval: false
p <- tabviz(data,
  label = "study",
  columns = list(
    viz_forest(point = "hr", lower = "lo", upper = "hi"),
    col_interval("hr", "lo", "hi", header = "HR (95% CI)")
  )
)

save_plot(p, "figures/forest.svg")
save_plot(p, "figures/forest.png", width = 1200, scale = 3)
save_plot(p, "figures/forest.pdf", width = 800)
```

:::{.callout-note}
## Backend dependencies

- SVG only: `V8` (always required).
- PNG / PDF: also requires the `rsvg` package.

There is no headless-browser dependency — exports run entirely in R via V8 and rsvg.
:::

## Related

- [`save_plot()`](save_plot.qmd) — the single-plot exporter and `save_split_table()` for split views
- [`split_table()`](split_table.qmd) — produce the `SplitForest` accepted by `save_split_table()`
- [Themes](themes.qmd) — theme presets aimed at print (`web_theme_jama()`, `web_theme_lancet()`, `web_theme_cochrane()`)
- [`web_interaction_publication()`](interaction.qmd#web_interaction_publication) — quiet, print-clean interaction preset
