---
title: "Export"
---

```{r}
#| include: false
library(tabviz)
```

Export tabviz widgets to static file formats (SVG, PDF, PNG).

The two exported entry points are:

| Function | Use case |
|---|---|
| [`save_plot()`](save_plot.qmd#save_plot) | Export a single widget / `WebSpec` to SVG, PDF, or PNG. |
| [`save_split_table()`](save_plot.qmd#save_split_table) | Export every sub-plot of a `SplitForest` to a directory tree. |

Both delegate to a shared SVG generator (the same one the browser uses), then convert to PDF / PNG via `rsvg`. See [`save_plot()`](save_plot.qmd) for the full reference, dependencies, and resolution guidance.

```{r}
#| eval: false
p <- tabviz(data,
  label = "study",
  columns = list(
    viz_forest(point = "hr", lower = "lo", upper = "hi"),
    col_interval("hr", "lo", "hi", header = "HR (95% CI)")
  )
)

save_plot(p, "figures/forest.svg")
save_plot(p, "figures/forest.png", width = 1200, scale = 3)
save_plot(p, "figures/forest.pdf", width = 800)
```

:::{.callout-note}
## Backend dependencies

- SVG only: `V8` (always required).
- PNG / PDF: also requires the `rsvg` package.
- Multi-page PDF (paginated): also requires the `qpdf` package, which merges the per-page PDFs. Missing `qpdf` falls back to a numbered series (`out_p01.pdf`, …) with a warning.

There is no headless-browser dependency — exports run entirely in R via V8, rsvg, and qpdf.
:::

## Pagination

Long tables can span multiple PDF pages. Attach a [`paginate_spec()`](columns.qmd#paginate) to the `tabviz()` call (or pass `paginate = ...` to `save_plot()` to override per render):

```{r}
#| eval: false
p <- tabviz(data, label = "study",
            columns = list(viz_forest(point = "hr", lower = "lo", upper = "hi")),
            paginate = paginate_letter())   # 30 rows / page

save_plot(p, "figures/forest.pdf")          # multi-page PDF via qpdf::pdf_combine
save_plot(p, "figures/forest.png")          # warns + flattens to one image
save_plot(p, "figures/forest.png",
          paginate = NULL)                  # silence the flatten warning
```

For `SplitForest` outputs, pagination cascades into every subview — `save_plot()` writes one combined multi-page PDF per split leaf. See [Export → Pagination](../guide/export.qmd#pagination) in the guide for the full surface, defaults, and `qpdf`-missing fallback details.

## Related

- [`save_plot()`](save_plot.qmd) — the single-plot exporter and `save_split_table()` for split views
- [`split_table()`](split_table.qmd) — produce the `SplitForest` accepted by `save_split_table()`
- [Themes](themes.qmd) — theme presets aimed at print (`web_theme_jama()`, `web_theme_lancet()`, `web_theme_cochrane()`)
- [`web_interaction_publication()`](interaction.qmd#web_interaction_publication) — quiet, print-clean interaction preset
