---
title: "forest_plot()"
---

```{r}
#| include: false
library(tabviz)
```

`forest_plot()` is a convenience wrapper around [`tabviz()`](tabviz.qmd) that auto-builds a forest column from `point` / `lower` / `upper` (or an `effects` list) and forwards everything else to `tabviz()`. Use it for quick forest plots; reach for `tabviz()` directly when you want full control over column ordering, multiple viz columns, or column groups.

## Signature

```r
forest_plot(
  x,
  point = NULL, lower = NULL, upper = NULL,
  effects = NULL,
  label = NULL,
  group = NULL,
  columns = NULL,
  scale = NULL,            # "linear" or "log"
  null_value = NULL,       # 0 (linear) or 1 (log) by default
  axis_label = NULL,
  axis_range = NULL,
  axis_ticks = NULL,
  axis_gridlines = NULL,
  theme = NULL,
  ...,                     # forwarded to tabviz()
  split_by = NULL,
  shared_axis = FALSE,
  width = NULL, height = NULL, elementId = NULL
)
```

## Arguments

| Argument | Type | Default | Description |
|---|---|---|---|
| `x` | `data.frame` / `WebSpec` / `SplitForest` | required | Input. A data frame builds a fresh widget; a `WebSpec` is rendered directly; a `SplitForest` is rendered as a split widget. |
| `point` | `character(1)` | `NULL` | Column name for point estimates. Required when `x` is a data frame and `effects` is not provided. |
| `lower` | `character(1)` | `NULL` | Lower-bound column. Required with `point` / `upper`. |
| `upper` | `character(1)` | `NULL` | Upper-bound column. |
| `effects` | `list` of `effect_forest()` | `NULL` | Multi-effect spec; when provided, `point` / `lower` / `upper` are ignored. |
| `label` | `character(1)` | `NULL` | Column name for row labels. |
| `group` | `character` / `list` | `NULL` | Same shape as `tabviz(group=)`. |
| `columns` | `list` | `NULL` | Extra `col_*()` / `viz_*()` columns added alongside the auto-built forest column. |
| `scale` | `character(1)` | `"linear"` | `"linear"` or `"log"`. |
| `null_value` | `numeric(1)` | `0` (linear) / `1` (log) | Reference value for the null line. |
| `axis_label` | `character(1)` | `"Effect"` | X-axis label. |
| `axis_range` | `numeric(2)` | `NULL` | Forwarded to the auto-built `viz_forest()`. |
| `axis_ticks` | `numeric` | `NULL` | Forwarded to the auto-built `viz_forest()`. |
| `axis_gridlines` | `logical(1)` | `FALSE` | Forwarded to the auto-built `viz_forest()`. |
| `theme` | `WebTheme` | `NULL` | Theme; `NULL` falls through to `tabviz()`'s default. |
| `...` | — | — | Forwarded to [`tabviz()`](tabviz.qmd). All `tabviz()` row/marker styling, titles, and interactivity arguments are accepted. |
| `split_by` | `character` | `NULL` | Splits the result into a `SplitForest`. |
| `shared_axis` | `logical(1)` | `FALSE` | Share axis range across split sub-plots. |
| `width`, `height`, `elementId` | — | `NULL` | Standard htmlwidget passthroughs. |

:::{.callout-note}
## `axis_range` / `axis_ticks` / `axis_gridlines` are native here

In `tabviz()`, these arguments were deprecated in 0.9.0 and now live on `viz_forest()`. `forest_plot()` keeps them as native arguments and forwards them into the auto-built forest column — that's the whole point of the wrapper.
:::

## Return value

An htmlwidget. When `split_by` is set or `x` is a `SplitForest`, returns the split-table widget.

## Examples

```{r}
data <- data.frame(
  study = c("Study A", "Study B", "Study C", "Study D"),
  hr    = c(0.72, 0.85, 0.91, 0.68),
  lower = c(0.55, 0.70, 0.75, 0.52),
  upper = c(0.95, 1.03, 1.10, 0.89)
)

# Quick forest from a data frame
forest_plot(data, point = "hr", lower = "lower", upper = "upper",
            label = "study")
```

```{r}
# Log scale for ratios (HR / OR / RR)
forest_plot(data, point = "hr", lower = "lower", upper = "upper",
            label = "study",
            scale = "log", null_value = 1, axis_label = "Hazard Ratio")
```

```{r}
# Add data columns alongside the auto-built forest column
data$n <- c(1250, 980, 1560, 1100)

forest_plot(data, point = "hr", lower = "lower", upper = "upper",
            label = "study",
            scale = "log", null_value = 1,
            columns = list(
              col_n("n"),
              col_interval("hr", "lower", "upper", header = "HR (95% CI)")
            ))
```

```{r}
# Multi-effect plot (e.g. ITT vs Per-Protocol)
multi <- data.frame(
  study  = c("TRIAL-1", "TRIAL-2", "TRIAL-3"),
  itt_hr = c(0.78, 0.82, 0.74),
  itt_lo = c(0.65, 0.68, 0.61),
  itt_hi = c(0.94, 0.99, 0.90),
  pp_hr  = c(0.72, 0.78, 0.70),
  pp_lo  = c(0.58, 0.63, 0.56),
  pp_hi  = c(0.89, 0.97, 0.88)
)

forest_plot(multi,
  label = "study",
  effects = list(
    effect_forest("itt_hr", "itt_lo", "itt_hi", label = "ITT",          color = "#2563eb"),
    effect_forest("pp_hr",  "pp_lo",  "pp_hi",  label = "Per-Protocol", color = "#16a34a")
  ),
  scale = "log", null_value = 1
)
```

```{r}
# Split view with shared axis
regional <- data.frame(
  study  = rep(c("Study 1", "Study 2", "Study 3"), 2),
  region = rep(c("North America", "Europe"), each = 3),
  hr     = c(0.72, 0.85, 0.78, 0.68, 0.82, 0.75),
  lo     = c(0.55, 0.70, 0.62, 0.52, 0.68, 0.58),
  hi     = c(0.95, 1.03, 0.98, 0.89, 0.99, 0.97)
)

forest_plot(regional, point = "hr", lower = "lo", upper = "hi",
            label = "study",
            scale = "log", null_value = 1,
            split_by = "region", shared_axis = TRUE)
```

## Equivalence with `tabviz()`

```r
forest_plot(data, point = "hr", lower = "lo", upper = "hi",
            label = "study", scale = "log", null_value = 1)

# Equivalent to:
tabviz(data, label = "study",
  columns = list(
    viz_forest(point = "hr", lower = "lo", upper = "hi",
               scale = "log", null_value = 1)
  )
)
```

## Related

- [`tabviz()`](tabviz.qmd) — full control over columns and layout
- [`viz_forest()`](columns.qmd#viz_forest) — the underlying forest column helper
- [`effect_forest()`](columns.qmd#effect_forest) — define one effect for multi-effect plots
- [`split_table()`](split_table.qmd) — build a `SplitForest` from a spec or widget
- [`save_plot()`](save_plot.qmd) — export to SVG/PNG/PDF
