---
title: "Interaction Settings"
---

```{r}
#| include: false
library(tabviz)
library(dplyr)
```

Control interactive behavior of tabviz widgets using `web_interaction()` and preset functions.

## Overview

All tabviz widgets are interactive by default. Use interaction settings to:

- Enable/disable specific interactions
- Configure hover tooltips
- Control theme selection menu
- Create static publication-ready output

## Quick Reference

| Function | Use Case |
|----------|----------|
| `web_interaction()` | Full control over all settings |
| `web_interaction_minimal()` | Hover only, no sorting/selection |
| `web_interaction_publication()` | Fully static for print |

---

## web_interaction()

Create a custom interaction specification:

```{r}
#| eval: false
tabviz(data, label = "study",
  interaction = web_interaction(
    show_legend = TRUE,
    enable_sort = TRUE,
    enable_collapse = TRUE,
    enable_select = TRUE,
    enable_hover = TRUE,
    enable_resize = TRUE,
    enable_export = TRUE,
    enable_filters = TRUE,
    enable_reorder_rows = TRUE,
    enable_reorder_columns = TRUE,
    enable_edit = TRUE,
    tooltip_fields = c("n", "pvalue"),
    enable_themes = "default"
  ),
  columns = list(...)
)
```

`web_interaction()` defaults enable the full interactive experience — sort, collapse, hover, resize, export, filters, row/column reorder, and inline editing all on. Pass individual flags to opt out of specific features, or use `web_interaction_minimal()` / `web_interaction_publication()` for quieter presets.

### Arguments

| Argument | Type | Default | Description |
|----------|------|---------|-------------|
| `show_legend` | logical | TRUE | Show legend for multi-effect plots |
| `enable_sort` | logical | TRUE | Click headers to sort |
| `enable_collapse` | logical | TRUE | Click group headers to collapse/expand |
| `enable_select` | logical | TRUE | Click rows to select |
| `enable_hover` | logical | TRUE | Highlight rows on hover |
| `enable_resize` | logical | TRUE | Drag column borders to resize |
| `enable_export` | logical | TRUE | Show download button on hover |
| `enable_filters` | logical | TRUE | Per-column filter popovers in headers |
| `enable_reorder_rows` | logical | TRUE | Drag rows within a group, group-siblings |
| `enable_reorder_columns` | logical | TRUE | Drag columns within a group, group-siblings |
| `enable_edit` | logical | TRUE | Double-click cells to edit; popover editor for forest numerics |
| `tooltip_fields` | character vector | NULL | Column names for hover tooltip |
| `enable_themes` | string, NULL, or list | "default" | Theme menu control |

### Theme Menu Control

The `enable_themes` argument accepts three types of values:

```r
# Default: Show menu with all package themes
web_interaction(enable_themes = "default")

# Disable: Hide theme menu entirely
web_interaction(enable_themes = NULL)

# Custom: Show menu with specific themes only
web_interaction(enable_themes = list(
  web_theme_jama(),
  web_theme_nature(),
  web_theme_modern()
))
```

---

## web_interaction_minimal()

Preset for viewing with minimal interaction—hover highlighting only:

```{r}
data(glp1_trials)

glp1_trials |>
  filter(row_type == "data") |>
  head(4) |>
  tabviz(label = "study",
    interaction = web_interaction_minimal(),
    columns = list(
      viz_forest(point = "hr", lower = "lower", upper = "upper",
                 scale = "log", null_value = 1)
    )
  )
```

Equivalent to:

```r
web_interaction(
  enable_sort = FALSE,
  enable_collapse = FALSE,
  enable_select = FALSE,
  enable_hover = TRUE,
  enable_resize = FALSE,
  enable_export = FALSE
)
```

---

## web_interaction_publication()

Preset for static, publication-ready output:

```{r}
#| eval: false
tabviz(data, label = "study",
  interaction = web_interaction_publication(),
  columns = list(
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1)
  )
)
```

Equivalent to:

```r
web_interaction(
  show_legend = FALSE,
  enable_sort = FALSE,
  enable_collapse = FALSE,
  enable_select = FALSE,
  enable_hover = FALSE,
  enable_resize = FALSE,
  enable_export = FALSE
)
```

Use this when:

- Exporting for print publications
- Embedding in static documents
- Screenshots for presentations

---

## Hover Tooltips

Add custom tooltips that appear when hovering over markers:

```{r}
data(glp1_trials)

tooltip_demo <- glp1_trials |>
  filter(row_type == "data") |>
  head(4)

tabviz(tooltip_demo, label = "study",
  interaction = web_interaction(
    tooltip_fields = c("drug", "n", "events", "pvalue")
  ),
  footnote = "Hover over markers to see details",
  columns = list(
    viz_forest(point = "hr", lower = "lower", upper = "upper",
               scale = "log", null_value = 1),
    col_interval("HR (95% CI)", point = "hr", lower = "lower", upper = "upper")
  )
)
```

::: {.callout-note}
Tooltips are opt-in. Without `tooltip_fields`, no tooltip appears on hover.
:::

---

## Examples

### Disable Sorting Only

```r
tabviz(data, label = "study",
  interaction = web_interaction(enable_sort = FALSE),
  columns = list(...)
)
```

### Disable Export Button

```r
tabviz(data, label = "study",
  interaction = web_interaction(enable_export = FALSE),
  columns = list(...)
)
```

### Custom Theme Menu

```r
tabviz(data, label = "study",
  interaction = web_interaction(
    enable_themes = list(
      web_theme_jama(),
      web_theme_lancet()
    )
  ),
  columns = list(...)
)
```

---

## See Also

- [Visualizations Overview](../guide/visualizations/index.qmd#interactivity) — Interactivity guide
- [Export](export.qmd) — Exporting to SVG/PNG/PDF
- [`tabviz()`](tabviz.qmd) — Main entry point
