---
title: "Interaction"
---

```{r}
#| include: false
library(tabviz)
```

`web_interaction()` builds an `InteractionSpec` controlling what users can do with a tabviz widget — sorting, filtering, reordering, editing, exporting, theme switching. Three presets cover common cases (`_full`, `_minimal`, `_publication`).

:::{.callout-tip}
## Naming convention: `show_*` vs `enable_*`

- **`show_*`** controls whether a piece of static UI chrome is rendered at all (e.g. `show_legend`, `show_group_counts`). Toggling it on/off doesn't expose a runtime control to bring it back.
- **`enable_*`** controls whether a user *capability* is on (sorting, filtering, resizing, editing, reordering, exporting, the theme menu). The associated UI is rendered when the capability is on, hidden when off.

When adding new arguments, follow the same rule: render-or-not is `show_`, can-the-user-do-it is `enable_`.
:::

---

## `web_interaction()` {#web_interaction}

```r
web_interaction(
  show_legend = TRUE,
  enable_sort = TRUE,
  enable_collapse = TRUE,
  enable_select = TRUE,
  enable_hover = TRUE,
  enable_resize = TRUE,
  enable_export = TRUE,
  enable_filters = TRUE,
  enable_reorder_rows = TRUE,
  enable_reorder_columns = TRUE,
  enable_edit = TRUE,
  show_group_counts = FALSE,
  tooltip_fields = NULL,
  enable_themes = getOption("tabviz.enable_themes", "default")
)
```

| Argument | Type | Default | Description |
|---|---|---|---|
| `show_legend` | `logical(1)` | `TRUE` | Render the legend (multi-effect plots). |
| `enable_sort` | `logical(1)` | `TRUE` | Click headers to sort. |
| `enable_collapse` | `logical(1)` | `TRUE` | Click row-group headers to collapse / expand. |
| `enable_select` | `logical(1)` | `TRUE` | Click rows to select. |
| `enable_hover` | `logical(1)` | `TRUE` | Highlight rows on hover. |
| `enable_resize` | `logical(1)` | `TRUE` | Drag column borders to resize. |
| `enable_export` | `logical(1)` | `TRUE` | Show download / export button on hover. |
| `enable_filters` | `logical(1)` | `TRUE` | Per-column filter popovers in headers. |
| `enable_reorder_rows` | `logical(1)` | `TRUE` | Drag rows within a group, and row-groups among siblings. |
| `enable_reorder_columns` | `logical(1)` | `TRUE` | Drag columns within a group, and column-groups among top-level siblings. |
| `enable_edit` | `logical(1)` | `TRUE` | Double-click cells to edit; popover editor for forest-cell numerics. Also enables inline editing of title / subtitle / caption / footnote (v0.15+). |
| `show_group_counts` | `logical(1)` | `FALSE` | Append `(n)` after each row-group label. |
| `tooltip_fields` | `character` | `NULL` | Column names shown in the hover tooltip. `NULL` disables tooltips. |
| `enable_themes` | `"default"` / `NULL` / `list` of `WebTheme` | `getOption("tabviz.enable_themes", "default")` | Curate the theme switcher (see below). |

**Returns:** an `InteractionSpec`.

:::{.callout-important}
## `show_filters` was renamed in 0.9.0

The argument is now `enable_filters` — the column-filter UI is a user *capability*, not a static visibility toggle. The old `show_filters` is still accepted for one cycle and emits `lifecycle::deprecate_warn()`.
:::

### `enable_themes`

Curates the in-widget theme switcher.

| Value | Effect |
|---|---|
| `"default"` | Show every theme returned by `package_themes()`. |
| `NULL` | Hide the switcher entirely. |
| `list(WebTheme, ...)` | Show exactly those themes. **Named** entries override the displayed name (e.g. `list(Classical = web_theme_jama())`). The spec's active theme is auto-included so users can always revert. |

```r
# Two-theme curated menu, custom display names
web_interaction(enable_themes = list(
  Classical = web_theme_jama(),
  Modern    = web_theme_modern()
))
```

The fluent equivalent on a built spec is [`selectable_themes()`](styling.qmd#selectable_themes).

---

## `web_interaction_full()` {#web_interaction_full}

All capabilities on. This is what `default_interaction_for_theme()` returns, and what `tabviz()` falls back to when `interaction = NULL`.

```r
web_interaction_full()
```

---

## `web_interaction_minimal()` {#web_interaction_minimal}

Read-only with hover highlighting only.

```r
web_interaction_minimal()
# Equivalent to web_interaction() with every enable_* set to FALSE
# except enable_hover = TRUE; show_legend kept TRUE.
```

---

## `web_interaction_publication()` {#web_interaction_publication}

Print-clean: nothing is interactive, no legend chrome.

```r
web_interaction_publication()
# Every show_* and enable_* set to FALSE.
```

Use when:

- Exporting screenshots for print publications.
- Embedding in static documents where any UI noise is unwelcome.

---

## `default_interaction_for_theme()` {#default_interaction_for_theme}

```r
default_interaction_for_theme(theme)
```

Returns `web_interaction_full()` regardless of theme — every shipped theme defaults to full interactivity. The function is exported for use in custom widget builders. Pass `interaction = web_interaction_publication()` (or `_minimal()`) explicitly for a quieter widget.

---

## Examples

```{r}
data(glp1_trials)

# Hover-only preset
glp1_trials |>
  dplyr::filter(row_type == "data") |>
  head(4) |>
  tabviz(label = "study",
    interaction = web_interaction_minimal(),
    columns = list(
      viz_forest(point = "hr", lower = "lower", upper = "upper",
                 scale = "log", null_value = 1)
    ))
```

```{r}
# Hover tooltips with selected fields
glp1_trials |>
  dplyr::filter(row_type == "data") |>
  head(4) |>
  tabviz(label = "study",
    interaction = web_interaction(tooltip_fields = c("drug", "n", "events", "pvalue")),
    footnote = "Hover over markers to see details",
    columns = list(
      viz_forest(point = "hr", lower = "lower", upper = "upper",
                 scale = "log", null_value = 1),
      col_interval("hr", "lower", "upper", header = "HR (95% CI)")
    ))
```

```r
# Disable just one capability
tabviz(data, label = "study",
  interaction = web_interaction(enable_sort = FALSE),
  columns = list(...))

# Curated theme menu via the constructor
tabviz(data, label = "study",
  interaction = web_interaction(enable_themes = list(
    web_theme_jama(), web_theme_lancet()
  )),
  columns = list(...))
```

---

## Related

- [`selectable_themes()`](styling.qmd#selectable_themes) — fluent equivalent for the `enable_themes` argument
- [`set_theme()`](styling.qmd#set_theme) — change the active theme
- [Themes](themes.qmd) — `package_themes()` and the `web_theme_*()` builders
- [`tabviz()`](tabviz.qmd) — the `interaction` argument
