---
title: "Shiny Integration"
---

```{r}
#| include: false
library(tabviz)
```

Functions for embedding tabviz plots in Shiny applications and updating them
in place from server code.

## Output and Render

### tabvizOutput {#tabvizOutput}

Create a Shiny output element for a tabviz plot.

```r
tabvizOutput(outputId, width = "100%", height = "400px")
```

| Argument | Description |
|----------|-------------|
| `outputId` | Output variable name |
| `width` | Widget width (CSS units) |
| `height` | Widget height (CSS units) |

### renderTabviz {#renderTabviz}

Render a tabviz plot in Shiny.

```r
renderTabviz(expr, env = parent.frame(), quoted = FALSE)
```

| Argument | Description |
|----------|-------------|
| `expr` | Expression that returns a widget from `tabviz()` |
| `env` | Environment for evaluation |
| `quoted` | Whether `expr` is already quoted |

## Proxy Updates

### tabviz_proxy {#tabviz_proxy}

Create a proxy to update a running tabviz widget without re-rendering it.
Pass the proxy to any fluent verb (`sort_rows()`, `filter_rows()`,
`add_column()`, `set_theme()`, …) to dispatch an update.

```r
tabviz_proxy(id, session = shiny::getDefaultReactiveDomain())
```

| Argument | Description |
|----------|-------------|
| `id` | The widget (output) ID |
| `session` | Shiny session (default: current reactive domain) |

Returns a `tabviz_proxy` object. Always create proxies inside a reactive
context (`observe`, `observeEvent`, …), never at the top of `server`.

### Verbs that accept a proxy

All fluent verbs dispatch on input type: static `WebSpec` / `htmlwidget`
objects are mutated in R; `tabviz_proxy` objects send a custom message to
the browser. See the full list in [Styling](styling.qmd) and the
[Fluent API guide](../guide/fluent-api.qmd).

| Category | Verbs |
|----------|-------|
| Columns  | `add_column()`, `remove_column()`, `move_column()`, `resize_column()`, `update_column()` |
| Rows     | `sort_rows()`, `filter_rows()`, `clear_filters()`, `select_rows()`, `move_row()` |
| Data     | `update_data()` (proxy requires a `WebSpec`) |
| Groups   | `toggle_group()` |
| Cells    | `set_cell()`, `set_row_label()`, `clear_edits()` (proxy-only) |
| Global   | `set_theme()`, `set_zoom()`, styling `set_*` verbs |

## Split-Table Bindings

### splitTabvizOutput {#splitTabvizOutput}

```r
splitTabvizOutput(outputId, width = "100%", height = "600px")
```

### renderSplitTabviz {#renderSplitTabviz}

```r
renderSplitTabviz(expr, env = parent.frame(), quoted = FALSE)
```

`expr` should return a widget created from a `split_table()` result.

### split_tabviz_proxy {#split_tabviz_proxy}

```r
split_tabviz_proxy(id, session = shiny::getDefaultReactiveDomain())
```

Returns a `split_tabviz_proxy` object.

### split_tabviz_select {#split_tabviz_select}

Select a sub-plot in a split-table tabviz.

```r
split_tabviz_select(proxy, key)
```

| Argument | Description |
|----------|-------------|
| `proxy` | A `split_tabviz_proxy` object |
| `key` | Key of the sub-plot (e.g., `"Americas"` or `"Female/Young"`) |

## Examples

### Basic Shiny app

```{r}
#| eval: false
library(shiny)
library(tabviz)

ui <- fluidPage(
  titlePanel("Forest Plot Demo"),
  sidebarLayout(
    sidebarPanel(
      selectInput("theme", "Theme:",
        choices = c("Default", "JAMA", "Lancet", "Modern")
      )
    ),
    mainPanel(
      tabvizOutput("plot", height = "500px")
    )
  )
)

server <- function(input, output, session) {
  data <- data.frame(
    study = c("Study A", "Study B", "Study C"),
    hr    = c(0.72, 0.85, 0.91),
    lower = c(0.55, 0.70, 0.75),
    upper = c(0.95, 1.03, 1.10)
  )

  output$plot <- renderTabviz({
    theme <- switch(input$theme,
      "Default" = web_theme_default(),
      "JAMA"    = web_theme_jama(),
      "Lancet"  = web_theme_lancet(),
      "Modern"  = web_theme_modern()
    )

    tabviz(data,
      label = "study",
      columns = list(
        viz_forest(point = "hr", lower = "lower", upper = "upper",
                   scale = "log", null_value = 1)
      ),
      theme = theme
    )
  })
}

shinyApp(ui, server)
```

### Proxy updates (no re-render)

```{r}
#| eval: false
library(shiny)
library(tabviz)

ui <- fluidPage(
  tabvizOutput("plot"),
  actionButton("filter_btn", "Filter HR < 1"),
  actionButton("clear_btn",  "Clear Filter"),
  actionButton("sort_btn",   "Sort HR desc")
)

server <- function(input, output, session) {
  data <- data.frame(
    study = c("Study A", "Study B", "Study C", "Study D"),
    hr    = c(0.72, 0.85, 1.15, 0.91),
    lower = c(0.55, 0.70, 0.90, 0.75),
    upper = c(0.95, 1.03, 1.45, 1.10)
  )

  output$plot <- renderTabviz({
    tabviz(data,
      label = "study",
      columns = list(
        viz_forest(point = "hr", lower = "lower", upper = "upper",
                   scale = "log", null_value = 1)
      )
    )
  })

  proxy <- tabviz_proxy("plot")

  observeEvent(input$filter_btn, filter_rows(proxy, "hr", "lt", 1))
  observeEvent(input$clear_btn,  clear_filters(proxy))
  observeEvent(input$sort_btn,   sort_rows(proxy, "hr", "desc"))
}
```

### Replacing the dataset

`update_data()` on a proxy requires a `WebSpec` (not a raw data frame) so
the new columns/types can be serialized. Build one with `tabviz(..., .spec_only = TRUE)`.

```{r}
#| eval: false
observeEvent(input$switch_dataset, {
  new_spec <- tabviz(
    get_dataset(input$dataset_name),
    label = "study",
    columns = list(
      viz_forest(point = "hr", lower = "lower", upper = "upper",
                 scale = "log", null_value = 1)
    ),
    .spec_only = TRUE
  )
  update_data(tabviz_proxy("plot"), new_spec)
})
```

## See Also

- [Shiny guide](../guide/shiny.qmd) — reactive patterns and a complete example.
- [Fluent API guide](../guide/fluent-api.qmd) — full verb list and static/proxy dispatch.
- [`tabviz()`](tabviz.qmd) — main entry point.
