---
title: "split_table()"
---

```{r}
#| include: false
library(tabviz)
```

`split_table()` slices a spec or widget into a `SplitForest` — one sub-plot per unique value (or combination of values) of the splitting column(s). The resulting object renders with sidebar navigation in the browser, can be exported one file per sub-plot via `save_split_table()`, and supports a shared-axis option for cross-subgroup comparison.

The same effect is available inline on `tabviz()` and `forest_plot()` via the `split_by` / `shared_axis` arguments — `split_table()` is the explicit constructor used when you want a `SplitForest` object to inspect, modify, or hand to `save_split_table()`.

## Signature

```r
split_table(x, by, shared_axis = FALSE, ...)
```

## Arguments

| Argument | Type | Default | Description |
|---|---|---|---|
| `x` | `WebSpec` / htmlwidget / `data.frame` | required | Source. A data frame is fed through `web_spec()` first. |
| `by` | `character` | required | Column name(s) to split on. A single string creates one plot per unique value; a vector creates a hierarchical navigation tree (outermost first). |
| `shared_axis` | `logical(1)` | `FALSE` | When `TRUE`, all sub-plots use the same x-axis range so subgroup effects can be compared at a glance. |
| `...` | — | — | Forwarded to `web_spec()` when `x` is a data frame. |

## Return value

A `SplitForest` containing one `WebSpec` per split combination, the navigation tree, the split variables, and (when `shared_axis = TRUE`) the shared axis range.

## `tabviz()` / `forest_plot()` shortcut

Both top-level entry points accept `split_by` and `shared_axis` directly. They build a `WebSpec` first, call `split_table()` internally, then render the result.

```r
# Equivalent
tabviz(data, label = "study",
       columns = list(viz_forest("or", "lo", "hi")),
       split_by = "region", shared_axis = TRUE)

tabviz(data, label = "study",
       columns = list(viz_forest("or", "lo", "hi")),
       .spec_only = TRUE) |>
  split_table(by = "region", shared_axis = TRUE) |>
  forest_plot()
```

## Examples

```{r}
set.seed(42)
data <- data.frame(
  study  = paste0("Study ", 1:30),
  region = sample(c("Americas", "Europe", "Asia"), 30, replace = TRUE),
  or     = exp(rnorm(30, log(0.8), 0.3))
)
data$lower <- data$or * exp(-1.96 * 0.25)
data$upper <- data$or * exp( 1.96 * 0.25)

forest_plot(data,
  point = "or", lower = "lower", upper = "upper",
  label = "study",
  split_by = "region",
  scale = "log", null_value = 1,
  theme = web_theme_modern()
)
```

```{r}
#| eval: false
# Hierarchical split with nested sidebar
forest_plot(data,
  point = "or", lower = "lower", upper = "upper",
  label = "study",
  split_by = c("sex", "age_group"),
  scale = "log", null_value = 1
)
```

```{r}
#| eval: false
# Build the SplitForest explicitly to export
sf <- tabviz(data, label = "study",
             columns = list(viz_forest("or", "lower", "upper")),
             .spec_only = TRUE) |>
  split_table(by = c("sex", "age_group"), shared_axis = TRUE)

sf                              # printed summary of the SplitForest
save_split_table(sf, "out/")    # one file per sub-plot
```

## Related

- [`tabviz()`](tabviz.qmd) and [`forest_plot()`](forest_plot.qmd) — both accept `split_by` directly
- [`save_split_table()`](save_plot.qmd#save_split_table) — export every sub-plot to a directory tree
- [Shiny bindings](shiny.qmd) — `splitTabvizOutput()`, `renderSplitTabviz()`, `split_tabviz_proxy()`
