---
title: "Styling modifiers"
---

```{r}
#| include: false
library(tabviz)
```

Fluent verbs that mutate a `WebSpec`, htmlwidget, or live Shiny `tabviz_proxy` to add row, column, marker, theme, and zoom styling. Every modifier returns the same kind of object it received, so they compose cleanly with `|>`.

For *theme*-level modifiers (palettes, typography, axis, layout) see [Themes](themes.qmd).

:::{.callout-tip}
## Constructor arguments vs fluent API

Most modifiers have an equivalent argument on `tabviz()` (e.g. `set_row_style(bold = ...)` ↔ `tabviz(row_bold = ...)`). Use the constructor argument when the value is known up front; reach for the fluent verb when styling depends on later computation, when you're producing variants, or when you want to layer styling on top of an existing widget.
:::

---

## `set_row_style()` {#set_row_style}

Apply row-level styling driven by columns in the data.

```r
set_row_style(
  x,
  bold = NULL,
  italic = NULL,
  color = NULL,
  bg = NULL,
  badge = NULL,
  icon = NULL,
  indent = NULL,
  type = NULL,
  weight = NULL,
  emphasis = NULL,
  muted = NULL,
  accent = NULL
)
```

| Argument | Type | Default | Description |
|---|---|---|---|
| `x` | `WebSpec` / htmlwidget | required | Spec or widget. |
| `bold` | column name | `NULL` | Logical column for bold rows. |
| `italic` | column name | `NULL` | Logical column for italic rows. |
| `color` | column name | `NULL` | CSS color column for row text. |
| `bg` | column name | `NULL` | CSS color column for row background. |
| `badge` | column name | `NULL` | Text column rendered as a badge on the label cell. |
| `icon` | column name | `NULL` | Emoji/unicode column rendered on the label cell. |
| `indent` | column name | `NULL` | Numeric column for label indentation. |
| `type` | column name | `NULL` | One of `"data"`, `"header"`, `"summary"`, `"spacer"`. |
| `weight` | column name | `NULL` | Numeric column driving marker size. Same shape as `marker_size`. |
| `emphasis` | column name | `NULL` | Logical column → bold + foreground color. |
| `muted` | column name | `NULL` | Logical column → muted styling. |
| `accent` | column name | `NULL` | Logical column → accent color. |

**Returns:** modified input (`WebSpec` or htmlwidget).

```{r}
styled <- data.frame(
  label = c("Primary Endpoint", "  CV Death", "  MI", "Summary"),
  hr = c(NA, 0.82, 0.79, 0.78),
  lo = c(NA, 0.72, 0.68, 0.65),
  hi = c(NA, 0.94, 0.92, 0.93),
  rtype = c("header", "data", "data", "summary"),
  rbold = c(TRUE, FALSE, FALSE, TRUE)
)

tabviz(styled, label = "label",
  columns = list(viz_forest(point = "hr", lower = "lo", upper = "hi",
                            scale = "log", null_value = 1))
) |>
  set_row_style(type = "rtype", bold = "rbold")
```

---

## `set_column_style()` {#set_column_style}

Per-cell styling targeting a single column.

```r
set_column_style(
  x, column,
  bold = NULL, italic = NULL,
  color = NULL, bg = NULL,
  badge = NULL, icon = NULL
)
```

| Argument | Type | Default | Description |
|---|---|---|---|
| `x` | `WebSpec` / htmlwidget | required | |
| `column` | `character(1)` | required | Field name of the target column. |
| `bold` / `italic` | column name | `NULL` | Logical-column driver. |
| `color` / `bg` | column name | `NULL` | CSS color column. |
| `badge` / `icon` | column name | `NULL` | Per-cell badge / icon. |

When both row and column styles target the same cell, the column style wins.

```r
tabviz(data, label = "study",
  columns = list(col_pvalue("pval"))
) |>
  set_column_style("pval", bold = "is_sig", color = "sig_color")
```

---

## `set_marker_style()` {#set_marker_style}

Style the primary effect's interval markers.

```r
set_marker_style(x,
                 color = NULL, shape = NULL,
                 opacity = NULL, size = NULL)
```

| Argument | Type | Default | Description |
|---|---|---|---|
| `color` | column name | `NULL` | CSS color column for marker fill. |
| `shape` | column name | `NULL` | Column of `"square"` / `"circle"` / `"diamond"` / `"triangle"`. |
| `opacity` | column name | `NULL` | Numeric column 0–1. |
| `size` | column name | `NULL` | Numeric weight-like column (e.g. study weights 0–100). The renderer normalizes via `0.5 + sqrt(w/100) * 1.5`. Pass raw weights, not a pre-normalized multiplier. |

For multi-effect plots, additional effects use the per-effect color/shape/opacity from `effect_forest()`.

---

## `set_zoom()` {#set_zoom}

Zoom level and container constraints. Works on htmlwidgets and live `tabviz_proxy` objects (it is **not** a `WebSpec` modifier — apply it after rendering).

```r
set_zoom(x,
         zoom = 1.0,
         auto_fit = TRUE,
         max_width = NULL,
         max_height = NULL,
         show_controls = TRUE)
```

| Argument | Type | Default | Description |
|---|---|---|---|
| `x` | htmlwidget / `tabviz_proxy` | required | |
| `zoom` | `numeric(1)` | `1.0` | 0.5–2.0. |
| `auto_fit` | `logical(1)` | `TRUE` | Shrink content to fit container if too large; never enlarges. |
| `max_width` | `numeric(1)` | `NULL` | Max container width in px (≥ 100). |
| `max_height` | `numeric(1)` | `NULL` | Max container height in px (≥ 100). |
| `show_controls` | `logical(1)` | `TRUE` | Show zoom controls on hover. |

:::{.callout-important}
## Replaces `tabviz()` zoom arguments

`tabviz(zoom = ...)`, `tabviz(auto_fit = ...)`, `tabviz(max_width = ...)`, `tabviz(max_height = ...)`, and `tabviz(show_zoom_controls = ...)` were deprecated in 0.9.0. Pipe to `set_zoom()` after constructing the widget instead.
:::

```r
tabviz(data, label = "study", columns = list(viz_forest("hr", "lo", "hi"))) |>
  set_zoom(zoom = 0.8, max_width = 900)
```

---

## `set_theme()` {#set_theme}

Swap the active theme on a `WebSpec`, htmlwidget, or `tabviz_proxy`. On a proxy, sends a message so the widget swaps themes without a full re-render.

```r
set_theme(x, theme)
```

| Argument | Type | Description |
|---|---|---|
| `x` | `WebSpec` / htmlwidget / `tabviz_proxy` | |
| `theme` | character or `WebTheme` | Either a built-in name or a `web_theme_*()` object. |

**Built-in names:** `"default"`, `"minimal"`, `"dark"`, `"jama"`, `"lancet"`, `"modern"`, `"presentation"`, `"cochrane"`, `"nature"`.

```r
tabviz(data, label = "study",
       columns = list(viz_forest(point = "hr", lower = "lo", upper = "hi"))) |>
  set_theme("jama")
```

---

## `selectable_themes()` {#selectable_themes}

Curate which themes appear in the in-widget theme switcher. This is the fluent sibling of the `enable_themes` argument of [`web_interaction()`](interaction.qmd).

```r
selectable_themes(x, themes)
```

| Argument | Type | Description |
|---|---|---|
| `x` | `WebSpec` or htmlwidget | Live proxy not supported. |
| `themes` | `NULL` / `"default"` / list of `WebTheme` | `NULL` hides the switcher; `"default"` shows all `package_themes()`; a list shows exactly those (named entries override displayed names). The active theme is always auto-included so the user can revert. |

```r
tabviz(data, label = "study") |>
  set_theme("jama") |>
  selectable_themes(list(
    Classical = web_theme_jama(),
    Modern    = web_theme_cochrane()
  ))
```

---

## Quick reference

| Modifier | Operates on | Purpose |
|---|---|---|
| `set_row_style()` | `WebSpec`, htmlwidget | Row-level styling and row types. |
| `set_column_style()` | `WebSpec`, htmlwidget | Per-cell styling for a single column. |
| `set_marker_style()` | `WebSpec`, htmlwidget | Marker color / shape / opacity / size driver. |
| `set_zoom()` | htmlwidget, `tabviz_proxy` | Zoom and container size constraints. |
| `set_theme()` | `WebSpec`, htmlwidget, `tabviz_proxy` | Swap the active theme. |
| `selectable_themes()` | `WebSpec`, htmlwidget | Curate the in-widget theme switcher. |

## Related

- [Themes](themes.qmd) — `set_inputs()`, `set_variants()`, `set_spacing()`, `set_theme_field()` for the v2 theme builder
- [Data ops](data-ops.qmd) — `add_column()`, `sort_rows()`, `filter_rows()`, …
- [Interaction](interaction.qmd) — `web_interaction()` and the `show_*` / `enable_*` rule
- [`tabviz()`](tabviz.qmd) — equivalent constructor arguments
