---
title: "tabviz()"
---

```{r}
#| include: false
library(tabviz)
```

`tabviz()` is the main entry point for building interactive table visualizations. It returns an htmlwidget that renders in R Markdown, Quarto, Shiny, or the RStudio Viewer. Use `tabviz()` whenever you want full control over which columns appear (text, numeric, badges, forest plots, sparklines, etc.). For a quick forest-only plot driven by `point` / `lower` / `upper` arguments, see [`forest_plot()`](forest_plot.qmd).

## Signature

```r
tabviz(
  data,
  label = NULL,
  label_header = "Study",
  group = NULL,
  columns = NULL,
  extra_columns = NULL,
  available_exclude = NULL,
  title = NULL, subtitle = NULL, caption = NULL, footnote = NULL,
  # Row styling (column name or formula)
  row_bold = NULL, row_italic = NULL,
  row_color = NULL, row_bg = NULL,
  row_badge = NULL, row_icon = NULL,
  row_indent = NULL, row_type = NULL,
  row_emphasis = NULL, row_muted = NULL, row_accent = NULL,
  # Marker styling (column name or formula)
  marker_color = NULL, marker_shape = NULL,
  marker_opacity = NULL, marker_size = NULL,
  # Theme & interaction
  theme = web_theme_default(),
  interaction = NULL,
  plot_position = NULL,
  row_height = NULL,
  banding = NULL,
  width = NULL, height = NULL, elementId = NULL,
  # Splitting
  split_by = NULL, shared_axis = FALSE,
  .spec_only = FALSE
)
```

## Arguments

### Data

| Argument | Type | Default | Description |
|---|---|---|---|
| `data` | `data.frame` | required | A data.frame, data.table, or tibble. |
| `label` | `character(1)` | `NULL` | Column name for row labels. When `NULL`, row numbers are used. |
| `label_header` | `character(1)` | `"Study"` | Header text for the label column. The literal default `"Study"` is auto-prettified from the `label` field name when applicable. |
| `group` | `character` / `list` | `NULL` | Grouping spec. A single column name (flat), a vector (hierarchical), or a list of `web_group()` objects. See [Grouping](#grouping). |
| `columns` | `list` | `NULL` | List of column specs from `col_*()` / `viz_*()` helpers. |
| `extra_columns` | `list` | `NULL` | Pre-configured columns hidden by default but available in the interactive column picker. |
| `available_exclude` | `character` | `NULL` | Data columns to omit from the column picker (e.g. internal/sensitive fields). |

### Titles

| Argument | Type | Default | Description |
|---|---|---|---|
| `title` | `character(1)` | `NULL` | Main title (above the plot). |
| `subtitle` | `character(1)` | `NULL` | Subtitle (below title). |
| `caption` | `character(1)` | `NULL` | Caption (below plot). |
| `footnote` | `character(1)` | `NULL` | Footnote (below caption, italicized). |

### Row styling

All row-styling arguments accept either a **column name (string)** that holds the per-row value, or a **formula** evaluated against `data` (e.g. `~ pvalue < 0.05`).

| Argument | Value type | Default | Description |
|---|---|---|---|
| `row_bold` | logical | `NULL` | Bold whole row. |
| `row_italic` | logical | `NULL` | Italicize row. |
| `row_color` | CSS color | `NULL` | Row text color. |
| `row_bg` | CSS color | `NULL` | Row background color. |
| `row_badge` | character | `NULL` | Badge text on the label cell. |
| `row_icon` | emoji/unicode | `NULL` | Icon on the label cell. |
| `row_indent` | numeric | `NULL` | Indentation level. |
| `row_type` | character | `NULL` | One of `"data"`, `"header"`, `"summary"`, `"spacer"`. |
| `row_emphasis` | logical | `NULL` | Bold + foreground color. |
| `row_muted` | logical | `NULL` | Muted (lighter, reduced prominence). |
| `row_accent` | logical | `NULL` | Theme accent color. |

### Marker styling

Same column-name-or-formula pattern as row styling:

| Argument | Value type | Default | Description |
|---|---|---|---|
| `marker_color` | CSS color | `NULL` | Marker fill color. |
| `marker_shape` | character | `NULL` | One of `"square"`, `"circle"`, `"diamond"`, `"triangle"`. |
| `marker_opacity` | numeric `[0,1]` | `NULL` | Marker opacity. |
| `marker_size` | numeric | `NULL` | Driver for marker size — pass raw weights (e.g. study weights 0–100). The renderer normalizes via `0.5 + sqrt(w/100) * 1.5`. |

### Theme, interaction, layout

| Argument | Type | Default | Description |
|---|---|---|---|
| `theme` | `WebTheme` | `web_theme_default()` | Theme object (use a `web_theme_*()` builder). |
| `interaction` | `InteractionSpec` | `NULL` | Use `web_interaction()`. When `NULL`, `default_interaction_for_theme(theme)` is used (full interactivity). |
| `plot_position` | `character(1)` | `NULL` | `"left"` or `"right"`. |
| `row_height` | `numeric(1)` | `NULL` | Row height in pixels. |
| `banding` | `character(1)` | `NULL` | Row banding mode: `"none"`, `"row"`, `"group"` (deepest level, default), or `"group-n"` (e.g. `"group-1"` for outermost). Overrides the banding on `theme`. Use the viewer's settings panel (gear icon) to adjust interactively. |
| `width` | character/numeric | `NULL` | Widget width (auto). |
| `height` | character/numeric | `NULL` | Widget height (auto). |
| `elementId` | `character(1)` | `NULL` | HTML element id. |

### Splitting

| Argument | Type | Default | Description |
|---|---|---|---|
| `split_by` | `character` | `NULL` | Column name(s) to split into one plot per unique value/combination, returning a `SplitForest` widget. |
| `shared_axis` | `logical(1)` | `FALSE` | When `split_by` is set, share an axis range across all sub-plots. |

### Other

| Argument | Type | Default | Description |
|---|---|---|---|
| `.spec_only` | `logical(1)` | `FALSE` | When `TRUE`, returns the `WebSpec` object instead of an htmlwidget — useful for programmatic manipulation before rendering. |

## Deprecated arguments

:::{.callout-important}
## Removed from the `tabviz()` surface in 0.9.0

The arguments below are still accepted for one cycle but emit `lifecycle::deprecate_warn()`. They have moved to dedicated, more discoverable homes.
:::

| Old argument | Replacement |
|---|---|
| `tabviz(axis_range = ...)` | Pass `axis_range` to `viz_forest()` on the forest column. |
| `tabviz(axis_ticks = ...)` | Pass `axis_ticks` to `viz_forest()` on the forest column. |
| `tabviz(axis_gridlines = ...)` | Pass `axis_gridlines` to `viz_forest()` on the forest column. |
| `tabviz(zoom = ...)` | Pipe to [`set_zoom(zoom = ...)`](styling.qmd#set_zoom). |
| `tabviz(auto_fit = ...)` | Pipe to `set_zoom(auto_fit = ...)`. |
| `tabviz(max_width = ...)` | Pipe to `set_zoom(max_width = ...)`. |
| `tabviz(max_height = ...)` | Pipe to `set_zoom(max_height = ...)`. |
| `tabviz(show_zoom_controls = ...)` | Pipe to `set_zoom(show_controls = ...)`. |
| `tabviz(weight = ...)` | Use `marker_size = ...`. |

## Return value

An htmlwidget. When `.spec_only = TRUE`, returns a `WebSpec` instead. When `split_by` is set, returns a split-table widget backed by a `SplitForest`.

## Grouping {#grouping}

Three modes of `group`:

```r
# Flat grouping
tabviz(data, group = "category", ...)

# Hierarchical grouping
tabviz(data, group = c("region", "country"), ...)

# Explicit group definitions
tabviz(data,
  group = list(
    web_group("cardio", label = "Cardiovascular"),
    web_group("renal",  label = "Renal Outcomes", collapsed = TRUE)
  ),
  ...)
```

## Examples

```{r}
meta <- data.frame(
  study = c("Smith 2020", "Jones 2021", "Lee 2022"),
  drug  = c("A", "A", "B"),
  hr    = c(0.72, 0.85, 0.91),
  lo    = c(0.55, 0.70, 0.75),
  hi    = c(0.95, 1.03, 1.10),
  n     = c(245, 180, 320)
)

# Minimal: forest column + a couple of data columns
tabviz(meta,
  label = "study",
  columns = list(
    col_text("drug", "Drug"),
    col_n("n"),
    viz_forest(point = "hr", lower = "lo", upper = "hi",
               scale = "log", null_value = 1, axis_label = "Hazard Ratio"),
    col_interval("hr", "lo", "hi", header = "HR (95% CI)")
  )
)
```

```{r}
# Row styling: bold significant rows, color the favourable ones green
tabviz(meta,
  label = "study",
  columns = list(
    viz_forest(point = "hr", lower = "lo", upper = "hi",
               scale = "log", null_value = 1)
  ),
  row_bold  = ~ hi < 1,
  row_color = ~ ifelse(hi < 1, "#16a34a", "#64748b")
)
```

```{r}
#| eval: false
# Spec-only: build, mutate fluently, then render
tabviz(meta,
  label = "study",
  columns = list(viz_forest(point = "hr", lower = "lo", upper = "hi")),
  .spec_only = TRUE
) |>
  set_theme("jama") |>
  set_marker_style(color = "drug") |>
  forest_plot()
```

## Related

- [`forest_plot()`](forest_plot.qmd) — convenience wrapper around `tabviz()` + `viz_forest()`
- [Columns](columns.qmd) — every `col_*()` and `viz_*()` helper
- [Styling modifiers](styling.qmd) — `set_row_style()`, `set_marker_style()`, `set_zoom()`
- [Themes](themes.qmd) — `web_theme_*()` builders and `set_*()` modifiers
- [Interaction](interaction.qmd) — `web_interaction()`
- [Export](export.qmd) — `save_plot()` for SVG / PNG / PDF
