---
title: "Themes"
---

```{r}
#| include: false
library(tabviz)
```

Built-in theme presets, the generic `web_theme()` constructor, fluent `set_*()` modifiers for tweaking a theme, and `package_themes()` for listing every shipped theme. Pass any theme to `tabviz(theme = ...)` or apply it after the fact with [`set_theme()`](styling.qmd#set_theme).

---

## Preset themes

`web_theme_*()` builders return a fully configured `WebTheme`. They take no arguments — chain `set_*()` modifiers to customize.

| Theme | Purpose | Notable defaults |
|---|---|---|
| [`web_theme_default()`](#web_theme_default) | Clean, modern starting point | Cyan accents, system font, rounded corners |
| [`web_theme_minimal()`](#web_theme_minimal) | Academic / publication ready | Pure B&W, Georgia serif, sharp corners |
| [`web_theme_dark()`](#web_theme_dark) | Dark mode | Catppuccin-Mocha palette, comfortable contrast |
| [`web_theme_jama()`](#web_theme_jama) | JAMA journals | Pure B&W, Arial 9pt, 18px rows, dense |
| [`web_theme_lancet()`](#web_theme_lancet) | Lancet journals | Lancet navy + gold, Georgia serif |
| [`web_theme_modern()`](#web_theme_modern) | Reports & dashboards | Vibrant blue, Inter font, 30px rows, 8px radius |
| [`web_theme_presentation()`](#web_theme_presentation) | Slides & posters | Oversized fonts, 12px markers, 36px rows |
| [`web_theme_cochrane()`](#web_theme_cochrane) | Cochrane reviews | Cochrane teal, Arial 12px, 20px rows |
| [`web_theme_nature()`](#web_theme_nature) | Nature journals | Nature blue, Helvetica Neue, 22px rows |

```r
tabviz(data, label = "study",
       columns = list(viz_forest("hr", "lo", "hi")),
       theme = web_theme_jama())
```

### `web_theme_default()` {#web_theme_default}
### `web_theme_minimal()` {#web_theme_minimal}
### `web_theme_dark()` {#web_theme_dark}
### `web_theme_jama()` {#web_theme_jama}
### `web_theme_lancet()` {#web_theme_lancet}
### `web_theme_modern()` {#web_theme_modern}
### `web_theme_presentation()` {#web_theme_presentation}
### `web_theme_cochrane()` {#web_theme_cochrane}
### `web_theme_nature()` {#web_theme_nature}

All preset builders take no arguments and return a `WebTheme`.

---

## `package_themes()` {#package_themes}

Returns a named list of every theme distributed with the package — handy for the `enable_themes` argument of [`web_interaction()`](interaction.qmd) or for iterating in tests.

```r
package_themes()
# $default, $minimal, $dark, $jama, $lancet, $modern, $presentation, $cochrane, $nature
```

---

## `web_theme()` {#web_theme}

Generic constructor: extend a base theme with a list of overrides per slot. Every override list mirrors the corresponding `set_*()` modifier's arguments.

```r
web_theme(
  name = "custom",
  colors = NULL,
  typography = NULL,
  spacing = NULL,
  shapes = NULL,
  axis = NULL,
  layout = NULL,
  group_headers = NULL,
  base_theme = web_theme_default()
)
```

| Argument | Type | Default | Description |
|---|---|---|---|
| `name` | `character(1)` | `"custom"` | Theme display name. |
| `colors` | named list | `NULL` | Overrides for the `ColorPalette` slot — see [`set_colors()`](#set_colors). |
| `typography` | named list | `NULL` | Overrides — see [`set_typography()`](#set_typography). |
| `spacing` | named list | `NULL` | Overrides — see [`set_spacing()`](#set_spacing). |
| `shapes` | named list | `NULL` | Overrides — see [`set_shapes()`](#set_shapes). |
| `axis` | named list | `NULL` | Overrides — see [`set_axis()`](#set_axis). |
| `layout` | named list | `NULL` | Overrides — see [`set_layout()`](#set_layout). |
| `group_headers` | named list | `NULL` | Overrides — see [`set_group_headers()`](#set_group_headers). |
| `base_theme` | `WebTheme` | `web_theme_default()` | Theme to inherit from. |

```r
my_theme <- web_theme(
  name = "house_style",
  colors = list(primary = "#1a365d", ci_marker_fill = "#1a365d"),
  typography = list(font_family = "Georgia, serif"),
  spacing = list(row_height = 22),
  base_theme = web_theme_minimal()
)
```

---

## Theme modifiers (`set_*()`)

Each modifier takes a `WebTheme`, returns a `WebTheme`, and composes with `|>`. Every argument is optional — unset arguments leave the existing value alone.

### `set_colors()` {#set_colors}

```r
set_colors(
  theme,
  background = NULL, foreground = NULL,
  primary = NULL, secondary = NULL,
  accent = NULL, muted = NULL,
  border = NULL,
  row_bg = NULL, alt_bg = NULL,
  ci_marker_fill = NULL,
  ci_line = NULL,
  summary_fill = NULL,
  summary_border = NULL
)
```

| Argument | Default | Description |
|---|---|---|
| `background` | `"#ffffff"` | Plot/container background. |
| `foreground` | `"#333333"` | Default text color. |
| `primary` | `"#0891b2"` | Primary accent. Cascades into `ci_marker_fill` and `summary_fill` if those are unset. |
| `secondary` | `"#64748b"` | Secondary text/UI. |
| `accent` | `"#8b5cf6"` | Palette slot referenced by the `accent` semantic bundle. |
| `muted` | `"#94a3b8"` | Palette slot referenced by the `muted` semantic bundle; also used for footnotes / captions. |
| `border` | `"#e2e8f0"` | Row dividers, container borders. |
| `row_bg` | inherits `background` | Even-row background. |
| `alt_bg` | inherits `row_bg` | Odd-row background (zebra stripes). |
| `ci_marker_fill` | `"#0891b2"` | Default CI marker fill. Cascades into `summary_fill`. Preempted by `shapes.effect_colors[0]` when that is set. |
| `ci_line` | `"#475569"` | Confidence-interval line color. |
| `summary_fill` | inherits `ci_marker_fill` | Summary diamond fill (per-row `row_type = "summary"` and the legacy `spec.data.overall` path). |
| `summary_border` | `"#0e7490"` | Summary diamond border. |

:::{.callout-important}
## CI color arguments were renamed in 0.9.0

The `interval*` family was renamed to `ci_marker_*` / `ci_line` for clarity.

| Old | New |
|---|---|
| `interval` | `ci_marker_fill` |
| `interval_line` | `ci_line` |

The old names still work for one cycle and emit `lifecycle::deprecate_warn()`.
:::

:::{.callout-note}
## Direction-aware markers were removed in 0.13.0

`ci_marker_positive`, `ci_marker_negative`, `ci_marker_neutral` (and the matching `interval_*` names) were never rendered — they were emitted as CSS vars but no code read them. The arguments are still accepted for back-compat and emit a deprecation warning. To color markers by direction, map a color column via `marker_color` in `viz_forest()`, or use per-effect literal colors.
:::

:::{.callout-tip}
## Cascading defaults
Setting `primary` automatically updates `ci_marker_fill` and `summary_fill` unless those are passed explicitly. Setting `background` cascades into `row_bg` and `alt_bg`. Override individually to break the cascade.
:::

```r
web_theme_default() |>
  set_colors(
    primary = "#0b3d91",
    ci_marker_fill = "#0b3d91",
    summary_fill = "#16a34a"
  )
```

### `set_typography()` {#set_typography}

```r
set_typography(
  theme,
  font_family = NULL,
  font_size_sm = NULL, font_size_base = NULL, font_size_lg = NULL,
  font_weight_normal = NULL, font_weight_medium = NULL, font_weight_bold = NULL,
  line_height = NULL,
  header_font_scale = NULL
)
```

| Argument | Default | Description |
|---|---|---|
| `font_family` | system stack | CSS font-family string. |
| `font_size_sm` | `"0.75rem"` | Small text. |
| `font_size_base` | `"0.875rem"` | Base text. |
| `font_size_lg` | `"1rem"` | Large text. |
| `font_weight_normal` | `400` | |
| `font_weight_medium` | `500` | |
| `font_weight_bold` | `600` | |
| `line_height` | `1.5` | Line-height multiplier. |
| `header_font_scale` | `1.05` | Header size relative to base. |

### `set_spacing()` {#set_spacing}

```r
set_spacing(
  theme,
  row_height = NULL, header_height = NULL,
  padding = NULL,
  container_padding = NULL, axis_gap = NULL,
  group_padding = NULL,
  cell_padding_x = NULL, cell_padding_y = NULL
)
```

| Argument | Default | Description |
|---|---|---|
| `row_height` | `24` | Data row height (px). |
| `header_height` | `32` | Header row height (px). |
| `padding` | `12` | Padding around the forest plot SVG (px). |
| `container_padding` | `0` | Outer container padding (px). |
| `axis_gap` | `12` | Gap between table content and axis (px). |
| `group_padding` | `8` | Column-group header horizontal padding (px). Note: applies to column-group headers, not row-group headers. |
| `cell_padding_x` | `10` | Horizontal cell padding (px). |
| `cell_padding_y` | `4` | Vertical cell padding (px). |

### `set_shapes()` {#set_shapes}

```r
set_shapes(
  theme,
  point_size = NULL, summary_height = NULL,
  line_width = NULL, border_radius = NULL,
  effect_colors = NULL,
  marker_shapes = NULL
)
```

| Argument | Default | Description |
|---|---|---|
| `point_size` | `6` | Marker radius (px). |
| `summary_height` | `10` | Summary diamond height (px). |
| `line_width` | `1.5` | CI line stroke width (px). |
| `border_radius` | `2` | Container/row rounding (px). |
| `effect_colors` | theme cycle | Colors used by multi-effect viz columns. |
| `marker_shapes` | `c("square","circle","diamond","triangle")` | Shape sequence for multi-effect forests. |

### `set_effect_colors()` {#set_effect_colors}

Shorthand for `set_shapes(effect_colors = ...)`.

```r
set_effect_colors(theme, colors)
```

### `set_marker_shapes()` {#set_marker_shapes}

Shorthand for `set_shapes(marker_shapes = ...)`.

```r
set_marker_shapes(theme, shapes)
```

### `set_axis()` {#set_axis}

```r
set_axis(
  theme,
  range_min = NULL, range_max = NULL,
  tick_count = NULL, tick_values = NULL,
  gridlines = NULL, gridline_style = NULL,
  ci_clip_factor = NULL,
  include_null = NULL, symmetric = NULL,
  null_tick = NULL,
  marker_margin = NULL
)
```

| Argument | Default | Description |
|---|---|---|
| `range_min`, `range_max` | `NA` | Axis endpoints. `NA` = auto-computed. |
| `tick_count` | `NA` | Target number of ticks. |
| `tick_values` | `NULL` | Explicit tick positions (overrides `tick_count`). |
| `gridlines` | `FALSE` | Show vertical gridlines. |
| `gridline_style` | `"dotted"` | `"solid"`, `"dashed"`, or `"dotted"`. |
| `ci_clip_factor` | `2.0` | Clip CIs extending beyond this multiple of the estimate range; `Inf` disables clipping. |
| `include_null` | `TRUE` | Always include null value in axis range. |
| `symmetric` | `NULL` | Force symmetric range around `null_value`. |
| `null_tick` | `TRUE` | Always show a tick at the null value. |
| `marker_margin` | `TRUE` | Add half-marker-width padding at edges. |

### `set_layout()` {#set_layout}

```r
set_layout(
  theme,
  plot_width = NULL,
  container_border = NULL,
  container_border_radius = NULL,
  banding = NULL
)
```

| Argument | Default | Description |
|---|---|---|
| `plot_width` | `"auto"` | `"auto"` or pixel width for the forest column. |
| `container_border` | `FALSE` | Show outer container border. |
| `container_border_radius` | `8` | Outer corner radius (px). Visible when `container_border = TRUE`. |
| `banding` | `"group"` | Row banding mode: `"none"`, `"row"`, `"group"` (default — alternates at deepest group level; falls back to `"row"` with no groups), or `"group-n"` (e.g. `"group-1"` for outermost). Uses `row_bg` / `alt_bg`. Also available as `tabviz(banding=)` with the same grammar. |

`set_layout()` also accepts `cell_padding_x` / `cell_padding_y` (deprecated — forwards to `set_spacing()`) and the long-removed `row_border` / `row_border_style` (silently ignored with a warning).

:::{.callout-note}
## `plot_position` / `table_width` / `section_gap` / `column_gap` removed in 0.13.0
These fields never reached any rendering code — they were declared, serialized, and exposed in the UI, but no frontend consumer read them. They have been dropped from the theme classes. Column order in `columns = list(...)` controls plot placement; let column widths auto-size from content, or use `plot_width` to pin the forest column.
:::

### `set_group_headers()` {#set_group_headers}

Per-level styling for nested row-group headers (level 1 = outermost).

```r
set_group_headers(
  theme,
  level1_font_size = NULL, level1_font_weight = NULL,
  level1_italic = NULL, level1_background = NULL, level1_border_bottom = NULL,
  level2_font_size = NULL, level2_font_weight = NULL,
  level2_italic = NULL, level2_background = NULL, level2_border_bottom = NULL,
  level3_font_size = NULL, level3_font_weight = NULL,
  level3_italic = NULL, level3_background = NULL, level3_border_bottom = NULL,
  indent_per_level = NULL
)
```

| Argument | Default | Description |
|---|---|---|
| `levelN_font_size` | level-dependent | Font size for level N (1, 2, 3). |
| `levelN_font_weight` | 700 / 500 / 400 | |
| `levelN_italic` | `FALSE` / `TRUE` / `FALSE` | |
| `levelN_background` | `NULL` (auto from `primary`) | Header background. |
| `levelN_border_bottom` | `FALSE` | Bottom border. |
| `indent_per_level` | `16` | Per-level indent (px). |

---

## Examples

```{r}
#| eval: false
# Chain modifiers
web_theme_jama() |>
  set_colors(primary = "#0066cc") |>
  set_typography(font_family = "Helvetica, Arial, sans-serif",
                 font_size_base = "10pt") |>
  set_spacing(row_height = 22) |>
  set_axis(gridlines = TRUE)

# Build a custom theme from scratch
publication_theme <- web_theme_minimal() |>
  set_colors(foreground = "#1a1a1a", ci_marker_fill = "#333333", ci_line = "#333333") |>
  set_typography(font_family = "'Times New Roman', serif", font_size_base = "11pt") |>
  set_spacing(row_height = 24)

# Distinct visual hierarchy on nested groups
web_theme_default() |>
  set_group_headers(
    level1_font_weight = 800,
    level1_background = "#e0f2fe",
    level2_italic = TRUE,
    level2_background = "#f0f9ff",
    indent_per_level = 24
  )
```

## Related

- [`set_theme()`](styling.qmd#set_theme) — apply a theme to a `WebSpec` / widget / proxy
- [`selectable_themes()`](styling.qmd#selectable_themes) — curate the in-widget theme switcher
- [`web_interaction()`](interaction.qmd) — `enable_themes` argument
- [`tabviz()`](tabviz.qmd) — `theme` constructor argument
