---
title: "webtable"
---

```{r}
#| include: false
library(webforest)
```

Create a table-only widget without the forest plot visualization.

## Usage

```r
webtable(
  x,
  ...,
  width_mode = c("fit", "fill", "responsive"),
  height_mode = c("auto", "scroll"),
  width = NULL,
  height = NULL,
  elementId = NULL
)
```

## Arguments

| Argument | Description |
|----------|-------------|
| `x` | A data frame or WebSpec object |
| `...` | Arguments passed to `web_spec()` when x is a data frame |
| `width_mode` | `"fit"`, `"fill"`, or `"responsive"` |
| `height_mode` | `"auto"` or `"scroll"` |
| `width` | Widget width |
| `height` | Widget height |
| `elementId` | HTML element ID |

## Value

An htmlwidget object displaying the table portion without the forest plot.

## Examples

### Basic Table

```{r}
data <- data.frame(
  study = c("Study A", "Study B", "Study C"),
  hr = c(0.72, 0.85, 0.91),
  lower = c(0.55, 0.70, 0.75),
  upper = c(0.95, 1.03, 1.10),
  n = c(245, 180, 320),
  pval = c(0.018, 0.092, 0.31)
)

webtable(data,
  point = "hr", lower = "lower", upper = "upper",
  label = "study",
  columns = list(
    col_numeric("n", "N"),
    col_interval("HR (95% CI)"),
    col_pvalue("pval", "P-value")
  )
)
```

## See Also

- [forest_plot()](forest_plot.qmd) for full forest plots
- [Columns](columns.qmd) for column types
